### 2018-05-22 inst/doc/command-line-example.Rmd

---
title: "LDATS Example"
author: "Juniper L. Simonis"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r, include=FALSE}
library(LDATS)
vers <- packageVersion("LDATS")
today <- Sys.Date()
```

This vignette walks through an example of **LDATS** at the command line and
was constructed using **LDATS** version `r vers` on `r today`.

## Installation

To obtain the most recent version of **LDATS**, install the most recent 
version from GitHub:

```{r, eval=FALSE}
install.packages("devtools")
devtools::install_github("weecology/LDATS")
```

## Data

For this vignette, we will be using rodent data from the control plots of the
[Portal Project](https://github.com/weecology/portaldata), which come with
the **LDATS** package

```{r}
data(rodents)
rodents
```

### 2018-06-22 previous_work/paper/markdown_analyses.Rmd
updating the previous work folder

---
title: "LDA kratplots results"
author: "Renata Diaz"
date: "5/8/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(topicmodels)
library(RCurl)
library(multipanelfigure)
library(reshape2)
library(LDATS)

source('rodent_data_for_LDA.r')
source('rodent_data_from_portalr.R')
source('AIC_model_selection.R')
source('LDA_figure_scripts.R')
source('changepointmodel.r')
source('LDA-distance.R')

```

## Eight plots, starting in 1988

### Prepare rodent data

```{r}

dat = get_exclosure_rodents(cont_or_exclosures = 'control', time_or_plots = 'plots')

# dates to go with count data
moondat = read.csv(text=getURL("https://raw.githubusercontent.com/weecology/PortalData/master/Rodents/moon_dates.csv"),stringsAsFactors = F)
moondat$date = as.Date(moondat$censusdate)

period_dates = filter(moondat,period %in% dat$period) %>% select(period,date)
dates = period_dates$date

dat = dat[,2:22]

```

### Select number of topics
```{r}
# ==================================================================
# 2a. select number of topics
# ==================================================================
Select LDA model using LDATS

# Changepoint models

# set up parameters for model
year_continuous = 1970 + as.integer(julian(dates)) / 365.25
x = data.frame(
  year_continuous = year_continuous,
  sin_year = sin(year_continuous * 2 * pi),
  cos_year = cos(year_continuous * 2 * pi)
)

# run models with 1, 2, 3, 4, 5 changepoints
cp_results_rodent_3topics = changepoint_model(ldamodel3, x, 1, weights = rep(1,length(year_continuous)))
cp_results_rodent2_3topics = changepoint_model(ldamodel3, x, 2, weights = rep(1,length(year_continuous)))
cp_results_rodent3_3topics = changepoint_model(ldamodel3, x, 3, weights = rep(1,length(year_continuous)))
cp_results_rodent4_3topics = changepoint_model(ldamodel3, x, 4, weights = rep(1,length(year_continuous)))
cp_results_rodent5_3topics = changepoint_model(ldamodel3, x, 5, weights = rep(1,length(year_continuous)))

# Save changepoint models
save(cp_results_rodent_3topics,
     cp_results_rodent2_3topics, 
     cp_results_rodent3_3topics, cp_results_rodent4_3topics, 
     cp_results_rodent5_3topics, 
     file = 'models/cp_8plots_3topics.Rdata')

# change point model selection
# mean deviance ( -2 * log likelihood) + 2*(#parameters)
mean(cp_results_rodent_3topics$saved_lls * -2) + 2*(3*(ntopics-1)*(1+1)+(1))
mean(cp_results_rodent2_3topics$saved_lls * -2)+ 2*(3*(ntopics-1)*(2+1)+(2))
mean(cp_results_rodent3_3topics$saved_lls * -2)+ 2*(3*(ntopics-1)*(3+1)+(3))
mean(cp_results_rodent4_3topics$saved_lls * -2)+ 2*(3*(ntopics-1)*(4+1)+(4))
mean(cp_results_rodent5_3topics$saved_lls * -2)+ 2*(3*(ntopics-1)*(5+1)+(5))

# 3 topics is best supported.

rm(list= c(cp_results_rodent2_3topics, cp_results_rodent_3topics,
           cp_results_rodent4_3topics, cp_results_rodent5_3topics))

# some quick histograms of changepoint model results
hist(year_continuous[cp_results_rodent3_3topics$saved[,1,]],breaks = seq(1988,2016,.25),xlab='',main='Changepoint Estimate')
annual_hist(cp_results_rodent3_3topics,year_continuous)

# turn changepoint results into data frame
df_3_3topics = as.data.frame(t(cp_results_rodent3_3topics$saved[,1,])) %>% melt()
df_3_3topics$value = year_continuous[df_3_3topics$value]

# find 95% confidence intervals on each changepoint:
quantile(df_3_3topics[df_3_3topics$variable=='V1','value'],probs=c(.025,.975)) %>% date_decimal() %>% format('%d-%m-%Y')
quantile(df_3_3topics[df_3_3topics$variable=='V2','value'],probs=c(.025,.975)) %>% date_decimal() %>% format('%d-%m-%Y')
quantile(df_3_3topics[df_3_3topics$variable=='V3','value'],probs=c(.025,.975)) %>% date_decimal() %>% format('%d-%m-%Y')

write.csv(df_3_3topics, file = 'models/df_3_3topics.csv', row.names =F)

```

#### Eight plots starting in 1988, 4 topics
```{r}
SEED_4 = unlist(best_seed_4)[1]

ntopics = 4

ldamodel4 = LDA(dat,ntopics, control = list(seed = SEED_4),method='VEM')

```

### 2018-06-22 previous_work/paper/NDVI_figure.R

@@ -1,0 +1,32 @@
+
+# Plot NDVI over time to show drought
+library(RCurl)
+library(dplyr)
+library(ggplot2)
+
+ndvi = read.csv('Monthly_Landsat_NDVI.csv',
+                   na.strings=c(""), stringsAsFactors = FALSE)
+
+ndvi$Date = as.Date(ndvi$Date,format='%Y-%M')
+ndvi$NDVI = as.numeric(ndvi$NDVI)
+ndvi$year = format(ndvi$date,'%Y') %>% as.numeric()
+
+plot(ndvi$date,ndvi$NDVI)
+lines(ndvi$date,ndvi$NDVI)
+
+
+# ===============
+# yearly avg - from 1984 there are >12 images per year (should go back and remove excessively cloudy images too)
+
+ndviyr = aggregate(ndvi$NDVI,by=list(year = ndvi$year),FUN=mean,na.rm=T) %>% filter(year>1983)
+
+
+# ggplot version - with long term mean
+ggplot(ndviyr,aes(x=as.integer(year),y=x)) +
+  geom_line(size=1.5) +  
+  geom_point(size=3) +
+  geom_hline(yintercept = mean(ndviyr$x),linetype=2) +
+  labs(x='',y='NDVI (yearly mean)') +
+  theme(axis.text=element_text(size=14),
+        axis.title=element_text(size=14))
+  

### 2018-06-22 previous_work/paper/abundance_plots.r

@@ -1,54 +1,67 @@
+library(fitdistrplus)
+library(stats)
+library(dplyr)
+library(RCurl)
+library(ggplot2)
 # scripts for making supporting figures in the extreme events/ LDA project
 
-# Figure 5 in manuscript
+
 
 # ==============================================
-# total abundance
+# Abundance figure in manuscript (Figure 2)
 
-dat = read.csv('inst/extdata/Rodent_table_dat.csv',na.strings = '',as.is=T)
-rdat = read.csv('~/PortalData/Rodents/Portal_rodent.csv',as.is=T)
-trappingdat = read.csv('~/PortalData/Rodents/Portal_rodent_trapping.csv',as.is=T,na.strings = '')
+# total abundance
+dat = read.csv('Rodent_table_dat.csv',na.strings = '',as.is=T)
+
+# find censuses that were complete
+trappingdat = read.csv(text=getURL("https://raw.githubusercontent.com/weecology/PortalData/master/Rodents/Portal_rodent_trapping.csv"),
+                       na.strings=c(""), as.is=T, stringsAsFactors = FALSE)
 trapdat = aggregate(trappingdat$sampled,by=list(period=trappingdat$period),FUN=sum)
 fullcensus = trapdat[trapdat$x>20,]
-perioddates = read.csv('~/PortalData/Rodents/moon_dates.csv',as.is=T,na.strings = '')
+perioddates = read.csv(text=getURL('https://raw.githubusercontent.com/weecology/PortalData/master/Rodents/moon_dates.csv'),as.is=T,na.strings = '')
 perioddates$censusdate = as.Date(perioddates$censusdate)
 fullcensus = merge(fullcensus,perioddates)
 
-# data frame
+# data frame of abundance by period - restricted to only complete censuses
 abund_dat = data.frame(period = 1:436, n = rowSums(dat))
 abund_dat = merge(abund_dat,fullcensus[,c('period','censusdate')])
+abund_dat$density = abund_dat$n/2 # density is in rodents/hectare: each plot is 1/4 hectare, there are 8 control plots, so total area covered by control plots is 2 hectares
 
-# plot
[…]

### 2018-06-22 previous_work/paper/earlier version/AIC_model_selection.R

@@ -1,0 +1,133 @@
+#' AIC model selection for LDA using VEM method
+#' 
+#' Runs LDA using different numbers of topics and VEM method, calculates AIC 
+#'   values for comparison
+#' 
+#' @param dat Table of integer data (species counts by period)
+#' @param SEED set seed to keep LDA model runs consistent (default 2010)
+#' @param topic_min lowest number of topics; must be >=2
+#' @param topic_max highest number of topics
+#' 
+#' @return data frame of number of topics (k) and aic value (aic)
+#' 
+#' @examples aic_values = aic_model(dat,2010,2,10)
+#' @export 
+
+aic_model = function(dat,SEED,topic_min,topic_max) {
+
+  aic_values <- data.frame(seed = integer(), k = integer(), aic = numeric())
+
+  for (k in seq(topic_min,topic_max)) {
+
+    # run LDA
+
+      VEM <- topicmodels::LDA(dat, k, control = list(seed = SEED),
+                              method = 'VEM')
+  
+    # get parameter estimates
+
+      z <- modeltools::posterior(VEM)
+      commun.plot <- z$topics
+      commun.spp <- z$terms
+  
+    # calculate AIC
+
+      # extract estimate of maximum loglikelihood
+
+        max.logl <- sum(VEM@loglikelihood) 
+
+      # number of parameters       
+
+      nparam <- (nrow(commun.plot) - 1) * (ncol(commun.plot)) + 
+                 nrow(commun.spp) * (ncol(commun.spp) - 1) 
+
+    aic <- 2*nparam - 2*max.logl   #aic calculation
+    aic_values <- rbind(aic_values,data.frame(SEED, k, aic))
+  }
+
+  return(aic_values)
+}
+
+
+
+
+
+#' AIC model selection for LDA using Gibbs sampler method
+#' 
+#' Runs LDA using different numbers of topics and Gibbs method, calculates 
+#'   AIC values for comparison
+#' 
+#' @param dat Table of integer data (species counts by period)
+#' @param ngibbs number of iterations for gibbs sampler -- must be greater 
+#'    than 200 (default 1000)
+#' @param topic_min lowest number of topics; must be >=2
+#' @param topic_max highest number of topics
+#' @param save_runs T/F whether to save each run of LDA for later retrieval 
+#'   (these take a long time to run)
+#' 
[…]

### 2018-06-22 previous_work/paper/earlier version/LDA-distance.R

@@ -1,8 +1,3 @@
-# This script assumes that `dat` is defined
-
-library(topicmodels)
-
-
 #' Compute hellinger distance
 #' 
 #' Hellinger distance is Euclidean distance between square roots, divided by
@@ -10,6 +5,7 @@ library(topicmodels)
 #' 
 #' @param a,b numeric vectors
 #' 
+#' @export 
 Hellinger = function(a, b){
   diff = sqrt(a) - sqrt(b)
   sqrt(sum(diff^2) / 2)
@@ -21,6 +17,7 @@ Hellinger = function(a, b){
 #' @param p1,p2 numeric vectors: species composition
 #' @param k number of topics
 #' 
+#' @export 
 min_H = function(p1, p2, k) {
   # Find the cost associated with each pairwise topic assignment
   costs = matrix(0, k, k)
@@ -46,20 +43,21 @@ min_H = function(p1, p2, k) {
 #'
 #' @param ldas list of results from LDA model
 #' @param seeds vector of seeds for which LDA models will be fit and results compared
-#' @param k number of topics
 #' 
 #' @return list including: seed associated with highest log likelihood
 #'                         mean distance between best model and all others
 #'                         max distuance between best model and worst
 #'
-calculate_LDA_distance = function(ldas,seeds,k) {
+#' @export 
+calculate_LDA_distance = function(ldas,seeds) {
   
-  # Calculate a bunch of LDAs with k topics
+  # Calculate a bunch of LDAs with 4 topics
+  k = 4
   ldas = purrr::map(seeds, 
-                    ~LDA(dat, k = k, method = "VEM", control = list(seed = .x)))
+                    ~ topicmodels::LDA(dat, k = k, method = "VEM", control = list(seed = .x)))
   
   # Log-likelihoods for each lda
-  lls = purrr:::map_dbl(ldas, logLik)
+  lls = purrr:::map_dbl(ldas, topicmodels::logLik)
   
   # Pick the LDA model with the highest log-likelihood
   best_lda = which.max(lls)
@@ -104,8 +102,8 @@ calculate_LDA_distance = function(ldas,seeds,k) {
   
   # between community distance
   comm_dist = c()
-  for (i in 1:(k-1)) {
-    for (j in (i+1):k) {
+  for (i in 1:3) {
+    for (j in (i+1):4) {
       comm_dist = append(comm_dist,Hellinger(best_ps[i,],best_ps[j,]))
     }
   }

### 2018-06-22 previous_work/paper/earlier version/LDA_figure_scripts.R

@@ -1,33 +1,19 @@
-# Scripts for making figures from LDA analyses
-
-library(ggplot2)
-library(gridExtra)
-library(dplyr)
-
-
-#cbPalette <- c( "#E69F00","#999999", "#56B4E9", "#0072B2", "#D55E00", "#F0E442", "#009E73", "#CC79A7")
-cbPalette <- c( "#e19c02","#999999", "#56B4E9", "#0072B2", "#D55E00", "#F0E442", "#009E73", "#CC79A7")
-
-
-
 
 #' Plot gamma
-#' 
+#'
 #' This plotting function plots the component communities over time
 #' It's used by the plot_component_communities function, but can also take
 #' any data frame as input as long as it's the right form -- used by the simulations
-#' 
+#'
 #' @param gamma_frame a data frame containing columns for date, relabund, and community
 #' @param ntopics number of topics
 #' @param ylab label for y axis (optional)
-#' 
+#'
 #' @return a ggplot object
-
-ltypes = c('solid','longdash','longdash','solid','solid')
-
+#' @export 
 
 plot_gamma = function(gamma_frame,ntopics,ylab='',colors=cbPalette) {
-  g = ggplot(gamma_frame, aes(x=date,y=relabund,colour=community)) + 
+  g = ggplot(gamma_frame, aes(x=date,y=relabund,colour=community)) +
     geom_point() +
     geom_line(aes(size=community)) +
     scale_size_manual(values=c(1,1,1.5,1.5,1), guide=FALSE) +
@@ -47,37 +33,39 @@ plot_gamma = function(gamma_frame,ntopics,ylab='',colors=cbPalette) {
 
 
 #' Plot component communities
-#' 
+#'
 #' Plots timeseries of component communities (topics)
-#' 
+#'
 #' @param ldamodel object of class LDA_VEM created by the function LDA in topicmodels package
 #' @param ntopics number of topics used in ldamodel
 #' @param xticks vector of dates for x-axis labels
 #' @param ylab y axis label (optional)
 #' @param topic_order order of topics (for color control)
-#' 
+#'
 #' @return ggplot object
-#' 
-#' @example plot_component_communities(ldamodel,ntopics,period_dates$date)
+#'
+#' @examples plot_component_communities(ldamodel,ntopics,period_dates$date)
+#' @export 
 
 plot_component_communities = function(ldamodel,ntopics,xticks,ylab='',topic_order = seq(ntopics),colors = cbPalette) {
-  
+
   z = posterior(ldamodel)
[…]

### 2018-06-22 previous_work/paper/readResults.R

@@ -1,0 +1,27 @@
+# read rodent changepoint/lda results without running the changepoint model again
+
+everything <- readRDS('/Users/renatadiaz/Documents/GitHub/Extreme-events-LDA/exclosuresResults.rds')
+dat <- everything[[1]]
+seeds <- everything[[2]]
+best_ntopic <- everything[[3]]
+seeds_3topics <- everything[[4]]
+best_seed <- everything[[5]]
+mean_dist <- everything[[6]]
+max_dist <- everything[[7]]
+ntopics <- everything[[8]]
+SEED <- everything[[9]]
+ldamodel <- everything[[10]]
+year_continuous <- everything[[11]]
+x <- everything[[12]]
+cp_results_rodent <- everything[[13]]
+cp_results_rodent2 <- everything[[14]]
+cp_results_rodent3 <- everything[[15]]
+cp_results_rodent4 <- everything[[16]]
+cp_results_rodent5 <- everything[[17]]
+df_1 <- everything[[18]]
+df_2 <- everything[[19]]
+df_3 <- everything[[20]]
+df_4 <- everything[[21]]
+df_5 <- everything[[22]]
+dates <- everything[[23]]
+rm(everything)

### 2018-06-22 previous_work/paper/rodent_LDA_analysis_exclosures.R

@@ -11,54 +11,72 @@
 
 library(topicmodels)
 library(RCurl)
-library(multipanelfigure)
+library(multipanelfigure) 
 library(reshape2)
-
-
-source('previous-work/rodent_data_for_LDA.r')
-source('rodent_data_from_portalr.R')
+library(portalr)
+library(dplyr)
+library(tidyr)
+#source('rodent_data_for_LDA.r')
+source('RodentAbundancesAdjustable.R')
+# try this
+# source('/Users/renatadiaz/Documents/GitHub/portalr/R/RodentAbundancesAdjustable.R') # not sure how to source the thing I added to portalr
 source('AIC_model_selection.R')
 source('LDA_figure_scripts.R')
 source('changepointmodel.r')
 source('LDA-distance.R')
 
+
+# to skip running the changepoint and just load results
+#source('readResults.R')
+
 # ===================================================================
-# 1. prepare rodent data
+# 1. prepare rodent data -- needs to be adjusted to work with new changes to portalr
 # ===================================================================
-# for controls:
-dat = create_rodent_table(period_first = 1,
-                          period_last = 436,
-                          selected_plots = c(2,4,8,11,12,14,17,22),
-                          selected_species = c('BA','DM','DO','DS','NA','OL','OT','PB','PE','PF','PH','PI','PL','PM','PP','RF','RM','RO','SF','SH','SO'))
+# 
+# dat.full <- abundance.adjustable(path = 'repo', level="treatment.adj",type="Rodents",
+#                                              length="longterm",unknowns=F,incomplete=T,
+#                                              shape="list",time="period", dates = T)
+# 
+# 
+# 
+# period_first = 1
+# period_last = 436
+# selected_treatment = 'exclosure'
+# 
+# dat = dat.full %>% 
+#   filter(period >= period_first, period <= period_last, 
+#                 treatment == selected_treatment) %>% 
+#   mutate(usual.n = (ceiling(mean(n)))) %>% 
+#   mutate(abundance.adj = round(abundance.perplot * usual.n)) %>% 
[…]

### 2018-06-22 previous_work/paper/rodent_data_from_portalr.R

@@ -1,0 +1,61 @@
+library(tidyverse)
+library(portalr)
+
+get_exclosure_rodents = function(cont_or_exclosures = 'control', time_or_plots = 'time') {
+  if (cont_or_exclosures == 'exclosure') {
+  
+  if (tolower(time_or_plots) == 'plots') {
+    length = 'all'
+    startperiod = 118
+    standardeffort = 8
+  }
+  if(tolower(time_or_plots) == 'time') {
+    length = 'longterm'
+    startperiod = 1
+    standardeffort = 4
+  }
+  }
+  if(cont_or_exclosures == 'control') {
+    length = 'all'
+    startperiod = 1
+    standardeffort = 8
+  }
+  
+  
+  dat <- abundance(path = "repo", clean = FALSE, 
+                   level = 'Plot', type = "Rodents", length = length, 
+                   unknowns = FALSE, fill_incomplete = F, shape = 'crosstab',
+                   time = 'period', effort = TRUE, min_plots = 0)
+  
+  
+  if (cont_or_exclosures == 'exclosure') {
+    dat2 <- dat %>%
+    filter(treatment == cont_or_exclosures, period %in% startperiod:436,
+           ntraps >= 1) %>%
+    mutate(effort = 1) %>%
+    group_by(period) %>%
+    summarise_at(c(colnames(dat)[5:25], 'effort'), sum)
+  } else {
+    dat2 <- dat %>%
+      filter(plot %in% c(2,4,8,11,12,14,17,22), 
+             period %in% startperiod:436,
+             ntraps >= 1) %>%
+      mutate(effort = 1) %>%
+      group_by(period) %>%
+      summarise_at(c(colnames(dat)[5:25], 'effort'), sum)
+  }
+  
+  datsums = vector(length =nrow(dat2))
+  
+  for(i in 1:nrow(dat2)) {
+    thiseffort = dat2[i, 'effort']
+    for (j in 2:22) {
+      dat2[i,j] = round((dat2[i, j] / thiseffort) * standardeffort)
+    }
+    datsums[i] = sum(dat2[i,2:22])
+  }
+  
+  dat2 = dat2[ which(datsums >= 1), 1:22]
+
+  return(dat2)
+}

### 2018-09-11 R/LDA.R
aligning notation and tidying code
working through from front to back
LDA functions now all included
need to include tests

@@ -1,0 +1,120 @@
+#' @title Run a set of Latent Dirichlet Allocation models
+#' 
+#' @description For a given dataset (counts of words across several 
+#'   documents), conduct multiple Latent Dirichlet Allocation (LDA) models
+#'   (using the Variational Expectation Maximization (VEM) algorithm) to 
+#'   account for [1] uncertainty in the number of latent topics and
+#'   [2] the impact of intial values in the estimation procedure 
+#'   
+#' @param MV Matrix of observation count data (rows: documents (\code{M}), 
+#'   columns: terms (\code{V})).
+#'
+#' @param topics Vector of the number of topics to evaluate.
+#'
+#' @param nseeds Integer number of seeds (replicate starts) to use for each 
+#'   value of \code{ntopics}.
+#'
+#' @param control Named list of control parameters to be used in 
+#'   \code{\link[topicmodels]{LDA}} (note that "seed" will be overwritten).
+#' 
+#' @return List (class: "\code{LDA_list}") of LDA models (class: 
+#'   "\code{LDA}").
+#' 
+#' @examples 
+#' \dontrun{
+#'   data(rodents)
+#'   lda_data <- dplyr::select(rodents, -c(newmoon, date, plots, traps))
+#'   r_LDA <- LDA_set(MV = lda_data, topics = 2, nseeds = 2)                         
+#' }
+#' 
+#' @export
+#'
+LDA_set <- function(MV, topics = 2, nseeds = 1, control = NULL){
+
+  check_MV(MV)
+  check_topics(topics)
+  mod_topics <- rep(topics, each = length(seq(2, nseeds * 2, 2)))
+  mod_seeds <- rep(seq(2, nseeds * 2, 2), length(topics))
+
+  nmods <- length(mod_topics)
+  mods <- vector("list", length = nmods)
+  for (i in 1:nmods){
+    control <- prep_LDA_control(seed = mod_seeds[i], control = control)
+    mods[[i]] <- LDA(x = MV, k = mod_topics[i], control = control)
+  }
+  names(mods) <- paste0("c: ", mod_topics, ", seed: ", mod_seeds)
+  class(mods) <- c("LDA_list", "list")  
+  return(mods)
+}
+
+#' @title Verify that MV matrix is proper
+#' 
+#' @description Verify that the matrix of observations is conformable to
+#'   a matrix of integers
+#'   
+#' @param MV Matrix of observation count data (rows: documents (\code{M}), 
+#'   columns: terms (\code{V})).
[…]

### 2018-09-11 R/LDA_plots.R

@@ -1,0 +1,222 @@
+#' @title Plot a set of LDATS LDA models
+#'
+#' @description Generalization of the plot function to work on a list of LDA 
+#'   topic models (class \code{LDA_list}). 
+#' 
+#' @param x An \code{LDA_list} object of LDA topic models.
+#' 
+#' @param ... Additional arguments to be passed to subfunctions.
+#' 
+#' @return Model-by-model plots.
+#' 
+#' @examples 
+#' \dontrun{
+#'   data(rodents)
+#'   lda_data <- select(rodents, -c(newmoon, date, plots, traps))
+#'   r_LDA <- parLDA(data = lda_data, ntopics = 2, nseeds = 2)
+#'   plot(r_LDA)
+#' }
+#' @export 
+#'
+plot.LDA_list <- function(x, ...){
+  devAskNewPage(TRUE)
+  lapply(x, plot, ...)
+  devAskNewPage(FALSE)
+}
+
+#' @title Plot the results of an LDATS LDA model
+#'
+#' @description Create an LDATS LDA summary plot, with a top panel showing
+#'   the topic proportions for each word and a bottom panel showing the topic
+#'   proportions of each document/over time. 
+#' 
+#' @param x Object of class \code{LDA}
+#'
+#' @param cols colors to be used to plot the topics
+#' 
+#' @param xtime Optional x values used to plot the topic proportions according
+#'   to a specific time value (rather than simply the order of observations).
+#'
+#' @param xname Optional name for the x values used in plotting the topic
+#'   proportions (otherwise defaults to "Document"). 
+#'
+#' @param option A character string indicating the colormap option to use if 
+#'   `cols == NULL`. Four options are available: "magma" (or "A"), "inferno" 
+#'   (or "B"), "plasma" (or "C"), "viridis" (or "D", the default option) and 
+#'   "cividis" (or "E").
+#'
+#' @param ... Not used, retained for alignment with base function.
+#'
+#' @return LDATS LDA model plots.
+#' 
+#' @examples 
+#' \dontrun{
+#'   data(rodents)
+#'   lda_data <- dplyr::select(rodents, -c(newmoon, date, plots, traps))
[…]

### 2018-09-11 working/copied/LDATS_functions.R

@@ -28,7 +28,7 @@ LDA_TS <- function(document_term_matrix = NULL,
                    formula = ~1, nchangepoints = 1, ...){
 
   weights <- doc_weights(document_term_matrix)
-  ldas <- parLDA(data = document_term_matrix, ...) 
+  ldas <- parLDA(data = document_term_matrix) 
   selected <- LDA_select(ldas, ...) 
   mtss <- selected %>%
           MTS_prep(document_covariate_matrix) %>%

### 2018-09-11 working/copied/LDATS_package.R

@@ -1,0 +1,30 @@
+#' @importFrom magrittr %>%
+#' @importFrom foreach %dopar% foreach
+#' @importFrom grDevices devAskNewPage rgb
+#' @importFrom graphics axis mtext par plot points rect text
+#' @importFrom methods is
+#' @importFrom stats as.formula logLik rgeom runif
+#' @importFrom utils globalVariables
+#' @importFrom stats sd median
+#' @importFrom parallel detectCores makeCluster stopCluster
+#' @importFrom doParallel registerDoParallel
+#' @importFrom dplyr select
+#' @importFrom nnet multinom
+#' @importFrom memoise memoise
+#' @importFrom progress progress_bar
+#' @importFrom coda HPDinterval as.mcmc autocorr.diag effectiveSize autocorr
+#' @importFrom topicmodels logLik LDA
+#' @importFrom viridis viridis
+#'
+
+#' @title Performs two-stage LDA-timeseries analyses
+#'
+#' @description This package is designed to analyze multivariate time series
+#'   using a combination of Latent Dirichlet Allocation and multinomial 
+#'   regression.
+#'
+#' @name LDATS
+#' @docType package
+#' @keywords package
+#'
+NULL

### 2018-10-26 inst/doc/LDATS.Rmd

--- 2018-05-22 inst/doc/command-line-example.Rmd
@@
 ---
-title: "LDATS Example"
+title: "Latent Dirichlet Allocation Time Series (LDATS)"
 author: "Juniper L. Simonis"
@@
 
-This vignette walks through an example of **LDATS** at the command line and
-was constructed using **LDATS** version `r vers` on `r today`.
+## Overview
+
+This vignette details the code base for the **LDATS** package. It was
+constructed using **LDATS** version `r vers` on `r today`.
 
@@
 
-## Data
+## Analytical Pipeline
 
-For this vignette, we will be using rodent data from the control plots of the
-[Portal Project](https://github.com/weecology/portaldata), which come with
-the **LDATS** package
-
-```{r}
-data(rodents)
-rodents
-```
+* `LDA_TS()`
+  * `LDA_set()`
+    * `LDA()`
+  * `select_LDA()`
+  * `TS_set_on_LDA()`
+  * `select_TS()`

### 2018-11-27 working/devspace.R
minor edits

@@ -21,24 +21,9 @@ formula<- gamma ~ newmoon
 changepoints <- NULL
 
 
-
-mts <- multinom_TS(data, formula, changepoints = c(5, 100))
-
-# multinom_TS and its set of underlying functions are all good to go at this 
-# point. so now let's jump back to TS and get going there!
-#  presently working within TS, have all of the prep work done, leading up to
-#  the main for loop
-
-
-
-xx <- multinom_TS(data, formula, changepoints = c(5, 100))[[1]][[1]]
-vcov(xx)
-
-
-system.time(
-xx<-prep_proposal_dist(6)
-)
-
-system.time(
-for(i in 1:1e6)
-xx<-prep_proposal_dist(6))
+#
+# working within the TS function
+#   everything is set and working great within est_changepts
+#   now need to include the unconditional estimates of the regressors
+#   also want to include tests on the all existing functions and probably
+#   want to break up the TS function script a bit soon...

### 2018-11-28 inst/doc/rodents-example.R
updating the data documentation

@@ -15,5 +15,6 @@ today <- Sys.Date()
 
 ## ------------------------------------------------------------------------
 data(rodents)
-head(rodents, 10)
+head(rodents$document_term_table, 10)
+head(rodents$document_covariate_table, 10)
 

### 2018-12-13 rmd-ignore/compare-christensen.Rmd
rmd-ignore folder contains comparison w ecology paper

---
title: "Comparison with Christensen paper"
author: "Renata Diaz"
date: "12/13/2018"
output: github_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

Side-by-side comparison of LDATS results (and Portal rodents vignette) with analysis from Christensen et al 2018.

## LDATS Installation

To obtain the most recent version of **LDATS**, install the most recent 
version from GitHub:

```{r, eval=FALSE}
install.packages("devtools")
devtools::install_github("weecology/LDATS")
```

```{r, eval = T}
library(LDATS)
```

## Christensen 2018 analysis

Download Christensen 2018 analysis scripts & data files from [Extreme-events-LDA repo:](https://github.com/emchristensen/Extreme-events-LDA)

Main Analysis Scripts:

  * rodent_LDA_analysis.R main script for analyzing rodent community change using LDA

  * rodent_data_for_LDA.R contains a function that creates the rodent data table used in analyses

  * AIC_model_selection.R contains functions for calculating AIC for different candidate LDA models

  * changepointmodel.r contains change-point model code

  * LDA-distance.R function for computing Hellinger distance analyses

Data:

  * Rodent_table_dat.csv table of rodent data, created by rodent_data_for_LDA.R

Figure scripts:

  * LDA_figure_scripts.R contains functions for making main plots in manuscript (Fig 1). Called from rodent_LDA_analysis.R

```{r download Christensen, eval = FALSE}

files_to_download <- c('rodent_LDA_analysis.r', 'rodent_data_for_LDA.r', 'AIC_model_selection.R', 'changepointmodel.r', 'LDA-distance.R', 'Rodent_table_dat.csv', 'LDA_figure_scripts.R')

for(i in 1:length(files_to_download)) {
  download.file(url = paste0("https://raw.githubusercontent.com/emchristensen/Extreme-events-LDA/master/", files_to_download[i]),
                destfile = paste0('christensen-ecology/', files_to_download[i]))
}

rm(files_to_download)
rm(i)

```

## Data

The Portal rodents control data is included in the LDATS package:

```{r LDATS data}

data(rodents)

ldats_dat <- rodents[[1]]

head(ldats_dat)

```

Load the data used in Christensen et al:

### 2018-12-13 rmd-ignore/christensen-ecology/LDA_figure_scripts.R
rmd-ignore folder contains comparison w ecology paper

@@ -1,0 +1,299 @@
+# Scripts for making figures from LDA analyses
+
+library(ggplot2)
+library(gridExtra)
+library(dplyr)
+
+
+#cbPalette <- c( "#E69F00","#999999", "#56B4E9", "#0072B2", "#D55E00", "#F0E442", "#009E73", "#CC79A7")
+cbPalette <- c( "#e19c02","#999999", "#56B4E9", "#0072B2", "#D55E00", "#F0E442", "#009E73", "#CC79A7")
+
+
+
+
+#' Plot gamma
+#' 
+#' This plotting function plots the component communities over time
+#' It's used by the plot_component_communities function, but can also take
+#' any data frame as input as long as it's the right form -- used by the simulations
+#' 
+#' @param gamma_frame a data frame containing columns for date, relabund, and community
+#' @param ntopics number of topics
+#' @param ylab label for y axis (optional)
+#' 
+#' @return a ggplot object
+
+ltypes = c('solid','longdash','longdash','solid','solid')
+
+
+plot_gamma = function(gamma_frame,ntopics,ylab='',colors=cbPalette) {
+  g = ggplot(gamma_frame, aes(x=date,y=relabund,colour=community)) + 
+    #geom_point() +
+    geom_line(aes(size=community)) +
+    scale_size_manual(values=c(1,1,1,1,1), guide=FALSE) +
+    scale_y_continuous(name=ylab,limits=c(0,1)) +
+    scale_x_date(name='') +
+    theme(axis.text=element_text(size=12),
+          axis.title=element_text(size=12),
+          #panel.background = element_blank(),
+          panel.border=element_rect(colour='black',fill=NA),
+          legend.position='none') +
+    scale_colour_manual(name="Component\nCommunity",
+                        breaks=as.character(seq(ntopics)),
+                        values=colors[1:ntopics],
+                        guide=FALSE)
+  return(g)
+}
+
+
+#' Plot component communities
+#' 
+#' Plots timeseries of component communities (topics)
+#' 
+#' @param ldamodel object of class LDA_VEM created by the function LDA in topicmodels package
+#' @param ntopics number of topics used in ldamodel
+#' @param xticks vector of dates for x-axis labels
+#' @param ylab y axis label (optional)
+#' @param topic_order order of topics (for color control)
+#' 
+#' @return ggplot object
+#' 
[…]

### 2018-12-13 rmd-ignore/christensen-ecology/rodent_data_for_LDA.r

@@ -1,0 +1,51 @@
+library(RCurl)
+library(dplyr)
+
+
+#' @title Create rodent species table
+#'
+#' @description Processes rodent capture data so it can be used for LDA analysis
+#' 
+#'
+#' @param period_first first trapping period desired
+#' @param period_last last trapping period desired
+#' @param selected_plots vector of plot numbers to be included
+#' @param selected_species vector of species codes to be included
+#' 
+#' @return Table of species counts per period
+#' 
+#' @example
+#' r_table = create_rodent_table(1,436,c(2,4,8,11,12,14,17,22),
+#                              selected_species = c('BA','DM','DO','DS','NA','OL','OT','PB','PE','PF','PH','PI','PL','PM','PP','RF','RM','RO','SF','SH','SO'))
+#'
+
+create_rodent_table = function(period_first,period_last,selected_plots,selected_species) {
+  
+  # retrieve current version of rodent data
+  rodents = read.csv(text=getURL("https://raw.githubusercontent.com/weecology/PortalData/master/Rodents/Portal_rodent.csv"),
+                     na.strings=c(""), colClasses=c('tag'='character'), stringsAsFactors = FALSE)
+  
+  # extract desired data by period, plot, and species
+  rod = filter(rodents, period>=period_first, period<=period_last,
+               plot %in% selected_plots,
+               species %in% selected_species)
+  
+  # create table of species counts by period
+  r_table = table(rod$period,rod$species)
+  
+  # retrieve data on number of plots trapped per month
+  trap_table = read.csv('https://raw.githubusercontent.com/weecology/PortalData/master/Rodents/Portal_rodent_trapping.csv')
+  trap_table_controls = filter(trap_table, plot %in% selected_plots)
+  nplots_controls = aggregate(trap_table_controls$sampled,by=list(period = trap_table_controls$period),FUN=sum)
+  
+  # adjust species counts by number of plots trapped that month
+  r_table_adjusted = as.data.frame.matrix(r_table)
+  for (n in 1:436) {
+    #divide by number of control plots actually trapped (should be 8) and multiply by 8 to estimate captures as if all plots were trapped
+    r_table_adjusted[n,] = round(r_table_adjusted[n,]/nplots_controls$x[n]*8)
+  }
+  
+  #write.table(r_table_adjusted,'Rodent_table_dat.csv',sep=',',row.names = F)
+  return(r_table_adjusted)
+}
+

### 2018-12-13 rmd-ignore/compare-christensen.R

@@ -1,0 +1,14 @@
+# Do the data tables match?
+
+ch_dat <- read.csv('rmd-ignore/previous-work/paper_dat.csv', 
+                    stringsAsFactors = F)
+
+library(LDATS)
+data(rodents)
+ldats_dat <- rodents$document_term_table
+
+compare <- ldats_dat == ch_dat
+
+which(!compare)
+ldats_dat[351,]
+ch_dat[351, ]

### 2018-12-13 paper-comparison/compare-christensen.Rmd
Rodent data table comparison

---
title: "Comparison with Christensen paper"
author: "Renata Diaz"
date: "12/13/2018"
output: github_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

Side-by-side comparison of LDATS results (and Portal rodents vignette) with analysis from Christensen et al 2018.

## LDATS Installation

To obtain the most recent version of **LDATS**, install the most recent 
version from GitHub:

```{r, eval=FALSE}
install.packages("devtools")
devtools::install_github("weecology/LDATS")
```

```{r, eval = T}
library(LDATS)
```

## Christensen 2018 analysis

Download Christensen 2018 analysis scripts & data files from [Extreme-events-LDA repo:](https://github.com/emchristensen/Extreme-events-LDA)

Main Analysis Scripts:

  * rodent_LDA_analysis.R main script for analyzing rodent community change using LDA

  * rodent_data_for_LDA.R contains a function that creates the rodent data table used in analyses

  * AIC_model_selection.R contains functions for calculating AIC for different candidate LDA models

  * changepointmodel.r contains change-point model code

  * LDA-distance.R function for computing Hellinger distance analyses

Data:

  * Rodent_table_dat.csv table of rodent data, created by rodent_data_for_LDA.R

Figure scripts:

  * LDA_figure_scripts.R contains functions for making main plots in manuscript (Fig 1). Called from rodent_LDA_analysis.R

```{r download Christensen, eval = FALSE}

files_to_download <- c('rodent_LDA_analysis.r', 'rodent_data_for_LDA.r', 'AIC_model_selection.R', 'changepointmodel.r', 'LDA-distance.R', 'Rodent_table_dat.csv', 'LDA_figure_scripts.R')

for(i in 1:length(files_to_download)) {
  download.file(url = paste0("https://raw.githubusercontent.com/emchristensen/Extreme-events-LDA/master/", files_to_download[i]),
                destfile = paste0('christensen-ecology-files/', files_to_download[i]))
}

rm(files_to_download)
rm(i)

```

## Data

The Portal rodents control data is included in the LDATS package:

```{r LDATS data}

data(rodents)

ldats_dat <- rodents[[1]]
ldats_dates <- rodents[[2]]
head(ldats_dat)

```

Load the data used in Christensen et al:

```{r Ch data}
source('christensen-ecology-files/rodent_data_for_LDA.r')
dat = create_rodent_table(period_first = 1,
                          period_last = 436,
                          selected_plots = c(2,4,8,11,12,14,17,22),
                          selected_species = c('BA','DM','DO','DS','NA','OL','OT','PB','PE','PF','PH','PI','PL','PM','PP','RF','RM','RO','SF','SH','SO'), diagnose = F)

# dates to go with count data
moondat = read.csv(text=getURL("https://raw.githubusercontent.com/weecology/PortalData/master/Rodents/moon_dates.csv"),stringsAsFactors = F)
moondat$date = as.Date(moondat$censusdate)

period_dates = filter(moondat,period %in% rownames(dat)) %>% select(period,date)
dates = period_dates$date

ch_dat <- dat

ch_dates <- dates

rm(list = c('dat', 'dates', 'period_dates', 'moondat',
            'create_rodent_table'))

```

Compare paper data to LDATS data:

```{r rodent data comparison}

compare <- ldats_dat == ch_dat

compare_rows <- vector(length = nrow(compare))

for(i in 1:nrow(compare)) {
  if(sum(compare[i, ]) == 21) {
    compare_rows[i] <- TRUE
  } else {
    compare_rows[i] <- FALSE
  }
}

length(which(compare_rows == F))
ldats_dat[ which(compare_rows == F), ]
ch_dat[ which(compare_rows == F), ]
```

There are 16 rows where the data included in LDATS differs from the paper data. This looks like it is because the LDATS data is not adjusted to account for trapping effort, but the paper data has divided all census counts by the actual number of plots trapped and multiplied by 8 to account for incompletely-trapped censuses.

Double-check:

```{r Ch rodent adjustment, eval = F}

source('christensen-ecology-files/rodent_data_for_LDA.r')
diagnose_dat = create_rodent_table(period_first = 1,
                          period_last = 436,
                          selected_plots = c(2,4,8,11,12,14,17,22),
                          selected_species = c('BA','DM','DO','DS','NA','OL','OT','PB','PE','PF','PH','PI','PL','PM','PP','RF','RM','RO','SF','SH','SO'), diagnose = T)

adjusted_table <- diagnose_dat[[1]]
raw_table <- diagnose_dat[[2]]
nplots <- diagnose_dat[[3]]
nplots <- nplots[ which(nplots$period <= 436), ]

compare_raw <- raw_table == ldats_dat
length(which(compare_raw == F))

which(compare_rows == F) 
which(nplots$x < 8)

```

The lines that are different are the lines that were adjusted; if you don't adjust the paper data you get the same table as the LDATS data.

```{r data adjustment cleanup, include = FALSE}
rm(list = c('diagnose_dat', 'nplots', 'compare', 'compare_raw', 'adjusted_table', 'i', 'compare_rows', 'raw_table', 'create_rodent_table'))
```

For now I will use the *adjusted rodent table* because this will avoid artificially low abundances for incompletely trapped periods.

```{r switch to adjusted rodent data}

ldats_dat <- ch_dat
rodents$document_term_table <- ch_dat

```

### 2018-12-14 paper-comparison/christensen-ecology-files/LDA-distance.R
Add LDA comparisons
LDATS chooses 6 topics but the paper had 4.

@@ -56,7 +56,7 @@ calculate_LDA_distance = function(ldas,seeds) {
   # Calculate a bunch of LDAs with 4 topics
   k = 4
   ldas = purrr::map(seeds, 
-                    ~LDA(dat, k = k, method = "VEM", control = list(seed = .x)))
+                    ~LDA(ch_dat, k = k, method = "VEM", control = list(seed = .x)))
   
   # Log-likelihoods for each lda
   lls = purrr:::map_dbl(ldas, logLik)
@@ -81,22 +81,22 @@ calculate_LDA_distance = function(ldas,seeds) {
   
   # Re-order the rows of the "farthest" model so it matches the "best" model
   assignment = min_H(best_ps, farthest_ps, k)$assignment
-  
-  # plot results
-  # "best" is in black, "farthest" is in red
-  par(mfrow = c(2, 2))
-  for (row in 1:k) {
-    plot(NULL, xlim = c(1, 21), ylim = c(0, 1),
-         axes = FALSE, main = paste("topic", row), ylab = "", xlab = "",
-         yaxs = "i")
-    abline(v = 1:ncol(dat), col = scales::alpha(1, .05))
-    points(best_ps[row, ], col = 1, pch = 16)
-    points(farthest_ps[assignment[row], ], col = 2, lwd = 2)
-    axis(2)
-    axis(1, 1:ncol(dat), colnames(dat), las = 2)
-  }
-  par(mfrow = c(1, 1))
-  
+  # 
+  # # plot results
+  # # "best" is in black, "farthest" is in red
+  # par(mfrow = c(2, 2))
+  # for (row in 1:k) {
+  #   plot(NULL, xlim = c(1, 21), ylim = c(0, 1),
+  #        axes = FALSE, main = paste("topic", row), ylab = "", xlab = "",
+  #        yaxs = "i")
+  #   abline(v = 1:ncol(dat), col = scales::alpha(1, .05))
+  #   points(best_ps[row, ], col = 1, pch = 16)
+  #   points(farthest_ps[assignment[row], ], col = 2, lwd = 2)
+  #   axis(2)
+  #   axis(1, 1:ncol(dat), colnames(dat), las = 2)
+  # }
+  # par(mfrow = c(1, 1))
+  # 
   # average distance between "best" model and others
   mean_distance = mean(unlist(lapply(minimum_distances,'[[','min_cost')))
   # maximum distance; between best model and worst

### 2018-12-14 paper-comparison/compare-christensen.Rmd
note diagnose argument
Add LDA comparisons
LDATS chooses 6 topics but the paper had 4.

@@
 
-## Christensen 2018 analysis
+## Christensen 2018 analysis files
 
@@
 
+I added an argument 'diagnose' to the create_rodent_table() function. If T, returns the raw table & # of plots tables as well as the adjusted table.
+
 The lines that are different are the lines that were adjusted; if you don't adjust the paper data you get the same table as the LDATS data.
@@
 
-ldats_dat <- ch_dat
-rodents$document_term_table <- ch_dat
+ldats_dat <- as.matrix(ch_dat)
+ldats_dat <- apply(ldats_dat, c(1,2), FUN = as.integer)
+rodents$document_term_table <- ldats_dat
+rm(ldats_dat)
+rm(ldats_dates)
 
 ```
+
+## LDA
+
+While LDATS can run start-to-finish with `LDATS::LDA_TS`, I'm first going to step through function-by-function to isolate differences with the paper.
+
+```{r LDATS LDAs}
+
+ldats_ldas <- LDATS::LDA_set(document_term_table = rodents$document_term_table, topics = c(2:6), nseeds = 100)
+ldats_lda_selected <- LDATS::select_LDA(LDA_models = ldats_ldas)
+
+```
+
+Paper LDAS:
+
+```{r paper LDAs}
+source('christensen-ecology-files/AIC_model_selection.R')
+source('christensen-ecology-files/LDA-distance.R')
+
+# Fit a bunch of LDA models with different seeds
+# Only use even numbers for seeds because consecutive seeds give identical results
+seeds = 2*seq(200)
+
+# repeat LDA model fit and AIC calculation with a bunch of different seeds to test robustness of the analysis
+best_ntopic = repeat_VEM(ch_dat,
+                         seeds,
+                         topic_min=2,
+                         topic_max=6)
+hist(best_ntopic$k,breaks=c(0.5,1.5,2.5,3.5,4.5,5.5,6.5,7.5,8.5,9.5),xlab='best # of topics', main='')
+
+# 2b. how different is species composition of 4 community-types when LDA is run with different seeds?
+# ==================================================================
+# get the best 100 seeds where 4 topics was the best LDA model
+seeds_4topics = best_ntopic %>% 
+  filter(k == 4) %>% 
+  arrange(aic) %>% 
+  head(100) %>% 
+  pull(SEED)
+
+# choose seed with highest log likelihood for all following analyses
+#    (also produces plot of community composition for 'best' run compared to 'worst')
+best_seed = calculate_LDA_distance(ch_dat,seeds_4topics)
+mean_dist = unlist(best_seed)[2]
+max_dist = unlist(best_seed)[3]
+
+# ==================================================================
+# 3. run LDA model
+# ==================================================================
+ntopics = 4
+SEED = unlist(best_seed)[1]  # For the paper, I use seed 206
+ldamodel = LDA(ch_dat,ntopics, control = list(seed = SEED),method='VEM')
+
+```
+
+```{r plot LDAs}
+# Paper
+LDA_plot_top_panel(ldamodel, cols = NULL, option = "D")
+
+# LDATS
+LDA_plot_top_panel(ldats_lda_selected[[1]], cols = NULL, option = "D")
+```
+
+## Changepoint

### 2018-12-14 paper-comparison/compare-christensen.Rmd
trying with exactly the same data

@@
 # repeat LDA model fit and AIC calculation with a bunch of different seeds to test robustness of the analysis
-best_ntopic = repeat_VEM(ch_dat,
+best_ntopic = repeat_VEM(rodents[[1]],
                          seeds,
@@
 #    (also produces plot of community composition for 'best' run compared to 'worst')
-best_seed = calculate_LDA_distance(ch_dat,seeds_4topics)
+best_seed = calculate_LDA_distance(rodents[[1]],seeds_4topics)
 mean_dist = unlist(best_seed)[2]
@@
 SEED = unlist(best_seed)[1]  # For the paper, I use seed 206
-ldamodel = LDA(ch_dat,ntopics, control = list(seed = SEED),method='VEM')
+ldamodel = LDA(rodents[[1]],ntopics, control = list(seed = SEED),method='VEM')
 

### 2018-12-17 paper-comparison/compare-christensen.Rmd
Paper cpt favors 5 cpts

@@
 
+```{r save LDAs}
+save(ldamodel, file = 'paper_lda.Rds')
+save(ldats_lda_selected, file = 'ldats_lda.Rds')
+```
+
 ## Changepoint
+
+<!-- skip for now due to error -->
+<!-- ```{r LDATS changepoint} -->
+
+<!-- ldats_cpt_paper_lda <- LDATS::TS_on_LDA(LDA_models = ldamodel, rodents[[2]], -->
+<!--                                         formulas =c( '~ sin_year + cos_year'), nchangepoints = c(2, 3, 4, 5, 6)) -->
+
+<!-- ``` -->
+
+```{r paper changepoint on LDATS lda}
+source('christensen-ecology-files/changepointmodel.r')
+# set up parameters for model
+year_continuous = 1970 + as.integer(julian(ch_dates)) / 365.25
+x = data.frame(
+  year_continuous = year_continuous,
+  sin_year = sin(year_continuous * 2 * pi),
+  cos_year = cos(year_continuous * 2 * pi)
+)
+
+# run models with 1, 2, 3, 4, 5 changepoints
+cp_results_rodent = changepoint_model(ldats_lda_selected[[1]], x, 1, weights = rep(1,length(year_continuous)))
+cp_results_rodent2 = changepoint_model(ldats_lda_selected[[1]], x, 2, weights = rep(1,length(year_continuous)))
+cp_results_rodent3 = changepoint_model(ldats_lda_selected[[1]], x, 3, weights = rep(1,length(year_continuous)))
+cp_results_rodent4 = changepoint_model(ldats_lda_selected[[1]], x, 4, weights = rep(1,length(year_continuous)))
+cp_results_rodent5 = changepoint_model(ldats_lda_selected[[1]], x, 5, weights = rep(1,length(year_continuous)))
+cp_results_rodent6 = changepoint_model(ldats_lda_selected[[1]], x, 6, weights = rep(1,length(year_continuous)))
+
+save.image(file = '/Users/renatadiaz/Desktop/paper_cpt_ldats_lda.Rdata')
+
+# change point model selection
+# mean deviance ( -2 * log likelihood) + 2*(#parameters)
+mean(cp_results_rodent$saved_lls * -2) + 2*(3*(ntopics-1)*(1+1)+(1))
+mean(cp_results_rodent2$saved_lls * -2)+ 2*(3*(ntopics-1)*(2+1)+(2))
+mean(cp_results_rodent3$saved_lls * -2)+ 2*(3*(ntopics-1)*(3+1)+(3))
+mean(cp_results_rodent4$saved_lls * -2)+ 2*(3*(ntopics-1)*(4+1)+(4))
+mean(cp_results_rodent5$saved_lls * -2)+ 2*(3*(ntopics-1)*(5+1)+(5))
+
+```
+
+The lowest deviance (1215.971) is for 5 changepoints!
+
+Running one with 6.

### 2018-12-17 paper-comparison/compare-christensen.Rmd
Updated with 6 changepoint paper model
5 changepoints is still the lowest-deviance

@@
 
-<!-- skip for now due to error -->
-<!-- ```{r LDATS changepoint} -->
+```{r LDATS changepoint}
 
-<!-- ldats_cpt_paper_lda <- LDATS::TS_on_LDA(LDA_models = ldamodel, rodents[[2]], -->
-<!--                                         formulas =c( '~ sin_year + cos_year'), nchangepoints = c(2, 3, 4, 5, 6)) -->
+ldats_cpt_paper_lda <- LDATS::TS_on_LDA(LDA_models = ldamodel, rodents[[2]],
+                                        formulas =c(~ sin_year + cos_year), nchangepoints = c(2, 3, 4, 5, 6))
 
-<!-- ``` -->
+```
 
@@
 
-save.image(file = '/Users/renatadiaz/Desktop/paper_cpt_ldats_lda.Rdata')
+save.image(file = '/Users/renatadiaz/Desktop/paper_cpt_ldats_lda.Rdata') # saving it to the LDATS directory breaks git...?
 
@@
 mean(cp_results_rodent5$saved_lls * -2)+ 2*(3*(ntopics-1)*(5+1)+(5))
+mean(cp_results_rodent6$saved_lls * -2)+ 2*(3*(ntopics-1)*(6+1)+(6))
 
@@
 The lowest deviance (1215.971) is for 5 changepoints!
-
-Running one with 6.

### 2018-12-17 paper-comparison/changepoint_error.R
Updated with 6 changepoint paper model
5 changepoints is still the lowest-deviance

@@ -2,44 +2,46 @@ library(LDATS)
 data(rodents)
 
 #### swap raw data for adjusted data (to match paper) ####
-# you can skip this
-# 
-# source('christensen-ecology-files/rodent_data_for_LDA.r')
-# dat = create_rodent_table(period_first = 1,
-#                           period_last = 436,
-#                           selected_plots = c(2,4,8,11,12,14,17,22),
-#                           selected_species = c('BA','DM','DO','DS','NA','OL','OT','PB','PE','PF','PH','PI','PL','PM','PP','RF','RM','RO','SF','SH','SO'), diagnose = F)
-# 
-# # dates to go with count data
-# moondat = read.csv(text=getURL("https://raw.githubusercontent.com/weecology/PortalData/master/Rodents/moon_dates.csv"),stringsAsFactors = F)
-# moondat$date = as.Date(moondat$censusdate)
-# 
-# period_dates = filter(moondat,period %in% rownames(dat)) %>% select(period,date)
-# dates = period_dates$date
-# 
-# ch_dat <- dat
-# 
-# ch_dates <- dates
-# 
-# rm(list = c('dat', 'dates', 'period_dates', 'moondat',
-#             'create_rodent_table'))
-# 
-# 
-# ldats_dat <- as.matrix(ch_dat)
-# ldats_dat <- apply(ldats_dat, c(1,2), FUN = as.integer)
-# rodents$document_term_table <- ldats_dat
-# rm(ldats_dat)
-# rm(ldats_dates)
+you can skip this
+
+source('/Users/renatadiaz/Documents/GitHub/weecology/LDATS/paper-comparison/christensen-ecology-files/rodent_data_for_LDA.r')
+dat = create_rodent_table(period_first = 1,
+                          period_last = 436,
+                          selected_plots = c(2,4,8,11,12,14,17,22),
+                          selected_species = c('BA','DM','DO','DS','NA','OL','OT','PB','PE','PF','PH','PI','PL','PM','PP','RF','RM','RO','SF','SH','SO'), diagnose = F)
+
+# dates to go with count data
+moondat = read.csv(text=getURL("https://raw.githubusercontent.com/weecology/PortalData/master/Rodents/moon_dates.csv"),stringsAsFactors = F)
+moondat$date = as.Date(moondat$censusdate)
+
[…]

### 2018-12-20 paper-comparison/compare-christensen.Rmd
knit through ldats on ldats

--- 2018-12-14 paper-comparison/compare-christensen.Rmd
@@
 
-```{r rodent data comparison}
+```{r rodent data comparison, eval = F}
 
@@
 
-```{r data adjustment cleanup, include = FALSE}
+```{r data adjustment cleanup, eval = F, include = FALSE}
 rm(list = c('diagnose_dat', 'nplots', 'compare', 'compare_raw', 'adjusted_table', 'i', 'compare_rows', 'raw_table', 'create_rodent_table'))
@@
 rm(ldats_dates)
-
+rm(ch_dates)
+rm(ch_dat)
 ```
@@
 
-```{r LDATS LDAs}
+```{r LDATS LDAs, eval = F}
 
@@
 
-```{r paper LDAs}
+```{r paper LDAs, eval = F}
 source('christensen-ecology-files/AIC_model_selection.R')
@@
 
+```{r save LDAs, eval = F, include = F}
+save(ldamodel, file = 'model-stash/paper_lda.Rds')
+save(ldats_lda_selected, file = 'model-stash/ldats_lda.Rds')
+```
+
+```{r reload LDAS, include = FALSE}
+
+load('model-stash/ldats_lda.Rds')
+load('model-stash/paper_lda.Rds')
+
+```
+
+### Plot LDAS
+
 ```{r plot LDAs}
 # Paper
-LDA_plot_top_panel(ldamodel, cols = NULL, option = "D")
+plot(ldamodel, cols = NULL, option = "D")
 
 # LDATS
-LDA_plot_top_panel(ldats_lda_selected[[1]], cols = NULL, option = "D")
+plot(ldats_lda_selected[[1]], cols = NULL, option = "D")
 ```
 
-## Changepoint
+## Changepoint models
+
+I am going to compare four combinations of LDA + changepoint models:
+
+* LDATS LDA + LDATS changepoint
+* LDATS LDA + paper changepoint
+* Paper LDA + LDATS changepoint
+* Paper LDA + paper changepoint
+
+There is the additional wrinkle of `document_term_weights`.
+The paper weighted all sample periods equally, wheras LDATS can weight sample periods according to how many individuals were captured.
+We now believe it is more appropriate to weight periods proportional to captures.
+However, for the purposes of comparison, I will continue to set all weights = 1 for both changepoint models.
+For an example of LDATS run with proportional weights, see the rodents vignette.
+[?]
+
+### LDATS LDA with LDATS changepoint
+```{r, include = F}
+# remove paper LDA (for memory)
+rm(ldamodel)
+```
+
+```{r LDATS LDA + LDATS cpt}
+# Run changepoint
+
+ldats_ldats <- LDATS::TS_on_LDA(LDA_models = ldats_lda_selected,
+                                document_covariate_table = rodents[[2]],
+                                formulas =c(~ sin_year + cos_year),
+                                nchangepoints = c(2, 3),
+                                weights = NULL)
+
+save(ldats_ldats, file = 'model-stash/ldats_ldats.Rds')
+
+# Select best changepoint
+
+ldats_ldats_selected <- LDATS::select_TS(TS_models = ldats_ldats)
+
+save(ldats_ldats_selected, file = 'model-stash/ldats_ldats_selected.Rds')
+
+```

### 2018-12-20 paper-comparison/compare-christensen.Rmd
Set up to run paper cpt on LDATS lda

@@
 rm(ldats_dates)
-rm(ch_dates)
 rm(ch_dat)
@@
 
-```{r LDATS LDA + LDATS cpt}
+```{r LDATS LDA + LDATS cpt, eval = F}
 # Run changepoint
@@
 
-save(ldats_ldats_selected, file = 'model-stash/ldats_ldats_selected.Rds')
+save(ldats_ldats_selected, file = '/Users/renatadiaz/Documents/model-stash/ldats_ldats_selected.Rds')
 
 ```
+
+```{r, eval = F}
+rm(ldats_ldats)
+rm(ldats_ldats_selected)
+```
+
+```{r paper cpt on LDATS LDA}
+source('christensen-ecology-files/changepointmodel.r')
+# set up parameters for model
+year_continuous = 1970 + as.integer(julian(ch_dates)) / 365.25
+x = data.frame(
+  year_continuous = year_continuous,
+  sin_year = sin(year_continuous * 2 * pi),
+  cos_year = cos(year_continuous * 2 * pi)
+)
+
+```
+
+```{r, eval = F}
+# run models with 1, 2, 3, 4, 5 changepoints
+cp_results_rodent = changepoint_model(ldats_lda_selected[[1]], x, 1, weights = rep(1,length(year_continuous)))
+cp_results_rodent2 = changepoint_model(ldats_lda_selected[[1]], x, 2, weights = rep(1,length(year_continuous)))
+cp_results_rodent3 = changepoint_model(ldats_lda_selected[[1]], x, 3, weights = rep(1,length(year_continuous)))
+cp_results_rodent4 = changepoint_model(ldats_lda_selected[[1]], x, 4, weights = rep(1,length(year_continuous)))
+cp_results_rodent5 = changepoint_model(ldats_lda_selected[[1]], x, 5, weights = rep(1,length(year_continuous)))
+cp_results_rodent6 = changepoint_model(ldats_lda_selected[[1]], x, 6, weights = rep(1,length(year_continuous)))
+
+paper_ldats <- list(cp_results_rodent, cp_results_rodent2,
+                    cp_results_rodent3, cp_results_rodent4,
+                    cp_results_rodent5, cp_results_rodent6)
+
+rm(list = c('cp_results_rodent', 'cp_results_rodent2',
+                    'cp_results_rodent3', 'cp_results_rodent4',
+                    'cp_results_rodent5', 'cp_results_rodent6'))
+
+save(paper_ldats, file = '/Users/renatadiaz/Documents/model-stash/paper_cpt_ldats_lda.Rds')

### 2018-12-20 paper-comparison/background-cpt.R
Set up to run paper cpt on LDATS lda

@@ -1,21 +1,0 @@
-library(LDATS)
-
-load('/Users/renatadiaz/Desktop/paper_cpt_ldats_lda.Rdata')
-source('/Users/renatadiaz/Documents/GitHub/weecology/LDATS/paper-comparison/christensen-ecology-files/changepointmodel.r')
-# set up parameters for model
-year_continuous = 1970 + as.integer(julian(ch_dates)) / 365.25
-x = data.frame(
-  year_continuous = year_continuous,
-  sin_year = sin(year_continuous * 2 * pi),
-  cos_year = cos(year_continuous * 2 * pi)
-)
-
-# run models with 1, 2, 3, 4, 5 changepoints
-cp_results_rodent = changepoint_model(ldats_lda_selected[[1]], x, 1, weights = rep(1,length(year_continuous)))
-cp_results_rodent2 = changepoint_model(ldats_lda_selected[[1]], x, 2, weights = rep(1,length(year_continuous)))
-cp_results_rodent3 = changepoint_model(ldats_lda_selected[[1]], x, 3, weights = rep(1,length(year_continuous)))
-cp_results_rodent4 = changepoint_model(ldats_lda_selected[[1]], x, 4, weights = rep(1,length(year_continuous)))
-cp_results_rodent5 = changepoint_model(ldats_lda_selected[[1]], x, 5, weights = rep(1,length(year_continuous)))
-cp_results_rodent6 = changepoint_model(ldats_lda_selected[[1]], x, 6, weights = rep(1,length(year_continuous)))
-
-save.image(file = '/Users/renatadiaz/Desktop/paper_cpt_ldats_lda.Rdata')

### 2018-12-26 paper-comparison/compare-christensen.Rmd
Updating with code to run the changepoint-LDA combinations
I'm saving the changepoint models to my computer, because if I put them in the repo it breaks git on my computer. I'm thinking of putting the .Rds files in Dropbox or some other workaround for large files.

@@
 
-### LDATS LDA with LDATS changepoint
+### run LDATS LDA with LDATS changepoint
 ```{r, include = F}
@@
 save(paper_ldats, file = '/Users/renatadiaz/Documents/model-stash/paper_cpt_ldats_lda.Rds') 
+```
+
+### run LDATS LDA with paper changepoint
+
+```{r ldats paper, eval = F}
+source('christensen-ecology-files/changepointmodel.r')
+# set up parameters for model
+year_continuous = 1970 + as.integer(julian(ch_dates)) / 365.25
+x = data.frame(
+  year_continuous = year_continuous,
+  sin_year = sin(year_continuous * 2 * pi),
+  cos_year = cos(year_continuous * 2 * pi)
+)
+
+# run models with 1, 2, 3, 4, 5 changepoints
+cp_results_rodent = changepoint_model(ldats_lda_selected[[1]], x, 1, weights = rep(1,length(year_continuous)))
+
+save(cp_results_rodent, file = '/Users/renatadiaz/Documents/model-stash/paper_cpt_ldats_lda1.Rds')
+rm(cp_results_rodent)
+
+cp_results_rodent2 = changepoint_model(ldats_lda_selected[[1]], x, 2, weights = rep(1,length(year_continuous)))
+
+save(cp_results_rodent2, file = '/Users/renatadiaz/Documents/model-stash/paper_cpt_ldats_lda2.Rds')
+rm(cp_results_rodent2)
+
+cp_results_rodent3 = changepoint_model(ldats_lda_selected[[1]], x, 3, weights = rep(1,length(year_continuous)))
+save(cp_results_rodent3, file = '/Users/renatadiaz/Documents/model-stash/paper_cpt_ldats_lda3.Rds')
+rm(cp_results_rodent3)
+
+cp_results_rodent4 = changepoint_model(ldats_lda_selected[[1]], x, 4, weights = rep(1,length(year_continuous)))
+save(cp_results_rodent4, file = '/Users/renatadiaz/Documents/model-stash/paper_cpt_ldats_lda4.Rds')
+rm(cp_results_rodent4)
+
+cp_results_rodent5 = changepoint_model(ldats_lda_selected[[1]], x, 5, weights = rep(1,length(year_continuous)))
+save(cp_results_rodent5, file = '/Users/renatadiaz/Documents/model-stash/paper_cpt_ldats_lda5.Rds')
+rm(cp_results_rodent5)
+
+cp_results_rodent6 = changepoint_model(ldats_lda_selected[[1]], x, 6, weights = rep(1,length(year_continuous)))
+save(cp_results_rodent6, file = '/Users/renatadiaz/Documents/model-stash/paper_cpt_ldats_lda6.Rds')
+rm(cp_results_rodent6)
+
+rm(ldats_lda_selected)
+```
+
+### run paper LDA with paper changepoint
+```{r paper paper, eval = F}
+source('christensen-ecology-files/changepointmodel.r')
+# set up parameters for model
+year_continuous = 1970 + as.integer(julian(ch_dates)) / 365.25
+x = data.frame(
+year_continuous = year_continuous,
+sin_year = sin(year_continuous * 2 * pi),
+cos_year = cos(year_continuous * 2 * pi)
+)
+# run models with 1, 2, 3, 4, 5 changepoints
+cp_results_rodent = changepoint_model(ldamodel, x, 1, weights = rep(1,length(year_continuous)))
+save(cp_results_rodent, file = '/Users/renatadiaz/Documents/model-stash/paper_cpt_paper_lda1.Rds')
+rm(cp_results_rodent)
+
+cp_results_rodent2 = changepoint_model(ldamodel, x, 2, weights = rep(1,length(year_continuous)))
+save(cp_results_rodent2, file = '/Users/renatadiaz/Documents/model-stash/paper_cpt_paper_lda2.Rds')
+rm(cp_results_rodent2)
+
+cp_results_rodent3 = changepoint_model(ldamodel, x, 3, weights = rep(1,length(year_continuous)))
+save(cp_results_rodent3, file = '/Users/renatadiaz/Documents/model-stash/paper_cpt_paper_lda3.Rds')
+rm(cp_results_rodent3)
+
+cp_results_rodent4 = changepoint_model(ldamodel, x, 4, weights = rep(1,length(year_continuous)))
+save(cp_results_rodent4, file = '/Users/renatadiaz/Documents/model-stash/paper_cpt_paper_lda4.Rds')
+rm(cp_results_rodent4)
+
+cp_results_rodent5 = changepoint_model(ldamodel, x, 5, weights = rep(1,length(year_continuous)))
+save(cp_results_rodent5, file = '/Users/renatadiaz/Documents/model-stash/paper_cpt_paper_lda5.Rds')
+rm(cp_results_rodent5)
+
+cp_results_rodent6 = changepoint_model(ldamodel, x, 6, weights = rep(1,length(year_continuous)))
+save(cp_results_rodent6, file = '/Users/renatadiaz/Documents/model-stash/paper_cpt_paper_lda6.Rds')
+rm(cp_results_rodent6)
+```
+
+### run paper LDA with LDATS changepoint
+```{r paper ldats, eval = F}
+# Run changepoint
+
+paper_ldats <- LDATS::TS_on_LDA(LDA_models = ldamodel,
+                                document_covariate_table = rodents[[2]],
+                                formulas =c(~ sin_year + cos_year),
+                                nchangepoints = c(2, 3, 4, 5, 6),
+                                weights = NULL)
+
+save(paper_ldats, file = 'model-stash/paper_ldats.Rds')
+
+# Select best changepoint
+
+paper_ldats_selected <- LDATS::select_TS(TS_models = paper_ldats)
+
+save(paper_ldats_selected, file = '/Users/renatadiaz/Documents/model-stash/paper_ldats_selected.Rds')
+
+```

### 2018-12-26 paper-comparison/compare-christensen.Rmd
Update compare-christensen.Rmd

@@
 
-### run LDATS LDA with LDATS changepoint
+#### run LDATS LDA with LDATS changepoint
 ```{r, include = F}
@@
 
-```{r paper cpt on LDATS LDA}
-source('christensen-ecology-files/changepointmodel.r')
-# set up parameters for model
-year_continuous = 1970 + as.integer(julian(ch_dates)) / 365.25
-x = data.frame(
-  year_continuous = year_continuous,
-  sin_year = sin(year_continuous * 2 * pi),
-  cos_year = cos(year_continuous * 2 * pi)
-)
-
-```
-
-```{r, eval = F}
-# run models with 1, 2, 3, 4, 5 changepoints
-cp_results_rodent = changepoint_model(ldats_lda_selected[[1]], x, 1, weights = rep(1,length(year_continuous)))
-cp_results_rodent2 = changepoint_model(ldats_lda_selected[[1]], x, 2, weights = rep(1,length(year_continuous)))
-cp_results_rodent3 = changepoint_model(ldats_lda_selected[[1]], x, 3, weights = rep(1,length(year_continuous)))
-cp_results_rodent4 = changepoint_model(ldats_lda_selected[[1]], x, 4, weights = rep(1,length(year_continuous)))
-cp_results_rodent5 = changepoint_model(ldats_lda_selected[[1]], x, 5, weights = rep(1,length(year_continuous)))
-cp_results_rodent6 = changepoint_model(ldats_lda_selected[[1]], x, 6, weights = rep(1,length(year_continuous)))
-
-paper_ldats <- list(cp_results_rodent, cp_results_rodent2,
-                    cp_results_rodent3, cp_results_rodent4,
-                    cp_results_rodent5, cp_results_rodent6)
-
-rm(list = c('cp_results_rodent', 'cp_results_rodent2',
-                    'cp_results_rodent3', 'cp_results_rodent4',
-                    'cp_results_rodent5', 'cp_results_rodent6'))
-
-save(paper_ldats, file = '/Users/renatadiaz/Documents/model-stash/paper_cpt_ldats_lda.Rds') 
-```
-
-### run LDATS LDA with paper changepoint
+#### run LDATS LDA with paper changepoint
 
@@
 
-### run paper LDA with paper changepoint
+#### run paper LDA with paper changepoint
 ```{r paper paper, eval = F}
@@
 
-### run paper LDA with LDATS changepoint
+#### run paper LDA with LDATS changepoint
 ```{r paper ldats, eval = F}

### 2018-12-27 paper-comparison/compare-christensen.Rmd
Files for side by side comparison of paper changepoint
Paper changepoint on paper lda vs LDATS lda.

At this point I find it easiest to open the image files in 'paper-changepoint-files' and put the LDATS lda ones side by side with the paper lda ones. I know this is... not ideal.

Both give same #'s of changepoints (4 or 6 - 6 has lower deviance, but I'm not sure if the changepoint is to be trusted for 6 cpts?) and roughly the same locations.

The LDATs LDA w/ 4 cpts has less bimodality in the 1990's  than the paper LDA.

@@
 #### run LDATS LDA with LDATS changepoint
-```{r, include = F}
-# remove paper LDA (for memory)
-rm(ldamodel)
-```
 
@@
 
+```{r paper paper selection}
+ntopics = ldamodel@k
+# Load models and compare deviance to select the best one
+load('/Users/renatadiaz/Documents/model-stash/paper_cpt_paper_lda1.Rds')
+load('/Users/renatadiaz/Documents/model-stash/paper_cpt_paper_lda2.Rds')
+load('/Users/renatadiaz/Documents/model-stash/paper_cpt_paper_lda3.Rds')
+load('/Users/renatadiaz/Documents/model-stash/paper_cpt_paper_lda4.Rds')
+load('/Users/renatadiaz/Documents/model-stash/paper_cpt_paper_lda5.Rds')
+load('/Users/renatadiaz/Documents/model-stash/paper_cpt_paper_lda6.Rds')
+
+# change point model selection
+# mean deviance ( -2 * log likelihood) + 2*(#parameters)
+mean(cp_results_rodent$saved_lls * -2) + 2*(3*(ntopics-1)*(1+1)+(1))
+mean(cp_results_rodent2$saved_lls * -2)+ 2*(3*(ntopics-1)*(2+1)+(2))
+mean(cp_results_rodent3$saved_lls * -2)+ 2*(3*(ntopics-1)*(3+1)+(3))
+mean(cp_results_rodent4$saved_lls * -2)+ 2*(3*(ntopics-1)*(4+1)+(4))
+mean(cp_results_rodent5$saved_lls * -2)+ 2*(3*(ntopics-1)*(5+1)+(5))
+mean(cp_results_rodent6$saved_lls * -2)+ 2*(3*(ntopics-1)*(5+1)+(5))
+
+```
+
 #### run paper LDA with LDATS changepoint
@@
 ```
+
+### Compare model outcomes

### 2018-12-27 paper-comparison/paper_changepoint.Rmd
Files for side by side comparison of paper changepoint
Paper changepoint on paper lda vs LDATS lda.

At this point I find it easiest to open the image files in 'paper-changepoint-files' and put the LDATS lda ones side by side with the paper lda ones. I know this is... not ideal.

Both give same #'s of changepoints (4 or 6 - 6 has lower deviance, but I'm not sure if the changepoint is to be trusted for 6 cpts?) and roughly the same locations.

The LDATs LDA w/ 4 cpts has less bimodality in the 1990's  than the paper LDA.

---
title: "Paper changepoint"
author: "Renata Diaz"
date: "12/26/2018"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
``` {r ldats lda}
library(LDATS)
#### Data ####

data(rodents)

ldats_dat <- rodents[[1]]
ldats_dates <- rodents[[2]]
head(ldats_dat)
source('christensen-ecology-files/rodent_data_for_LDA.r')
dat = create_rodent_table(period_first = 1,
                          period_last = 436,
                          selected_plots = c(2,4,8,11,12,14,17,22),
                          selected_species = c('BA','DM','DO','DS','NA','OL','OT','PB','PE','PF','PH','PI','PL','PM','PP','RF','RM','RO','SF','SH','SO'), diagnose = F)

# dates to go with count data
moondat = read.csv(text=getURL("https://raw.githubusercontent.com/weecology/PortalData/master/Rodents/moon_dates.csv"),stringsAsFactors = F)
moondat$date = as.Date(moondat$censusdate)

period_dates = filter(moondat,period %in% rownames(dat)) %>% select(period,date)
dates = period_dates$date

ch_dat <- dat

ch_dates <- dates

rm(list = c('dat', 'dates', 'period_dates', 'moondat',
            'create_rodent_table'))

ldats_dat <- as.matrix(ch_dat)
ldats_dat <- apply(ldats_dat, c(1,2), FUN = as.integer)
rodents$document_term_table <- ldats_dat
rm(ldats_dat)
rm(ldats_dates)
rm(ch_dat)

#### Load LDA ####
load('/Users/renatadiaz/Documents/model-stash/ldats_lda.Rds')

#### Run changepoint ####
source('christensen-ecology-files/changepointmodel.r')
# set up parameters for model
year_continuous = 1970 + as.integer(julian(ch_dates)) / 365.25
x = data.frame(
  year_continuous = year_continuous,
  sin_year = sin(year_continuous * 2 * pi),
  cos_year = cos(year_continuous * 2 * pi)
)

# # run models with 1, 2, 3, 4, 5 changepoints
# cp_results_rodent = changepoint_model(ldats_lda_selected[[1]], x, 1, weights = rep(1,length(year_continuous)))
# save(cp_results_rodent, file = '/Users/renatadiaz/Documents/model-stash/paper_cpt_ldats_lda1.Rds')
# rm(cp_results_rodent)
# 
# cp_results_rodent2 = changepoint_model(ldats_lda_selected[[1]], x, 2, weights = rep(1,length(year_continuous)))
# save(cp_results_rodent2, file = '/Users/renatadiaz/Documents/model-stash/paper_cpt_ldats_lda2.Rds')
# rm(cp_results_rodent2)
# 
# cp_results_rodent3 = changepoint_model(ldats_lda_selected[[1]], x, 3, weights = rep(1,length(year_continuous)))
# save(cp_results_rodent3, file = '/Users/renatadiaz/Documents/model-stash/paper_cpt_ldats_lda3.Rds')
# rm(cp_results_rodent3)
# 
# cp_results_rodent4 = changepoint_model(ldats_lda_selected[[1]], x, 4, weights = rep(1,length(year_continuous)))
# save(cp_results_rodent4, file = '/Users/renatadiaz/Documents/model-stash/paper_cpt_ldats_lda4.Rds')
# rm(cp_results_rodent4)
# 
# cp_results_rodent5 = changepoint_model(ldats_lda_selected[[1]], x, 5, weights = rep(1,length(year_continuous)))
# save(cp_results_rodent5, file = '/Users/renatadiaz/Documents/model-stash/paper_cpt_ldats_lda5.Rds')
# rm(cp_results_rodent5)
# 
# cp_results_rodent6 = changepoint_model(ldats_lda_selected[[1]], x, 6, weights = rep(1,length(year_continuous)))
# save(cp_results_rodent6, file = '/Users/renatadiaz/Documents/model-stash/paper_cpt_ldats_lda6.Rds')
# rm(cp_results_rodent6)
#

#### Changepoint model selection ####
load('/Users/renatadiaz/Documents/model-stash/paper_cpt_ldats_lda1.Rds')
load('/Users/renatadiaz/Documents/model-stash/paper_cpt_ldats_lda2.Rds')
load('/Users/renatadiaz/Documents/model-stash/paper_cpt_ldats_lda3.Rds')
load('/Users/renatadiaz/Documents/model-stash/paper_cpt_ldats_lda4.Rds')
load('/Users/renatadiaz/Documents/model-stash/paper_cpt_ldats_lda5.Rds')
load('/Users/renatadiaz/Documents/model-stash/paper_cpt_ldats_lda6.Rds')

ntopics = ldats_lda_selected[[1]]@k

# change point model selection
# mean deviance ( -2 * log likelihood) + 2*(#parameters)
mean(cp_results_rodent$saved_lls * -2) + 2*(3*(ntopics-1)*(1+1)+(1))
mean(cp_results_rodent2$saved_lls * -2)+ 2*(3*(ntopics-1)*(2+1)+(2))
mean(cp_results_rodent3$saved_lls * -2)+ 2*(3*(ntopics-1)*(3+1)+(3))
mean(cp_results_rodent4$saved_lls * -2)+ 2*(3*(ntopics-1)*(4+1)+(4))
mean(cp_results_rodent5$saved_lls * -2)+ 2*(3*(ntopics-1)*(5+1)+(5))
mean(cp_results_rodent6$saved_lls * -2)+ 2*(3*(ntopics-1)*(5+1)+(5))

# lowest deviance is for 4 (or 6) changepoints

rm(list = c('cp_results_rodent', 'cp_results_rodent2', 
            'cp_results_rodent3', 'cp_results_rodent5'))
hist(year_continuous[cp_results_rodent4$saved[,1,]],breaks = seq(1977,2016,.25),xlab='',main='Changepoint Estimate')
hist(year_continuous[cp_results_rodent6$saved[,1,]],breaks = seq(1977,2016,.25),xlab='',main='Changepoint Estimate')

annual_hist(cp_results_rodent4,year_continuous)

# turn changepoint results into data frame
df_4 = as.data.frame(t(cp_results_rodent4$saved[,1,])) %>% melt()
df_4$value = year_continuous[df_4$value]

# find 95% confidence intervals on each changepoint:
quantile(df_4[df_4$variable=='V1','value'],probs=c(.025,.975)) %>% date_decimal() %>% format('%d-%m-%Y')
quantile(df_4[df_4$variable=='V2','value'],probs=c(.025,.975)) %>% date_decimal() %>% format('%d-%m-%Y')
quantile(df_4[df_4$variable=='V3','value'],probs=c(.025,.975)) %>% date_decimal() %>% format('%d-%m-%Y')
quantile(df_4[df_4$variable=='V4','value'],probs=c(.025,.975)) %>% date_decimal() %>% format('%d-%m-%Y')

## 6 cpts 
annual_hist(cp_results_rodent6,year_continuous)
# turn changepoint results into data frame
df_6 = as.data.frame(t(cp_results_rodent6$saved[,1,])) %>% melt()
df_6$value = year_continuous[df_6$value]

# find 95% confidence intervals on each changepoint:
quantile(df_6[df_6$variable=='V1','value'],probs=c(.025,.975)) %>% date_decimal() %>% format('%d-%m-%Y')
quantile(df_6[df_6$variable=='V2','value'],probs=c(.025,.975)) %>% date_decimal() %>% format('%d-%m-%Y')
quantile(df_6[df_6$variable=='V3','value'],probs=c(.025,.975)) %>% date_decimal() %>% format('%d-%m-%Y')
quantile(df_6[df_6$variable=='V4','value'],probs=c(.025,.975)) %>% date_decimal() %>% format('%d-%m-%Y')
quantile(df_6[df_6$variable=='V5','value'],probs=c(.025,.975)) %>% date_decimal() %>% format('%d-%m-%Y')
quantile(df_6[df_6$variable=='V6','value'],probs=c(.025,.975)) %>% date_decimal() %>% format('%d-%m-%Y')

rm(list = ls())
```

```{r paper lda}
library(LDATS)
#### Data ####

data(rodents)

ldats_dat <- rodents[[1]]
ldats_dates <- rodents[[2]]
head(ldats_dat)
source('christensen-ecology-files/rodent_data_for_LDA.r')
dat = create_rodent_table(period_first = 1,
                          period_last = 436,
                          selected_plots = c(2,4,8,11,12,14,17,22),
                          selected_species = c('BA','DM','DO','DS','NA','OL','OT','PB','PE','PF','PH','PI','PL','PM','PP','RF','RM','RO','SF','SH','SO'), diagnose = F)

# dates to go with count data
moondat = read.csv(text=getURL("https://raw.githubusercontent.com/weecology/PortalData/master/Rodents/moon_dates.csv"),stringsAsFactors = F)
moondat$date = as.Date(moondat$censusdate)

period_dates = filter(moondat,period %in% rownames(dat)) %>% select(period,date)
dates = period_dates$date

ch_dat <- dat

ch_dates <- dates

rm(list = c('dat', 'dates', 'period_dates', 'moondat',
            'create_rodent_table'))

ldats_dat <- as.matrix(ch_dat)
ldats_dat <- apply(ldats_dat, c(1,2), FUN = as.integer)
rodents$document_term_table <- ldats_dat
rm(ldats_dat)
rm(ldats_dates)
rm(ch_dat)

#### Load LDA ####
load('/Users/renatadiaz/Documents/model-stash/paper_lda.Rds')

#### Run changepoint ####
source('christensen-ecology-files/changepointmodel.r')
# set up parameters for model
year_continuous = 1970 + as.integer(julian(ch_dates)) / 365.25
x = data.frame(
  year_continuous = year_continuous,
  sin_year = sin(year_continuous * 2 * pi),
  cos_year = cos(year_continuous * 2 * pi)
)

# # run models with 1, 2, 3, 4, 5 changepoints
# cp_results_rodent = changepoint_model(ldamodel, x, 1, weights = rep(1,length(year_continuous)))
# save(cp_results_rodent, file = '/Users/renatadiaz/Documents/model-stash/paper_cpt_paper_lda1.Rds')
# rm(cp_results_rodent)
# 
# 
# cp_results_rodent2 = changepoint_model(ldamodel, x, 2, weights = rep(1,length(year_continuous)))
# save(cp_results_rodent2, file = '/Users/renatadiaz/Documents/model-stash/paper_cpt_paper_lda2.Rds')
# rm(cp_results_rodent2)
# 
# cp_results_rodent3 = changepoint_model(ldamodel, x, 3, weights = rep(1,length(year_continuous)))
# save(cp_results_rodent3, file = '/Users/renatadiaz/Documents/model-stash/paper_cpt_paper_lda3.Rds')
# rm(cp_results_rodent3)
# 
# cp_results_rodent4 = changepoint_model(ldamodel, x, 4, weights = rep(1,length(year_continuous)))
# save(cp_results_rodent4, file = '/Users/renatadiaz/Documents/model-stash/paper_cpt_paper_lda4.Rds')
# rm(cp_results_rodent4)
# 
# cp_results_rodent5 = changepoint_model(ldamodel, x, 5, weights = rep(1,length(year_continuous)))
# save(cp_results_rodent5, file = '/Users/renatadiaz/Documents/model-stash/paper_cpt_paper_lda5.Rds')
# rm(cp_results_rodent5)
# 
# cp_results_rodent6 = changepoint_model(ldamodel, x, 6, weights = rep(1,length(year_continuous)))
# save(cp_results_rodent6, file = '/Users/renatadiaz/Documents/model-stash/paper_cpt_paper_lda6.Rds')
# rm(cp_results_rodent6)

## ----paper paper selection-----------------------------------------------
ntopics = ldamodel@k
# Load models and compare deviance to select the best one
load('/Users/renatadiaz/Documents/model-stash/paper_cpt_paper_lda1.Rds')
load('/Users/renatadiaz/Documents/model-stash/paper_cpt_paper_lda2.Rds')
load('/Users/renatadiaz/Documents/model-stash/paper_cpt_paper_lda3.Rds')
load('/Users/renatadiaz/Documents/model-stash/paper_cpt_paper_lda4.Rds')
load('/Users/renatadiaz/Documents/model-stash/paper_cpt_paper_lda5.Rds')
load('/Users/renatadiaz/Documents/model-stash/paper_cpt_paper_lda6.Rds')

# change point model selection
# mean deviance ( -2 * log likelihood) + 2*(#parameters)
mean(cp_results_rodent$saved_lls * -2) + 2*(3*(ntopics-1)*(1+1)+(1))
mean(cp_results_rodent2$saved_lls * -2)+ 2*(3*(ntopics-1)*(2+1)+(2))
mean(cp_results_rodent3$saved_lls * -2)+ 2*(3*(ntopics-1)*(3+1)+(3))
mean(cp_results_rodent4$saved_lls * -2)+ 2*(3*(ntopics-1)*(4+1)+(4))
mean(cp_results_rodent5$saved_lls * -2)+ 2*(3*(ntopics-1)*(5+1)+(5))
mean(cp_results_rodent6$saved_lls * -2)+ 2*(3*(ntopics-1)*(5+1)+(5))

# lowest deviance is again for 4 (or 6) changepoints

rm(list = c('cp_results_rodent', 'cp_results_rodent2', 
            'cp_results_rodent3', 'cp_results_rodent5'))
hist(year_continuous[cp_results_rodent4$saved[,1,]],breaks = seq(1977,2016,.25),xlab='',main='Changepoint Estimate')
hist(year_continuous[cp_results_rodent6$saved[,1,]],breaks = seq(1977,2016,.25),xlab='',main='Changepoint Estimate')

annual_hist(cp_results_rodent4,year_continuous)

# turn changepoint results into data frame
df_4 = as.data.frame(t(cp_results_rodent4$saved[,1,])) %>% melt()
df_4$value = year_continuous[df_4$value]

# find 95% confidence intervals on each changepoint:
quantile(df_4[df_4$variable=='V1','value'],probs=c(.025,.975)) %>% date_decimal() %>% format('%d-%m-%Y')
quantile(df_4[df_4$variable=='V2','value'],probs=c(.025,.975)) %>% date_decimal() %>% format('%d-%m-%Y')
quantile(df_4[df_4$variable=='V3','value'],probs=c(.025,.975)) %>% date_decimal() %>% format('%d-%m-%Y')
quantile(df_4[df_4$variable=='V4','value'],probs=c(.025,.975)) %>% date_decimal() %>% format('%d-%m-%Y')

## 6 cpts 
annual_hist(cp_results_rodent6,year_continuous)
# turn changepoint results into data frame
df_6 = as.data.frame(t(cp_results_rodent6$saved[,1,])) %>% melt()
df_6$value = year_continuous[df_6$value]

# find 95% confidence intervals on each changepoint:
quantile(df_6[df_6$variable=='V1','value'],probs=c(.025,.975)) %>% date_decimal() %>% format('%d-%m-%Y')
quantile(df_6[df_6$variable=='V2','value'],probs=c(.025,.975)) %>% date_decimal() %>% format('%d-%m-%Y')
quantile(df_6[df_6$variable=='V3','value'],probs=c(.025,.975)) %>% date_decimal() %>% format('%d-%m-%Y')
quantile(df_6[df_6$variable=='V4','value'],probs=c(.025,.975)) %>% date_decimal() %>% format('%d-%m-%Y')
quantile(df_6[df_6$variable=='V5','value'],probs=c(.025,.975)) %>% date_decimal() %>% format('%d-%m-%Y')
quantile(df_6[df_6$variable=='V6','value'],probs=c(.025,.975)) %>% date_decimal() %>% format('%d-%m-%Y')
```

### 2018-12-27 paper-comparison/paper_lda_paper_changepoint.R

@@ -1,0 +1,125 @@
+library(LDATS)
+#### Data #### 
+
+data(rodents)
+
+ldats_dat <- rodents[[1]]
+ldats_dates <- rodents[[2]]
+head(ldats_dat)
+source('paper-comparison/christensen-ecology-files/rodent_data_for_LDA.r')
+dat = create_rodent_table(period_first = 1,
+                          period_last = 436,
+                          selected_plots = c(2,4,8,11,12,14,17,22),
+                          selected_species = c('BA','DM','DO','DS','NA','OL','OT','PB','PE','PF','PH','PI','PL','PM','PP','RF','RM','RO','SF','SH','SO'), diagnose = F)
+
+# dates to go with count data
+moondat = read.csv(text=getURL("https://raw.githubusercontent.com/weecology/PortalData/master/Rodents/moon_dates.csv"),stringsAsFactors = F)
+moondat$date = as.Date(moondat$censusdate)
+
+period_dates = filter(moondat,period %in% rownames(dat)) %>% select(period,date)
+dates = period_dates$date
+
+ch_dat <- dat
+
+ch_dates <- dates
+
+rm(list = c('dat', 'dates', 'period_dates', 'moondat',
+            'create_rodent_table'))
+
+ldats_dat <- as.matrix(ch_dat)
+ldats_dat <- apply(ldats_dat, c(1,2), FUN = as.integer)
+rodents$document_term_table <- ldats_dat
+rm(ldats_dat)
+rm(ldats_dates)
+rm(ch_dat)
+
+#### Load LDA ####
+load('/Users/renatadiaz/Documents/model-stash/paper_lda.Rds')
+
+#### Run changepoint ####
+source('paper-comparison/christensen-ecology-files/changepointmodel.r')
+# set up parameters for model
+year_continuous = 1970 + as.integer(julian(ch_dates)) / 365.25
+x = data.frame(
+  year_continuous = year_continuous,
+  sin_year = sin(year_continuous * 2 * pi),
+  cos_year = cos(year_continuous * 2 * pi)
+)
+
+# # run models with 1, 2, 3, 4, 5 changepoints
+# cp_results_rodent = changepoint_model(ldamodel, x, 1, weights = rep(1,length(year_continuous)))
[…]

### 2019-01-08 paper-comparison/ldats_lda_lda_changepoint.R
Results of LDATS cpt on LDATS LDA and paper LDA

@@ -1,0 +1,69 @@
+library(LDATS)
+#### Data #### 
+
+data(rodents)
+
+ldats_dat <- rodents[[1]]
+ldats_dates <- rodents[[2]]
+head(ldats_dat)
+source('paper-comparison/christensen-ecology-files/rodent_data_for_LDA.r')
+dat = create_rodent_table(period_first = 1,
+                          period_last = 436,
+                          selected_plots = c(2,4,8,11,12,14,17,22),
+                          selected_species = c('BA','DM','DO','DS','NA','OL','OT','PB','PE','PF','PH','PI','PL','PM','PP','RF','RM','RO','SF','SH','SO'), diagnose = F)
+
+# dates to go with count data
+moondat = read.csv(text=getURL("https://raw.githubusercontent.com/weecology/PortalData/master/Rodents/moon_dates.csv"),stringsAsFactors = F)
+moondat$date = as.Date(moondat$censusdate)
+
+period_dates = filter(moondat,period %in% rownames(dat)) %>% select(newmoonnumber, period,date)
+dates = period_dates$date
+
+ch_dat <- dat
+
+ch_dates <- dates
+ch_newmoons <- period_dates$newmoonnumber
+
+rm(list = c('dat', 'dates', 'period_dates', 'moondat',
+            'create_rodent_table'))
+
+ldats_dat <- as.matrix(ch_dat)
+ldats_dat <- apply(ldats_dat, c(1,2), FUN = as.integer)
+rodents$document_term_table <- ldats_dat
+rm(ldats_dat)
+rm(ldats_dates)
+rm(ch_dat)
+
+#### Load LDA ####
+load('/Users/renatadiaz/Documents/model-stash/ldats_lda.Rds')
+year_continuous = 1970 + as.integer(julian(ch_dates)) / 365.25
+x = data.frame(
+  year_continuous = year_continuous,
+  sin_year = sin(year_continuous * 2 * pi),
+  cos_year = cos(year_continuous * 2 * pi)
+)
+
+x$newmoon <- ch_newmoons
+x <- select(x, newmoon, sin_year, cos_year)
+
+#### Run LDATS changepoint ####
+
+ldats_ldats_cpt <- TS_on_LDA(LDA_models = ldats_lda_selected, 
+                             document_covariate_table = x,
+                             formulas = ~ sin_year + cos_year,
+                             nchangepoints = 1:6,
+                             weights = NULL)
+
+
[…]

### 2019-01-08 paper-comparison/ldats_changepoint.Rmd
Results of LDATS cpt on LDATS LDA and paper LDA

---
title: "LDATS changepoint"
author: "Renata Diaz"
date: "1/8/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(LDATS)
```

## LDATS LDA

### Setup
To run changepoint models: 
```{r setup ldats lda + ldats cpt, eval = T}
#### Data ####

data(rodents)

ldats_dat <- rodents[[1]]
ldats_dates <- rodents[[2]]
head(ldats_dat)
source('christensen-ecology-files/rodent_data_for_LDA.r')
dat = create_rodent_table(period_first = 1,
                          period_last = 436,
                          selected_plots = c(2,4,8,11,12,14,17,22),
                          selected_species = c('BA','DM','DO','DS','NA','OL','OT','PB','PE','PF','PH','PI','PL','PM','PP','RF','RM','RO','SF','SH','SO'), diagnose = F)

# dates to go with count data
moondat = read.csv(text=getURL("https://raw.githubusercontent.com/weecology/PortalData/master/Rodents/moon_dates.csv"),stringsAsFactors = F)
moondat$date = as.Date(moondat$censusdate)

period_dates = filter(moondat,period %in% rownames(dat)) %>% select(newmoonnumber, period,date)
dates = period_dates$date

ch_dat <- dat

ch_dates <- dates
ch_newmoons <- period_dates$newmoonnumber

rm(list = c('dat', 'dates', 'period_dates', 'moondat',
            'create_rodent_table'))

ldats_dat <- as.matrix(ch_dat)
ldats_dat <- apply(ldats_dat, c(1,2), FUN = as.integer)
rodents$document_term_table <- ldats_dat
rm(ldats_dat)
rm(ldats_dates)
rm(ch_dat)

#### Load LDA ####
load('/Users/renatadiaz/Documents/model-stash/ldats_lda.Rds')
year_continuous = 1970 + as.integer(julian(ch_dates)) / 365.25
x = data.frame(
  year_continuous = year_continuous,
  sin_year = sin(year_continuous * 2 * pi),
  cos_year = cos(year_continuous * 2 * pi)
)

x$newmoon <- ch_newmoons
x <- select(x, newmoon, sin_year, cos_year)
```

### Run the TS models
```{r run ldats lda + ldats cpt, eval = F}
#### Run LDATS changepoint ####

ldats_ldats_cpt <- TS_on_LDA(LDA_models = ldats_lda_selected, 
                             document_covariate_table = x,
                             formulas = ~ sin_year + cos_year,
                             nchangepoints = 1:6,
                             weights = NULL)

save(ldats_ldats_cpt, file = '/Users/renatadiaz/Documents/model-stash/ldats_lda_ldats_cpt.Rds')
rm(ldats_ldats_cpt)
rm(ldats_lda_selected)
```

```{r load ldats lda + ldats cpt results, echo = F}
load('/Users/renatadiaz/Documents/model-stash/ldats_lda_ldats_cpt.Rds')
```

### Select model

```{r select ldats lda + ldats cpt}

ldats_ldats_cpt_selected <- select_TS(ldats_ldats_cpt)
```

### Show results

```{r show ldats lda + ldats cpt}
plot(ldats_ldats_cpt_selected)
ldats_ldats_cpt_selected$formula
ldats_ldats_cpt_selected$nchangepoints
ldats_ldats_cpt_selected$rho_summary
```

```{r clean up ldats lda + ldats cpt, echo = F}
rm(ldats_ldats_cpt)
rm(ldats_ldats_cpt_selected)
rm(ldats_lda_selected)

```

## Paper LDA

### Load LDA
```{r load paper LDA}
#### Load paper LDA ####
load('/Users/renatadiaz/Documents/model-stash/paper_lda.Rds')
```

### Run TS models

```{r run paper LDA + ldats cpt, eval = F}
paper_ldats_cpt <- TS_on_LDA(LDA_models = ldamodel, document_covariate_table = x, formulas = ~ sin_year + cos_year,nchangepoints = 1:6, weights = NULL)
save(paper_ldats_cpt, file = '/Users/renatadiaz/Documents/model-stash/paper_lda_ldats_cpt.Rds')
```

```{r load paper LDA + ldats cpt, echo = F}
load('/Users/renatadiaz/Documents/model-stash/paper_lda_ldats_cpt.Rds')
```

```{r select paper lda + ldats cpt}
paper_ldats_cpt_selected <- select_TS(paper_ldats_cpt)
```

### Show results

```{r show paper lda + ldats cpt}
plot(paper_ldats_cpt_selected)
paper_ldats_cpt_selected$formula
paper_ldats_cpt_selected$nchangepoints
paper_ldats_cpt_selected$rho_summary
```

### 2019-01-08 paper-comparison/ldats_changepoint.Rmd
Report changepoint estimates as dates

@@
 
-rm(list = c('dat', 'dates', 'period_dates', 'moondat',
+rm(list = c('dat', 'dates',
             'create_rodent_table'))
@@
 ldats_ldats_cpt_selected$nchangepoints
-ldats_ldats_cpt_selected$rho_summary
+cpt_means <- ldats_ldats_cpt_selected$rho_summary$Mean
+
+cpt_means <- floor(cpt_means)
+cpt_means <- as.data.frame(cpt_means)
+colnames(cpt_means) <- 'newmoonnumber'
+cpt_means$cpt <- 1:ldats_ldats_cpt_selected$nchangepoints
+
+cpt_dates <- as.data.frame(moondat) %>%
+  dplyr::select(newmoonnumber, newmoondate) %>%
+  dplyr::left_join(cpt_means, by = 'newmoonnumber') %>%
+  dplyr::filter(!is.na(cpt))
+
+cpt_dates
+
 ```
@@
 paper_ldats_cpt_selected$nchangepoints
-paper_ldats_cpt_selected$rho_summary
+cpt_means <- paper_ldats_cpt_selected$rho_summary$Mean
+
+cpt_means <- floor(cpt_means)
+cpt_means <- as.data.frame(cpt_means)
+colnames(cpt_means) <- 'newmoonnumber'
+cpt_means$cpt <- 1:paper_ldats_cpt_selected$nchangepoints
+
+cpt_dates <- as.data.frame(moondat) %>%
+  dplyr::select(newmoonnumber, newmoondate) %>%
+  dplyr::left_join(cpt_means, by = 'newmoonnumber') %>%
+  dplyr::filter(!is.na(cpt))
+
+cpt_dates
 ```

### 2019-01-08 paper-comparison/paper_changepoint.Rmd
Report mean cpt estimates for paper changepoint

@@
 df_4$value = year_continuous[df_4$value]
+
+# find means on each changepoint:
+mean(df_4[df_4$variable=='V1','value']) %>% date_decimal() %>% format('%d-%m-%Y')
+mean(df_4[df_4$variable=='V2','value']) %>% date_decimal() %>% format('%d-%m-%Y')
+mean(df_4[df_4$variable=='V3','value']) %>% date_decimal() %>% format('%d-%m-%Y')
+mean(df_4[df_4$variable=='V4','value']) %>% date_decimal() %>% format('%d-%m-%Y')
 
@@
 
+# find means on each changepoint:
+mean(df_4[df_4$variable=='V1','value']) %>% date_decimal() %>% format('%d-%m-%Y')
+mean(df_4[df_4$variable=='V2','value']) %>% date_decimal() %>% format('%d-%m-%Y')
+mean(df_4[df_4$variable=='V3','value']) %>% date_decimal() %>% format('%d-%m-%Y')
+mean(df_4[df_4$variable=='V4','value']) %>% date_decimal() %>% format('%d-%m-%Y')
+
 # find 95% confidence intervals on each changepoint:

### 2019-01-08 paper-comparison/paper_changepoint.Rmd
table summarizing cpt outcomes

@@
 mean(df_4[df_4$variable=='V4','value']) %>% date_decimal() %>% format('%d-%m-%Y')
+
+cpt_dates <- vector(length = 4)
+cpt_dates[1] <- mean(df_4[df_4$variable=='V1','value']) %>% date_decimal() %>% format('%d-%m-%Y')
+cpt_dates[2] <-mean(df_4[df_4$variable=='V2','value']) %>% date_decimal() %>% format('%d-%m-%Y')
+cpt_dates[3] <-mean(df_4[df_4$variable=='V3','value']) %>% date_decimal() %>% format('%d-%m-%Y')
+cpt_dates[4] <-mean(df_4[df_4$variable=='V4','value']) %>% date_decimal() %>% format('%d-%m-%Y')
+
+cpt_dates <- as.data.frame(cpt_dates)
+write.csv(cpt_dates, 'ldats_paper_dates.csv')
 
@@
 # find means on each changepoint:
+
+# find means on each changepoint:
 mean(df_4[df_4$variable=='V1','value']) %>% date_decimal() %>% format('%d-%m-%Y')
@@
 mean(df_4[df_4$variable=='V4','value']) %>% date_decimal() %>% format('%d-%m-%Y')
+
+cpt_dates <- vector(length = 4)
+cpt_dates[1] <- mean(df_4[df_4$variable=='V1','value']) %>% date_decimal() %>% format('%d-%m-%Y')
+cpt_dates[2] <-mean(df_4[df_4$variable=='V2','value']) %>% date_decimal() %>% format('%d-%m-%Y')
+cpt_dates[3] <-mean(df_4[df_4$variable=='V3','value']) %>% date_decimal() %>% format('%d-%m-%Y')
+cpt_dates[4] <-mean(df_4[df_4$variable=='V4','value']) %>% date_decimal() %>% format('%d-%m-%Y')
+
+cpt_dates <- as.data.frame(cpt_dates)
+write.csv(cpt_dates, 'paper_paper_dates.csv')
 

### 2019-01-08 paper-comparison/compare-christensen.Rmd
table summarizing cpt outcomes

--- 2018-12-26 paper-comparison/compare-christensen.Rmd
@@
 
-#### run LDATS LDA with LDATS changepoint
-```{r, include = F}
-# remove paper LDA (for memory)
-rm(ldamodel)
-```
+For narratives running all these models, see `ldats_changepoint.md` and `paper_changepoint.md`.
 
-```{r LDATS LDA + LDATS cpt, eval = F}
-# Run changepoint
+All the models found 4 changepoints (although the paper model actually finds lower deviances for 6 changepoints, 6 was outside the range considered by the paper/anticipated by LDATS, so I'm sticking with 4 for now)
 
-ldats_ldats <- LDATS::TS_on_LDA(LDA_models = ldats_lda_selected,
-                                document_covariate_table = rodents[[2]],
-                                formulas =c(~ sin_year + cos_year),
-                                nchangepoints = c(2, 3),
-                                weights = NULL)
+### Compare model outcomes
 
-save(ldats_ldats, file = 'model-stash/ldats_ldats.Rds')
+```{r read model results}
+ldats_ldats <- read.csv('ldats_ldats_dates.csv')
+ldats_paper <- read.csv('ldats_paper_dates.csv')
+paper_ldats <- read.csv('paper_ldats_dates.csv')
+paper_paper <- read.csv('paper_paper_dates.csv')
 
-# Select best changepoint
+estimates <- c(1:4)
+estimates <- as.data.frame(estimates)
 
-ldats_ldats_selected <- LDATS::select_TS(TS_models = ldats_ldats)
+# Naming convention: ldamethod_cptmethod
+# so ldats_paper is the ldats lda and paper cpt, etc
+estimates$ldats_ldats <- ldats_ldats$newmoondate
+estimates$ldats_paper <- ldats_paper$cpt_dates
+estimates$paper_ldats <- paper_ldats$newmoondate
+estimates$paper_paper <- paper_paper$cpt_dates
 
-save(ldats_ldats_selected, file = '/Users/renatadiaz/Documents/model-stash/ldats_ldats_selected.Rds')
+estimates
 
 ```
-
-```{r, eval = F}
-rm(ldats_ldats)
-rm(ldats_ldats_selected)
-```
-
-#### run LDATS LDA with paper changepoint
-
-```{r ldats paper, eval = F}
-source('christensen-ecology-files/changepointmodel.r')
-# set up parameters for model
-year_continuous = 1970 + as.integer(julian(ch_dates)) / 365.25
-x = data.frame(
-  year_continuous = year_continuous,
-  sin_year = sin(year_continuous * 2 * pi),
-  cos_year = cos(year_continuous * 2 * pi)
-)
-
-# run models with 1, 2, 3, 4, 5 changepoints
-cp_results_rodent = changepoint_model(ldats_lda_selected[[1]], x, 1, weights = rep(1,length(year_continuous)))
-
-save(cp_results_rodent, file = '/Users/renatadiaz/Documents/model-stash/paper_cpt_ldats_lda1.Rds')
-rm(cp_results_rodent)
-
-cp_results_rodent2 = changepoint_model(ldats_lda_selected[[1]], x, 2, weights = rep(1,length(year_continuous)))
-
-save(cp_results_rodent2, file = '/Users/renatadiaz/Documents/model-stash/paper_cpt_ldats_lda2.Rds')
-rm(cp_results_rodent2)
-
-cp_results_rodent3 = changepoint_model(ldats_lda_selected[[1]], x, 3, weights = rep(1,length(year_continuous)))
-save(cp_results_rodent3, file = '/Users/renatadiaz/Documents/model-stash/paper_cpt_ldats_lda3.Rds')
-rm(cp_results_rodent3)
-
-cp_results_rodent4 = changepoint_model(ldats_lda_selected[[1]], x, 4, weights = rep(1,length(year_continuous)))
-save(cp_results_rodent4, file = '/Users/renatadiaz/Documents/model-stash/paper_cpt_ldats_lda4.Rds')
-rm(cp_results_rodent4)
-
-cp_results_rodent5 = changepoint_model(ldats_lda_selected[[1]], x, 5, weights = rep(1,length(year_continuous)))
-save(cp_results_rodent5, file = '/Users/renatadiaz/Documents/model-stash/paper_cpt_ldats_lda5.Rds')
-rm(cp_results_rodent5)
-
-cp_results_rodent6 = changepoint_model(ldats_lda_selected[[1]], x, 6, weights = rep(1,length(year_continuous)))
-save(cp_results_rodent6, file = '/Users/renatadiaz/Documents/model-stash/paper_cpt_ldats_lda6.Rds')
-rm(cp_results_rodent6)
-
-rm(ldats_lda_selected)
-```
-
-#### run paper LDA with paper changepoint
-```{r paper paper, eval = F}
-source('christensen-ecology-files/changepointmodel.r')
-# set up parameters for model
-year_continuous = 1970 + as.integer(julian(ch_dates)) / 365.25
-x = data.frame(
-year_continuous = year_continuous,
-sin_year = sin(year_continuous * 2 * pi),
-cos_year = cos(year_continuous * 2 * pi)
-)
-# run models with 1, 2, 3, 4, 5 changepoints
-cp_results_rodent = changepoint_model(ldamodel, x, 1, weights = rep(1,length(year_continuous)))
-save(cp_results_rodent, file = '/Users/renatadiaz/Documents/model-stash/paper_cpt_paper_lda1.Rds')
-rm(cp_results_rodent)
-
-cp_results_rodent2 = changepoint_model(ldamodel, x, 2, weights = rep(1,length(year_continuous)))
-save(cp_results_rodent2, file = '/Users/renatadiaz/Documents/model-stash/paper_cpt_paper_lda2.Rds')
-rm(cp_results_rodent2)
-
-cp_results_rodent3 = changepoint_model(ldamodel, x, 3, weights = rep(1,length(year_continuous)))
-save(cp_results_rodent3, file = '/Users/renatadiaz/Documents/model-stash/paper_cpt_paper_lda3.Rds')
-rm(cp_results_rodent3)
-
-cp_results_rodent4 = changepoint_model(ldamodel, x, 4, weights = rep(1,length(year_continuous)))
-save(cp_results_rodent4, file = '/Users/renatadiaz/Documents/model-stash/paper_cpt_paper_lda4.Rds')
-rm(cp_results_rodent4)
-
-cp_results_rodent5 = changepoint_model(ldamodel, x, 5, weights = rep(1,length(year_continuous)))
-save(cp_results_rodent5, file = '/Users/renatadiaz/Documents/model-stash/paper_cpt_paper_lda5.Rds')
-rm(cp_results_rodent5)
-
-cp_results_rodent6 = changepoint_model(ldamodel, x, 6, weights = rep(1,length(year_continuous)))
-save(cp_results_rodent6, file = '/Users/renatadiaz/Documents/model-stash/paper_cpt_paper_lda6.Rds')
-rm(cp_results_rodent6)
-```
-
-#### run paper LDA with LDATS changepoint
-```{r paper ldats, eval = F}
-# Run changepoint
-
-paper_ldats <- LDATS::TS_on_LDA(LDA_models = ldamodel,
-                                document_covariate_table = rodents[[2]],
-                                formulas =c(~ sin_year + cos_year),
-                                nchangepoints = c(2, 3, 4, 5, 6),
-                                weights = NULL)
-
-save(paper_ldats, file = 'model-stash/paper_ldats.Rds')
-
-# Select best changepoint
-
-paper_ldats_selected <- LDATS::select_TS(TS_models = paper_ldats)
-
-save(paper_ldats_selected, file = '/Users/renatadiaz/Documents/model-stash/paper_ldats_selected.Rds')
-
-```

### 2019-01-09 paper-comparison/compare-christensen.Rmd
update compare-christensen
Correct typo in paper changepoint model selection
6 cpts is not, in fact, ever the best model.

@@
 
-All the models found 4 changepoints (although the paper model actually finds lower deviances for 6 changepoints, 6 was outside the range considered by the paper/anticipated by LDATS, so I'm sticking with 4 for now)
+All the models found 4 changepoints.
 
@@
 ```
+
+All of the outcomes are reasonably close.
+
+The choice of changepoint model seems to matter less than the choice of LDA model.

### 2019-01-09 paper-comparison/paper_changepoint.Rmd
Correct typo in paper changepoint model selection
6 cpts is not, in fact, ever the best model.

@@
 mean(cp_results_rodent5$saved_lls * -2)+ 2*(3*(ntopics-1)*(5+1)+(5))
-mean(cp_results_rodent6$saved_lls * -2)+ 2*(3*(ntopics-1)*(5+1)+(5))
+mean(cp_results_rodent6$saved_lls * -2)+ 2*(3*(ntopics-1)*(6+1)+(6))
 
-# lowest deviance is for 4 (or 6) changepoints
+# lowest deviance is for 4 changepoints
 
 rm(list = c('cp_results_rodent', 'cp_results_rodent2', 
-            'cp_results_rodent3', 'cp_results_rodent5'))
+            'cp_results_rodent3', 'cp_results_rodent5',
+            'cp_results_rodent6'))
 hist(year_continuous[cp_results_rodent4$saved[,1,]],breaks = seq(1977,2016,.25),xlab='',main='Changepoint Estimate')
-hist(year_continuous[cp_results_rodent6$saved[,1,]],breaks = seq(1977,2016,.25),xlab='',main='Changepoint Estimate')
 
@@
 quantile(df_4[df_4$variable=='V4','value'],probs=c(.025,.975)) %>% date_decimal() %>% format('%d-%m-%Y')
-
-## 6 cpts 
-annual_hist(cp_results_rodent6,year_continuous)
-# turn changepoint results into data frame
-df_6 = as.data.frame(t(cp_results_rodent6$saved[,1,])) %>% melt()
-df_6$value = year_continuous[df_6$value]
-
-# find 95% confidence intervals on each changepoint:
-quantile(df_6[df_6$variable=='V1','value'],probs=c(.025,.975)) %>% date_decimal() %>% format('%d-%m-%Y')
-quantile(df_6[df_6$variable=='V2','value'],probs=c(.025,.975)) %>% date_decimal() %>% format('%d-%m-%Y')
-quantile(df_6[df_6$variable=='V3','value'],probs=c(.025,.975)) %>% date_decimal() %>% format('%d-%m-%Y')
-quantile(df_6[df_6$variable=='V4','value'],probs=c(.025,.975)) %>% date_decimal() %>% format('%d-%m-%Y')
-quantile(df_6[df_6$variable=='V5','value'],probs=c(.025,.975)) %>% date_decimal() %>% format('%d-%m-%Y')
-quantile(df_6[df_6$variable=='V6','value'],probs=c(.025,.975)) %>% date_decimal() %>% format('%d-%m-%Y')
 
@@
 mean(cp_results_rodent5$saved_lls * -2)+ 2*(3*(ntopics-1)*(5+1)+(5))
-mean(cp_results_rodent6$saved_lls * -2)+ 2*(3*(ntopics-1)*(5+1)+(5))
+mean(cp_results_rodent6$saved_lls * -2)+ 2*(3*(ntopics-1)*(6+1)+(6))
 
-# lowest deviance is again for 4 (or 6) changepoints
+# lowest deviance is again for 4 changepoints
 
 rm(list = c('cp_results_rodent', 'cp_results_rodent2', 
-            'cp_results_rodent3', 'cp_results_rodent5'))
+            'cp_results_rodent3', 'cp_results_rodent5',
+            'cp_results_rodent6'))
 hist(year_continuous[cp_results_rodent4$saved[,1,]],breaks = seq(1977,2016,.25),xlab='',main='Changepoint Estimate')
-hist(year_continuous[cp_results_rodent6$saved[,1,]],breaks = seq(1977,2016,.25),xlab='',main='Changepoint Estimate')
 
@@
 
-## 6 cpts 
-annual_hist(cp_results_rodent6,year_continuous)
-# turn changepoint results into data frame
-df_6 = as.data.frame(t(cp_results_rodent6$saved[,1,])) %>% melt()
-df_6$value = year_continuous[df_6$value]
-
-# find 95% confidence intervals on each changepoint:
-quantile(df_6[df_6$variable=='V1','value'],probs=c(.025,.975)) %>% date_decimal() %>% format('%d-%m-%Y')
-quantile(df_6[df_6$variable=='V2','value'],probs=c(.025,.975)) %>% date_decimal() %>% format('%d-%m-%Y')
-quantile(df_6[df_6$variable=='V3','value'],probs=c(.025,.975)) %>% date_decimal() %>% format('%d-%m-%Y')
-quantile(df_6[df_6$variable=='V4','value'],probs=c(.025,.975)) %>% date_decimal() %>% format('%d-%m-%Y')
-quantile(df_6[df_6$variable=='V5','value'],probs=c(.025,.975)) %>% date_decimal() %>% format('%d-%m-%Y')
-quantile(df_6[df_6$variable=='V6','value'],probs=c(.025,.975)) %>% date_decimal() %>% format('%d-%m-%Y')
 ```

### 2019-01-14 paper-comparison/paper_lda_paper_changepoint.R
Begin LDA comparison and fix model stash path
start to compare ldas with adjusted/nonadjusted data
store model objects in dropbox

@@ -34,7 +34,7 @@ rm(ldats_dates)
 rm(ch_dat)
 
 #### Load LDA ####
-load('/Users/renatadiaz/Documents/model-stash/paper_lda.Rds')
+load('/~/Dropbox/ldats-models/paper_lda.Rds')
 
 #### Run changepoint ####
 source('paper-comparison/christensen-ecology-files/changepointmodel.r')
@@ -48,39 +48,39 @@ x = data.frame(
 
 # # run models with 1, 2, 3, 4, 5 changepoints
 # cp_results_rodent = changepoint_model(ldamodel, x, 1, weights = rep(1,length(year_continuous)))
-# save(cp_results_rodent, file = '/Users/renatadiaz/Documents/model-stash/paper_cpt_paper_lda1.Rds')
+# save(cp_results_rodent, file = '/~/Dropbox/ldats-models/paper_cpt_paper_lda1.Rds')
 # rm(cp_results_rodent)
 # 
 # 
 # cp_results_rodent2 = changepoint_model(ldamodel, x, 2, weights = rep(1,length(year_continuous)))
-# save(cp_results_rodent2, file = '/Users/renatadiaz/Documents/model-stash/paper_cpt_paper_lda2.Rds')
+# save(cp_results_rodent2, file = '/~/Dropbox/ldats-models/paper_cpt_paper_lda2.Rds')
 # rm(cp_results_rodent2)
 # 
 # cp_results_rodent3 = changepoint_model(ldamodel, x, 3, weights = rep(1,length(year_continuous)))
-# save(cp_results_rodent3, file = '/Users/renatadiaz/Documents/model-stash/paper_cpt_paper_lda3.Rds')
+# save(cp_results_rodent3, file = '/~/Dropbox/ldats-models/paper_cpt_paper_lda3.Rds')
 # rm(cp_results_rodent3)
 # 
 # cp_results_rodent4 = changepoint_model(ldamodel, x, 4, weights = rep(1,length(year_continuous)))
-# save(cp_results_rodent4, file = '/Users/renatadiaz/Documents/model-stash/paper_cpt_paper_lda4.Rds')
+# save(cp_results_rodent4, file = '/~/Dropbox/ldats-models/paper_cpt_paper_lda4.Rds')
 # rm(cp_results_rodent4)
 # 
 # cp_results_rodent5 = changepoint_model(ldamodel, x, 5, weights = rep(1,length(year_continuous)))
[…]

### 2019-01-14 paper-comparison/compare-christensen.R
Begin LDA comparison and fix model stash path
start to compare ldas with adjusted/nonadjusted data
store model objects in dropbox

@@ -12,17 +12,17 @@ knitr::opts_chunk$set(
 library(LDATS)
 
 ## ----download Christensen, eval = FALSE----------------------------------
-## 
+##
 ## files_to_download <- c('rodent_LDA_analysis.r', 'rodent_data_for_LDA.r', 'AIC_model_selection.R', 'changepointmodel.r', 'LDA-distance.R', 'Rodent_table_dat.csv', 'LDA_figure_scripts.R')
-## 
+##
 ## for(i in 1:length(files_to_download)) {
 ##   download.file(url = paste0("https://raw.githubusercontent.com/emchristensen/Extreme-events-LDA/master/", files_to_download[i]),
 ##                 destfile = paste0('christensen-ecology-files/', files_to_download[i]))
 ## }
-## 
+##
 ## rm(files_to_download)
 ## rm(i)
-## 
+##
 
 ## ----LDATS data----------------------------------------------------------
 
@@ -56,11 +56,11 @@ rm(list = c('dat', 'dates', 'period_dates', 'moondat',
 
 
 ## ----rodent data comparison, eval = F------------------------------------
-## 
+##
 ## compare <- ldats_dat == ch_dat
-## 
+##
 ## compare_rows <- vector(length = nrow(compare))
-## 
+##
 ## for(i in 1:nrow(compare)) {
 ##   if(sum(compare[i, ]) == 21) {
 ##     compare_rows[i] <- TRUE
@@ -68,64 +68,157 @@ rm(list = c('dat', 'dates', 'period_dates', 'moondat',
 ##     compare_rows[i] <- FALSE
 ##   }
 ## }
-## 
+##
 ## length(which(compare_rows == F))
 ## ldats_dat[ which(compare_rows == F), ]
 ## ch_dat[ which(compare_rows == F), ]
 
 ## ----Ch rodent adjustment, eval = F--------------------------------------
-## 
+##
 ## source('christensen-ecology-files/rodent_data_for_LDA.r')
 ## diagnose_dat = create_rodent_table(period_first = 1,
 ##                           period_last = 436,
 ##                           selected_plots = c(2,4,8,11,12,14,17,22),
 ##                           selected_species = c('BA','DM','DO','DS','NA','OL','OT','PB','PE','PF','PH','PI','PL','PM','PP','RF','RM','RO','SF','SH','SO'), diagnose = T)
-## 
+##
[…]

### 2019-01-14 paper-comparison/compare-christensen.Rmd
Begin LDA comparison and fix model stash path
start to compare ldas with adjusted/nonadjusted data
store model objects in dropbox

@@
 
-For now I will use the *adjusted rodent table* because this will avoid artificially low abundances for incompletely trapped periods.
+```{r adjusted rodent data}
 
-```{r switch to adjusted rodent data}
+ldats_dat_adj <- as.matrix(ch_dat)
+ldats_dat_adj <- apply(ldats_dat, c(1,2), FUN = as.integer)
+rodents_adj <- rodents
+rodents_adj$document_term_table <- ldats_dat_adj
+rm(ldats_dat_adj)
+rm(ch_dat)
+```
 
-ldats_dat <- as.matrix(ch_dat)
-ldats_dat <- apply(ldats_dat, c(1,2), FUN = as.integer)
-rodents$document_term_table <- ldats_dat
-rm(ldats_dat)
-rm(ldats_dates)
-rm(ch_dat)
+In principle, whether we use the adjusted or the non-adjusted rodent table should not affect the LDA results.
+This is because the LDA operates on proportional abundance, not total abundance; a community of abundances 1, 1, 2 would have the same proportions (25/25/50) as a community of abundances 2, 2, 4.
+However, it's possible that rounding errors could introduce discrepancies between LDAs run on adjusted vs.
+non-adjusted data.
+
+Exploring/illustrating this:
+
+```{r adj vs non adj LDAS, eval = F}
+
+# paper LDA w adjusted data
+source('christensen-ecology-files/AIC_model_selection.R')
+source('christensen-ecology-files/LDA-distance.R')
+
+# Fit a bunch of LDA models with different seeds
+# Only use even numbers for seeds because consecutive seeds give identical results
+seeds = 2*seq(200)
+
+# repeat LDA model fit and AIC calculation with a bunch of different seeds to test robustness of the analysis
+best_ntopic = repeat_VEM(rodents_adj[[1]],
+                         seeds,
+                         topic_min=2,
+                         topic_max=6)
+hist(best_ntopic$k,breaks=c(0.5,1.5,2.5,3.5,4.5,5.5,6.5,7.5,8.5,9.5),xlab='best # of topics', main='')
+
+# 2b. how different is species composition of 4 community-types when LDA is run with different seeds?
+# ==================================================================
+# get the best 100 seeds where 4 topics was the best LDA model
+seeds_4topics = best_ntopic %>% 
+  filter(k == 4) %>% 
+  arrange(aic) %>% 
+  head(100) %>% 
+  pull(SEED)
+
+# choose seed with highest log likelihood for all following analyses
+#    (also produces plot of community composition for 'best' run compared to 'worst')
+best_seed = calculate_LDA_distance(rodents[[1]],seeds_4topics)
+mean_dist = unlist(best_seed)[2]
+max_dist = unlist(best_seed)[3]
+
+# ==================================================================
+# 3. run LDA model
+# ==================================================================
+ntopics = 4
+SEED = unlist(best_seed)[1]  # For the paper, I use seed 206
+ldamodel_adj = LDA(rodents[[1]],ntopics, control = list(seed = SEED),method='VEM')
+
+plot(ldamodel_adj)
+
+# paper LDA w nonadjusted data
+
+# Fit a bunch of LDA models with different seeds
+# Only use even numbers for seeds because consecutive seeds give identical results
+seeds = 2*seq(200)
+
+# repeat LDA model fit and AIC calculation with a bunch of different seeds to test robustness of the analysis
+best_ntopic = repeat_VEM(rodents[[1]],
+                         seeds,
+                         topic_min=2,
+                         topic_max=6)
+hist(best_ntopic$k,breaks=c(0.5,1.5,2.5,3.5,4.5,5.5,6.5,7.5,8.5,9.5),xlab='best # of topics', main='')
+
+# 2b. how different is species composition of 4 community-types when LDA is run with different seeds?
+# ==================================================================
+# get the best 100 seeds where 4 topics was the best LDA model
+seeds_4topics = best_ntopic %>% 
+  filter(k == 4) %>% 
+  arrange(aic) %>% 
+  head(100) %>% 
+  pull(SEED)
+
+# choose seed with highest log likelihood for all following analyses
+#    (also produces plot of community composition for 'best' run compared to 'worst')
+best_seed = calculate_LDA_distance(rodents[[1]],seeds_4topics)
+mean_dist = unlist(best_seed)[2]
+max_dist = unlist(best_seed)[3]
+
+# ==================================================================
+# 3. run LDA model
+# ==================================================================
+ntopics = 4
+SEED = unlist(best_seed)[1]  # For the paper, I use seed 206
+ldamodel_nonadj = LDA(rodents[[1]],ntopics, control = list(seed = SEED),method='VEM')
+
+plot(ldamodel_nonadj)
+
+# ldats LDA w adjusted data
+
+ldats_adj_ldas <- LDATS::LDA_set(document_term_table = rodents_adj$document_term_table, topics = c(2:6), nseeds = 100)
+ldats_adj_lda_selected <- LDATS::select_LDA(LDA_models = ldats_adj_ldas)
+
+plot(ldats_adj_lda_selected)
+
+# ldats LDA w nonadjusted data
+
+ldats_nonadj_ldas <- LDATS::LDA_set(document_term_table = rodents$document_term_table, topics = c(2:6), nseeds = 100)
+ldats_nonadj_lda_selected <- LDATS::select_LDA(LDA_models = ldats_ldas)
+
+plot(ldats_nonadj_lda_selected)
+
 ```
@@
 ```{r save LDAs, eval = F, include = F}
-save(ldamodel, file = 'model-stash/paper_lda.Rds')
-save(ldats_lda_selected, file = 'model-stash/ldats_lda.Rds')
+save(ldamodel, file = '~/Dropbox/ldats-models/paper_lda.Rds')
+save(ldats_lda_selected, file = '~/Dropbox/ldats-models/ldats_lda.Rds')
 ```
@@
 
-load('model-stash/ldats_lda.Rds')
-load('model-stash/paper_lda.Rds')
+load('~/Dropbox/ldats-models/ldats_lda.Rds')
+load('~/Dropbox/ldats-models/paper_lda.Rds')
 

### 2019-01-14 paper-comparison/ldats_changepoint.Rmd
table summarizing cpt outcomes
Begin LDA comparison and fix model stash path
start to compare ldas with adjusted/nonadjusted data
store model objects in dropbox

@@
 #### Load LDA ####
-load('/Users/renatadiaz/Documents/model-stash/ldats_lda.Rds')
+load('~/Dropbox/ldats-models/ldats_lda.Rds')
 year_continuous = 1970 + as.integer(julian(ch_dates)) / 365.25
@@
 
-save(ldats_ldats_cpt, file = '/Users/renatadiaz/Documents/model-stash/ldats_lda_ldats_cpt.Rds')
+save(ldats_ldats_cpt, file = '~/Dropbox/ldats-models/ldats_lda_ldats_cpt.Rds')
 rm(ldats_ldats_cpt)
@@
 ```{r load ldats lda + ldats cpt results, echo = F}
-load('/Users/renatadiaz/Documents/model-stash/ldats_lda_ldats_cpt.Rds')
+load('~/Dropbox/ldats-models/ldats_lda_ldats_cpt.Rds')
 ```
@@
 
+write.csv(cpt_dates, 'ldats_ldats_dates.csv')
+
 ```
@@
 #### Load paper LDA ####
-load('/Users/renatadiaz/Documents/model-stash/paper_lda.Rds')
+load('~/Dropbox/ldats-models/paper_lda.Rds')
 ```
@@
 paper_ldats_cpt <- TS_on_LDA(LDA_models = ldamodel, document_covariate_table = x, formulas = ~ sin_year + cos_year,nchangepoints = 1:6, weights = NULL)
-save(paper_ldats_cpt, file = '/Users/renatadiaz/Documents/model-stash/paper_lda_ldats_cpt.Rds')
+save(paper_ldats_cpt, file = '~/Dropbox/ldats-models/paper_lda_ldats_cpt.Rds')
 ```
@@
 ```{r load paper LDA + ldats cpt, echo = F}
-load('/Users/renatadiaz/Documents/model-stash/paper_lda_ldats_cpt.Rds')
+load('~/Dropbox/ldats-models/paper_lda_ldats_cpt.Rds')
 ```
@@
 cpt_dates
+
+write.csv(cpt_dates, 'paper_ldats_dates.csv')
+
 ```

### 2019-01-14 paper-comparison/paper_changepoint.Rmd
Begin LDA comparison and fix model stash path
start to compare ldas with adjusted/nonadjusted data
store model objects in dropbox

@@
 #### Load LDA ####
-load('/Users/renatadiaz/Documents/model-stash/ldats_lda.Rds')
+load('~/Dropbox/ldats-models/ldats_lda.Rds')
 
@@
 # cp_results_rodent = changepoint_model(ldats_lda_selected[[1]], x, 1, weights = rep(1,length(year_continuous)))
-# save(cp_results_rodent, file = '/Users/renatadiaz/Documents/model-stash/paper_cpt_ldats_lda1.Rds')
+# save(cp_results_rodent, file = '~/Dropbox/ldats-models/paper_cpt_ldats_lda1.Rds')
 # rm(cp_results_rodent)
@@
 # cp_results_rodent2 = changepoint_model(ldats_lda_selected[[1]], x, 2, weights = rep(1,length(year_continuous)))
-# save(cp_results_rodent2, file = '/Users/renatadiaz/Documents/model-stash/paper_cpt_ldats_lda2.Rds')
+# save(cp_results_rodent2, file = '~/Dropbox/ldats-models/paper_cpt_ldats_lda2.Rds')
 # rm(cp_results_rodent2)
@@
 # cp_results_rodent3 = changepoint_model(ldats_lda_selected[[1]], x, 3, weights = rep(1,length(year_continuous)))
-# save(cp_results_rodent3, file = '/Users/renatadiaz/Documents/model-stash/paper_cpt_ldats_lda3.Rds')
+# save(cp_results_rodent3, file = '~/Dropbox/ldats-models/paper_cpt_ldats_lda3.Rds')
 # rm(cp_results_rodent3)
@@
 # cp_results_rodent4 = changepoint_model(ldats_lda_selected[[1]], x, 4, weights = rep(1,length(year_continuous)))
-# save(cp_results_rodent4, file = '/Users/renatadiaz/Documents/model-stash/paper_cpt_ldats_lda4.Rds')
+# save(cp_results_rodent4, file = '~/Dropbox/ldats-models/paper_cpt_ldats_lda4.Rds')
 # rm(cp_results_rodent4)
@@
 # cp_results_rodent5 = changepoint_model(ldats_lda_selected[[1]], x, 5, weights = rep(1,length(year_continuous)))
-# save(cp_results_rodent5, file = '/Users/renatadiaz/Documents/model-stash/paper_cpt_ldats_lda5.Rds')
+# save(cp_results_rodent5, file = '~/Dropbox/ldats-models/paper_cpt_ldats_lda5.Rds')
 # rm(cp_results_rodent5)
@@
 # cp_results_rodent6 = changepoint_model(ldats_lda_selected[[1]], x, 6, weights = rep(1,length(year_continuous)))
-# save(cp_results_rodent6, file = '/Users/renatadiaz/Documents/model-stash/paper_cpt_ldats_lda6.Rds')
+# save(cp_results_rodent6, file = '~/Dropbox/ldats-models/paper_cpt_ldats_lda6.Rds')
 # rm(cp_results_rodent6)
@@
 #### Changepoint model selection ####
-load('/Users/renatadiaz/Documents/model-stash/paper_cpt_ldats_lda1.Rds')
-load('/Users/renatadiaz/Documents/model-stash/paper_cpt_ldats_lda2.Rds')
-load('/Users/renatadiaz/Documents/model-stash/paper_cpt_ldats_lda3.Rds')
-load('/Users/renatadiaz/Documents/model-stash/paper_cpt_ldats_lda4.Rds')
-load('/Users/renatadiaz/Documents/model-stash/paper_cpt_ldats_lda5.Rds')
-load('/Users/renatadiaz/Documents/model-stash/paper_cpt_ldats_lda6.Rds')
+load('~/Dropbox/ldats-models/paper_cpt_ldats_lda1.Rds')
+load('~/Dropbox/ldats-models/paper_cpt_ldats_lda2.Rds')
+load('~/Dropbox/ldats-models/paper_cpt_ldats_lda3.Rds')
+load('~/Dropbox/ldats-models/paper_cpt_ldats_lda4.Rds')
+load('~/Dropbox/ldats-models/paper_cpt_ldats_lda5.Rds')
+load('~/Dropbox/ldats-models/paper_cpt_ldats_lda6.Rds')
 
@@
 #### Load LDA ####
-load('/Users/renatadiaz/Documents/model-stash/paper_lda.Rds')
+load('~/Dropbox/ldats-models/paper_lda.Rds')
 
@@
 # cp_results_rodent = changepoint_model(ldamodel, x, 1, weights = rep(1,length(year_continuous)))
-# save(cp_results_rodent, file = '/Users/renatadiaz/Documents/model-stash/paper_cpt_paper_lda1.Rds')
+# save(cp_results_rodent, file = '~/Dropbox/ldats-models/paper_cpt_paper_lda1.Rds')
 # rm(cp_results_rodent)
@@
 # cp_results_rodent2 = changepoint_model(ldamodel, x, 2, weights = rep(1,length(year_continuous)))
-# save(cp_results_rodent2, file = '/Users/renatadiaz/Documents/model-stash/paper_cpt_paper_lda2.Rds')
+# save(cp_results_rodent2, file = '~/Dropbox/ldats-models/paper_cpt_paper_lda2.Rds')
 # rm(cp_results_rodent2)
@@
 # cp_results_rodent3 = changepoint_model(ldamodel, x, 3, weights = rep(1,length(year_continuous)))
-# save(cp_results_rodent3, file = '/Users/renatadiaz/Documents/model-stash/paper_cpt_paper_lda3.Rds')
+# save(cp_results_rodent3, file = '~/Dropbox/ldats-models/paper_cpt_paper_lda3.Rds')
 # rm(cp_results_rodent3)
@@
 # cp_results_rodent4 = changepoint_model(ldamodel, x, 4, weights = rep(1,length(year_continuous)))
-# save(cp_results_rodent4, file = '/Users/renatadiaz/Documents/model-stash/paper_cpt_paper_lda4.Rds')
+# save(cp_results_rodent4, file = '~/Dropbox/ldats-models/paper_cpt_paper_lda4.Rds')
 # rm(cp_results_rodent4)
@@
 # cp_results_rodent5 = changepoint_model(ldamodel, x, 5, weights = rep(1,length(year_continuous)))
-# save(cp_results_rodent5, file = '/Users/renatadiaz/Documents/model-stash/paper_cpt_paper_lda5.Rds')
+# save(cp_results_rodent5, file = '~/Dropbox/ldats-models/paper_cpt_paper_lda5.Rds')
 # rm(cp_results_rodent5)
@@
 # cp_results_rodent6 = changepoint_model(ldamodel, x, 6, weights = rep(1,length(year_continuous)))
-# save(cp_results_rodent6, file = '/Users/renatadiaz/Documents/model-stash/paper_cpt_paper_lda6.Rds')
+# save(cp_results_rodent6, file = '~/Dropbox/ldats-models/paper_cpt_paper_lda6.Rds')
 # rm(cp_results_rodent6)
@@
 # Load models and compare deviance to select the best one
-load('/Users/renatadiaz/Documents/model-stash/paper_cpt_paper_lda1.Rds')
-load('/Users/renatadiaz/Documents/model-stash/paper_cpt_paper_lda2.Rds')
-load('/Users/renatadiaz/Documents/model-stash/paper_cpt_paper_lda3.Rds')
-load('/Users/renatadiaz/Documents/model-stash/paper_cpt_paper_lda4.Rds')
-load('/Users/renatadiaz/Documents/model-stash/paper_cpt_paper_lda5.Rds')
-load('/Users/renatadiaz/Documents/model-stash/paper_cpt_paper_lda6.Rds')
+load('~/Dropbox/ldats-models/paper_cpt_paper_lda1.Rds')
+load('~/Dropbox/ldats-models/paper_cpt_paper_lda2.Rds')
+load('~/Dropbox/ldats-models/paper_cpt_paper_lda3.Rds')
+load('~/Dropbox/ldats-models/paper_cpt_paper_lda4.Rds')
+load('~/Dropbox/ldats-models/paper_cpt_paper_lda5.Rds')
+load('~/Dropbox/ldats-models/paper_cpt_paper_lda6.Rds')
 

### 2019-01-14 paper-comparison/lda-data.R
Paper LDA adj v nonadj results
Adjusted v. nonadjusted data frames give qualitatively but not exactly the same results with the paper LDA

@@ -59,7 +59,7 @@ seeds_4topics = best_ntopic %>%
 
 # choose seed with highest log likelihood for all following analyses
 #    (also produces plot of community composition for 'best' run compared to 'worst')
-best_seed = calculate_LDA_distance(rodents[[1]],seeds_4topics)
+best_seed = calculate_LDA_distance(rodents_adj[[1]],seeds_4topics)
 mean_dist = unlist(best_seed)[2]
 max_dist = unlist(best_seed)[3]
 
@@ -68,14 +68,9 @@ max_dist = unlist(best_seed)[3]
 # ==================================================================
 ntopics = 4
 SEED = unlist(best_seed)[1]  # For the paper, I use seed 206
-ldamodel_adj = LDA(rodents[[1]],ntopics, control = list(seed = SEED),method='VEM')
+ldamodel_adj = LDA(rodents_adj[[1]],ntopics, control = list(seed = SEED),method='VEM')
 
 save(ldamodel_adj, file = '~/Dropbox/ldats-models/paper_lda_adj_data.Rds')
-
-plot(ldamodel_adj)
-ldamodel_adj@k
-head(ldamodel_adj@gamma)
-
 
 # paper LDA w nonadjusted data
 
@@ -113,7 +108,6 @@ SEED = unlist(best_seed)[1]  # For the paper, I use seed 206
 ldamodel_nonadj = LDA(rodents[[1]],ntopics, control = list(seed = SEED),method='VEM')
 
 save(ldamodel_nonadj, file = '~/Dropbox/ldats-models/paper_lda_nonadj_data.Rds')
-
 plot(ldamodel_nonadj)
 ldamodel_nonadj@k
 head(ldamodel_nonadj@gamma)

### 2019-01-14 paper-comparison/compare-christensen.Rmd
Paper LDA adj v nonadj results
Adjusted v. nonadjusted data frames give qualitatively but not exactly the same results with the paper LDA

@@
 
-```{r adj vs non adj LDAS, eval = F}
-
+```{r run adj vs non adj LDAS, eval = F}
 # paper LDA w adjusted data
@@
 #    (also produces plot of community composition for 'best' run compared to 'worst')
-best_seed = calculate_LDA_distance(rodents[[1]],seeds_4topics)
+best_seed = calculate_LDA_distance(rodents_adj[[1]],seeds_4topics)
 mean_dist = unlist(best_seed)[2]
@@
 SEED = unlist(best_seed)[1]  # For the paper, I use seed 206
-ldamodel_adj = LDA(rodents[[1]],ntopics, control = list(seed = SEED),method='VEM')
+ldamodel_adj = LDA(rodents_adj[[1]],ntopics, control = list(seed = SEED),method='VEM')
 
-plot(ldamodel_adj)
+save(ldamodel_adj, file = '~/Dropbox/ldats-models/paper_lda_adj_data.Rds')
 
@@
 
-plot(ldamodel_nonadj)
+save(ldamodel_nonadj, file = '~/Dropbox/ldats-models/paper_lda_nonadj_data.Rds')
 
@@
 
-plot(ldats_adj_lda_selected)
+save(ldats_adj_lda_selected, file = '~/Dropbox/ldats-models/ldats_lda_adj_data.Rds')
 
@@
 ldats_nonadj_ldas <- LDATS::LDA_set(document_term_table = rodents$document_term_table, topics = c(2:6), nseeds = 100)
-ldats_nonadj_lda_selected <- LDATS::select_LDA(LDA_models = ldats_ldas)
+ldats_nonadj_lda_selected <- LDATS::select_LDA(LDA_models = ldats_nonadj_ldas)
 
-plot(ldats_nonadj_lda_selected)
+save(ldats_nonadj_lda_selected, file = '~/Dropbox/ldats-models/ldats_lda_nonadj_data.Rds')
+
+```
+
+```{r paper adj lda}
+
+load('~/Dropbox/ldats-models/paper_lda_adj_data.Rds')
+load('~/Dropbox/ldats-models/paper_lda_nonadj_data.Rds')
+
+# Adjusted data
+plot(ldamodel_adj)
+
+ldamodel_adj@k
+```
+
+```{r paper nonadj lda}
+# Nonadjusted data
+plot(ldamodel_nonadj)
+
+ldamodel_nonadj@k
+
+paper_gamma_comparison <- ldamodel_adj@gamma == ldamodel_nonadj@gamma
+unique(paper_gamma_comparison)
 

### 2019-01-17 paper-comparison/compare-christensen.Rmd
Adding ldats lda adjusted/nonadjusted data comparisons
Correct rodents to rodents_adj in second half of vignette
The model objects referenced past the lda with adjusted/nonadjusted comparison were generated using the adjusted data, which I renamed to rodents_adjusted when I did the adj/nonadj comparison. Rather than rerun things, for now I am renaming the object later in the narrative for consistency.

@@
 
-paper_gamma_comparison <- ldamodel_adj@gamma == ldamodel_nonadj@gamma
-unique(paper_gamma_comparison)
+```
+
+```{r ldats adj lda}
+load('~/Dropbox/ldats-models/ldats_lda_adj_data.Rds')
+
+plot(ldats_adj_lda_selected[[1]])
+
+ldats_adj_lda_selected[[1]]@k
+
+```
+
+```{r ldats nonadj lda}
+load('~/Dropbox/ldats-models/ldats_lda_nonadj_data.Rds')
+
+plot(ldats_nonadj_lda_selected[[1]])
+
+ldats_nonadj_lda_selected[[1]]@k
 
@@
 
-ldats_ldas <- LDATS::LDA_set(document_term_table = rodents$document_term_table, topics = c(2:6), nseeds = 100)
+ldats_ldas <- LDATS::LDA_set(document_term_table = rodents_adj$document_term_table, topics = c(2:6), nseeds = 100)
 ldats_lda_selected <- LDATS::select_LDA(LDA_models = ldats_ldas)
@@
 # repeat LDA model fit and AIC calculation with a bunch of different seeds to test robustness of the analysis
-best_ntopic = repeat_VEM(rodents[[1]],
+best_ntopic = repeat_VEM(rodents_adj[[1]],
                          seeds,
@@
 SEED = unlist(best_seed)[1]  # For the paper, I use seed 206
-ldamodel = LDA(rodents[[1]],ntopics, control = list(seed = SEED),method='VEM')
+ldamodel = LDA(rodents_adj[[1]],ntopics, control = list(seed = SEED),method='VEM')
 

### 2019-01-25 vignette/rodents-example.Rmd
change vignettes directoirectory to vignette

---
title: "LDATS Rodents Example"
author: "Renata Diaz and Juniper L. Simonis"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{rodentsexample}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r, include=FALSE}
library(LDATS)
vers <- packageVersion("LDATS")
today <- Sys.Date()
```

This vignette walks through an example of **LDATS** at the command line and
was constructed using **LDATS** version `r vers` on `r today`.

## Installation

To obtain the most recent version of **LDATS**, install the most recent 
version from GitHub:

```{r, eval=FALSE}
install.packages("devtools")
devtools::install_github("weecology/LDATS")
```

## Data

For this vignette, we will be using rodent data from the control plots of the
[Portal Project](https://github.com/weecology/portaldata), which come with
the **LDATS** package (`data(rodents)`).

`rodents` contains two data tables, a `document_term_table` and a `document_covariate_table`.

The `document_term_table` is a matrix of species (term) counts by sampling period (document). Importantly, the `document_term_table` contains raw counts of individual rodent captures, not catch per unit effort. The LDA models operate on proportions and weight the information from each sampling period according to total catch (document size). Similarly, the TS models use the LDA compositional breakdown and weight the information from each period according to the total catch. We therefore give both models un-adjusted data. The values in the `document_term_table` must be integers.

The `document_covariate_table` contains the covariates we would like to use for the time series models. In this case, we have a column for `newmoon`, which is the time step for each of our sampling periods. This column can include nonconsecutive time steps (i.e., `2, 4, 5`) if sampling periods are skipped or at unequal intervals. If there is no time column, `LDATS` will assume all time steps are equidistant. We also include columns `sin_year` and `cos_year` so we can account for seasonal dynamics.

```{r}
data(rodents)
head(rodents$document_term_table, 10)
head(rodents$document_covariate_table, 10)
```

## Stage 1: LDA models

We use `LDA_set()` to run replicate LDA models (each with its own seed) with varying numbers of topics (`2:6`) and `select_LDA()` to select the best model.

We use `LDA_controls_list())` to pass controls to the LDA function. In this case, we can set `quiet = TRUE` to make the model run quietly.

```{r lda_set, eval =F}
lda_model_set <- LDA_set(document_term_table = rodents$document_term_table,
                         topics = c(2:6),
                         nseeds = 100,
                        control = LDA_controls_list(quiet = TRUE))

```

If we do not pass any controls, by default, `quiet = FALSE` (here run with only `2:3` topics and  `10` seeds, to keep output short):

```{r lda set not quiet}
lda_model_set2 <- LDA_set(document_term_table = rodents$document_term_table,
                         topics = c(2:3),
                         nseeds = 10)
```

`LDA_set()` returns a list of LDA models. We use `select_LDA()` to identify the best number of topics and choice of seed from our set of models. By default, we will choose models based on minimum AIC. To use different selection criteria, define the appropriate functions and specify them using `LDA_controls_list(measurer = [measurer function], selector = [max, min, etc])`.

```{r load lda model set, include = F}
load(here::here('vignettes', 'rodents-example-files', 'lda_model_set.Rds'))
rm(lda_model_set2)
```

```{r select LDA}
selected_lda_model <- select_LDA(lda_model_set)
```

We can access the results of the model:

```{r LDA results}
# Number of topics:

selected_lda_model[[1]]@k

# Topic composition of communities at each time step
# Columns are topics; rows are time steps
head(selected_lda_model[[1]]@gamma)

```

`LDATS` includes flexible plot functionality for LDAs and time series. The top panel illustrates topic composition by species, and the bottom panel shows the proportion of the community made up of each topic over time. For all the available plot options see `?plot.LDA_VEM`.

```{r plot lda, fig.width=7, fig.height=6}
plot(selected_lda_model[[1]])

```

## Stage 2: TS changepoint models

We use `TS_on_LDA()` to run LDATS changepoint models with `0:6` changepoints, and then use `select_TS()` to find the best-fit model of these.

Here, `TS_on_LDA()` predicts the `gamma` (the proportion of the community made of up each topic) from our LDA model(s) as a function of `sin_year` and `cos_year` in the `document_covariate_table`. We use `document_weights()` to weight the information from each time step according to the total number of rodents captured at that time step.

```{r ts on lda, eval = F}
changepoint_models <- TS_on_LDA(LDA_models = selected_lda_model, 
                                document_covariate_table = rodents$document_covariate_table,
                                formulas = ~ sin_year + cos_year,
                                nchangepoints = c(0:6), 
                                weights = document_weights(rodents$document_term_table),
                                control = TS_controls_list(nit = 1000, timename = "newmoon"))

```

We can adjust options for both TS functions using `TS_controls_list()`. For a full list see `?TS_controls_list`. Here we illustrate adjusting the number of ptMCMC iterations - the default is 10000, but it is convenient to use fewer iterations for code development.

Also, it is important to note that by default the TS functions take the name of the time-step column from the `document_covariate_table` to be `"time"`. To pass a different column name, use the `timename` argument in `TS_controls_list()`.

`select_TS()` will identify the best-fit changepoint model of the models from `TS_on_LDA()`. As with `select_LDA()`, we can adjust the `measurer` and `selector` using `TS_controls_list()`.

```{r reload ts, include = F}
load(here::here('vignettes', 'rodents-example-files', 'changepoint_models.Rds'))
```

```{r select ts}
selected_changepoint_model <- select_TS(changepoint_models)
```

We can access the results of the selected changepoint model:

```{r cpt results}
# Number of changepoints
selected_changepoint_model$nchangepoints

# Summary of timesteps (newmoon values) for each changepoint
selected_changepoint_model$rho_summary

# Raw estimates for timesteps for each changepoint
# Changepoints are columns
head(selected_changepoint_model$rhos)

```

LDATS will plot the results of a changepoint model:

```{r plot cpt, fig.width=7, fig.height=6}
plot(selected_changepoint_model)
```

## Full analysis with `LDA_TS`

Finally, we can perform an entire LDATS analysis, including all of the above steps, using the `LDA_TS()` function and passing options to the LDA and TS functions using `LDA_TS_controls_list()`.

```{r lda_ts, eval = F}
lda_ts_results <- LDA_TS(document_term_table = rodents$document_term_table,
                         document_covariate_table = rodents$document_covariate_table,
                         nseeds = 100, 
                         topics = 2:6,
                         formulas = ~ sin_year + cos_year,
                         nchangepoints= 0:6,
                         weights = document_weights(rodents$document_term_table),
                         control = LDA_TS_controls_list(
                           TS_control = TS_controls_list(nit = 1000, timename = "newmoon")))
```

```{r load ldats results, include = F}
load(here::here('vignettes', 'rodents-example-files', 'lda_ts_results.Rds'))
```

`LDA_TS()` returns a list of all the model objects, and we can access their contents as above:

```{r LDA_TS results}
names(lda_ts_results)

# Number of topics
lda_ts_results$`Selected LDA model`$k@k

# Number of changepoints
lda_ts_results$`Selected TS model`$nchangepoints

# Summary of changepoint locations
lda_ts_results$`Selected TS model`$rho_summary
```

Finally, we can plot the `LDA_TS` results.

```{r plot LDA_TS results, fig.height = 16, fig.width = 7}
plot(lda_ts_results)
```

### 2019-01-25 vignette/paper-comparison.Rmd
change vignettes directoirectory to vignette

---
title: "Paper comparison (cleaned up)"
author: "Renata Diaz"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

Side-by-side comparison of LDATS results with analysis from Christensen et al 2018.

## LDATS Installation

To obtain the most recent version of **LDATS**, install the most recent 
version from GitHub:

```{r, eval=FALSE}
install.packages("devtools")
devtools::install_github("weecology/LDATS")
install.packages('here')
```

```{r, eval = T}
library(LDATS)

```

## Christensen 2018 analysis files

Download Christensen 2018 analysis scripts & data files from [Extreme-events-LDA repo:](https://github.com/emchristensen/Extreme-events-LDA)

Main Analysis Scripts:

  * rodent_LDA_analysis.R main script for analyzing rodent community change using LDA

  * rodent_data_for_LDA.R contains a function that creates the rodent data table used in analyses

  * AIC_model_selection.R contains functions for calculating AIC for different candidate LDA models

  * changepointmodel.r contains change-point model code

  * LDA-distance.R function for computing Hellinger distance analyses

Data:

  * Rodent_table_dat.csv table of rodent data, created by rodent_data_for_LDA.R

Figure scripts:

  * LDA_figure_scripts.R contains functions for making main plots in manuscript (Fig 1). Called from rodent_LDA_analysis.R

```{r download Christensen}

paper_filepath <- here::here('vignette', 'christensenetal-comparison-files', 'christensenetal-2018-files')
test_filepath <- here::here('vignette', 'christensenetal-ristensenetal-comparison-files', 'christensenetal-2018-files', 'rodent_LDA_analysis.r')

if(!file.exists(test_filepath)) {

files_to_download <- c('rodent_LDA_analysis.r', 'rodent_data_for_LDA.r', 'AIC_model_selection.R', 'changepointmodel.r', 'LDA-distance.R', 'Rodent_table_dat.csv', 'LDA_figure_scripts.R')

for(i in 1:length(files_to_download)) {
  download.file(url = paste0("https://raw.githubusercontent.com/emchristensen/Extreme-events-LDA/master/", files_to_download[i]),
                destfile = paste0(paper_filepath, '/', files_to_download[i]))
}

rm(files_to_download)
rm(i)

}

rm(test_filepath)
rm(paper_filepath)
```

## Data

The Portal rodents control data is included in the LDATS package:

```{r LDATS data}

data(rodents)

head(rodents[[1]])

```

Load the data used in Christensen et al:

```{r Paper data}
source(here::here('vignette', 'christensenetal-comparison-files', 'christensenetal-2018-files', 'rodent_data_for_LDA.r'))

dat = create_rodent_table(period_first = 1,
                          period_last = 436,
                          selected_plots = c(2,4,8,11,12,14,17,22),
                          selected_species = c('BA','DM','DO','DS','NA','OL','OT','PB','PE','PF','PH','PI','PL','PM','PP','RF','RM','RO','SF','SH','SO'))

# dates to go with count data
moondat = read.csv(text=getURL("https://raw.githubusercontent.com/weecology/PortalData/master/Rodents/moon_dates.csv"),stringsAsFactors = F)
moondat$date = as.Date(moondat$censusdate)

period_dates = filter(moondat,period %in% rownames(dat)) %>% select(period,date)
dates = period_dates$date

paper_dat <- dat

paper_dates <- dates

rm(list = c('dat', 'dates', 'period_dates',
            'create_rodent_table'))

```

Compare paper data to LDATS data:

```{r rodent data comparison}

compare <- rodents[[1]] == paper_dat

length(which(rowSums(compare) < ncol(compare)))
```

There are 16 rows where the data included in LDATS differs from the paper data. This is because the LDATS data is not adjusted to account for trapping effort, but the paper data has divided all census counts by the actual number of plots trapped and multiplied by 8 to account for incompletely-trapped censuses.

To confirm this, refer to lines 36-46 in `rodent_data_for_LDA.r`:

```
# retrieve data on number of plots trapped per month
  trap_table = read.csv('https://raw.githubusercontent.com/weecology/PortalData/master/Rodents/Portal_rodent_trapping.csv')
  trap_table_controls = filter(trap_table, plot %in% selected_plots)
  nplots_controls = aggregate(trap_table_controls$sampled,by=list(period = trap_table_controls$period),FUN=sum)

  # adjust species counts by number of plots trapped that month
  r_table_adjusted = as.data.frame.matrix(r_table)
  for (n in 1:436) {
    #divide by number of control plots actually trapped (should be 8) and multiply by 8 to estimate captures as if all plots were trapped
    r_table_adjusted[n,] = round(r_table_adjusted[n,]/nplots_controls$x[n]*8)
  }

    return(r_table_adjusted)

```

Running the same procedure on the LDATS data, we modify 16 rows, and finish with a data frame that matches the one from the paper.

```{r adjust LDATS data after Christensen et al}

trap_table = read.csv('https://raw.githubusercontent.com/weecology/PortalData/master/Rodents/Portal_rodent_trapping.csv')
  trap_table_controls =dplyr::filter(trap_table, plot %in% c(2,4,8,11,12,14,17,22))
  nplots_controls = aggregate(trap_table_controls$sampled,by=list(period = trap_table_controls$period),FUN=sum)

   # adjust species counts by number of plots trapped that month
  ldats_rodents_adjusted = as.data.frame.matrix(rodents[[1]])
  for (n in 1:436) {
    #divide by number of control plots actually trapped (should be 8) and multiply by 8 to estimate captures as if all plots were trapped
    ldats_rodents_adjusted[n,] = round(ldats_rodents_adjusted[n,]/nplots_controls$x[n]*8)
  }

compare_raw <- rodents[[1]] == ldats_rodents_adjusted
length(which(rowSums(compare_raw) < ncol(compare_raw)))

compare_adjusted <- ldats_rodents_adjusted == paper_dat
length(which(rowSums(compare_adjusted) < ncol(compare_adjusted)))

```

```{r data adjustment cleanup, include = FALSE}
rm(list = c('compare', 'compare_adjusted', 'compare_raw', 'n', 'ldats_rodents_adjusted', 'nplots_controls', 'trap_table', 'trap_table_controls'))
```

Because the LDA procedure weights the information from documents (census periods) according to the number of words (rodents captured), we now believe it is most appropriate to run the LDA on _unadjusted_ trapping data, and we recommend that users of LDATS do so. However, to maintain consistency with Christensen et al 2018, we will proceed using the _adjusted_ rodent table in this vignette.

```{r switch to adjusted rodents}
rodents[[1]] <- paper_dat
```

The LDATS rodent data comes with a `document_covariate_table`, which we will use later as the predictor variables for the changepoint models. In this table, time is expressed as new moon numbers. Later we will want to be able to interpret the results in terms of census dates. We will add a column to the `document_covariate_table` to convert new moon numbers to census dates. We will not reference this column in any of the formulas we pass to the changepoint models, so it will be ignored until we need it.

```{r add dates to covariate table}

head(rodents$document_covariate_table)

moondat <- dplyr::select(moondat, newmoonnumber, censusdate)
colnames(moondat) <- c('newmoon', 'censusdate')

new_cov_table <- dplyr::left_join(rodents$document_covariate_table, moondat, by = 'newmoon')

rodents$document_covariate_table <- new_cov_table

```

```{r cleanup dates, include = F}
rm(moondat)
rm(new_cov_table)
```

## Stage 1: LDA

While LDATS can run start-to-finish with `LDATS::LDA_TS`, here we will work through the process function-by-function to isolate differences with the paper. For an illustration of `LDA_TS`, see the `codebase` vignette.

```{r LDATS LDAs, eval = F}

ldats_ldas <- LDATS::LDA_set(document_term_table = rodents$document_term_table, topics = c(2:6), nseeds = 100)
ldats_lda_selected <- LDATS::select_LDA(LDA_models = ldats_ldas)

save(ldats_ldas, file = here::here('vignette', 'christensenetal-comparison-files', 'model-cache', 'ldats_ldas.Rds'))
save(ldats_lda_selected, file = here::here('vignette', 'christensenetal-comparison-files', 'model-cache', 'ldats_lda_selected.Rds'))

```

```{r rm LDATS LDAS, include = F, eval = F}
rm(list = c('ldats_ldas', 'ldats_lda_selected'))
```

Paper LDAS:

```{r create dat for paper lda, include = F}
dat = paper_dat
```

```{r paper LDAs, eval = F}
source(here::here('vignette', 'christensenetal-comparison-files', 'christensenetal-2018-files', 'AIC_model_selection.R'))
source(here::here('vignette', 'christensenetal-comparison-files', 'christensenetal-2018-files', 'LDA-distance.R'))

# Fit a bunch of LDA models with different seeds
# Only use even numbers for seeds because consecutive seeds give identical results
seeds = 2*seq(200)

# repeat LDA model fit and AIC calculation with a bunch of different seeds to test robustness of the analysis
best_ntopic = repeat_VEM(paper_dat,
                         seeds,
                         topic_min=2,
                         topic_max=6)
hist(best_ntopic$k,breaks=c(0.5,1.5,2.5,3.5,4.5,5.5,6.5,7.5,8.5,9.5),xlab='best # of topics', main='')

# 2b. how different is species composition of 4 community-types when LDA is run with different seeds?
# ==================================================================
# get the best 100 seeds where 4 topics was the best LDA model
seeds_4topics = best_ntopic %>% 
  filter(k == 4) %>% 
  arrange(aic) %>% 
  head(100) %>% 
  pull(SEED)

# choose seed with highest log likelihood for all following analyses
#    (also produces plot of community composition for 'best' run compared to 'worst')
best_seed = calculate_LDA_distance(paper_dat,seeds_4topics)
mean_dist = unlist(best_seed)[2]
max_dist = unlist(best_seed)[3]

# ==================================================================
# 3. run LDA model
# ==================================================================
ntopics = 4
SEED = unlist(best_seed)[1]  # For the paper, use seed 206
ldamodel = LDA(paper_dat,ntopics, control = list(seed = SEED),method='VEM')

save(ldamodel,file = here::here('vignette', 'christensenetal-comparison-files', 'model-cache', 'paper_ldamodel.Rds'))
```

```{r cleanup paper LDAS, include = F, eval = F}
rm(list = c('dat', 'ldamodel', 'max_dist', 'mean_dist',
            'ntopics', 'SEED', 'seeds', 'seeds_4topics', 'aic_model', 'aic_model_gibbs', 'calculate_LDA_distance', 'Hellinger', 'min_H', 'repeat_VEM', 'best_ntopic', 'best_seed'))
```

```{r reload LDAS, include = FALSE}

load(here::here('vignette', 'christensenetal-comparison-files', 'model-cache', 'ldats_lda_selected.Rds'))
load(here::here('vignette', 'christensenetal-comparison-files', 'model-cache', 'paper_ldamodel.Rds'))

```

### Plot LDAS

```{r plot LDAs}
# Paper
plot(ldamodel, cols = NULL, option = "D")

# LDATS
plot(ldats_lda_selected[[1]], cols = NULL, option = "D")
```

```{r remove LDAS after plotting, include = F}
rm(ldamodel)
rm(ldats_lda_selected)
```

The paper method finds 4 topics and LDATS finds 6. This is because of an update to the model selection procedure. The paper conservatively overestimates the number of parameters and therefore overpenalizes AIC for models with more topics. For this vignette, we will compare the results from using both LDA models.

## Changepoint models

We will compare four combinations of LDA + changepoint models:

* LDATS LDA + LDATS changepoint
* LDATS LDA + paper changepoint
* Paper LDA + LDATS changepoint
* Paper LDA + paper changepoint

The paper changepoint model weighted all sample periods equally, wheras LDATS can weight sample periods according to how many individuals were captured (controlled by the `document_term_weights`). We now believe it is more appropriate to weight periods proportional to captures, and this is what we recommend for LDATS users. For the purposes of comparison, we will continue set all weights = 1 for both changepoint models. For an example of LDATS run with proportional weights, see the rodents vignette. [?]

### Running paper changepoint models

#### LDATS LDA and paper changepoint

```{r load LDATS LDA for paper cpt}

#### Load LDA ####
load(here::here('vignette', 'christensenetal-comparison-files', 'model-cache', 'ldats_lda_selected.Rds'))

```

```{r run LDATS LDA and paper cpt, eval = F}
#### Run changepoint ####
source(here::here('vignette', 'christensenetal-comparison-files', 'christensenetal-2018-files', 'changepointmodel.r'))

# set up parameters for model
year_continuous = 1970 + as.integer(julian(paper_dates)) / 365.25
x = data.frame(
  year_continuous = year_continuous,
  sin_year = sin(year_continuous * 2 * pi),
  cos_year = cos(year_continuous * 2 * pi)
)

# run models with 1, 2, 3, 4, 5, 6 changepoints
cp_results_rodent = changepoint_model(ldats_lda_selected[[1]], x, 1, weights = rep(1,length(year_continuous)))
save(cp_results_rodent, file = here::here('vignette', 'christensenetal-comparison-files', 'model-cache', 'ldats_paper_1.Rds'))
rm(cp_results_rodent)

cp_results_rodent2 = changepoint_model(ldats_lda_selected[[1]], x, 2, weights = rep(1,length(year_continuous)))
save(cp_results_rodent2, file = here::here('vignette', 'christensenetal-comparison-files', 'model-cache', 'ldats_paper_2.Rds'))
rm(cp_results_rodent2)

cp_results_rodent3 = changepoint_model(ldats_lda_selected[[1]], x, 3, weights = rep(1,length(year_continuous)))
save(cp_results_rodent3, file = here::here('vignette', 'christensenetal-comparison-files', 'model-cache', 'ldats_paper_3.Rds'))
rm(cp_results_rodent3)

cp_results_rodent4 = changepoint_model(ldats_lda_selected[[1]], x, 4, weights = rep(1,length(year_continuous)))
save(cp_results_rodent4, file = here::here('vignette', 'christensenetal-comparison-files', 'model-cache', 'ldats_paper_4.Rds'))
rm(cp_results_rodent4)

cp_results_rodent5 = changepoint_model(ldats_lda_selected[[1]], x, 5, weights = rep(1,length(year_continuous)))
save(cp_results_rodent5, file = here::here('vignette', 'christensenetal-comparison-files', 'model-cache', 'ldats_paper_5.Rds'))
rm(cp_results_rodent5)

cp_results_rodent6 = changepoint_model(ldats_lda_selected[[1]], x, 6, weights = rep(1,length(year_continuous)))
save(cp_results_rodent6, file = here::here('vignette', 'christensenetal-comparison-files', 'model-cache', 'ldats_paper_6.Rds'))
rm(cp_results_rodent6)
```

```{r select LDATS LDA and paper cpt}
#### Changepoint model selection ####
load(here::here('vignette', 'christensenetal-comparison-files', 'model-cache', 'ldats_paper_1.Rds'))
load(here::here('vignette', 'christensenetal-comparison-files', 'model-cache', 'ldats_paper_2.Rds'))
load(here::here('vignette', 'christensenetal-comparison-files', 'model-cache', 'ldats_paper_3.Rds'))
load(here::here('vignette', 'christensenetal-comparison-files', 'model-cache', 'ldats_paper_4.Rds'))
load(here::here('vignette', 'christensenetal-comparison-files', 'model-cache', 'ldats_paper_5.Rds'))
load(here::here('vignette', 'christensenetal-comparison-files', 'model-cache', 'ldats_paper_6.Rds'))

ntopics = ldats_lda_selected[[1]]@k

# change point model selection
# mean deviance ( -2 * log likelihood) + 2*(#parameters)
mean(cp_results_rodent$saved_lls * -2) + 2*(3*(ntopics-1)*(1+1)+(1))
mean(cp_results_rodent2$saved_lls * -2)+ 2*(3*(ntopics-1)*(2+1)+(2))
mean(cp_results_rodent3$saved_lls * -2)+ 2*(3*(ntopics-1)*(3+1)+(3))
mean(cp_results_rodent4$saved_lls * -2)+ 2*(3*(ntopics-1)*(4+1)+(4))
mean(cp_results_rodent5$saved_lls * -2)+ 2*(3*(ntopics-1)*(5+1)+(5))
mean(cp_results_rodent6$saved_lls * -2)+ 2*(3*(ntopics-1)*(6+1)+(6))

# The lowest deviance is for 4 changepoints.

df_4 = as.data.frame(t(cp_results_rodent4$saved[,1,])) %>% reshape::melt()
year_continuous = 1970 + as.integer(julian(paper_dates)) / 365.25
df_4$value = year_continuous[df_4$value]

ldats_paper_cpt_dates <- vector(length = 4)
ldats_paper_cpt_dates[1] <- mean(df_4[df_4$variable=='V1','value']) %>% lubridate::date_decimal() %>% format('%Y-%m-%d')
ldats_paper_cpt_dates[2] <-mean(df_4[df_4$variable=='V2','value']) %>% lubridate::date_decimal() %>% format('%Y-%m-%d')
ldats_paper_cpt_dates[3] <-mean(df_4[df_4$variable=='V3','value']) %>% lubridate::date_decimal() %>% format('%Y-%m-%d')
ldats_paper_cpt_dates[4] <-mean(df_4[df_4$variable=='V4','value']) %>% lubridate::date_decimal() %>% format('%Y-%m-%d')

ldats_paper_cpt_dates

```

```{r clean up LDATS LDA and paper changepoint, include = F}
rm(list = c('cp_results_rodent', 'cp_results_rodent2', 'cp_results_rodent3', 'cp_results_rodent4', 'cp_results_rodent5', 'cp_results_rodent6', 'ntopics', 'ldats_lda_selected', 'df_4'))
```

#### Paper LDA and paper changepoint

```{r load paper LDA for paper cpt}

#### Load LDA ####
load(here::here('vignette', 'christensenetal-comparison-files', 'model-cache', 'paper_ldamodel.Rds'))

```

```{r run paper LDA and paper cpt, eval = F}
#### Run changepoint ####
source(here::here('vignette', 'christensenetal-comparison-files', 'christensenetal-2018-files', 'changepointmodel.r'))

# set up parameters for model
year_continuous = 1970 + as.integer(julian(paper_dates)) / 365.25
x = data.frame(
  year_continuous = year_continuous,
  sin_year = sin(year_continuous * 2 * pi),
  cos_year = cos(year_continuous * 2 * pi)
)

# run models with 1, 2, 3, 4, 5, 6 changepoints
cp_results_rodent = changepoint_model(ldamodel, x, 1, weights = rep(1,length(year_continuous)))
save(cp_results_rodent, file = here::here('vignette', 'christensenetal-comparison-files', 'model-cache', 'paper_paper1.Rds'))
rm(cp_results_rodent)

cp_results_rodent2 = changepoint_model(ldamodel, x, 2, weights = rep(1,length(year_continuous)))
save(cp_results_rodent2, file = here::here('vignette', 'christensenetal-comparison-files', 'model-cache', 'paper_paper2.Rds'))
rm(cp_results_rodent2)

cp_results_rodent3 = changepoint_model(ldamodel, x, 3, weights = rep(1,length(year_continuous)))
save(cp_results_rodent3, file = here::here('vignette', 'christensenetal-comparison-files', 'model-cache', 'paper_paper3.Rds'))
rm(cp_results_rodent3)

cp_results_rodent4 = changepoint_model(ldamodel, x, 4, weights = rep(1,length(year_continuous)))
save(cp_results_rodent4, file = here::here('vignette', 'christensenetal-comparison-files', 'model-cache', 'paper_paper4.Rds'))
rm(cp_results_rodent4)

cp_results_rodent5 = changepoint_model(ldamodel, x, 5, weights = rep(1,length(year_continuous)))
save(cp_results_rodent5, file = here::here('vignette', 'christensenetal-comparison-files', 'model-cache', 'paper_paper5.Rds'))
rm(cp_results_rodent5)

cp_results_rodent6 = changepoint_model(ldamodel, x, 6, weights = rep(1,length(year_continuous)))
save(cp_results_rodent6, file = here::here('vignette', 'christensenetal-comparison-files', 'model-cache', 'paper_paper6.Rds'))
rm(cp_results_rodent6)
```

```{r select paper LDA and paper cpt}
#### Changepoint model selection ####
load(here::here('vignette', 'christensenetal-comparison-files', 'model-cache', 'paper_paper_1.Rds'))
load(here::here('vignette', 'christensenetal-comparison-files', 'model-cache', 'paper_paper_2.Rds'))
load(here::here('vignette', 'christensenetal-comparison-files', 'model-cache', 'paper_paper_3.Rds'))
load(here::here('vignette', 'christensenetal-comparison-files', 'model-cache', 'paper_paper_4.Rds'))
load(here::here('vignette', 'christensenetal-comparison-files', 'model-cache', 'paper_paper_5.Rds'))
load(here::here('vignette', 'christensenetal-comparison-files', 'model-cache', 'paper_paper_6.Rds'))

ntopics = ldamodel@k

# change point model selection
# mean deviance ( -2 * log likelihood) + 2*(#parameters)
mean(cp_results_rodent$saved_lls * -2) + 2*(3*(ntopics-1)*(1+1)+(1))
mean(cp_results_rodent2$saved_lls * -2)+ 2*(3*(ntopics-1)*(2+1)+(2))
mean(cp_results_rodent3$saved_lls * -2)+ 2*(3*(ntopics-1)*(3+1)+(3))
mean(cp_results_rodent4$saved_lls * -2)+ 2*(3*(ntopics-1)*(4+1)+(4))
mean(cp_results_rodent5$saved_lls * -2)+ 2*(3*(ntopics-1)*(5+1)+(5))
mean(cp_results_rodent6$saved_lls * -2)+ 2*(3*(ntopics-1)*(6+1)+(6))

# The lowest deviance is for 4 changepoints.

df_4 = as.data.frame(t(cp_results_rodent4$saved[,1,])) %>%  reshape::melt()
df_4$value = year_continuous[df_4$value]

paper_paper_cpt_dates <- vector(length = 4)
paper_paper_cpt_dates[1] <- mean(df_4[df_4$variable=='V1','value']) %>% lubridate::date_decimal() %>% format('%Y-%m-%d')
paper_paper_cpt_dates[2] <-mean(df_4[df_4$variable=='V2','value']) %>% lubridate::date_decimal() %>% format('%Y-%m-%d')
paper_paper_cpt_dates[3] <-mean(df_4[df_4$variable=='V3','value']) %>% lubridate::date_decimal() %>% format('%Y-%m-%d')
paper_paper_cpt_dates[4] <-mean(df_4[df_4$variable=='V4','value']) %>% lubridate::date_decimal() %>% format('%Y-%m-%d')

paper_paper_cpt_dates

```

```{r clean up paper LDA and paper changepoint, include = F}
rm(list = c('cp_results_rodent', 'cp_results_rodent2', 'cp_results_rodent3', 'cp_results_rodent4', 'cp_results_rodent5', 'cp_results_rodent6', 'ntopics', 'ldamodel', 'df_4'))
```

### Running LDATS changepoint models

#### LDATS LDA and LDATS changepoint

```{r load LDATS LDA for LDATS cpt}

#### Load LDA ####
load(here::here('vignette', 'christensenetal-comparison-files', 'model-cache', 'ldats_lda_selected.Rds'))

```
#### Run the TS models

```{r run LDATS LDA and LDATS changepoint, eval = F}
#### Run LDATS changepoint ####

ldats_ldats_cpt <- TS_on_LDA(LDA_models = ldats_lda_selected, 
                             document_covariate_table = rodents$document_covariate_table,
                             formulas = ~ sin_year + cos_year,
                             nchangepoints = 1:6,
                             weights = NULL)

save(ldats_ldats_cpt, file = here::here('vignette', 'christensenetal-comparison-files', 'model-cache', 'ldats_ldats.Rds'))
rm(ldats_ldats_cpt)
rm(ldats_lda_selected)
```

```{r load LDATS LDA and LDATS changepoint, include = F}
load(here::here('vignette', 'christensenetal-comparison-files', 'model-cache', 'ldats_ldats.Rds'))
```

### Select model

```{r select ldats lda + ldats cpt}

ldats_ldats_cpt_selected <- select_TS(ldats_ldats_cpt)

ldats_ldats_cpt_selected$nchangepoints

```

### Get dates associated with changepoints

```{r get ldats lda + ldats cpt dates}
ldats_ldats_cpt_dates <- ldats_ldats_cpt_selected$rho_summary$Mean
ldats_ldats_cpt_dates <- round(ldats_ldats_cpt_dates)

ldats_ldats_cpt_dates <- as.data.frame(ldats_ldats_cpt_dates)
colnames(ldats_ldats_cpt_dates) <- 'newmoon'
ldats_ldats_cpt_dates <- dplyr::left_join(ldats_ldats_cpt_dates, rodents$document_covariate_table, by = 'newmoon')
ldats_ldats_cpt_dates <- ldats_ldats_cpt_dates$censusdate

```

```{r cleanup LDATS LDA and LDATS changepoint, include = F}

save(ldats_ldats_cpt_selected, file = here::here('vignette', 'christensenetal-comparison-files', 'model-cache', 'ldats_ldats_selected.Rds'))

rm(ldats_ldats_cpt_selected)
rm(ldats_ldats_cpt)
```

### Paper LDA and LDATS changepoint

```{r load paper LDA for LDATS cpt}

#### Load LDA ####
load(here::here('vignette', 'christensenetal-comparison-files', 'model-cache', 'paper_ldamodel.Rds'))

```
#### Run the TS models

```{r run paper LDA and LDATS changepoint, eval = F}
#### Run LDATS changepoint ####

paper_ldats_cpt <- TS_on_LDA(LDA_models = ldamodel, 
                             document_covariate_table = rodents$document_covariate_table,
                             formulas = ~ sin_year + cos_year,
                             nchangepoints = 1:6,
                             weights = NULL)

save(paper_ldats_cpt, file = here::here('vignette', 'christensenetal-comparison-files', 'model-cache', 'paper_ldats.Rds'))
rm(paper_ldats_cpt)
rm(ldamodel)
```

```{r load paper LDA and LDATS changepoint, include = F}
load(here::here('vignette', 'christensenetal-comparison-files', 'model-cache', 'paper_ldats.Rds'))
```

### Select model

```{r select paper lda + ldats cpt}

paper_ldats_cpt_selected <- select_TS(paper_ldats_cpt)

paper_ldats_cpt_selected$nchangepoints

```

#### Get dates

Unlike the paper changepoint model, LDATS can recognize that sampling periods may not be equidistant, and can place changepoint estimates at new moons if they fall between nonconsecutive sampling periods. We can estimate the dates corresponding to those new moons.

```{r get paper lda + ldats cpt dates}
paper_ldats_cpt_dates <- paper_ldats_cpt_selected$rho_summary$Mean
paper_ldats_cpt_dates <- round(paper_ldats_cpt_dates)

paper_ldats_cpt_dates <- as.data.frame(paper_ldats_cpt_dates)
colnames(paper_ldats_cpt_dates) <- 'newmoon'
paper_ldats_cpt_dates <- dplyr::left_join(paper_ldats_cpt_dates, rodents$document_covariate_table, by = 'newmoon')

halfinterval = as.integer((as.Date(rodents$document_covariate_table$censusdate[ which(rodents$document_covariate_table$newmoon == paper_ldats_cpt_dates$newmoon[1] + 1)]) - as.Date(rodents$document_covariate_table$censusdate[ which(rodents$document_covariate_table$newmoon == paper_ldats_cpt_dates$newmoon[1] - 1)]))/2)

newdate <- lubridate::ymd(rodents$document_covariate_table$censusdate[ which(rodents$document_covariate_table$newmoon == paper_ldats_cpt_dates$newmoon[1] -1 )]) + lubridate::days(halfinterval)

paper_ldats_cpt_dates$censusdate[1] <- as.character(newdate)

halfinterval =as.integer((as.Date(rodents$document_covariate_table$censusdate[ which(rodents$document_covariate_table$newmoon == paper_ldats_cpt_dates$newmoon[3] + 2)]) - as.Date(rodents$document_covariate_table$censusdate[ which(rodents$document_covariate_table$newmoon == paper_ldats_cpt_dates$newmoon[3] - 1)]))/3)

newdate <- lubridate::ymd(rodents$document_covariate_table$censusdate[ which(rodents$document_covariate_table$newmoon == paper_ldats_cpt_dates$newmoon[3] -2 )]) + lubridate::days(halfinterval)

paper_ldats_cpt_dates$censusdate[3] <- as.character(newdate)

paper_ldats_cpt_dates <- paper_ldats_cpt_dates$censusdate
```

```{r cleanup paper LDA and LDATS changepoint, include = F}

save(paper_ldats_cpt_selected, file = here::here('vignette', 'christensenetal-comparison-files', 'model-cache', 'paper_ldats_selected.Rds'))

rm(paper_ldats_cpt_selected)
rm(paper_ldats_cpt)
rm(list = c('newdate', 'halfinterval'))
```

All of the models find four changepoints.

```{r report cpt dates, include = F}

cpt_dates <- as.data.frame(paper_paper_cpt_dates)
cpt_dates <- cbind(cpt_dates, ldats_paper_cpt_dates)
cpt_dates <- cbind(cpt_dates, ldats_ldats_cpt_dates)
cpt_dates <- cbind(cpt_dates, paper_ldats_cpt_dates)

colnames(cpt_dates) <- c('paperLDA_papercpt', 'ldatsLDA_papercpt', 'ldatsLDA_ldatscpt', 'paperLDA_ldatscpt')

```

```{r print cpt dates}
cpt_dates
```

The choice of LDA has more influence on the changepoint locations than the choice of changepoint model - probably because the LDATS LDA has 6 topics, and the paper LDA has 4.

```{r cleanup cpt_dates, include = F}
rm(list= 'ldats_ldats_cpt_dates', 'ldats_paper_cpt_dates',  'paper_ldats_cpt_dates', 'paper_paper_cpt_dates', 'cpt_dates')
```

### Plot changepoint models

```{r load cpts to plot, include = F}
load(here::here('vignette', 'christensenetal-comparison-files', 'model-cache', 'paper_ldats_selected.Rds'))
load(here::here('vignette', 'christensenetal-comparison-files', 'model-cache', 'ldats_ldats_selected.Rds'))
load(here::here('vignette', 'christensenetal-comparison-files', 'model-cache', 'paper_paper_4.Rds'))
paper_paper_cpt_selected <- cp_results_rodent4
load(here::here('vignette', 'christensenetal-comparison-files', 'model-cache', 'ldats_paper_4.Rds'))
ldats_paper_cpt_selected <- cp_results_rodent4

year_continuous = 1970 + as.integer(julian(paper_dates)) / 365.25
x = data.frame(
  year_continuous = year_continuous,
  sin_year = sin(year_continuous * 2 * pi),
  cos_year = cos(year_continuous * 2 * pi)
)
ntopics = 4

```

```{r plot cpts}

plot(paper_ldats_cpt_selected)

plot(ldats_ldats_cpt_selected)

source(here::here('vignette', 'christensenetal-comparison-files', 'christensenetal-2018-files', 'LDA_figure_scripts.R'))
source(here::here('vignette', 'christensenetal-comparison-files', 'christensenetal-2018-files', 'changepointmodel.r'))
paper_cpts = find_changepoint_location(paper_paper_cpt_selected)
paper_cpt_plot = get_ll_non_memoized_plot(ldamodel,x,paper_cpts,make_plot=T,weights=rep(1,length(year_continuous)))
annual_hist(paper_paper_cpt_selected, year_continuous)
paper_cpt_plot

ntopics = 6
ldats_cpts = find_changepoint_location(ldats_paper_cpt_selected)
ldats_cpt_plot = get_ll_non_memoized_plot(ldats_lda_selected[[1]],x,ldats_cpts,make_plot=T,weights=rep(1,length(year_continuous)))
annual_hist(ldats_paper_cpt_selected, year_continuous)
ldats_cpt_plot
```

### 2019-01-25 vignette/rodents-example.Rmd
change vignettes directoirectory to vignette

@@
 title: "LDATS Rodents Example"
-author: "Renata Diaz and Juniper L. Simonis"
+author: "Juniper L. Simonis"
 output: rmarkdown::html_vignette
 vignette: >
-  %\VignetteIndexEntry{rodentsexample}
+  %\VignetteIndexEntry{Vignette Title}
   %\VignetteEngine{knitr::rmarkdown}
@@
 [Portal Project](https://github.com/weecology/portaldata), which come with
-the **LDATS** package (`data(rodents)`).
-
-`rodents` contains two data tables, a `document_term_table` and a `document_covariate_table`.
-
-The `document_term_table` is a matrix of species (term) counts by sampling period (document).
-Importantly, the `document_term_table` contains raw counts of individual rodent captures, not catch per unit effort.
-The LDA models operate on proportions and weight the information from each sampling period according to total catch (document size).
-Similarly, the TS models use the LDA compositional breakdown and weight the information from each period according to the total catch.
-We therefore give both models un-adjusted data.
-The values in the `document_term_table` must be integers.
-
-The `document_covariate_table` contains the covariates we would like to use for the time series models.
-In this case, we have a column for `newmoon`, which is the time step for each of our sampling periods.
-This column can include nonconsecutive time steps (i.e., `2, 4, 5`) if sampling periods are skipped or at unequal intervals.
-If there is no time column, `LDATS` will assume all time steps are equidistant.
-We also include columns `sin_year` and `cos_year` so we can account for seasonal dynamics.
+the **LDATS** package
 
@@
 ```
-
-## Stage 1: LDA models
-
-We use `LDA_set()` to run replicate LDA models (each with its own seed) with varying numbers of topics (`2:6`) and `select_LDA()` to select the best model.
-
-We use `LDA_controls_list())` to pass controls to the LDA function. In this case, we can set `quiet = TRUE` to make the model run quietly.
-
-```{r lda_set, eval =F}
-lda_model_set <- LDA_set(document_term_table = rodents$document_term_table,
-                         topics = c(2:6),
-                         nseeds = 100,
-                        control = LDA_controls_list(quiet = TRUE))
-
-```
-
-If we do not pass any controls, by default, `quiet = FALSE` (here run with only `2:3` topics and  `10` seeds, to keep output short):
-
-```{r lda set not quiet}
-lda_model_set2 <- LDA_set(document_term_table = rodents$document_term_table,
-                         topics = c(2:3),
-                         nseeds = 10)
-```
-
-`LDA_set()` returns a list of LDA models.
-We use `select_LDA()` to identify the best number of topics and choice of seed from our set of models.
-By default, we will choose models based on minimum AIC.
-To use different selection criteria, define the appropriate functions and specify them using `LDA_controls_list(measurer = [measurer function], selector = [max, min, etc])`.
-
-```{r load lda model set, include = F}
-load(here::here('vignettes', 'rodents-example-files', 'lda_model_set.Rds'))
-rm(lda_model_set2)
-```
-
-```{r select LDA}
-selected_lda_model <- select_LDA(lda_model_set)
-```
-
-We can access the results of the model:
-
-```{r LDA results}
-# Number of topics:
-
-selected_lda_model[[1]]@k
-
-# Topic composition of communities at each time step
-# Columns are topics; rows are time steps
-head(selected_lda_model[[1]]@gamma)
-
-```
-
-`LDATS` includes flexible plot functionality for LDAs and time series.
-The top panel illustrates topic composition by species, and the bottom panel shows the proportion of the community made up of each topic over time.
-For all the available plot options see `?plot.LDA_VEM`.
-
-```{r plot lda, fig.width=7, fig.height=6}
-plot(selected_lda_model[[1]])
-
-```
-
-## Stage 2: TS changepoint models
-
-We use `TS_on_LDA()` to run LDATS changepoint models with `0:6` changepoints, and then use `select_TS()` to find the best-fit model of these.
-
-Here, `TS_on_LDA()` predicts the `gamma` (the proportion of the community made of up each topic) from our LDA model(s) as a function of `sin_year` and `cos_year` in the `document_covariate_table`.
-We use `document_weights()` to weight the information from each time step according to the total number of rodents captured at that time step.
-
-```{r ts on lda, eval = F}
-changepoint_models <- TS_on_LDA(LDA_models = selected_lda_model, 
-                                document_covariate_table = rodents$document_covariate_table,
-                                formulas = ~ sin_year + cos_year,
-                                nchangepoints = c(0:6), 
-                                weights = document_weights(rodents$document_term_table),
-                                control = TS_controls_list(nit = 1000, timename = "newmoon"))
-
-```
-
-We can adjust options for both TS functions using `TS_controls_list()`.
-For a full list see `?TS_controls_list`.
-Here we illustrate adjusting the number of ptMCMC iterations - the default is 10000, but it is convenient to use fewer iterations for code development.
-
-Also, it is important to note that by default the TS functions take the name of the time-step column from the `document_covariate_table` to be `"time"`.
-To pass a different column name, use the `timename` argument in `TS_controls_list()`.
-
-`select_TS()` will identify the best-fit changepoint model of the models from `TS_on_LDA()`. As with `select_LDA()`, we can adjust the `measurer` and `selector` using `TS_controls_list()`.
-
-```{r reload ts, include = F}
-load(here::here('vignettes', 'rodents-example-files', 'changepoint_models.Rds'))
-```
-
-```{r select ts}
-selected_changepoint_model <- select_TS(changepoint_models)
-```
-
-We can access the results of the selected changepoint model:
-
-```{r cpt results}
-# Number of changepoints
-selected_changepoint_model$nchangepoints
-
-# Summary of timesteps (newmoon values) for each changepoint
-selected_changepoint_model$rho_summary
-
-# Raw estimates for timesteps for each changepoint
-# Changepoints are columns
-head(selected_changepoint_model$rhos)
-
-```
-
-LDATS will plot the results of a changepoint model:
-
-```{r plot cpt, fig.width=7, fig.height=6}
-plot(selected_changepoint_model)
-```
-
-## Full analysis with `LDA_TS`
-
-Finally, we can perform an entire LDATS analysis, including all of the above steps, using the `LDA_TS()` function and passing options to the LDA and TS functions using `LDA_TS_controls_list()`.
-
-```{r lda_ts, eval = F}
-lda_ts_results <- LDA_TS(document_term_table = rodents$document_term_table,
-                         document_covariate_table = rodents$document_covariate_table,
-                         nseeds = 100, 
-                         topics = 2:6,
-                         formulas = ~ sin_year + cos_year,
-                         nchangepoints= 0:6,
-                         weights = document_weights(rodents$document_term_table),
-                         control = LDA_TS_controls_list(
-                           TS_control = TS_controls_list(nit = 1000, timename = "newmoon")))
-```
-
-```{r load ldats results, include = F}
-load(here::here('vignettes', 'rodents-example-files', 'lda_ts_results.Rds'))
-```
-
-`LDA_TS()` returns a list of all the model objects, and we can access their contents as above:
-
-```{r LDA_TS results}
-names(lda_ts_results)
-
-# Number of topics
-lda_ts_results$`Selected LDA model`$k@k
-
-# Number of changepoints
-lda_ts_results$`Selected TS model`$nchangepoints
-
-# Summary of changepoint locations
-lda_ts_results$`Selected TS model`$rho_summary
-```
-
-Finally, we can plot the `LDA_TS` results.
-
-```{r plot LDA_TS results, fig.height = 16, fig.width = 7}
-plot(lda_ts_results)
-```

### 2019-02-05 doc/paper-comparison.Rmd
build vignettes and push doc

--- 2019-01-25 vignette/paper-comparison.Rmd
@@
 ---
-title: "Paper comparison (cleaned up)"
+title: "Comparison to Christensen et al. 2018"
 author: "Renata Diaz"
@@
 vignette: >
-  %\VignetteIndexEntry{Vignette Title}
+  %\VignetteIndexEntry{paper-comparison}
   %\VignetteEngine{knitr::rmarkdown}
@@
 devtools::install_github("weecology/LDATS")
-install.packages('here')
 ```
@@
 
-paper_filepath <- here::here('vignette', 'christensenetal-comparison-files', 'christensenetal-2018-files')
-test_filepath <- here::here('vignette', 'christensenetal-ristensenetal-comparison-files', 'christensenetal-2018-files', 'rodent_LDA_analysis.r')
+paper_filepath <- here::here('vignettes', 'christensenetal-comparison-files', 'christensenetal-2018-files')
+test_filepath <- here::here('vignettes', 'christensenetal-ristensenetal-comparison-files', 'christensenetal-2018-files', 'rodent_LDA_analysis.r')
 
@@
 ```{r Paper data}
-source(here::here('vignette', 'christensenetal-comparison-files', 'christensenetal-2018-files', 'rodent_data_for_LDA.r'))
+source(here::here('vignettes', 'christensenetal-comparison-files', 'christensenetal-2018-files', 'rodent_data_for_LDA.r'))
 
@@
 While LDATS can run start-to-finish with `LDATS::LDA_TS`, here we will work through the process function-by-function to isolate differences with the paper.
-For an illustration of `LDA_TS`, see the `codebase` vignette.
+For a breakdown of the `LDA_TS` pipeline, see the [`codebase` vignette](https://weecology.github.io/LDATS/articles/LDATS_codebase.html).
 
@@
 
-save(ldats_ldas, file = here::here('vignette', 'christensenetal-comparison-files', 'model-cache', 'ldats_ldas.Rds'))
-save(ldats_lda_selected, file = here::here('vignette', 'christensenetal-comparison-files', 'model-cache', 'ldats_lda_selected.Rds'))
+save(ldats_ldas, file = here::here('vignettes', 'christensenetal-comparison-files', 'model-cache', 'ldats_ldas.Rds'))
+save(ldats_lda_selected, file = here::here('vignettes', 'christensenetal-comparison-files', 'model-cache', 'ldats_lda_selected.Rds'))
 
@@
 ```{r paper LDAs, eval = F}
-source(here::here('vignette', 'christensenetal-comparison-files', 'christensenetal-2018-files', 'AIC_model_selection.R'))
-source(here::here('vignette', 'christensenetal-comparison-files', 'christensenetal-2018-files', 'LDA-distance.R'))
+source(here::here('vignettes', 'christensenetal-comparison-files', 'christensenetal-2018-files', 'AIC_model_selection.R'))
+source(here::here('vignettes', 'christensenetal-comparison-files', 'christensenetal-2018-files', 'LDA-distance.R'))
 
@@
 
-save(ldamodel,file = here::here('vignette', 'christensenetal-comparison-files', 'model-cache', 'paper_ldamodel.Rds'))
+save(ldamodel,file = here::here('vignettes', 'christensenetal-comparison-files', 'model-cache', 'paper_ldamodel.Rds'))
 ```
@@
 
-load(here::here('vignette', 'christensenetal-comparison-files', 'model-cache', 'ldats_lda_selected.Rds'))
-load(here::here('vignette', 'christensenetal-comparison-files', 'model-cache', 'paper_ldamodel.Rds'))
+load(here::here('vignettes', 'christensenetal-comparison-files', 'model-cache', 'ldats_lda_selected.Rds'))
+load(here::here('vignettes', 'christensenetal-comparison-files', 'model-cache', 'paper_ldamodel.Rds'))
 
@@
 
-```{r plot LDAs}
+#### Paper LDA
+```{r plot paper LDA, fig.width=6, fig.height=6}
 # Paper
@@
 
+```
+
+#### LDATS LDA
+```{r plot LDATS LDA, fig.width=6, fig.height=6}
 # LDATS
@@
 This is because of an update to the model selection procedure.
-The paper conservatively overestimates the number of parameters and therefore overpenalizes AIC for models with more topics.
+The paper conservatively overestimates the number of parameters (by counting all of the variational parameteres) and therefore overpenalizes the AIC for models with more topics.
+Comparatively, the LDATS method now uses the number of parameters remaining after the variational approximation, as returned by the LDA object.
 For this vignette, we will compare the results from using both LDA models.
@@
 
-The paper changepoint model weighted all sample periods equally, wheras LDATS can weight sample periods according to how many individuals were captured (controlled by the `document_term_weights`).
-We now believe it is more appropriate to weight periods proportional to captures, and this is what we recommend for LDATS users.
-For the purposes of comparison, we will continue set all weights = 1 for both changepoint models.
-For an example of LDATS run with proportional weights, see the rodents vignette.
-[?]
+Having divided the data to generate catch-per-effort, the paper changepoint model weighted all sample periods equally.
+In comparison, LDATS does not force an equal weighting, but assumes that as default, and can weight sample periods according to how many individuals were captured (controlled by the `weights` argument to `LDA_TS`, and easily calculated for a document-term matrix using `document_term_weights`.
+We now believe it is more appropriate to weight periods proportional to captures in the time series (despite the LDA function returning only proportions of each topic), and this is what we recommend for LDATS users.
+For the purposes of comparison, however, we will continue set all weights = 1 for both changepoint models.
+For an example of LDATS run with proportional weights, see the [rodents vignette](https://weecology.github.io/LDATS/articles/rodents-example.html).
 
@@
 #### Load LDA ####
-load(here::here('vignette', 'christensenetal-comparison-files', 'model-cache', 'ldats_lda_selected.Rds'))
+load(here::here('vignettes', 'christensenetal-comparison-files', 'model-cache', 'ldats_lda_selected.Rds'))
 
@@
 #### Run changepoint ####
-source(here::here('vignette', 'christensenetal-comparison-files', 'christensenetal-2018-files', 'changepointmodel.r'))
+source(here::here('vignettes', 'christensenetal-comparison-files', 'christensenetal-2018-files', 'changepointmodel.r'))
 
@@
 cp_results_rodent = changepoint_model(ldats_lda_selected[[1]], x, 1, weights = rep(1,length(year_continuous)))
-save(cp_results_rodent, file = here::here('vignette', 'christensenetal-comparison-files', 'model-cache', 'ldats_paper_1.Rds'))
+save(cp_results_rodent, file = here::here('vignettes', 'christensenetal-comparison-files', 'model-cache', 'ldats_paper_1.Rds'))
 rm(cp_results_rodent)
@@
 cp_results_rodent2 = changepoint_model(ldats_lda_selected[[1]], x, 2, weights = rep(1,length(year_continuous)))
-save(cp_results_rodent2, file = here::here('vignette', 'christensenetal-comparison-files', 'model-cache', 'ldats_paper_2.Rds'))
+save(cp_results_rodent2, file = here::here('vignettes', 'christensenetal-comparison-files', 'model-cache', 'ldats_paper_2.Rds'))
 rm(cp_results_rodent2)
@@
 cp_results_rodent3 = changepoint_model(ldats_lda_selected[[1]], x, 3, weights = rep(1,length(year_continuous)))
-save(cp_results_rodent3, file = here::here('vignette', 'christensenetal-comparison-files', 'model-cache', 'ldats_paper_3.Rds'))
+save(cp_results_rodent3, file = here::here('vignettes', 'christensenetal-comparison-files', 'model-cache', 'ldats_paper_3.Rds'))
 rm(cp_results_rodent3)
@@
 cp_results_rodent4 = changepoint_model(ldats_lda_selected[[1]], x, 4, weights = rep(1,length(year_continuous)))
-save(cp_results_rodent4, file = here::here('vignette', 'christensenetal-comparison-files', 'model-cache', 'ldats_paper_4.Rds'))
+save(cp_results_rodent4, file = here::here('vignettes', 'christensenetal-comparison-files', 'model-cache', 'ldats_paper_4.Rds'))
 rm(cp_results_rodent4)
@@
 cp_results_rodent5 = changepoint_model(ldats_lda_selected[[1]], x, 5, weights = rep(1,length(year_continuous)))
-save(cp_results_rodent5, file = here::here('vignette', 'christensenetal-comparison-files', 'model-cache', 'ldats_paper_5.Rds'))
+save(cp_results_rodent5, file = here::here('vignettes', 'christensenetal-comparison-files', 'model-cache', 'ldats_paper_5.Rds'))
 rm(cp_results_rodent5)
@@
 cp_results_rodent6 = changepoint_model(ldats_lda_selected[[1]], x, 6, weights = rep(1,length(year_continuous)))
-save(cp_results_rodent6, file = here::here('vignette', 'christensenetal-comparison-files', 'model-cache', 'ldats_paper_6.Rds'))
+save(cp_results_rodent6, file = here::here('vignettes', 'christensenetal-comparison-files', 'model-cache', 'ldats_paper_6.Rds'))
 rm(cp_results_rodent6)
@@
 #### Changepoint model selection ####
-load(here::here('vignette', 'christensenetal-comparison-files', 'model-cache', 'ldats_paper_1.Rds'))
-load(here::here('vignette', 'christensenetal-comparison-files', 'model-cache', 'ldats_paper_2.Rds'))
-load(here::here('vignette', 'christensenetal-comparison-files', 'model-cache', 'ldats_paper_3.Rds'))
-load(here::here('vignette', 'christensenetal-comparison-files', 'model-cache', 'ldats_paper_4.Rds'))
-load(here::here('vignette', 'christensenetal-comparison-files', 'model-cache', 'ldats_paper_5.Rds'))
-load(here::here('vignette', 'christensenetal-comparison-files', 'model-cache', 'ldats_paper_6.Rds'))
+load(here::here('vignettes', 'christensenetal-comparison-files', 'model-cache', 'ldats_paper_1.Rds'))
+load(here::here('vignettes', 'christensenetal-comparison-files', 'model-cache', 'ldats_paper_2.Rds'))
+load(here::here('vignettes', 'christensenetal-comparison-files', 'model-cache', 'ldats_paper_3.Rds'))
+load(here::here('vignettes', 'christensenetal-comparison-files', 'model-cache', 'ldats_paper_4.Rds'))
+load(here::here('vignettes', 'christensenetal-comparison-files', 'model-cache', 'ldats_paper_5.Rds'))
+load(here::here('vignettes', 'christensenetal-comparison-files', 'model-cache', 'ldats_paper_6.Rds'))
 
@@
 #### Load LDA ####
-load(here::here('vignette', 'christensenetal-comparison-files', 'model-cache', 'paper_ldamodel.Rds'))
+load(here::here('vignettes', 'christensenetal-comparison-files', 'model-cache', 'paper_ldamodel.Rds'))
 
@@
 cp_results_rodent = changepoint_model(ldamodel, x, 1, weights = rep(1,length(year_continuous)))
-save(cp_results_rodent, file = here::here('vignette', 'christensenetal-comparison-files', 'model-cache', 'paper_paper1.Rds'))
+save(cp_results_rodent, file = here::here('vignettes', 'christensenetal-comparison-files', 'model-cache', 'paper_paper1.Rds'))
 rm(cp_results_rodent)
@@
 cp_results_rodent2 = changepoint_model(ldamodel, x, 2, weights = rep(1,length(year_continuous)))
-save(cp_results_rodent2, file = here::here('vignette', 'christensenetal-comparison-files', 'model-cache', 'paper_paper2.Rds'))
+save(cp_results_rodent2, file = here::here('vignettes', 'christensenetal-comparison-files', 'model-cache', 'paper_paper2.Rds'))
 rm(cp_results_rodent2)
@@
 cp_results_rodent3 = changepoint_model(ldamodel, x, 3, weights = rep(1,length(year_continuous)))
-save(cp_results_rodent3, file = here::here('vignette', 'christensenetal-comparison-files', 'model-cache', 'paper_paper3.Rds'))
+save(cp_results_rodent3, file = here::here('vignettes', 'christensenetal-comparison-files', 'model-cache', 'paper_paper3.Rds'))
 rm(cp_results_rodent3)
@@
 cp_results_rodent4 = changepoint_model(ldamodel, x, 4, weights = rep(1,length(year_continuous)))
-save(cp_results_rodent4, file = here::here('vignette', 'christensenetal-comparison-files', 'model-cache', 'paper_paper4.Rds'))
+save(cp_results_rodent4, file = here::here('vignettes', 'christensenetal-comparison-files', 'model-cache', 'paper_paper4.Rds'))
 rm(cp_results_rodent4)
@@
 cp_results_rodent5 = changepoint_model(ldamodel, x, 5, weights = rep(1,length(year_continuous)))
-save(cp_results_rodent5, file = here::here('vignette', 'christensenetal-comparison-files', 'model-cache', 'paper_paper5.Rds'))
+save(cp_results_rodent5, file = here::here('vignettes', 'christensenetal-comparison-files', 'model-cache', 'paper_paper5.Rds'))
 rm(cp_results_rodent5)
@@
 cp_results_rodent6 = changepoint_model(ldamodel, x, 6, weights = rep(1,length(year_continuous)))
-save(cp_results_rodent6, file = here::here('vignette', 'christensenetal-comparison-files', 'model-cache', 'paper_paper6.Rds'))
+save(cp_results_rodent6, file = here::here('vignettes', 'christensenetal-comparison-files', 'model-cache', 'paper_paper6.Rds'))
 rm(cp_results_rodent6)
@@
 #### Changepoint model selection ####
-load(here::here('vignette', 'christensenetal-comparison-files', 'model-cache', 'paper_paper_1.Rds'))
-load(here::here('vignette', 'christensenetal-comparison-files', 'model-cache', 'paper_paper_2.Rds'))
-load(here::here('vignette', 'christensenetal-comparison-files', 'model-cache', 'paper_paper_3.Rds'))
-load(here::here('vignette', 'christensenetal-comparison-files', 'model-cache', 'paper_paper_4.Rds'))
-load(here::here('vignette', 'christensenetal-comparison-files', 'model-cache', 'paper_paper_5.Rds'))
-load(here::here('vignette', 'christensenetal-comparison-files', 'model-cache', 'paper_paper_6.Rds'))
+load(here::here('vignettes', 'christensenetal-comparison-files', 'model-cache', 'paper_paper_1.Rds'))
+load(here::here('vignettes', 'christensenetal-comparison-files', 'model-cache', 'paper_paper_2.Rds'))
+load(here::here('vignettes', 'christensenetal-comparison-files', 'model-cache', 'paper_paper_3.Rds'))
+load(here::here('vignettes', 'christensenetal-comparison-files', 'model-cache', 'paper_paper_4.Rds'))
+load(here::here('vignettes', 'christensenetal-comparison-files', 'model-cache', 'paper_paper_5.Rds'))
+load(here::here('vignettes', 'christensenetal-comparison-files', 'model-cache', 'paper_paper_6.Rds'))
 
@@
 
-save(ldats_ldats_cpt, file = here::here('vignette', 'christensenetal-comparison-files', 'model-cache', 'ldats_ldats.Rds'))
+save(ldats_ldats_cpt, file = here::here('vignettes', 'christensenetal-comparison-files', 'model-cache', 'ldats_ldats.Rds'))
 rm(ldats_ldats_cpt)
@@
 ```{r load LDATS LDA and LDATS changepoint, include = F}
-load(here::here('vignette', 'christensenetal-comparison-files', 'model-cache', 'ldats_ldats.Rds'))
+load(here::here('vignettes', 'christensenetal-comparison-files', 'model-cache', 'ldats_ldats.Rds'))
 ```
@@
 
-save(ldats_ldats_cpt_selected, file = here::here('vignette', 'christensenetal-comparison-files', 'model-cache', 'ldats_ldats_selected.Rds'))
+save(ldats_ldats_cpt_selected, file = here::here('vignettes', 'christensenetal-comparison-files', 'model-cache', 'ldats_ldats_selected.Rds'))
 
@@
 
-save(paper_ldats_cpt, file = here::here('vignette', 'christensenetal-comparison-files', 'model-cache', 'paper_ldats.Rds'))
+save(paper_ldats_cpt, file = here::here('vignettes', 'christensenetal-comparison-files', 'model-cache', 'paper_ldats.Rds'))
 rm(paper_ldats_cpt)
@@
 ```{r load paper LDA and LDATS changepoint, include = F}
-load(here::here('vignette', 'christensenetal-comparison-files', 'model-cache', 'paper_ldats.Rds'))
+load(here::here('vignettes', 'christensenetal-comparison-files', 'model-cache', 'paper_ldats.Rds'))
 ```
@@
 Unlike the paper changepoint model, LDATS can recognize that sampling periods may not be equidistant, and can place changepoint estimates at new moons if they fall between nonconsecutive sampling periods.
-We can estimate the dates corresponding to those new moons.
+We can estimate the dates corresponding to those new moons, extrapolating from the census dates for adjacent census periods.
 
@@
 
-save(paper_ldats_cpt_selected, file = here::here('vignette', 'christensenetal-comparison-files', 'model-cache', 'paper_ldats_selected.Rds'))
+save(paper_ldats_cpt_selected, file = here::here('vignettes', 'christensenetal-comparison-files', 'model-cache', 'paper_ldats_selected.Rds'))
 
@@
 
+### Plot changepoint models
+
+```{r load cpts to plot, include = F}
+load(here::here('vignettes', 'christensenetal-comparison-files', 'model-cache', 'paper_ldats_selected.Rds'))
+load(here::here('vignettes', 'christensenetal-comparison-files', 'model-cache', 'ldats_ldats_selected.Rds'))
+load(here::here('vignettes', 'christensenetal-comparison-files', 'model-cache', 'paper_paper_4.Rds'))
+paper_paper_cpt_selected <- cp_results_rodent4
+load(here::here('vignettes', 'christensenetal-comparison-files', 'model-cache', 'ldats_paper_4.Rds'))
+ldats_paper_cpt_selected <- cp_results_rodent4
+
+year_continuous = 1970 + as.integer(julian(paper_dates)) / 365.25
+x = data.frame(
+  year_continuous = year_continuous,
+  sin_year = sin(year_continuous * 2 * pi),
+  cos_year = cos(year_continuous * 2 * pi)
+)
+ntopics = 4
+
+```
+
+#### Paper LDA and LDATS changepoint
+
+```{r plot paper LDA and LDATS cpts, fig.width=6, fig.height=6}
+
+plot(paper_ldats_cpt_selected)
+
+```
+
+#### LDATS LDA and LDATS changepoint
+```{r plot ldats LDA and LDATS cpt, fig.width=6, fig.height=6}
+plot(ldats_ldats_cpt_selected)
+```
+
+#### Paper LDA and paper changepoint
+```{r plot paper LDA and paper cpt, fig.width=6, fig.height=6}
+
+source(here::here('vignettes', 'christensenetal-comparison-files', 'christensenetal-2018-files', 'LDA_figure_scripts.R'))
+source(here::here('vignettes', 'christensenetal-comparison-files', 'christensenetal-2018-files', 'changepointmodel.r'))
+paper_cpts = find_changepoint_location(paper_paper_cpt_selected)
+paper_cpt_plot = get_ll_non_memoized_plot(ldamodel,x,paper_cpts,make_plot=T,weights=rep(1,length(year_continuous)))
+annual_hist(paper_paper_cpt_selected, year_continuous)
+paper_cpt_plot
+
+```
+
+#### LDATS LDA and paper changepoint
+```{r plot LDATS lda and paper cpt, fig.width=6, fig.height=6}
+ntopics = 6
+ldats_cpts = find_changepoint_location(ldats_paper_cpt_selected)
+ldats_cpt_plot = get_ll_non_memoized_plot(ldats_lda_selected[[1]],x,ldats_cpts,make_plot=T,weights=rep(1,length(year_continuous)))
+annual_hist(ldats_paper_cpt_selected, year_continuous)
+ldats_cpt_plot
+```
+
+The results of the changepoint model appear extremely robust to both choice of LDA model and choice of changepoint model.
+
+### Report changepoint dates
 ```{r report cpt dates, include = F}
@@
 The choice of LDA has more influence on the changepoint locations than the choice of changepoint model - probably because the LDATS LDA has 6 topics, and the paper LDA has 4.
+However, all of the models agree to within 6 months in most cases, and a year for the broader early 1990s changepoint.
 
@@
 ```
-
-### Plot changepoint models
-
-```{r load cpts to plot, include = F}
-load(here::here('vignette', 'christensenetal-comparison-files', 'model-cache', 'paper_ldats_selected.Rds'))
-load(here::here('vignette', 'christensenetal-comparison-files', 'model-cache', 'ldats_ldats_selected.Rds'))
-load(here::here('vignette', 'christensenetal-comparison-files', 'model-cache', 'paper_paper_4.Rds'))
-paper_paper_cpt_selected <- cp_results_rodent4
-load(here::here('vignette', 'christensenetal-comparison-files', 'model-cache', 'ldats_paper_4.Rds'))
-ldats_paper_cpt_selected <- cp_results_rodent4
-
-year_continuous = 1970 + as.integer(julian(paper_dates)) / 365.25
-x = data.frame(
-  year_continuous = year_continuous,
-  sin_year = sin(year_continuous * 2 * pi),
-  cos_year = cos(year_continuous * 2 * pi)
-)
-ntopics = 4
-
-```
-
-```{r plot cpts}
-
-plot(paper_ldats_cpt_selected)
-
-plot(ldats_ldats_cpt_selected)
-
-source(here::here('vignette', 'christensenetal-comparison-files', 'christensenetal-2018-files', 'LDA_figure_scripts.R'))
-source(here::here('vignette', 'christensenetal-comparison-files', 'christensenetal-2018-files', 'changepointmodel.r'))
-paper_cpts = find_changepoint_location(paper_paper_cpt_selected)
-paper_cpt_plot = get_ll_non_memoized_plot(ldamodel,x,paper_cpts,make_plot=T,weights=rep(1,length(year_continuous)))
-annual_hist(paper_paper_cpt_selected, year_continuous)
-paper_cpt_plot
-
-ntopics = 6
-ldats_cpts = find_changepoint_location(ldats_paper_cpt_selected)
-ldats_cpt_plot = get_ll_non_memoized_plot(ldats_lda_selected[[1]],x,ldats_cpts,make_plot=T,weights=rep(1,length(year_continuous)))
-annual_hist(ldats_paper_cpt_selected, year_continuous)
-ldats_cpt_plot
-```

### 2019-02-05 doc/rodents-example.R

@@ -1,0 +1,114 @@
+## ----setup, include=FALSE------------------------------------------------
+knitr::opts_chunk$set(
+  collapse = TRUE,
+  comment = "#>"
+)
+
+## ---- include=FALSE------------------------------------------------------
+library(LDATS)
+vers <- packageVersion("LDATS")
+today <- Sys.Date()
+
+## ---- eval=FALSE---------------------------------------------------------
+#  install.packages("devtools")
+#  devtools::install_github("weecology/LDATS")
+
+## ------------------------------------------------------------------------
+data(rodents)
+head(rodents$document_term_table, 10)
+head(rodents$document_covariate_table, 10)
+
+## ----lda_set, eval =F----------------------------------------------------
+#  lda_model_set <- LDA_set(document_term_table = rodents$document_term_table,
+#                           topics = c(2:6),
+#                           nseeds = 100,
+#                          control = LDA_controls_list(quiet = TRUE))
+#  
+
+## ----lda set not quiet---------------------------------------------------
+lda_model_set2 <- LDA_set(document_term_table = rodents$document_term_table,
+                         topics = c(2:3),
+                         nseeds = 10)
+
+## ----load lda model set, include = F-------------------------------------
+load(here::here('vignettes', 'rodents-example-files', 'lda_model_set.Rds'))
+rm(lda_model_set2)
+
+## ----select LDA----------------------------------------------------------
+selected_lda_model <- select_LDA(lda_model_set)
+
+## ----LDA results---------------------------------------------------------
+# Number of topics:
+
+selected_lda_model[[1]]@k
+
+# Topic composition of communities at each time step
+# Columns are topics; rows are time steps
+head(selected_lda_model[[1]]@gamma)
+
+
+## ----plot lda, fig.width=7, fig.height=6---------------------------------
+plot(selected_lda_model[[1]])
+
+
+## ----ts on lda, eval = F-------------------------------------------------
+#  changepoint_models <- TS_on_LDA(LDA_models = selected_lda_model,
+#                                  document_covariate_table = rodents$document_covariate_table,
+#                                  formulas = ~ sin_year + cos_year,
+#                                  nchangepoints = c(0:6),
+#                                  weights = document_weights(rodents$document_term_table),
+#                                                             control = TS_controls_list(nit = 1000))
+#  
+#  save(changepoint_models, file = here::here('vignettes', 'rodents-example-files', 'changepoint_models.Rds'))
+#  
+
[…]

### 2019-02-07 vignettes/rodents-example.R
fix DESCRIPTION and update tname in rodents-example

@@ -30,3 +30,82 @@ lda_model_set2 <- LDA_set(document_term_table = rodents$document_term_table,
                          topics = c(2:3),
                          nseeds = 10)
 
+## ----load lda model set, include = F-------------------------------------
+load(here::here('vignettes', 'rodents-example-files', 'lda_model_set.Rds'))
+rm(lda_model_set2)
+
+## ----select LDA----------------------------------------------------------
+selected_lda_model <- select_LDA(lda_model_set)
+
+## ----LDA results---------------------------------------------------------
+# Number of topics:
+
+selected_lda_model[[1]]@k
+
+# Topic composition of communities at each time step
+# Columns are topics; rows are time steps
+head(selected_lda_model[[1]]@gamma)
+
+
+## ----plot lda, fig.width=7, fig.height=6---------------------------------
+plot(selected_lda_model[[1]])
+
+
+## ----ts on lda, eval = F-------------------------------------------------
+#  changepoint_models <- TS_on_LDA(LDA_models = selected_lda_model,
+#                                  document_covariate_table = rodents$document_covariate_table,
+#                                  formulas = ~ sin_year + cos_year,
+#                                  nchangepoints = c(0:6),
+#                                  weights = document_weights(rodents$document_term_table),
+#                                                             control = TS_controls_list(nit = 1000))
+#  
+#  save(changepoint_models, file = here::here('vignettes', 'rodents-example-files', 'changepoint_models.Rds'))
+#  
+
+## ----reload ts, include = F----------------------------------------------
+load(here::here('vignettes', 'rodents-example-files', 'changepoint_models.Rds'))
+
+## ----select ts-----------------------------------------------------------
+selected_changepoint_model <- select_TS(changepoint_models)
+
+## ----cpt results---------------------------------------------------------
+# Number of changepoints
+selected_changepoint_model$nchangepoints
+
+# Summary of timesteps (newmoon values) for each changepoint
+selected_changepoint_model$rho_summary
+
+# Raw estimates for timesteps for each changepoint
+# Changepoints are columns
+head(selected_changepoint_model$rhos)
+
+
+## ----plot cpt, fig.width=7, fig.height=6---------------------------------
+plot(selected_changepoint_model)
+
+## ----lda_ts, eval = F----------------------------------------------------
+#  lda_ts_results <- LDA_TS(document_term_table = rodents$document_term_table,
+#                           document_covariate_table = rodents$document_covariate_table,
[…]

### 2019-02-07 doc/rodents-example.Rmd
rebuild vignettes, trying to resolve rebase

---
title: "LDATS Rodents Example"
author: "Renata Diaz and Juniper L. Simonis"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{rodentsexample}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r, include=FALSE}
library(LDATS)
vers <- packageVersion("LDATS")
today <- Sys.Date()
```

This vignette walks through an example of **LDATS** at the command line and
was constructed using **LDATS** version `r vers` on `r today`.

## Installation

To obtain the most recent version of **LDATS**, install the most recent 
version from GitHub:

```{r, eval=FALSE}
install.packages("devtools")
devtools::install_github("weecology/LDATS")
```

## Data

For this vignette, we will be using rodent data from the control plots of the
[Portal Project](https://github.com/weecology/portaldata), which come with
the **LDATS** package (`data(rodents)`).

`rodents` contains two data tables, a `document_term_table` and a `document_covariate_table`.

The `document_term_table` is a matrix of species (term) counts by sampling period (document). Importantly, the `document_term_table` contains raw counts of individual rodent captures, not catch per unit effort. The LDA models operate on proportions and weight the information from each sampling period according to total catch (document size). Similarly, the TS models use the LDA compositional breakdown and weight the information from each period according to the total catch. We therefore give both models un-adjusted data. The values in the `document_term_table` must be integers.

The `document_covariate_table` contains the covariates we would like to use for the time series models. In this case, we have a column for `newmoon`, which is the time step for each of our sampling periods. This column can include nonconsecutive time steps (i.e., `2, 4, 5`) if sampling periods are skipped or at unequal intervals. If there is no time column, `LDATS` will assume all time steps are equidistant. We also include columns `sin_year` and `cos_year` so we can account for seasonal dynamics.

```{r}
data(rodents)
head(rodents$document_term_table, 10)
head(rodents$document_covariate_table, 10)
```

## Stage 1: LDA models

We use `LDA_set()` to run replicate LDA models (each with its own seed) with varying numbers of topics (`2:6`) and `select_LDA()` to select the best model.

We use `LDA_controls_list())` to pass controls to the LDA function. In this case, we can set `quiet = TRUE` to make the model run quietly.

```{r lda_set, eval =F}
lda_model_set <- LDA_set(document_term_table = rodents$document_term_table,
                         topics = c(2:6),
                         nseeds = 100,
                        control = LDA_controls_list(quiet = TRUE))

```

If we do not pass any controls, by default, `quiet = FALSE` (here run with only `2:3` topics and  `10` seeds, to keep output short):

```{r lda set not quiet}
lda_model_set2 <- LDA_set(document_term_table = rodents$document_term_table,
                         topics = c(2:3),
                         nseeds = 10)
```

`LDA_set()` returns a list of LDA models. We use `select_LDA()` to identify the best number of topics and choice of seed from our set of models. By default, we will choose models based on minimum AIC. To use different selection criteria, define the appropriate functions and specify them using `LDA_controls_list(measurer = [measurer function], selector = [max, min, etc])`.

```{r load lda model set, include = F}
load(here::here('vignettes', 'rodents-example-files', 'lda_model_set.Rds'))
rm(lda_model_set2)
```

```{r select LDA}
selected_lda_model <- select_LDA(lda_model_set)
```

We can access the results of the model:

```{r LDA results}
# Number of topics:

selected_lda_model[[1]]@k

# Topic composition of communities at each time step
# Columns are topics; rows are time steps
head(selected_lda_model[[1]]@gamma)

```

`LDATS` includes flexible plot functionality for LDAs and time series. The top panel illustrates topic composition by species, and the bottom panel shows the proportion of the community made up of each topic over time. For all the available plot options see `?plot.LDA_VEM`.

```{r plot lda, fig.width=7, fig.height=6}
plot(selected_lda_model[[1]])

```

## Stage 2: TS changepoint models

We use `TS_on_LDA()` to run LDATS changepoint models with `0:6` changepoints, and then use `select_TS()` to find the best-fit model of these.

Here, `TS_on_LDA()` predicts the `gamma` (the proportion of the community made of up each topic) from our LDA model(s) as a function of `sin_year` and `cos_year` in the `document_covariate_table`. We use `document_weights()` to weight the information from each time step according to the total number of rodents captured at that time step.

```{r ts on lda, eval = F}
changepoint_models <- TS_on_LDA(LDA_models = selected_lda_model, 
                                document_covariate_table = rodents$document_covariate_table,
                                formulas = ~ sin_year + cos_year,
                                nchangepoints = c(0:6), 
                                weights = document_weights(rodents$document_term_table),
                                control = TS_controls_list(nit = 1000, timename = "newmoon"))

```

We can adjust options for both TS functions using `TS_controls_list()`. For a full list see `?TS_controls_list`. Here we illustrate adjusting the number of ptMCMC iterations - the default is 10000, but it is convenient to use fewer iterations for code development.

Also, it is important to note that by default the TS functions take the name of the time-step column from the `document_covariate_table` to be `"time"`. To pass a different column name, use the `timename` argument in `TS_controls_list()`.

`select_TS()` will identify the best-fit changepoint model of the models from `TS_on_LDA()`. As with `select_LDA()`, we can adjust the `measurer` and `selector` using `TS_controls_list()`.

```{r reload ts, include = F}
load(here::here('vignettes', 'rodents-example-files', 'changepoint_models.Rds'))
```

```{r select ts}
selected_changepoint_model <- select_TS(changepoint_models)
```

We can access the results of the selected changepoint model:

```{r cpt results}
# Number of changepoints
selected_changepoint_model$nchangepoints

# Summary of timesteps (newmoon values) for each changepoint
selected_changepoint_model$rho_summary

# Raw estimates for timesteps for each changepoint
# Changepoints are columns
head(selected_changepoint_model$rhos)

```

LDATS will plot the results of a changepoint model:

```{r plot cpt, fig.width=7, fig.height=6}
plot(selected_changepoint_model)
```

## Full analysis with `LDA_TS`

Finally, we can perform an entire LDATS analysis, including all of the above steps, using the `LDA_TS()` function and passing options to the LDA and TS functions using `LDA_TS_controls_list()`.

```{r lda_ts, eval = F}
lda_ts_results <- LDA_TS(document_term_table = rodents$document_term_table,
                         document_covariate_table = rodents$document_covariate_table,
                         nseeds = 100, 
                         topics = 2:6,
                         formulas = ~ sin_year + cos_year,
                         nchangepoints= 0:6,
                         weights = document_weights(rodents$document_term_table),
                         control = LDA_TS_controls_list(
                           TS_control = TS_controls_list(nit = 1000, timename = "newmoon")))
```

```{r load ldats results, include = F}
load(here::here('vignettes', 'rodents-example-files', 'lda_ts_results.Rds'))
```

`LDA_TS()` returns a list of all the model objects, and we can access their contents as above:

```{r LDA_TS results}
names(lda_ts_results)

# Number of topics
lda_ts_results$`Selected LDA model`$k@k

# Number of changepoints
lda_ts_results$`Selected TS model`$nchangepoints

# Summary of changepoint locations
lda_ts_results$`Selected TS model`$rho_summary
```

Finally, we can plot the `LDA_TS` results.

```{r plot LDA_TS results, fig.height = 16, fig.width = 7, echo = F}
plot(lda_ts_results, control = LDA_TS_controls_list(
                           TS_control = TS_controls_list(timename = "newmoon")))
```

### 2019-02-07 doc/paper-comparison.R

@@ -75,7 +75,7 @@ length(which(rowSums(compare) < ncol(compare)))
 trap_table = read.csv('https://raw.githubusercontent.com/weecology/PortalData/master/Rodents/Portal_rodent_trapping.csv')
   trap_table_controls =dplyr::filter(trap_table, plot %in% c(2,4,8,11,12,14,17,22))
   nplots_controls = aggregate(trap_table_controls$sampled,by=list(period = trap_table_controls$period),FUN=sum)
-  
+
    # adjust species counts by number of plots trapped that month
   ldats_rodents_adjusted = as.data.frame.matrix(rodents[[1]])
   for (n in 1:436) {
@@ -83,7 +83,7 @@ trap_table = read.csv('https://raw.githubusercontent.com/weecology/PortalData/ma
     ldats_rodents_adjusted[n,] = round(ldats_rodents_adjusted[n,]/nplots_controls$x[n]*8)
   }
 
-  
+
 compare_raw <- rodents[[1]] == ldats_rodents_adjusted
 length(which(rowSums(compare_raw) < ncol(compare_raw)))
 
@@ -253,7 +253,7 @@ mean(cp_results_rodent4$saved_lls * -2)+ 2*(3*(ntopics-1)*(4+1)+(4))
 mean(cp_results_rodent5$saved_lls * -2)+ 2*(3*(ntopics-1)*(5+1)+(5))
 mean(cp_results_rodent6$saved_lls * -2)+ 2*(3*(ntopics-1)*(6+1)+(6))
 
-# The lowest deviance is for 4 changepoints. 
+# The lowest deviance is for 4 changepoints.
 
 
 df_4 = as.data.frame(t(cp_results_rodent4$saved[,1,])) %>% reshape::melt()
@@ -336,7 +336,7 @@ mean(cp_results_rodent4$saved_lls * -2)+ 2*(3*(ntopics-1)*(4+1)+(4))
 mean(cp_results_rodent5$saved_lls * -2)+ 2*(3*(ntopics-1)*(5+1)+(5))
 mean(cp_results_rodent6$saved_lls * -2)+ 2*(3*(ntopics-1)*(6+1)+(6))
 
-# The lowest deviance is for 4 changepoints. 
+# The lowest deviance is for 4 changepoints.
 
 df_4 = as.data.frame(t(cp_results_rodent4$saved[,1,])) %>%  reshape::melt()
 df_4$value = year_continuous[df_4$value]

### 2019-02-07 doc/rodents-example.R

@@ -57,9 +57,7 @@ plot(selected_lda_model[[1]])
 #                                  formulas = ~ sin_year + cos_year,
 #                                  nchangepoints = c(0:6),
 #                                  weights = document_weights(rodents$document_term_table),
-#                                                             control = TS_controls_list(nit = 1000))
-#  
-#  save(changepoint_models, file = here::here('vignettes', 'rodents-example-files', 'changepoint_models.Rds'))
+#                                  control = TS_controls_list(nit = 1000, timename = "newmoon"))
 #  
 
 ## ----reload ts, include = F----------------------------------------------
@@ -92,7 +90,7 @@ plot(selected_changepoint_model)
 #                           nchangepoints= 0:6,
 #                           weights = document_weights(rodents$document_term_table),
 #                           control = LDA_TS_controls_list(
-#                             TS_control = TS_controls_list(nit = 1000)))
+#                             TS_control = TS_controls_list(nit = 1000, timename = "newmoon")))
 
 ## ----load ldats results, include = F-------------------------------------
 load(here::here('vignettes', 'rodents-example-files', 'lda_ts_results.Rds'))
@@ -109,6 +107,7 @@ lda_ts_results$`Selected TS model`$nchangepoints
 # Summary of changepoint locations
 lda_ts_results$`Selected TS model`$rho_summary
 
-## ----plot LDA_TS results, fig.height = 16, fig.width = 7-----------------
-plot(lda_ts_results)
+## ----plot LDA_TS results, fig.height = 16, fig.width = 7, echo = F-------
+plot(lda_ts_results, control = LDA_TS_controls_list(
+                           TS_control = TS_controls_list(timename = "newmoon")))
 

### 2019-02-07 vignettes/paper-comparison.R

@@ -1,523 +1,0 @@
-## ----setup, include=FALSE------------------------------------------------
-knitr::opts_chunk$set(
-  collapse = TRUE,
-  comment = "#>"
-)
-
-## ---- eval=FALSE---------------------------------------------------------
-#  install.packages("devtools")
-#  devtools::install_github("weecology/LDATS")
-
-## ---- eval = T-----------------------------------------------------------
-library(LDATS)
-
-
-## ----download Christensen------------------------------------------------
-
-paper_filepath <- here::here('vignettes', 'christensenetal-comparison-files', 'christensenetal-2018-files')
-test_filepath <- here::here('vignettes', 'christensenetal-ristensenetal-comparison-files', 'christensenetal-2018-files', 'rodent_LDA_analysis.r')
-
-if(!file.exists(test_filepath)) {
-
-files_to_download <- c('rodent_LDA_analysis.r', 'rodent_data_for_LDA.r', 'AIC_model_selection.R', 'changepointmodel.r', 'LDA-distance.R', 'Rodent_table_dat.csv', 'LDA_figure_scripts.R')
-
-for(i in 1:length(files_to_download)) {
-  download.file(url = paste0("https://raw.githubusercontent.com/emchristensen/Extreme-events-LDA/master/", files_to_download[i]),
-                destfile = paste0(paper_filepath, '/', files_to_download[i]))
-}
-
-rm(files_to_download)
-rm(i)
-
-}
-
-rm(test_filepath)
-rm(paper_filepath)
-
-## ----LDATS data----------------------------------------------------------
-
-data(rodents)
-
-head(rodents[[1]])
-
-
-## ----Paper data----------------------------------------------------------
-source(here::here('vignettes', 'christensenetal-comparison-files', 'christensenetal-2018-files', 'rodent_data_for_LDA.r'))
-
-dat = create_rodent_table(period_first = 1,
-                          period_last = 436,
-                          selected_plots = c(2,4,8,11,12,14,17,22),
-                          selected_species = c('BA','DM','DO','DS','NA','OL','OT','PB','PE','PF','PH','PI','PL','PM','PP','RF','RM','RO','SF','SH','SO'))
-
-# dates to go with count data
[…]

### 2019-02-11 vignettes/rodents-example.R
Resolve conflicts, fix ambiguous naming of changepoint objects, prepare to consolidate all .Rds files into one file to load

@@ -1,0 +1,113 @@
+## ----setup, include=FALSE------------------------------------------------
+knitr::opts_chunk$set(
+  collapse = TRUE,
+  comment = "#>"
+)
+
+## ---- include=FALSE------------------------------------------------------
+library(LDATS)
+vers <- packageVersion("LDATS")
+today <- Sys.Date()
+
+## ---- eval=FALSE---------------------------------------------------------
+## install.packages("devtools")
+## devtools::install_github("weecology/LDATS")
+
+## ------------------------------------------------------------------------
+data(rodents)
+head(rodents$document_term_table, 10)
+head(rodents$document_covariate_table, 10)
+
+## ----lda_set, eval =F----------------------------------------------------
+## lda_model_set <- LDA_set(document_term_table = rodents$document_term_table,
+##                          topics = c(2:6),
+##                          nseeds = 100,
+##                         control = LDA_controls_list(quiet = TRUE))
+## 
+
+## ----lda set not quiet---------------------------------------------------
+lda_model_set2 <- LDA_set(document_term_table = rodents$document_term_table,
+                         topics = c(2:3),
+                         nseeds = 10)
+
+## ----load lda model set, include = F-------------------------------------
+load(here::here('vignettes', 'rodents-example-files', 'lda_model_set.Rds'))
+rm(lda_model_set2)
+
+## ----select LDA----------------------------------------------------------
+selected_lda_model <- select_LDA(lda_model_set)
+
+## ----LDA results---------------------------------------------------------
+# Number of topics:
+
+selected_lda_model[[1]]@k
+
+# Topic composition of communities at each time step
+# Columns are topics; rows are time steps
+head(selected_lda_model[[1]]@gamma)
+
+
+## ----plot lda, fig.width=7, fig.height=6---------------------------------
+plot(selected_lda_model[[1]])
+
+
+## ----ts on lda, eval = F-------------------------------------------------
+## changepoint_models <- TS_on_LDA(LDA_models = selected_lda_model,
+##                                 document_covariate_table = rodents$document_covariate_table,
+##                                 formulas = ~ sin_year + cos_year,
+##                                 nchangepoints = c(0:6),
+##                                 weights = document_weights(rodents$document_term_table),
+##                                 control = TS_controls_list(nit = 1000, timename = "newmoon"))
+## 
+
+## ----reload ts, include = F----------------------------------------------
[…]

### 2019-02-11 doc/paper-comparison.Rmd
Build vignettes on RMD's machine

@@
 )
+
 ```
@@
 Side-by-side comparison of LDATS results with analysis from Christensen et al 2018.
+
+## Summary
+
+| Step | Changes from Christensen et al 2018 to `LDATS` | Effect on comparison | Recommendations for future users |
+|:---------------|:------------------------|:-----------------------|:----------------------------|
+|Data|Paper adjusts abundances according to unit effort, while `LDATS` uses raw capture numbers.|None: run comparison using adjusted data|Use raw, unweighted abundances.|
+|LDA model selection|Paper conservatively overestimated the number of parameters for calculating AIC for model selection.
+`LDATS` calculates AIC appropriately.|Paper LDA selects 4 topics, while `LDATS` finds 6.
+Compare all combinations of paper and `LDATS` LDA models and changepoint models| Use `LDATS` AIC calculation.|
+|Changepoint model document weights|Paper weighted all sampling periods equally regardless of abundance.
+`LDATS` by default weights the information from each sampling period according to the number of individuals captured (i.e.
+the amount of information gleaned about the underlying community composition).|None; use `weights = NULL` to set all weights equal to 1 for LDATS| Weight sampling periods according to abundance|
+|Overall LDA + changepoint results| All combinations of LDA + changepoint model find 4 changepoints at approximately the same time steps|Choice of LDA model has more of an effect than choice of changepoint model|`LDATS` reflects best practices, but the paper methods produced qualitatively similar results.|
 
@@
 library(LDATS)
-
+library(dplyr)
+load(here::here('vignettes', 'christensenetal-comparison-files',  'data_files.Rdata'))
+load(here::here('vignettes', 'christensenetal-comparison-files', 'model-cache',  'ldas.RData'))
+load(here::here('vignettes', 'christensenetal-comparison-files', 'model-cache', 'ldats_changepoints.RData'))
+load(here::here('vignettes', 'christensenetal-comparison-files', 'model-cache', 'paper_changepoints.RData'))
 ```
@@
 
-```{r download Christensen}
+```{r download Christensen, eval = F}
 
 paper_filepath <- here::here('vignettes', 'christensenetal-comparison-files', 'christensenetal-2018-files')
-test_filepath <- here::here('vignettes', 'christensenetal-ristensenetal-comparison-files', 'christensenetal-2018-files', 'rodent_LDA_analysis.r')
+test_filepath <- here::here('vignettes', 'christensenetal-comparison-files', 'christensenetal-2018-files', 'rodent_LDA_analysis.r')
 
@@
 ```{r LDATS data}
-
 data(rodents)
@@
 head(rodents[[1]])
-
 ```
@@
 
-```{r Paper data}
+```{r Paper data, eval = F}
 source(here::here('vignettes', 'christensenetal-comparison-files', 'christensenetal-2018-files', 'rodent_data_for_LDA.r'))
@@
 
-```{r adjust LDATS data after Christensen et al}
+```{r adjust LDATS data after Christensen et al, eval = F}
 
 trap_table = read.csv('https://raw.githubusercontent.com/weecology/PortalData/master/Rodents/Portal_rodent_trapping.csv')
+
+```
+
+```{r continue adjust LDATS after paper}
  trap_table_controls =dplyr::filter(trap_table, plot %in% c(2,4,8,11,12,14,17,22))
   nplots_controls = aggregate(trap_table_controls$sampled,by=list(period = trap_table_controls$period),FUN=sum)
-
    # adjust species counts by number of plots trapped that month
@@
 
-```{r rm LDATS LDAS, include = F, eval = F}
-rm(list = c('ldats_ldas', 'ldats_lda_selected'))
-```
-
 Paper LDAS:
@@
 
-```{r cleanup paper LDAS, include = F, eval = F}
-rm(list = c('dat', 'ldamodel', 'max_dist', 'mean_dist',
-            'ntopics', 'SEED', 'seeds', 'seeds_4topics', 'aic_model', 'aic_model_gibbs', 'calculate_LDA_distance', 'Hellinger', 'min_H', 'repeat_VEM', 'best_ntopic', 'best_seed'))
-```
-
-```{r reload LDAS, include = FALSE}
-
-load(here::here('vignettes', 'christensenetal-comparison-files', 'model-cache', 'ldats_lda_selected.Rds'))
-load(here::here('vignettes', 'christensenetal-comparison-files', 'model-cache', 'paper_ldamodel.Rds'))
-
-```
-
 ### Plot LDAS
@@
 #### Paper LDA
-```{r plot paper LDA, fig.width=6, fig.height=6}
+```{r plot paper LDA, fig.width=7, fig.height=6}
 # Paper
@@
 #### LDATS LDA
-```{r plot LDATS LDA, fig.width=6, fig.height=6}
+```{r plot LDATS LDA, fig.width=7, fig.height=6}
 # LDATS
 plot(ldats_lda_selected[[1]], cols = NULL, option = "D")
-```
-
-```{r remove LDAS after plotting, include = F}
-rm(ldamodel)
-rm(ldats_lda_selected)
 ```
@@
 
-```{r load LDATS LDA for paper cpt}
-
-#### Load LDA ####
-load(here::here('vignettes', 'christensenetal-comparison-files', 'model-cache', 'ldats_lda_selected.Rds'))
-
-```
-
 ```{r run LDATS LDA and paper cpt, eval = F}
@@
 # run models with 1, 2, 3, 4, 5, 6 changepoints
-cp_results_rodent = changepoint_model(ldats_lda_selected[[1]], x, 1, weights = rep(1,length(year_continuous)))
-save(cp_results_rodent, file = here::here('vignettes', 'christensenetal-comparison-files', 'model-cache', 'ldats_paper_1.Rds'))
-rm(cp_results_rodent)
+ldats_paper_cp_results_rodent = changepoint_model(ldats_lda_selected[[1]], x, 1, weights = rep(1,length(year_continuous)))
+save(ldats_paper_cp_results_rodent, file = here::here('vignettes', 'christensenetal-comparison-files', 'model-cache', 'ldats_paper_1.Rds'))
+rm(ldats_paper_cp_results_rodent)
 
-cp_results_rodent2 = changepoint_model(ldats_lda_selected[[1]], x, 2, weights = rep(1,length(year_continuous)))
-save(cp_results_rodent2, file = here::here('vignettes', 'christensenetal-comparison-files', 'model-cache', 'ldats_paper_2.Rds'))
-rm(cp_results_rodent2)
+ldats_paper_cp_results_rodent2 = changepoint_model(ldats_lda_selected[[1]], x, 2, weights = rep(1,length(year_continuous)))
+save(ldats_paper_cp_results_rodent2, file = here::here('vignettes', 'christensenetal-comparison-files', 'model-cache', 'ldats_paper_2.Rds'))
+rm(ldats_paper_cp_results_rodent2)
 
-cp_results_rodent3 = changepoint_model(ldats_lda_selected[[1]], x, 3, weights = rep(1,length(year_continuous)))
-save(cp_results_rodent3, file = here::here('vignettes', 'christensenetal-comparison-files', 'model-cache', 'ldats_paper_3.Rds'))
-rm(cp_results_rodent3)
+ldats_paper_cp_results_rodent3 = changepoint_model(ldats_lda_selected[[1]], x, 3, weights = rep(1,length(year_continuous)))
+save(ldats_paper_cp_results_rodent3, file = here::here('vignettes', 'christensenetal-comparison-files', 'model-cache', 'ldats_paper_3.Rds'))
+rm(ldats_paper_cp_results_rodent3)
 
-cp_results_rodent4 = changepoint_model(ldats_lda_selected[[1]], x, 4, weights = rep(1,length(year_continuous)))
-save(cp_results_rodent4, file = here::here('vignettes', 'christensenetal-comparison-files', 'model-cache', 'ldats_paper_4.Rds'))
-rm(cp_results_rodent4)
+ldats_paper_cp_results_rodent4 = changepoint_model(ldats_lda_selected[[1]], x, 4, weights = rep(1,length(year_continuous)))
+save(ldats_paper_cp_results_rodent4, file = here::here('vignettes', 'christensenetal-comparison-files', 'model-cache', 'ldats_paper_4.Rds'))
+rm(ldats_paper_cp_results_rodent4)
 
-cp_results_rodent5 = changepoint_model(ldats_lda_selected[[1]], x, 5, weights = rep(1,length(year_continuous)))
-save(cp_results_rodent5, file = here::here('vignettes', 'christensenetal-comparison-files', 'model-cache', 'ldats_paper_5.Rds'))
-rm(cp_results_rodent5)
+ldats_paper_cp_results_rodent5 = changepoint_model(ldats_lda_selected[[1]], x, 5, weights = rep(1,length(year_continuous)))
+save(ldats_paper_cp_results_rodent5, file = here::here('vignettes', 'christensenetal-comparison-files', 'model-cache', 'ldats_paper_5.Rds'))
+rm(ldats_paper_cp_results_rodent5)
 
-cp_results_rodent6 = changepoint_model(ldats_lda_selected[[1]], x, 6, weights = rep(1,length(year_continuous)))
-save(cp_results_rodent6, file = here::here('vignettes', 'christensenetal-comparison-files', 'model-cache', 'ldats_paper_6.Rds'))
-rm(cp_results_rodent6)
+ldats_paper_cp_results_rodent6 = changepoint_model(ldats_lda_selected[[1]], x, 6, weights = rep(1,length(year_continuous)))
+save(ldats_paper_cp_results_rodent6, file = here::here('vignettes', 'christensenetal-comparison-files', 'model-cache', 'ldats_paper_6.Rds'))
+rm(ldats_paper_cp_results_rodent6)
 ```
 
-```{r select LDATS LDA and paper cpt}
+```{r select LDATS LDA and paper cpt, eval = F}
 #### Changepoint model selection ####
@@
 
+```
+
+```{r select LDATS-paper, eval = T}
 ntopics = ldats_lda_selected[[1]]@k
@@
 # mean deviance ( -2 * log likelihood) + 2*(#parameters)
-mean(cp_results_rodent$saved_lls * -2) + 2*(3*(ntopics-1)*(1+1)+(1))
-mean(cp_results_rodent2$saved_lls * -2)+ 2*(3*(ntopics-1)*(2+1)+(2))
-mean(cp_results_rodent3$saved_lls * -2)+ 2*(3*(ntopics-1)*(3+1)+(3))
-mean(cp_results_rodent4$saved_lls * -2)+ 2*(3*(ntopics-1)*(4+1)+(4))
-mean(cp_results_rodent5$saved_lls * -2)+ 2*(3*(ntopics-1)*(5+1)+(5))
-mean(cp_results_rodent6$saved_lls * -2)+ 2*(3*(ntopics-1)*(6+1)+(6))
+mean(ldats_paper_cp_results_rodent$saved_lls * -2) + 2*(3*(ntopics-1)*(1+1)+(1))
+mean(ldats_paper_cp_results_rodent2$saved_lls * -2)+ 2*(3*(ntopics-1)*(2+1)+(2))
+mean(ldats_paper_cp_results_rodent3$saved_lls * -2)+ 2*(3*(ntopics-1)*(3+1)+(3))
+mean(ldats_paper_cp_results_rodent4$saved_lls * -2)+ 2*(3*(ntopics-1)*(4+1)+(4))
+mean(ldats_paper_cp_results_rodent5$saved_lls * -2)+ 2*(3*(ntopics-1)*(5+1)+(5))
+mean(ldats_paper_cp_results_rodent6$saved_lls * -2)+ 2*(3*(ntopics-1)*(6+1)+(6))
 
@@
 
-df_4 = as.data.frame(t(cp_results_rodent4$saved[,1,])) %>% reshape::melt()
+ldats_paper_cpt_selected = ldats_paper_cp_results_rodent4
+
+df_4 = as.data.frame(t(ldats_paper_cpt_selected$saved[,1,])) %>% reshape::melt()
 year_continuous = 1970 + as.integer(julian(paper_dates)) / 365.25
@@
 
-```{r clean up LDATS LDA and paper changepoint, include = F}
-rm(list = c('cp_results_rodent', 'cp_results_rodent2', 'cp_results_rodent3', 'cp_results_rodent4', 'cp_results_rodent5', 'cp_results_rodent6', 'ntopics', 'ldats_lda_selected', 'df_4'))
-```
-
 #### Paper LDA and paper changepoint
 
-```{r load paper LDA for paper cpt}
+```{r load paper LDA for paper cpt, eval = F}
 
@@
 # run models with 1, 2, 3, 4, 5, 6 changepoints
-cp_results_rodent = changepoint_model(ldamodel, x, 1, weights = rep(1,length(year_continuous)))
-save(cp_results_rodent, file = here::here('vignettes', 'christensenetal-comparison-files', 'model-cache', 'paper_paper1.Rds'))
-rm(cp_results_rodent)
+paper_paper_cp_results_rodent = changepoint_model(ldamodel, x, 1, weights = rep(1,length(year_continuous)))
+save(paper_paper_cp_results_rodent, file = here::here('vignettes', 'christensenetal-comparison-files', 'model-cache', 'paper_paper_1.Rds'))
+rm(paper_paper_cp_results_rodent)
 
-cp_results_rodent2 = changepoint_model(ldamodel, x, 2, weights = rep(1,length(year_continuous)))
-save(cp_results_rodent2, file = here::here('vignettes', 'christensenetal-comparison-files', 'model-cache', 'paper_paper2.Rds'))
-rm(cp_results_rodent2)
+paper_paper_cp_results_rodent2 = changepoint_model(ldamodel, x, 2, weights = rep(1,length(year_continuous)))
+save(paper_paper_cp_results_rodent2, file = here::here('vignettes', 'christensenetal-comparison-files', 'model-cache', 'paper_paper_2.Rds'))
+rm(paper_paper_cp_results_rodent2)
 
-cp_results_rodent3 = changepoint_model(ldamodel, x, 3, weights = rep(1,length(year_continuous)))
-save(cp_results_rodent3, file = here::here('vignettes', 'christensenetal-comparison-files', 'model-cache', 'paper_paper3.Rds'))
-rm(cp_results_rodent3)
+paper_paper_cp_results_rodent3 = changepoint_model(ldamodel, x, 3, weights = rep(1,length(year_continuous)))
+save(paper_paper_cp_results_rodent3, file = here::here('vignettes', 'christensenetal-comparison-files', 'model-cache', 'paper_paper_3.Rds'))
+rm(paper_paper_cp_results_rodent3)
 
-cp_results_rodent4 = changepoint_model(ldamodel, x, 4, weights = rep(1,length(year_continuous)))
-save(cp_results_rodent4, file = here::here('vignettes', 'christensenetal-comparison-files', 'model-cache', 'paper_paper4.Rds'))
-rm(cp_results_rodent4)
+paper_paper_cp_results_rodent4 = changepoint_model(ldamodel, x, 4, weights = rep(1,length(year_continuous)))
+save(paper_paper_cp_results_rodent4, file = here::here('vignettes', 'christensenetal-comparison-files', 'model-cache', 'paper_paper_4.Rds'))
+rm(paper_paper_cp_results_rodent4)
 
-cp_results_rodent5 = changepoint_model(ldamodel, x, 5, weights = rep(1,length(year_continuous)))
-save(cp_results_rodent5, file = here::here('vignettes', 'christensenetal-comparison-files', 'model-cache', 'paper_paper5.Rds'))
-rm(cp_results_rodent5)
+paper_paper_cp_results_rodent5 = changepoint_model(ldamodel, x, 5, weights = rep(1,length(year_continuous)))
+save(paper_paper_cp_results_rodent5, file = here::here('vignettes', 'christensenetal-comparison-files', 'model-cache', 'paper_paper_5.Rds'))
+rm(paper_paper_cp_results_rodent5)
 
-cp_results_rodent6 = changepoint_model(ldamodel, x, 6, weights = rep(1,length(year_continuous)))
-save(cp_results_rodent6, file = here::here('vignettes', 'christensenetal-comparison-files', 'model-cache', 'paper_paper6.Rds'))
-rm(cp_results_rodent6)
+paper_paper_cp_results_rodent6 = changepoint_model(ldamodel, x, 6, weights = rep(1,length(year_continuous)))
+save(paper_paper_cp_results_rodent6, file = here::here('vignettes', 'christensenetal-comparison-files', 'model-cache', 'paper_paper_6.Rds'))
+rm(paper_paper_cp_results_rodent6)
 ```
 
-```{r select paper LDA and paper cpt}
+```{r select paper LDA and paper cpt, eval = F}
 #### Changepoint model selection ####
@@
 load(here::here('vignettes', 'christensenetal-comparison-files', 'model-cache', 'paper_paper_6.Rds'))
+```
 
+```{r select paper-paper, eval = T}
 ntopics = ldamodel@k
@@
 # mean deviance ( -2 * log likelihood) + 2*(#parameters)
-mean(cp_results_rodent$saved_lls * -2) + 2*(3*(ntopics-1)*(1+1)+(1))
-mean(cp_results_rodent2$saved_lls * -2)+ 2*(3*(ntopics-1)*(2+1)+(2))
-mean(cp_results_rodent3$saved_lls * -2)+ 2*(3*(ntopics-1)*(3+1)+(3))
-mean(cp_results_rodent4$saved_lls * -2)+ 2*(3*(ntopics-1)*(4+1)+(4))
-mean(cp_results_rodent5$saved_lls * -2)+ 2*(3*(ntopics-1)*(5+1)+(5))
-mean(cp_results_rodent6$saved_lls * -2)+ 2*(3*(ntopics-1)*(6+1)+(6))
+mean(paper_paper_cp_results_rodent$saved_lls * -2) + 2*(3*(ntopics-1)*(1+1)+(1))
+mean(paper_paper_cp_results_rodent2$saved_lls * -2)+ 2*(3*(ntopics-1)*(2+1)+(2))
+mean(paper_paper_cp_results_rodent3$saved_lls * -2)+ 2*(3*(ntopics-1)*(3+1)+(3))
+mean(paper_paper_cp_results_rodent4$saved_lls * -2)+ 2*(3*(ntopics-1)*(4+1)+(4))
+mean(paper_paper_cp_results_rodent5$saved_lls * -2)+ 2*(3*(ntopics-1)*(5+1)+(5))
+mean(paper_paper_cp_results_rodent6$saved_lls * -2)+ 2*(3*(ntopics-1)*(6+1)+(6))
 
@@
 
-df_4 = as.data.frame(t(cp_results_rodent4$saved[,1,])) %>%  reshape::melt()
+paper_paper_cpt_selected = paper_paper_cp_results_rodent4
+
+df_4 = as.data.frame(t(paper_paper_cpt_selected$saved[,1,])) %>%  reshape::melt()
 df_4$value = year_continuous[df_4$value]
@@
 
-```{r clean up paper LDA and paper changepoint, include = F}
-rm(list = c('cp_results_rodent', 'cp_results_rodent2', 'cp_results_rodent3', 'cp_results_rodent4', 'cp_results_rodent5', 'cp_results_rodent6', 'ntopics', 'ldamodel', 'df_4'))
-```
-
 ### Running LDATS changepoint models
@@
 
-```{r load LDATS LDA for LDATS cpt}
-
-#### Load LDA ####
-load(here::here('vignettes', 'christensenetal-comparison-files', 'model-cache', 'ldats_lda_selected.Rds'))
-
-```
 #### Run the TS models
@@
                              nchangepoints = 1:6,
-                             weights = NULL)
+                             weights = NULL,
+                             control = TS_controls_list(timename = 'newmoon'))
 
@@
 rm(ldats_lda_selected)
-```
-
-```{r load LDATS LDA and LDATS changepoint, include = F}
-load(here::here('vignettes', 'christensenetal-comparison-files', 'model-cache', 'ldats_ldats.Rds'))
 ```
@@
 
-```{r cleanup LDATS LDA and LDATS changepoint, include = F}
-
-save(ldats_ldats_cpt_selected, file = here::here('vignettes', 'christensenetal-comparison-files', 'model-cache', 'ldats_ldats_selected.Rds'))
-
-rm(ldats_ldats_cpt_selected)
-rm(ldats_ldats_cpt)
-```
-
 ### Paper LDA and LDATS changepoint
 
-```{r load paper LDA for LDATS cpt}
+```{r load paper LDA for LDATS cpt, eval = F}
 
@@
                              nchangepoints = 1:6,
-                             weights = NULL)
+                             weights = NULL,
+                              control = TS_controls_list(timename = 'newmoon'))
 
@@
 rm(ldamodel)
-```
-
-```{r load paper LDA and LDATS changepoint, include = F}
-load(here::here('vignettes', 'christensenetal-comparison-files', 'model-cache', 'paper_ldats.Rds'))
 ```
@@
 
-```{r cleanup paper LDA and LDATS changepoint, include = F}
-
-save(paper_ldats_cpt_selected, file = here::here('vignettes', 'christensenetal-comparison-files', 'model-cache', 'paper_ldats_selected.Rds'))
-
-rm(paper_ldats_cpt_selected)
-rm(paper_ldats_cpt)
-rm(list = c('newdate', 'halfinterval'))
-```
-
 All of the models find four changepoints.
@@
 ```{r load cpts to plot, include = F}
-load(here::here('vignettes', 'christensenetal-comparison-files', 'model-cache', 'paper_ldats_selected.Rds'))
-load(here::here('vignettes', 'christensenetal-comparison-files', 'model-cache', 'ldats_ldats_selected.Rds'))
-load(here::here('vignettes', 'christensenetal-comparison-files', 'model-cache', 'paper_paper_4.Rds'))
-paper_paper_cpt_selected <- cp_results_rodent4
-load(here::here('vignettes', 'christensenetal-comparison-files', 'model-cache', 'ldats_paper_4.Rds'))
-ldats_paper_cpt_selected <- cp_results_rodent4
 
@@
 
-```{r plot paper LDA and LDATS cpts, fig.width=6, fig.height=6}
+```{r plot paper LDA and LDATS cpts, fig.width=7, fig.height=6}
 
@@
 #### LDATS LDA and LDATS changepoint
-```{r plot ldats LDA and LDATS cpt, fig.width=6, fig.height=6}
+```{r plot ldats LDA and LDATS cpt, fig.width=7, fig.height=6}
 plot(ldats_ldats_cpt_selected)
@@
 #### Paper LDA and paper changepoint
-```{r plot paper LDA and paper cpt, fig.width=6, fig.height=6}
+```{r plot paper LDA and paper cpt, fig.width=7, fig.height=6}
 
@@
 #### LDATS LDA and paper changepoint
-```{r plot LDATS lda and paper cpt, fig.width=6, fig.height=6}
+```{r plot LDATS lda and paper cpt, fig.width=7, fig.height=6}
 ntopics = 6
@@
 However, all of the models agree to within 6 months in most cases, and a year for the broader early 1990s changepoint.
-
-```{r cleanup cpt_dates, include = F}
-rm(list= 'ldats_ldats_cpt_dates', 'ldats_paper_cpt_dates',  'paper_ldats_cpt_dates', 'paper_paper_cpt_dates', 'cpt_dates')
-```

### 2019-02-11 vignettes/paper-comparison.R

@@ -1,523 +1,0 @@
-## ----setup, include=FALSE------------------------------------------------
-knitr::opts_chunk$set(
-  collapse = TRUE,
-  comment = "#>"
-)
-
-## ---- eval=FALSE---------------------------------------------------------
-## install.packages("devtools")
-## devtools::install_github("weecology/LDATS")
-
-## ---- eval = T-----------------------------------------------------------
-library(LDATS)
-
-
-## ----download Christensen------------------------------------------------
-
-paper_filepath <- here::here('vignettes', 'christensenetal-comparison-files', 'christensenetal-2018-files')
-test_filepath <- here::here('vignettes', 'christensenetal-ristensenetal-comparison-files', 'christensenetal-2018-files', 'rodent_LDA_analysis.r')
-
-if(!file.exists(test_filepath)) {
-
-files_to_download <- c('rodent_LDA_analysis.r', 'rodent_data_for_LDA.r', 'AIC_model_selection.R', 'changepointmodel.r', 'LDA-distance.R', 'Rodent_table_dat.csv', 'LDA_figure_scripts.R')
-
-for(i in 1:length(files_to_download)) {
-  download.file(url = paste0("https://raw.githubusercontent.com/emchristensen/Extreme-events-LDA/master/", files_to_download[i]),
-                destfile = paste0(paper_filepath, '/', files_to_download[i]))
-}
-
-rm(files_to_download)
-rm(i)
-
-}
-
-rm(test_filepath)
-rm(paper_filepath)
-
-## ----LDATS data----------------------------------------------------------
-
-data(rodents)
-
-head(rodents[[1]])
-
-
-## ----Paper data----------------------------------------------------------
-source(here::here('vignettes', 'christensenetal-comparison-files', 'christensenetal-2018-files', 'rodent_data_for_LDA.r'))
-
-dat = create_rodent_table(period_first = 1,
-                          period_last = 436,
-                          selected_plots = c(2,4,8,11,12,14,17,22),
-                          selected_species = c('BA','DM','DO','DS','NA','OL','OT','PB','PE','PF','PH','PI','PL','PM','PP','RF','RM','RO','SF','SH','SO'))
-
-# dates to go with count data
[…]

### 2019-02-12 doc/paper-comparison.Rmd
Make a trivial change to trigger a Travis build

@@
 
-Side-by-side comparison of LDATS results with analysis from Christensen et al 2018.
+This document prodives a side-by-side comparison of LDATS (v0.1.0) results with analysis from [Christensen et al. 2018](https://esajournals.onlinelibrary.wiley.com/doi/full/10.1002/ecy.2373).
 
@@
 the amount of information gleaned about the underlying community composition).|None; use `weights = NULL` to set all weights equal to 1 for LDATS| Weight sampling periods according to abundance|
-|Overall LDA + changepoint results| All combinations of LDA + changepoint model find 4 changepoints at approximately the same time steps|Choice of LDA model has more of an effect than choice of changepoint model|`LDATS` reflects best practices, but the paper methods produced qualitatively similar results.|
+|Overall LDA + changepoint results| All combinations of LDA + changepoint model find 4 changepoints at approximately the same time steps|Choice of LDA model has more of an effect than choice of changepoint model|`LDATS` reflects best practices, but the paper methods will produce qualitatively similar results.|
 

### 2019-02-14 doc/rodents-example.Rmd
adding the library call to the example vignette

@@
 
-To obtain the most recent version of **LDATS**, install the most recent 
+To obtain the most recent version of **LDATS**, install and load the most recent 
 version from GitHub:
@@
 devtools::install_github("weecology/LDATS")
+library(LDATS)
 ```

### 2019-07-04 doc/LDATS_codebase.Rmd
elevating timename to a main argument
pulling it out of the TS controls list

---
title: "Latent Dirichlet Allocation Time Series (LDATS)"
author: "Juniper L. Simonis"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{LDATScodebase}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r, include=FALSE}
library(LDATS)
vers <- packageVersion("LDATS")
today <- Sys.Date()
```

## Overview

This vignette outlines the code base for the **LDATS** package. It was
constructed using **LDATS** version `r vers` on `r today`.

## Installation

To obtain the most recent version of **LDATS**, install the most recent 
version from GitHub:

```{r, eval=FALSE}
install.packages("devtools")
devtools::install_github("weecology/LDATS")
```

## Analytical Pipeline

The full LDATS analysis can be executed using the `LDA_TS` function,
which is the top-level of the coding pipeline's hierarchy of
functions:

* `LDA_TS()`
  * `check_LDA_TS_inputs()`
    * `check_timename()`
    * `check_formulas()`
    * `check_nchangepoints()`
    * `check_weights()`
    * `check_control()`
    * `check_document_term_table()`
    * `check_document_covariate_table()`	
    * `check_topics()`
    * `check_seeds()`
  * `LDA_set()`
    * `check_LDA_TS_inputs()`
      * `check_control()`
      * `check_document_term_table()`
      * `check_topics()`
      * `check_seeds()`
    * loop over models
      * `prep_LDA_control()`
      * `LDA()`
  * `select_LDA()`
    * applies the `measurer` and `selector` functions supplied via `LDA_controls_list()`
  * `TS_on_LDA()`
    * `check_LDA_TS_inputs()`
      * `check_LDA_models()`
      * `check_formulas()`
      * `check_nchangepoints()`
      * `check_timename()`
      * `check_control()`
      * `check_weights()`
      * `check_document_covariate_table()`	
    * `expand_TS()`
    * loop over models
      * `prep_TS_data()`
      * `TS()`
        * `check_TS_inputs()`
          * `check_formulas()`
          * `check_nchangepoints()`
          * `check_control()`
          * `check_weights()`
        * `est_changepoints()`
          * `prep_saves()`
          * `prep_ptMCMC_inputs()`
          * `prep_cpts()`
            * loop over chains
              * `multinom_TS()`
          * `prep_ids()`
          * `prep_pbar()`
          * loop over iterations
            * `update_pbar()`
            * `step_chains()`
              * `propose_step()`
                * `proposed_step_mods()`
                  * loop over chains
                    * `multinom_TS()`
              * `eval_step()`
              * `take_step()`
            * `swap_chains()`
                * loop over chain neighbors
            * `update_saves()`
            * `update_cpts()`
            * `update_ids()`
        * `est_regressors()`
          * loop over unique realizations of change point locations
            * `multinom_TS()`
            * loop over chunks
              * `mirror_vcov()`
              * `rmvnorm()`
        * `summarize_TS()`
          * `diagnose_ptMCMC()`
            * `count_trips()`
          * `summarize_rhos()`
          * `measure_rho_vcov()`
          * `summarize_etas()`
          * `measure_eta_vcov()`
    * `select_TS()`
      * applies the `measurer` and `selector` functions supplied via `TS_controls_list()`
    * `package_LDA_TS()`

Each component model's function (`LDA_set()` and `TS()`) can be run
independently, as well.

## Controls Lists

To minimize the length of argument lists and facilitate simple default usage
throughout the pipeline, we implement an options/controls list approach, where
each of the main functions (`LDA_TS`, `LDA_set`, and `TS`) has a controls 
list (although `LDA_set` is referenced as `LDA` in the naming scheme)

* `LDA_TS_controls_list()`
* `LDA_controls_list()`
* `TS_controls_list()`

### 2019-07-04 doc/rodents-example.Rmd
elevating timename to a main argument
pulling it out of the TS controls list

@@
 
-We use `LDA_set()` to run replicate LDA models (each with its own seed) with varying numbers of topics (`2:6`) and `select_LDA()` to select the best model.
+We use `LDA_set()` to run replicate LDA models (each with its own seed) with varying numbers of topics (`2:5`) and `select_LDA()` to select the best model.
 
@@
 lda_model_set <- LDA_set(document_term_table = rodents$document_term_table,
-                         topics = c(2:6),
-                         nseeds = 100,
+                         topics = c(2:5),
+                         nseeds = 10,
                          control = LDA_controls_list(quiet = TRUE))
@@
 
-If we do not pass any controls, by default, `quiet = FALSE` (here run with only `2:3` topics and  `10` seeds, to keep output short):
+If we do not pass any controls, by default, `quiet = FALSE` (here run with only `2:3` topics and `2` seeds, to keep output short):
 
@@
                          topics = c(2:3),
-                         nseeds = 10)
+                         nseeds = 2)
 ```
@@
                                 formulas = ~ sin_year + cos_year,
-                                nchangepoints = c(0:6), 
+                                nchangepoints = c(0:1), 
                                 weights = document_weights(rodents$document_term_table),
-                                control = TS_controls_list(nit = 1000, timename = "newmoon"))
+                                timename = "newmoon",
+                                control = TS_controls_list(nit = 1000))
 
@@
 Also, it is important to note that by default the TS functions take the name of the time-step column from the `document_covariate_table` to be `"time"`.
-To pass a different column name, use the `timename` argument in `TS_controls_list()`.
+To pass a different column name, use the `timename` argument in `TS_on_LDA()`.
 
@@
                          document_covariate_table = rodents$document_covariate_table,
-                         nseeds = 100, 
-                         topics = 2:6,
+                         nseeds = 10, 
+                         topics = 2:5,
                          formulas = ~ sin_year + cos_year,
-                         nchangepoints= 0:6,
+                         nchangepoints= 0:1,
                          weights = document_weights(rodents$document_term_table),
+                         timename = "newmoon",
                          control = LDA_TS_controls_list(
-                           TS_control = TS_controls_list(nit = 1000, timename = "newmoon")))
+                           TS_control = TS_controls_list(nit = 1000)))
 ```
@@
 ```{r plot LDA_TS results, fig.height = 16, fig.width = 7, echo = F}
-plot(lda_ts_results, control = LDA_TS_controls_list(
-                           TS_control = TS_controls_list(timename = "newmoon")))
+plot(lda_ts_results)
 ```

### 2019-07-04 R/LDA_TS.R

@@ -54,6 +54,9 @@
 #'   For most models, a scaling of the weights (so that the average is 1) is
 #'   most appropriate, and this is accomplished using \code{document_weights}.
 #' 
+#' @param timename \code{character} element indicating the time variable
+#'   used in the time series. Defaults to \code{"time"}.
+#'
 #' @param control Class \code{LDA_TS_controls} list (as generated by
 #'   \code{\link{LDA_TS_controls_list}}) that contains 
 #'   \code{LDA_controls} (\code{\link{LDA_controls_list}}), 
@@ -93,12 +96,9 @@
 #'   data(rodents)
 #'   document_term_table <- rodents$document_term_table
 #'   document_covariate_table <- rodents$document_covariate_table
-#'   TS_controls <- TS_controls_list(timename = "newmoon")
-#'   controls <- LDA_TS_controls_list(TS_control = TS_controls)
 #'   mod <- LDA_TS(document_term_table, document_covariate_table,
 #'                 topics = 2, nseeds = 1, formulas = ~ 1, nchangepoints = 1,
-#'                 weights = document_weights(document_term_table), 
-#'                 control = controls)
+#'                 weights = document_weights(document_term_table), "newmoon")
 #' }
 #'
 #' @export
@@ -106,16 +106,16 @@
 LDA_TS <- function(document_term_table, document_covariate_table,
                    topics = 2, nseeds = 1, formulas = ~ 1, nchangepoints = 0,
                    weights = document_weights(document_term_table), 
-                   control = LDA_TS_controls_list()){
+                   timename = "time", control = LDA_TS_controls_list()){
   check_LDA_TS_inputs(document_term_table, document_covariate_table,
                       topics, nseeds, formulas, nchangepoints, weights, 
-                      control)
+                      timename, control)
   qprint("Latent Dirichlet Allocation", "----", control$quiet)
   LDAs <- LDA_set(document_term_table, topics, nseeds, control$LDA_control)
   sel_LDA <- select_LDA(LDAs, control$LDA_control)
   qprint("Time Series Models", "----", control$quiet)
   TSs <- TS_on_LDA(sel_LDA, document_covariate_table, formulas, nchangepoints,
-                   weights, control$TS_control)
+                   weights, timename, control$TS_control)
   sel_TSs <- select_TS(TSs, control$TS_control)
   package_LDA_TS(LDAs, sel_LDA, TSs, sel_TSs)
 }
[…]

### 2019-07-04 R/TS.R

@@ -12,7 +12,7 @@
 #'   estimate regressors.
 #'
 #' @param data \code{data.frame} including [1] the time variable (indicated 
-#'   in \code{control$timename}), [2] the predictor variables (required by
+#'   in \code{timename}), [2] the predictor variables (required by
 #'   \code{formula}) and [3], the multinomial response variable (indicated in
 #'   \code{formula}) as verified by \code{\link{check_timename}} and 
 #'   \code{\link{check_formula}}. Note that the response variables should be
@@ -42,6 +42,9 @@
 #'   proportions, which does not account for size differences among documents.
 #'   For most models, a scaling of the weights (so that the average is 1) is
 #'   most appropriate, and this is accomplished using \code{document_weights}.
+#'
+#' @param timename \code{character} element indicating the time variable
+#'   used in the time series. Defaults to \code{"time"}.
 #'
 #' @param control Class \code{TS_controls} list, holding control parameters
 #'   for the Time Series model including the parallel tempering Markov Chain 
@@ -128,20 +131,21 @@
 #'   data <- document_covariate_table
 #'   data$gamma <- LDA_models@gamma
 #'   weights <- document_weights(document_term_table)
-#'   controls <- TS_controls_list(timename = "newmoon")
-#'   TSmod <- TS(data, gamma ~ 1, nchangepoints = 1, weights, controls)
+#'   TSmod <- TS(data, gamma ~ 1, nchangepoints = 1, weights, "newmoon")
 #' }
 #'
 #' @export
 #'
-TS <- function(data, formula, nchangepoints, weights, 
+TS <- function(data, formula, nchangepoints, weights, timename = "time",
                control = TS_controls_list()){
-  check_TS_inputs(data, formula, nchangepoints, weights, control)
+  check_TS_inputs(data, formula, nchangepoints, weights, timename, control)
   set.seed(control$seed)
-  data <- data[order(data[,control$timename]), ]
-  rho_dist <- est_changepoints(data, formula, nchangepoints, weights, control)
-  eta_dist <- est_regressors(rho_dist, data, formula, weights, control)
-  summarize_TS(data, formula, weights, control, rho_dist, eta_dist)
+  data <- data[order(data[,timename]), ]
+  rho_dist <- est_changepoints(data, formula, nchangepoints, weights, 
+                               timename, control)
+  eta_dist <- est_regressors(rho_dist, data, formula, weights, 
+                             timename, control)
[…]

### 2019-07-04 R/TS_plots.R

@@ -356,13 +356,14 @@ pred_gamma_TS_plot <- function(x, selection = "median", cols, xlab = NULL){
   } else{
     spec_rhos <- NULL
   }
-  seg_mods <- multinom_TS(x$data, x$formula, spec_rhos, x$weights, x$control)
+  seg_mods <- multinom_TS(x$data, x$formula, spec_rhos, x$weights, 
+                          x$timename, x$control)
   nsegs <- length(seg_mods[[1]])
-  t1 <- min(x$data[, x$control$timename])
-  t2 <- max(x$data[, x$control$timename])
+  t1 <- min(x$data[, x$timename])
+  t2 <- max(x$data[, x$timename])
 
   if (is.null(xlab)){
-    xlab <- x$control$timename
+    xlab <- x$timename
   }
   par(mar = c(4.5, 5, 1, 1))
   plot(1, 1, type = "n", bty = "L", xlab = xlab, ylab = "", xaxt = "n", 
@@ -437,7 +438,7 @@ rho_hist <- function(x, cols, bin_width = 1, xlab = NULL, border = NA){
   rhos <- x$rhos
   nrhos <- ncol(rhos)
   niter <- nrow(rhos)
-  timeobs <- x$data[, x$control$timename]
+  timeobs <- x$data[, x$timename]
   timerange <- range(timeobs)
   timevals <- seq(timerange[1], timerange[2], 1)
   ntimes <- length(timevals) 

### 2019-07-04 doc/paper-comparison.R

@@ -345,7 +345,7 @@ paper_paper_cpt_dates
 #                               formulas = ~ sin_year + cos_year,
 #                               nchangepoints = 1:6,
 #                               weights = NULL,
-#                               control = TS_controls_list(timename = 'newmoon'))
+#                               timename = 'newmoon')
 #  
 #  
 #  save(ldats_ldats_cpt, file = here::here('vignettes', 'christensenetal-comparison-files', 'model-cache', 'ldats_ldats.Rds'))
@@ -383,7 +383,7 @@ ldats_ldats_cpt_dates <- ldats_ldats_cpt_dates$censusdate
 #                               formulas = ~ sin_year + cos_year,
 #                               nchangepoints = 1:6,
 #                               weights = NULL,
-#                                control = TS_controls_list(timename = 'newmoon'))
+#                               timename = 'newmoon')
 #  
 #  
 #  save(paper_ldats_cpt, file = here::here('vignettes', 'christensenetal-comparison-files', 'model-cache', 'paper_ldats.Rds'))

### 2019-07-04 doc/rodents-example.R

@@ -21,15 +21,15 @@ head(rodents$document_covariate_table, 10)
 
 ## ----lda_set, eval =F----------------------------------------------------
 #  lda_model_set <- LDA_set(document_term_table = rodents$document_term_table,
-#                           topics = c(2:6),
-#                           nseeds = 100,
-#                          control = LDA_controls_list(quiet = TRUE))
+#                           topics = c(2:5),
+#                           nseeds = 10,
+#                           control = LDA_controls_list(quiet = TRUE))
 #  
 
 ## ----lda set not quiet---------------------------------------------------
 lda_model_set2 <- LDA_set(document_term_table = rodents$document_term_table,
                          topics = c(2:3),
-                         nseeds = 10)
+                         nseeds = 2)
 
 ## ----load lda model set, include = F-------------------------------------
 load(here::here('vignettes', 'rodents-example-files', 'lda_model_set.Rds'))
@@ -56,9 +56,10 @@ plot(selected_lda_model[[1]])
 #  changepoint_models <- TS_on_LDA(LDA_models = selected_lda_model,
 #                                  document_covariate_table = rodents$document_covariate_table,
 #                                  formulas = ~ sin_year + cos_year,
-#                                  nchangepoints = c(0:6),
+#                                  nchangepoints = c(0:1),
 #                                  weights = document_weights(rodents$document_term_table),
-#                                  control = TS_controls_list(nit = 1000, timename = "newmoon"))
+#                                  timename = "newmoon",
+#                                  control = TS_controls_list(nit = 1000))
 #  
 
 ## ----reload ts, include = F----------------------------------------------
@@ -85,13 +86,14 @@ plot(selected_changepoint_model)
 ## ----lda_ts, eval = F----------------------------------------------------
 #  lda_ts_results <- LDA_TS(document_term_table = rodents$document_term_table,
 #                           document_covariate_table = rodents$document_covariate_table,
-#                           nseeds = 100,
-#                           topics = 2:6,
+#                           nseeds = 10,
+#                           topics = 2:5,
 #                           formulas = ~ sin_year + cos_year,
-#                           nchangepoints= 0:6,
+#                           nchangepoints= 0:1,
 #                           weights = document_weights(rodents$document_term_table),
+#                           timename = "newmoon",
 #                           control = LDA_TS_controls_list(
[…]

### 2019-07-04 tests/testthat/test-04-LDA_TS.R

@@ -9,9 +9,9 @@ test_that("LDA_TS on 0 changepoints", {
   mod0 <- LDA_TS(document_term_table, document_covariate_table,
                  topics = 2, nseeds = 1, formulas = ~ 1, nchangepoints = 0,
                  weights = document_weights(document_term_table), 
+                 timename = "newmoon",
                  control = LDA_TS_controls_list(
-                           TS_control = TS_controls_list(nit = 100,
-                                                     timename = "newmoon")))
+                           TS_control = TS_controls_list(nit = 100)))
   expect_is(mod0, "LDA_TS")
   expect_equal(length(names(mod0)), 4)
   expect_is(mod0[[4]], "TS_fit")
@@ -21,9 +21,9 @@ test_that("LDA_TS on 1 changepoints", {
   mod1 <- LDA_TS(document_term_table, document_covariate_table,
                  topics = 2, nseeds = 1, formulas = ~ 1, nchangepoints = 1,
                  weights = document_weights(document_term_table), 
+                 timename = "newmoon",
                  control = LDA_TS_controls_list(
-                           TS_control = TS_controls_list(nit = 100,
-                                                     timename = "newmoon")))
+                           TS_control = TS_controls_list(nit = 100)))
   expect_is(mod1, "LDA_TS")
   expect_equal(length(names(mod1)), 4)
   expect_is(mod1[[4]], "TS_fit")
@@ -33,9 +33,9 @@ test_that("check print on LDA_TS", {
   mod1 <- LDA_TS(document_term_table, document_covariate_table,
                  topics = 2, nseeds = 1, formulas = ~ 1, nchangepoints = 1,
                  weights = document_weights(document_term_table), 
+                 timename = "newmoon",
                  control = LDA_TS_controls_list(
-                           TS_control = TS_controls_list(nit = 100,
-                                                     timename = "newmoon")))
+                           TS_control = TS_controls_list(nit = 100)))
   expect_output(print(mod1))
 })
 
@@ -53,12 +53,12 @@ test_that("Check package_LDA_TS", {
   formulas <- ~ 1
   nchangepoints <- 1
   weights <- document_weights(document_term_table)
-  control <- LDA_TS_controls_list(TS_control = TS_controls_list(nit = 100,
-                                                     timename = "newmoon"))
+  timename <-  "newmoon"
[…]

### 2019-07-04 tests/testthat/test-06-multinom_TS.R

@@ -6,17 +6,18 @@ lda <- LDA_set(lda_data, c(4), nseeds = 1, LDA_controls_list(quiet = TRUE))
 dct <- rodents$document_covariate_table
 mts_data <- data.frame(dct)
 mts_data$gamma <- lda[[1]]@gamma
-control <- TS_controls_list(timename = "newmoon")
+timename <- "newmoon"
 
 test_that("check packaging of chunk fits", {
   TS_chunk_memo <- memoise_fun(multinom_TS_chunk, control$memoise)
   chunks <- prep_chunks(data = mts_data, changepoints = c(20,50), 
-                        timename = control$timename)
+                        timename = timename)
   nchunks <- nrow(chunks)
   fits <- vector("list", length = nchunks)
   for (i in 1:nchunks){
     fits[[i]] <- TS_chunk_memo(data = mts_data, formula = gamma ~ 1, 
                                chunk = chunks[i, ], weights = NULL,
+                               timename = timename,
                                control = control)
   }
   packaged <- package_chunk_fits(chunks, fits)
@@ -27,7 +28,7 @@ test_that("check packaging of chunk fits", {
 test_that("check logLik for multinom_TS_fit", {
   mts <- multinom_TS(data = mts_data, formula = gamma~1, 
            changepoints = c(20,50), weights = NULL, 
-           control = TS_controls_list(timename = "newmoon"))
+           timename = "newmoon")
   expect_is(logLik(mts), "logLik")
   expect_equal(round(as.numeric(logLik(mts))), -517)
 })
@@ -35,7 +36,7 @@ test_that("check logLik for multinom_TS_fit", {
 test_that("check good output from multinom_TS", {
   mts <- multinom_TS(data = mts_data, formula = gamma~1, 
            changepoints = c(20,50), weights = NULL, 
-           control = TS_controls_list(timename = "newmoon"))
+           timename = "newmoon")
   expect_is(mts, "list")
   expect_is(mts, "multinom_TS_fit")
   expect_equal(length(mts), 3)
@@ -54,7 +55,7 @@ test_that("check check_changepoints", {
 test_that("check failed output from multinom_TS", {
   mts <- multinom_TS(data = mts_data, formula = gamma~1, 
            changepoints = c(50,40), weights = NULL, 
-           control = TS_controls_list(timename = "newmoon"))
+           timename = "newmoon")
[…]

### 2019-07-04 tests/testthat/test-07-ptMCMC.R

@@ -15,11 +15,11 @@ nseeds <- 1
 formulas <- ~ 1
 nchangepoints <- 1
 weights <- document_weights(document_term_table)
-TS_controls <- TS_controls_list(timename = "newmoon")
-control <- LDA_TS_controls_list(TS_control = TS_controls)
+timename <- "newmoon"
+control <- LDA_TS_controls_list()
 LDAs <- LDA_set(document_term_table, topics, nseeds, control$LDA_control)
 LDA_models <- select_LDA(LDAs, control$LDA_control)
-control <- TS_controls_list(nit = 1e3, seed = 1, timename = "newmoon")
+control <- TS_controls_list(nit = 1e3, seed = 1)
 mods <- expand_TS(LDA_models, formulas, nchangepoints)
 formula <- mods$formula[[1]]
 nchangepoints <- mods$nchangepoints[1]
@@ -27,14 +27,17 @@ data <- prep_TS_data(document_covariate_table, LDA_models, mods, 1)
 
 set.seed(1)
 rho_dist0 <- est_changepoints(data, formula, nchangepoints = 0, weights, 
-                              control)
-rho_dist <- est_changepoints(data, formula, nchangepoints, weights, control)
-eta_dist <- est_regressors(rho_dist, data, formula, weights, control)
+                              timename, control)
+rho_dist <- est_changepoints(data, formula, nchangepoints, weights, 
+                             timename, control)
+eta_dist <- est_regressors(rho_dist, data, formula, weights, timename, 
+                            control)
 
 
 saves <- prep_saves(nchangepoints, control)
-inputs <- prep_ptMCMC_inputs(data, formula, nchangepoints, weights, control)
-cpts <- prep_cpts(data, formula, nchangepoints, weights, control)
+inputs <- prep_ptMCMC_inputs(data, formula, nchangepoints, weights, 
+                             timename, control)
+cpts <- prep_cpts(data, formula, nchangepoints, weights, timename, control)
 ids <- prep_ids(control)
 
 test_that("check prep_proposal_dist", {
@@ -51,21 +54,25 @@ test_that("check prep_proposal_dist", {
 
 
 test_that("check prep_ptMCMC_inputs", {
-  inpts <- prep_ptMCMC_inputs(data, formula, nchangepoints, weights, control)
+  inpts <- prep_ptMCMC_inputs(data, formula, nchangepoints, weights, 
+                              timename, control)
   expect_is(inpts, "ptMCMC_inputs")
[…]

### 2019-07-04 tests/testthat/test-09-TS_on_LDA.R

@@ -9,8 +9,8 @@ nseeds <- 1
 formulas <- ~ 1
 nchangepoints <- 0
 weights <- document_weights(document_term_table)
-TS_controls <- TS_controls_list(timename = "newmoon")
-control <- LDA_TS_controls_list(TS_control = TS_controls)
+timename <- "newmoon"
+control <- LDA_TS_controls_list()
 LDAs <- LDA_set(document_term_table, topics, nseeds, control$LDA_control)
 LDA_models <- select_LDA(LDAs, control$LDA_control)
 control <- control$TS_control
@@ -39,12 +39,12 @@ test_that("check prep_TS_data", {
 
 test_that("check select_TS", {
   mods <- TS_on_LDA(LDA_models, document_covariate_table, formulas, 
-                  nchangepoints = 0:1, weights, 
-                  control = TS_controls_list(nit = 1e3, timename = "newmoon"))
+                  nchangepoints = 0:1, weights, timename,
+                  control = TS_controls_list(nit = 1e3))
   sel_mod <- select_TS(mods, control)
 
   expect_equal(length(mods), 2)
-  expect_equal(length(sel_mod), 16)
+  expect_equal(length(sel_mod), 17)
   expect_is(mods, "TS_on_LDA")
   expect_is(sel_mod, "TS_fit")
   expect_error(select_TS(mods, LDA_controls_list()))
@@ -64,9 +64,8 @@ test_that("check package_TS_on_LDA", {
     formula_i <- mods$formula[[i]]
     nchangepoints_i <- mods$nchangepoints[i]
     data_i <- prep_TS_data(document_covariate_table, LDA_models, mods, i)
-    TSmods[[i]] <- TS(data_i, formula_i, nchangepoints_i, weights, 
-                      control = TS_controls_list(nit = 1e3, 
-                                                 timename = "newmoon"))
+    TSmods[[i]] <- TS(data_i, formula_i, nchangepoints_i, weights, timename,
+                      control = TS_controls_list(nit = 1e3))
   }
   expect_is(package_TS_on_LDA(TSmods, LDA_models, mods), "TS_on_LDA")
   expect_is(package_TS_on_LDA(TSmods, LDA_models[[1]], mods), "TS_on_LDA")
@@ -78,9 +77,8 @@ test_that("check package_TS_on_LDA", {
 
 test_that("check TS_on_LDA", {
   mods <- TS_on_LDA(LDA_models, document_covariate_table, formulas, 
-                  nchangepoints = 0:1, weights, 
[…]

### 2019-07-04 tests/testthat/test-10-TS_plots.R

@@ -13,12 +13,13 @@ weights <- document_weights(document_term_table)
 control <- LDA_TS_controls_list()
 LDAs <- LDA_set(document_term_table, topics, nseeds, control$LDA_control)
 LDA_models <- select_LDA(LDAs, control$LDA_control)
-control <- TS_controls_list(nit = 1e3, seed = 1, timename = "newmoon")
+control <- TS_controls_list(nit = 1e3, seed = 1)
+timename <- "newmoon"
 mods <- expand_TS(LDA_models, formulas, nchangepoints)
 formula <- mods$formula[[1]]
 nchangepoints <- mods$nchangepoints[1]
 data <- prep_TS_data(document_covariate_table, LDA_models, mods, 1)
-TSmod <- TS(data, formula, nchangepoints, weights, control)
+TSmod <- TS(data, formula, nchangepoints, weights, timename, control)
 
 
 test_that("check rho_hist color generator", {
@@ -147,7 +148,7 @@ test_that("check plotting of TS_fit", {
 })
 
 test_that("check TS_diagnostics_plot", {
-  TSmod0 <- TS(data, formula, nchangepoints = 0, weights, control)
+  TSmod0 <- TS(data, formula, nchangepoints = 0, weights, timename, control)
   if (tenv == "cran"){
     expect_silent(TS_diagnostics_plot(TSmod0, interactive = FALSE))
   } else{

### 2019-07-04 tests/testthat/test-11-utilities.R

@@ -47,15 +47,18 @@ test_that("check mirror_vcov", {
   nchangepoints <- 2
   weights <- document_weights(document_term_table)
   control <- LDA_TS_controls_list()
+  timename <- "newmoon"
   LDAs <- LDA_set(document_term_table, topics, nseeds, control$LDA_control)
   LDA_models <- select_LDA(LDAs, control$LDA_control)
-  control <- TS_controls_list(nit = 1e2, seed = 1, timename = "newmoon")
+  control <- TS_controls_list(nit = 1e2, seed = 1)
   mods <- expand_TS(LDA_models, formulas, nchangepoints)
   formula <- mods$formula[[1]]
   nchangepoints <- mods$nchangepoints[1]
   data <- prep_TS_data(document_covariate_table, LDA_models, mods, 1)
-  rho_dist <- est_changepoints(data, formula, nchangepoints, weights, control)
-  mod <- multinom_TS(data, formula, changepoints = NULL, weights, control)
+  rho_dist <- est_changepoints(data, formula, nchangepoints, weights, 
+                               timename, control)
+  mod <- multinom_TS(data, formula, changepoints = NULL, weights, 
+                     timename, control)
   expect_equal(isSymmetric(vcov(mod[[1]][[1]])), FALSE)
   expect_equal(isSymmetric(mirror_vcov(mod[[1]][[1]])), TRUE)
 

### 2019-07-05 R/utilities.R
combining data structure in LDA_TS
LDA_TS now takes a single "data" input that can be either a document term table or a list containing at least a document term table and optionally also a covariate table. now if a covariate table isn't provided in LDA_TS, the assumption is made that the data are equispersed in time

@@ -1,3 +1,24 @@
+#' @title Replace if TRUE
+#'
+#' @description If the focal input is \code{TRUE}, replace it with 
+#'   alternative. 
+#'
+#' @param x Focal input.
+#'
+#' @param alt Alternative value.
+#'
+#' @return \code{x} if not \code{TRUE}, \code{alt} otherwise.
+#' 
+#' @export 
+#'
+iftrue <- function(x = TRUE, alt = NULL){
+  if (is.logical(x) && x){
+    x <- alt
+  }
+  x
+}
+
+
 #' @title Determine the mode of a distribution
 #'
 #' @description Find the most common entry in a vector. Ties are not allowed,

### 2019-07-05 doc/rodents-example.R

@@ -57,8 +57,8 @@ plot(selected_lda_model[[1]])
 #                                  document_covariate_table = rodents$document_covariate_table,
 #                                  formulas = ~ sin_year + cos_year,
 #                                  nchangepoints = c(0:1),
-#                                  weights = document_weights(rodents$document_term_table),
 #                                  timename = "newmoon",
+#                                  weights = document_weights(rodents$document_term_table),
 #                                  control = TS_controls_list(nit = 1000))
 #  
 
@@ -84,13 +84,11 @@ head(selected_changepoint_model$rhos)
 plot(selected_changepoint_model)
 
 ## ----lda_ts, eval = F----------------------------------------------------
-#  lda_ts_results <- LDA_TS(document_term_table = rodents$document_term_table,
-#                           document_covariate_table = rodents$document_covariate_table,
+#  lda_ts_results <- LDA_TS(data = rodents,
 #                           nseeds = 10,
 #                           topics = 2:5,
 #                           formulas = ~ sin_year + cos_year,
 #                           nchangepoints= 0:1,
-#                           weights = document_weights(rodents$document_term_table),
 #                           timename = "newmoon",
 #                           control = LDA_TS_controls_list(
 #                             TS_control = TS_controls_list(nit = 1000)))

### 2019-07-05 tests/testthat/test-04-LDA_TS.R

@@ -6,9 +6,8 @@ document_term_table <- rodents$document_term_table
 document_covariate_table <- rodents$document_covariate_table
 
 test_that("LDA_TS on 0 changepoints", {
-  mod0 <- LDA_TS(document_term_table, document_covariate_table,
+  mod0 <- LDA_TS(rodents,
                  topics = 2, nseeds = 1, formulas = ~ 1, nchangepoints = 0,
-                 weights = document_weights(document_term_table), 
                  timename = "newmoon",
                  control = LDA_TS_controls_list(
                            TS_control = TS_controls_list(nit = 100)))
@@ -18,9 +17,8 @@ test_that("LDA_TS on 0 changepoints", {
 })
 
 test_that("LDA_TS on 1 changepoints", {
-  mod1 <- LDA_TS(document_term_table, document_covariate_table,
+  mod1 <- LDA_TS(rodents,
                  topics = 2, nseeds = 1, formulas = ~ 1, nchangepoints = 1,
-                 weights = document_weights(document_term_table), 
                  timename = "newmoon",
                  control = LDA_TS_controls_list(
                            TS_control = TS_controls_list(nit = 100)))
@@ -30,9 +28,8 @@ test_that("LDA_TS on 1 changepoints", {
 })
 
 test_that("check print on LDA_TS", {
-  mod1 <- LDA_TS(document_term_table, document_covariate_table,
+  mod1 <- LDA_TS(rodents,
                  topics = 2, nseeds = 1, formulas = ~ 1, nchangepoints = 1,
-                 weights = document_weights(document_term_table), 
                  timename = "newmoon",
                  control = LDA_TS_controls_list(
                            TS_control = TS_controls_list(nit = 100)))
@@ -58,7 +55,7 @@ test_that("Check package_LDA_TS", {
   LDAs <- LDA_set(document_term_table, topics, nseeds, control$LDA_control)
   sel_LDA <- select_LDA(LDAs, control$LDA_control)
   TSs <- TS_on_LDA(sel_LDA, document_covariate_table, formulas, nchangepoints,
-                   weights, timename, control$TS_control)
+                   timename, weights, control$TS_control)
   sel_TSs <- select_TS(TSs, control$TS_control)
   
   expect_is(package_LDA_TS(LDAs, sel_LDA, TSs, sel_TSs), "LDA_TS")

### 2019-07-05 tests/testthat/test-09-TS_on_LDA.R

@@ -39,7 +39,7 @@ test_that("check prep_TS_data", {
 
 test_that("check select_TS", {
   mods <- TS_on_LDA(LDA_models, document_covariate_table, formulas, 
-                  nchangepoints = 0:1, weights, timename,
+                  nchangepoints = 0:1, timename, weights, 
                   control = TS_controls_list(nit = 1e3))
   sel_mod <- select_TS(mods, control)
 
@@ -77,7 +77,7 @@ test_that("check package_TS_on_LDA", {
 
 test_that("check TS_on_LDA", {
   mods <- TS_on_LDA(LDA_models, document_covariate_table, formulas, 
-                  nchangepoints = 0:1, weights, timename, 
+                  nchangepoints = 0:1, timename, weights, 
                   control = TS_controls_list(nit = 1e3))
   expect_is(mods, "TS_on_LDA")
   expect_equal(length(mods), 2)
@@ -86,25 +86,25 @@ test_that("check TS_on_LDA", {
 
   expect_error(TS_on_LDA())
   expect_error(TS_on_LDA("ok", document_covariate_table, formulas, 
-                  nchangepoints, weights, timename, control ))
+                  nchangepoints, timename, weights, control ))
   expect_error(TS_on_LDA(LDA_models, "ok", formulas, 
-                  nchangepoints, weights, timename, control ))
+                  nchangepoints, timename, weights, control ))
   expect_error(TS_on_LDA(LDA_models, document_covariate_table, "ok", 
-                  nchangepoints, weights, timename, control ))
+                  nchangepoints, timename, weights, control ))
   expect_error(TS_on_LDA(LDA_models, document_covariate_table, formulas, 
-                  "ok", weights, timename, control ))
+                  "ok", timename, weights, control ))
   expect_error(TS_on_LDA(LDA_models, document_covariate_table, formulas, 
-                  nchangepoints, "ok", timename, control ))
+                  nchangepoints, timename, "ok", control ))
   expect_error(TS_on_LDA(LDA_models, document_covariate_table, formulas, 
-                  nchangepoints, weights, "ok", control))
+                  nchangepoints, "ok", weights, control))
   expect_error(TS_on_LDA(LDA_models, document_covariate_table, formulas, 
-                  nchangepoints, weights, timename, "ok"))
+                  nchangepoints, timename, weights, "ok"))
 })
 
 
 test_that("check printing for TS_on_LDA", {
[…]

### 2019-07-05 tests/testthat/test-11-utilities.R

@@ -1,5 +1,12 @@
 context("Check utilities")
 
+test_that("check iftrue", {
+  expect_equal(iftrue(TRUE,1), 1)
+  expect_equal(iftrue(1,2), 1)
+})
+
+
+
 test_that("check modalvalue", {
   xx <- c(1, 2, 3, 4, 5, 4, 3, 2, 1, 2)
   expect_equal(modalvalue(xx), 2)

### 2019-07-08 doc/rodents-example.Rmd
simplification of the control lists
The control lists have been made easier to interact with. Primarily, the arguments that previously required `LDA_controls_list`, `TS_controls_list`, or `LDA_TS_controls_list` inputs now take general `list` inputs (so  `LDA_TS` does not need to have a nested set of control functions). Each control list is passed through a function (`LDA_set_control`, `TS_control`, or `LDA_TS_control`) to set any non-input values to their defaults. This also allows the removal of those controls list class definitions. ([addresses issue 130](https://github.com/weecology/LDATS/i

@@
 
-We use `LDA_controls_list())` to pass controls to the LDA function. In this case, we can set `quiet = TRUE` to make the model run quietly.
+We use the `control` argument to pass controls to the LDA function via a `list`. In this case, we can set `quiet = TRUE` to make the model run quietly.
 
@@
                          nseeds = 10,
-                         control = LDA_controls_list(quiet = TRUE))
+                         control = list(quiet = TRUE))
 
@@
 By default, we will choose models based on minimum AIC.
-To use different selection criteria, define the appropriate functions and specify them using `LDA_controls_list(measurer = [measurer function], selector = [max, min, etc])`.
+To use different selection criteria, define the appropriate functions and specify them by passing `list(measurer = [measurer function], selector = [max, min, etc])` to the `control` argument.
 
@@
                                 nchangepoints = c(0:1), 
+                                timename = "newmoon",
                                 weights = document_weights(rodents$document_term_table),
-                                timename = "newmoon",
-                                control = TS_controls_list(nit = 1000))
+                                control = list(nit = 1000))
 
@@
 
-We can adjust options for both TS functions using `TS_controls_list()`.
-For a full list see `?TS_controls_list`.
+We can adjust options (default settings can be seen using `TS_control()`) for both TS functions by passing a list to the `control` argument.
+For a full list see `?TS_control`.
 Here we illustrate adjusting the number of ptMCMC iterations - the default is 10000, but it is convenient to use fewer iterations for code development.
@@
 
-`select_TS()` will identify the best-fit changepoint model of the models from `TS_on_LDA()`. As with `select_LDA()`, we can adjust the `measurer` and `selector` using `TS_controls_list()`.
+`select_TS()` will identify the best-fit changepoint model of the models from `TS_on_LDA()`.
+As with `select_LDA()`, we can adjust the `measurer` and `selector` functions using the `control` argument list.
 
@@
 
-Finally, we can perform an entire LDATS analysis, including all of the above steps, using the `LDA_TS()` function and passing options to the LDA and TS functions using `LDA_TS_controls_list()`.
+Finally, we can perform an entire LDATS analysis, including all of the above steps, using the `LDA_TS()` function and passing options to the LDA and TS functions as a `list` to the `control` argument.
+The default is for `LDA_TS` to weight the time series model based on the document sizes, so we do not need to tell it to do so.
 
 ```{r lda_ts, eval = F}
-lda_ts_results <- LDA_TS(document_term_table = rodents$document_term_table,
-                         document_covariate_table = rodents$document_covariate_table,
+lda_ts_results <- LDA_TS(data = rodents,
                          nseeds = 10, 
@@
                          nchangepoints= 0:1,
-                         weights = document_weights(rodents$document_term_table),
                          timename = "newmoon",
-                         control = LDA_TS_controls_list(
-                           TS_control = TS_controls_list(nit = 1000)))
+                         control = list(nit = 1000))
 ```

### 2019-07-08 R/LDA_TS.R

@@ -64,11 +64,9 @@
 #'   \code{NULL} for an equal weighting among documents or a \code{numeric}
 #'   vector.
 #'
-#' @param control Class \code{LDA_TS_controls} list (as generated by
-#'   \code{\link{LDA_TS_controls_list}}) that contains 
-#'   \code{LDA_controls} (\code{\link{LDA_controls_list}}), 
-#'   \code{TS_controls} (\code{\link{TS_controls_list}}), and the single 
-#'   top-level control \code{quiet}.
+#' @param control A \code{list} of parameters to control the running and 
+#'   selecting of LDA and TS models. Values not input assume default values 
+#'   set by \code{\link{LDA_TS_control}}. 
 #'
 #' @return 
 #'   \code{LDA_TS} returns a class \code{LDA_TS} list object including all 
@@ -109,16 +107,17 @@
 #'
 LDA_TS <- function(data, topics = 2, nseeds = 1, formulas = ~ 1,
                    nchangepoints = 0, timename = "time", weights = TRUE, 
-                   control = LDA_TS_controls_list()){
+                   control = list()){
   check_LDA_TS_inputs(data, topics, nseeds, formulas, nchangepoints,  
                       timename, weights, control)
+  control <- do.call("LDA_TS_control", control)
   data <- conform_LDA_TS_data(data, control$quiet)
   dtt <- data$document_term_table
   dct <- data$document_covariate_table
   weights <- iftrue(weights, document_weights(dtt))
   qprint("Latent Dirichlet Allocation", "----", control$quiet)
-  LDAs <- LDA_set(dtt, topics, nseeds, control$LDA_control)
-  sel_LDA <- select_LDA(LDAs, control$LDA_control)
+  LDAs <- LDA_set(dtt, topics, nseeds, control$LDA_set_control)
+  sel_LDA <- select_LDA(LDAs, control$LDA_set_control)
   qprint("Time Series Models", "----", control$quiet)
   TSs <- TS_on_LDA(sel_LDA, dct, formulas, nchangepoints, timename, weights, 
                    control$TS_control)
@@ -180,6 +179,8 @@ check_LDA_TS_inputs <- function(data = NULL,
                               timename = "time", 
                               weights = TRUE, 
                               control = LDA_TS_controls_list()){
+  check_control(control)
+  control <- do.call("LDA_TS_control", control)
   data <- conform_LDA_TS_data(data)
   weights <- iftrue(weights, document_weights(data$document_term_table))
   check_document_covariate_table(data$document_covariate_table, 
[…]

### 2019-07-08 R/LDA_TS_plots.R

@@ -45,8 +45,7 @@
 #'
 #'   mod <- LDA_TS(document_term_table, document_covariate_table,
 #'                 topics = 2, nseeds = 1, formulas = ~ 1, nchangepoints = 1,
-#'                 weights = document_weights(document_term_table),
-#'                 control = LDA_TS_controls_list())
+#'                 weights = document_weights(document_term_table))
 #'   plot(mod, binwidth = 5, xlab = "New moon")
 #' }
 #'

### 2019-07-08 R/TS.R

@@ -34,6 +34,9 @@
 #'   time series into chunks fit with separate models dictated by 
 #'   \code{formula}.
 #'
+#' @param timename \code{character} element indicating the time variable
+#'   used in the time series. Defaults to \code{"time"}.
+#'
 #' @param weights Optional class \code{numeric} vector of weights for each 
 #'   document. Defaults to \code{NULL}, translating to an equal weight for
 #'   each document. When using \code{multinom_TS} in a standard LDATS 
@@ -43,13 +46,10 @@
 #'   For most models, a scaling of the weights (so that the average is 1) is
 #'   most appropriate, and this is accomplished using \code{document_weights}.
 #'
-#' @param timename \code{character} element indicating the time variable
-#'   used in the time series. Defaults to \code{"time"}.
-#'
-#' @param control Class \code{TS_controls} list, holding control parameters
-#'   for the Time Series model including the parallel tempering Markov Chain 
-#'   Monte Carlo (ptMCMC) controls, generated by 
-#'   \code{\link{TS_controls_list}}.
+#' @param control A \code{list} of parameters to control the fitting of the
+#'   Time Series model including the parallel tempering Markov Chain 
+#'   Monte Carlo (ptMCMC) controls. Values not input assume defaults set by 
+#'   \code{\link{TS_control}}.
 #'
 #' @return \code{TS_fit}-class list containing the following elements, many of
 #'   which are hidden for \code{print}ing, but are accessible:
@@ -131,21 +131,22 @@
 #'   data <- document_covariate_table
 #'   data$gamma <- LDA_models@gamma
 #'   weights <- document_weights(document_term_table)
-#'   TSmod <- TS(data, gamma ~ 1, nchangepoints = 1, weights, "newmoon")
+#'   TSmod <- TS(data, gamma ~ 1, nchangepoints = 1, "newmoon", weights)
 #' }
 #'
 #' @export
 #'
-TS <- function(data, formula, nchangepoints, weights, timename = "time",
-               control = TS_controls_list()){
-  check_TS_inputs(data, formula, nchangepoints, weights, timename, control)
+TS <- function(data, formula, nchangepoints, timename = "time",  
+               weights = NULL, control = list()){
+  check_TS_inputs(data, formula, nchangepoints, timename, weights, control)
+  control <- do.call("TS_control", control)
   set.seed(control$seed)
   data <- data[order(data[,timename]), ]
-  rho_dist <- est_changepoints(data, formula, nchangepoints, weights, 
[…]

### 2019-07-08 R/TS_on_LDA.R

@@ -48,10 +48,10 @@
 #'   For most models, a scaling of the weights (so that the average is 1) is
 #'   most appropriate, and this is accomplished using \code{document_weights}.
 #'
-#' @param control Class \code{TS_controls} list, holding control parameters
-#'   for the Time Series model including the parallel tempering Markov Chain 
-#'   Monte Carlo (ptMCMC) controls, generated by 
-#'   \code{\link{TS_controls_list}}.
+#' @param control A \code{list} of parameters to control the fitting of the
+#'   Time Series model including the parallel tempering Markov Chain 
+#'   Monte Carlo (ptMCMC) controls. Values not input assume defaults set by 
+#'   \code{\link{TS_control}}.
 #'
 #' @return Class \code{TS_on_LDA} list of results from \code{\link{TS}} 
 #'   applied for each model on each LDA model input.
@@ -66,17 +66,17 @@
 #'   weights <- document_weights(document_term_table)
 #'   formulas <- c(~ 1, ~ newmoon)
 #'   mods <- TS_on_LDA(LDA_models, document_covariate_table, formulas,
-#'                     nchangepoints = 0:1, timename = "newmoon", weights, 
-#'                     controls)
+#'                     nchangepoints = 0:1, timename = "newmoon", weights)
 #' }
 #'
 #' @export
 #'
 TS_on_LDA <- function(LDA_models, document_covariate_table, formulas = ~ 1, 
                       nchangepoints = 0, timename = "time", weights = NULL, 
-                      control = TS_controls_list()){
+                      control = list()){
   check_TS_on_LDA_inputs(LDA_models, document_covariate_table, formulas, 
                          nchangepoints, timename, weights, control)
+  control <- do.call("TS_control", control)
   mods <- expand_TS(LDA_models, formulas, nchangepoints)
   nmods <- nrow(mods)
   TSmods <- vector("list", nmods)
@@ -86,8 +86,8 @@ TS_on_LDA <- function(LDA_models, document_covariate_table, formulas = ~ 1,
     formula_i <- mods$formula[[i]]
     nchangepoints_i <- mods$nchangepoints[i]
     data_i <- prep_TS_data(document_covariate_table, LDA_models, mods, i)
-    TSmods[[i]] <- TS(data_i, formula_i, nchangepoints_i, weights, 
-                      timename, control)
+    TSmods[[i]] <- TS(data_i, formula_i, nchangepoints_i, timename, weights, 
+                      control)
   }
   package_TS_on_LDA(TSmods, LDA_models, mods)
 
[…]

### 2019-07-08 R/multinom_TS.R

@@ -25,6 +25,9 @@
 #'   checked by \code{\link{check_changepoints}} and
 #'   \code{\link{verify_changepoint_locations}}.
 #'
+#' @param timename \code{character} element indicating the time variable
+#'   used in the time series. Defaults to \code{"time"}.
+#'
 #' @param weights Optional class \code{numeric} vector of weights for each 
 #'   document. Defaults to \code{NULL}, translating to an equal weight for
 #'   each document. When using \code{multinom_TS} in a standard LDATS 
@@ -35,13 +38,10 @@
 #'   most appropriate, and this is accomplished using 
 #'   \code{\link{document_weights}}.
 #'
-#' @param timename \code{character} element indicating the time variable
-#'   used in the time series. Defaults to \code{"time"}.
-#'
-#' @param control Class \code{TS_controls} list, holding control parameters
-#'   for the Time Series model including the parallel tempering Markov Chain 
-#'   Monte Carlo (ptMCMC) controls, generated by 
-#'   \code{\link{TS_controls_list}}.
+#' @param control A \code{list} of parameters to control the fitting of the
+#'   Time Series model including the parallel tempering Markov Chain 
+#'   Monte Carlo (ptMCMC) controls. Values not input assume defaults set by 
+#'   \code{\link{TS_control}}.
 #'
 #' @return Object of class \code{multinom_TS_fit}, which is a list of [1]
 #'   chunk-level model fits (\code{"chunk models"}), [2] the total log 
@@ -57,22 +57,23 @@
 #' \dontrun{
 #'   data(rodents)
 #'   dtt <- rodents$document_term_table
-#'   lda <- LDA_set(dtt, 4, 1, LDA_controls_list(quiet = TRUE))
+#'   lda <- LDA_set(dtt, 4, 1, list(quiet = TRUE))
 #'   dct <- rodents$document_covariate_table
 #'   dct$gamma <- lda[[1]]@gamma
 #'   weights <- document_weights(dtt)
 #'   mts <- multinom_TS(dct, formula = gamma ~ 1, changepoints = c(20,50),
-#'                      weights = weights, timename = "newmoon") 
+#'                      timename = "newmoon", weights = weights) 
 #' }
 #'
 #' @export 
 #'
 multinom_TS <- function(data, formula, changepoints = NULL, 
-                        weights = NULL, timename = "time",
-                        control = TS_controls_list()){
+                        timename = "time", weights = NULL, 
+                        control = list()){
 
[…]

### 2019-07-08 R/ptMCMC.R

@@ -330,8 +330,8 @@ proposed_step_mods <- function(prop_changepts, inputs){
   timename <- inputs$timename
   out <- vector("list", length = ntemps)
   for (i in 1:ntemps){
-    out[[i]] <- TS_memo(data, formula, prop_changepts[ , i], weights, 
-                        timename, control)
+    out[[i]] <- TS_memo(data, formula, prop_changepts[ , i], timename, 
+                        weights, control)
   }
   out
 }
@@ -349,16 +349,17 @@ proposed_step_mods <- function(prop_changepts, inputs){
 #'   and would work within any general ptMCMC as long as \code{control},
 #'   \code{ids}, and \code{swaps} are formatted properly.
 #'
-#' @param control Class \code{TS_controls} list, holding control parameters
-#'   for the Time Series model including the parallel tempering Markov Chain 
-#'   Monte Carlo (ptMCMC) controls, generated by 
-#'   \code{\link{TS_controls_list}}. 
+#' @param control A \code{list} of parameters to control the fitting of the
+#'   Time Series model including the parallel tempering Markov Chain 
+#'   Monte Carlo (ptMCMC) controls. Values not input assume defaults set by 
+#'   \code{\link{TS_control}}.
 #'
 #' @return The vector of chain ids.
 #'
 #' @export
 #'
-prep_ids <- function(control){
+prep_ids <- function(control = list()){
+  control <- do.call("TS_control", control)
   if (!is.numeric(control$ntemps) || any(control$ntemps %% 1 != 0)){
     stop("ntemps must be integer-valued")
   }
@@ -405,6 +406,9 @@ update_ids <- function(ids, swaps){
 #'   number of change points is used to dictate the segementation of the data  
 #'   for each continuous model and each LDA model.
 #'
+#' @param timename \code{character} element indicating the time variable
+#'   used in the time series.
+#'
 #' @param weights Optional class \code{numeric} vector of weights for each 
 #'   document. Defaults to \code{NULL}, translating to an equal weight for
 #'   each document. When using \code{multinom_TS} in a standard LDATS 
@@ -415,25 +419,24 @@ update_ids <- function(ids, swaps){
 #'   most appropriate, and this is accomplished using 
 #'   \code{\link{document_weights}}.
 #'
-#' @param timename \code{character} element indicating the time variable
-#'   used in the time series. 
-#'
-#' @param control Class \code{TS_controls} list, holding control parameters
[…]

### 2019-07-08 R/utilities.R

@@ -38,7 +38,6 @@ iftrue <- function(x = TRUE, alt = NULL){
   x
 }
 
-
 #' @title Determine the mode of a distribution
 #'
 #' @description Find the most common entry in a vector. Ties are not allowed,
@@ -198,9 +197,9 @@ memoise_fun <- function(fun, memoise_tf){
 #'
 #' @export
 #'
-check_control <- function(control, eclass = "TS_controls"){
+check_control <- function(control, eclass = "list"){
   if (!(eclass %in% class(control))){
-    stop(paste0("control is not of class ", eclass))
+    stop(paste0("control is not a ", eclass))
   }
 }
 

### 2019-07-08 doc/rodents-example.R

@@ -23,7 +23,7 @@ head(rodents$document_covariate_table, 10)
 #  lda_model_set <- LDA_set(document_term_table = rodents$document_term_table,
 #                           topics = c(2:5),
 #                           nseeds = 10,
-#                           control = LDA_controls_list(quiet = TRUE))
+#                           control = list(quiet = TRUE))
 #  
 
 ## ----lda set not quiet---------------------------------------------------
@@ -59,7 +59,7 @@ plot(selected_lda_model[[1]])
 #                                  nchangepoints = c(0:1),
 #                                  timename = "newmoon",
 #                                  weights = document_weights(rodents$document_term_table),
-#                                  control = TS_controls_list(nit = 1000))
+#                                  control = list(nit = 1000))
 #  
 
 ## ----reload ts, include = F----------------------------------------------
@@ -90,8 +90,7 @@ plot(selected_changepoint_model)
 #                           formulas = ~ sin_year + cos_year,
 #                           nchangepoints= 0:1,
 #                           timename = "newmoon",
-#                           control = LDA_TS_controls_list(
-#                             TS_control = TS_controls_list(nit = 1000)))
+#                           control = list(nit = 1000))
 
 ## ----load ldats results, include = F-------------------------------------
 load(here::here('vignettes', 'rodents-example-files', 'lda_ts_results.Rds'))

### 2019-07-08 tests/testthat/test-03-lda_plot.R

@@ -3,7 +3,7 @@ tenv <- "cran"
 
 data(rodents)
 lda_data <- rodents$document_term_table
-ldas <- LDA_set(lda_data, c(2, 4), 2, LDA_controls_list(quiet = TRUE))
+ldas <- LDA_set(lda_data, c(2, 4), 2, list(quiet = TRUE))
 lda <- ldas[[1]]
 xtime <- rodents$newmoon
 

### 2019-07-08 tests/testthat/test-04-LDA_TS.R

@@ -9,8 +9,7 @@ test_that("LDA_TS on 0 changepoints", {
   mod0 <- LDA_TS(rodents,
                  topics = 2, nseeds = 1, formulas = ~ 1, nchangepoints = 0,
                  timename = "newmoon",
-                 control = LDA_TS_controls_list(
-                           TS_control = TS_controls_list(nit = 100)))
+                 control = list(nit = 100))
   expect_is(mod0, "LDA_TS")
   expect_equal(length(names(mod0)), 4)
   expect_is(mod0[[4]], "TS_fit")
@@ -20,8 +19,7 @@ test_that("LDA_TS on 1 changepoints", {
   mod1 <- LDA_TS(rodents,
                  topics = 2, nseeds = 1, formulas = ~ 1, nchangepoints = 1,
                  timename = "newmoon",
-                 control = LDA_TS_controls_list(
-                           TS_control = TS_controls_list(nit = 100)))
+                 control = list(nit = 100))
   expect_is(mod1, "LDA_TS")
   expect_equal(length(names(mod1)), 4)
   expect_is(mod1[[4]], "TS_fit")
@@ -31,17 +29,13 @@ test_that("check print on LDA_TS", {
   mod1 <- LDA_TS(rodents,
                  topics = 2, nseeds = 1, formulas = ~ 1, nchangepoints = 1,
                  timename = "newmoon",
-                 control = LDA_TS_controls_list(
-                           TS_control = TS_controls_list(nit = 100)))
+                 control = list(nit = 100))
   expect_output(print(mod1))
 })
 
 test_that("Check LDA_TS_controls_list", {
-  expect_is(LDA_TS_controls_list(), "LDA_TS_controls")
-  expect_equal(length(LDA_TS_controls_list()), 3)
-  expect_error(LDA_TS_controls_list(LDA_control = "ok"))
-  expect_error(LDA_TS_controls_list(TS_control = "ok"))
-  expect_error(LDA_TS_controls_list(quiet = "ok"))
+  expect_is(LDA_TS_control(), "list")
+  expect_equal(length(LDA_TS_control()), 3)
 })
 
 test_that("Check conform_LDA_TS_data", {
@@ -65,9 +59,10 @@ test_that("Check package_LDA_TS", {
   nchangepoints <- 1
   weights <- document_weights(document_term_table)
   timename <-  "newmoon"
[…]

### 2019-07-08 tests/testthat/test-05-LDA_TS_plots.R

@@ -9,15 +9,11 @@ document_covariate_table <- rodents$document_covariate_table
 mod <- LDA_TS(rodents,
               topics = 2, nseeds = 1, formulas = ~ 1, nchangepoints = 1,
               timename = "newmoon",
-              control = LDA_TS_controls_list(
-                        TS_control = TS_controls_list(nit = 100, seed = 1)))
+              control = list(nit = 100, seed = 1))
 
 test_that("check plot for LDA_TS", {
   if (tenv == "cran"){
-    expect_silent(plot(mod, 
-                       control = LDA_TS_controls_list(
-                                       TS_control = TS_controls_list(
-                                                     nit = 100, seed = 1)),
+    expect_silent(plot(mod, control = list(nit = 100, seed = 1),
                            interactive = FALSE))
   } else{
     plot(mod, interactive = FALSE)

### 2019-07-08 tests/testthat/test-07-ptMCMC.R

@@ -15,28 +15,27 @@ formulas <- ~ 1
 nchangepoints <- 1
 weights <- document_weights(document_term_table)
 timename <- "newmoon"
-control <- LDA_TS_controls_list()
-LDAs <- LDA_set(document_term_table, topics, nseeds, control$LDA_control)
-LDA_models <- select_LDA(LDAs, control$LDA_control)
-control <- TS_controls_list(nit = 1e3, seed = 1)
+LDAs <- LDA_set(document_term_table, topics, nseeds)
+LDA_models <- select_LDA(LDAs)
+control <- list(nit = 1e3, seed = 1)
 mods <- expand_TS(LDA_models, formulas, nchangepoints)
 formula <- mods$formula[[1]]
 nchangepoints <- mods$nchangepoints[1]
 data <- prep_TS_data(document_covariate_table, LDA_models, mods, 1)
 
 set.seed(1)
-rho_dist0 <- est_changepoints(data, formula, nchangepoints = 0, weights, 
-                              timename, control)
-rho_dist <- est_changepoints(data, formula, nchangepoints, weights, 
-                             timename, control)
-eta_dist <- est_regressors(rho_dist, data, formula, weights, timename, 
-                            control)
+rho_dist0 <- est_changepoints(data, formula, nchangepoints = 0, timename, 
+                              weights, control)
+rho_dist <- est_changepoints(data, formula, nchangepoints, timename, 
+                              weights, control)
+eta_dist <- est_regressors(rho_dist, data, formula, timename, weights, 
+                           control)
 
 
 saves <- prep_saves(nchangepoints, control)
-inputs <- prep_ptMCMC_inputs(data, formula, nchangepoints, weights, 
-                             timename, control)
-cpts <- prep_cpts(data, formula, nchangepoints, weights, timename, control)
+inputs <- prep_ptMCMC_inputs(data, formula, nchangepoints, timename, weights, 
+                             control)
+cpts <- prep_cpts(data, formula, nchangepoints, timename, weights, control)
 ids <- prep_ids(control)
 
 test_that("check prep_proposal_dist", {
@@ -53,36 +52,36 @@ test_that("check prep_proposal_dist", {
 
 
 test_that("check prep_ptMCMC_inputs", {
-  inpts <- prep_ptMCMC_inputs(data, formula, nchangepoints, weights, 
[…]

### 2019-07-08 tests/testthat/test-09-TS_on_LDA.R

@@ -10,10 +10,9 @@ formulas <- ~ 1
 nchangepoints <- 0
 weights <- document_weights(document_term_table)
 timename <- "newmoon"
-control <- LDA_TS_controls_list()
-LDAs <- LDA_set(document_term_table, topics, nseeds, control$LDA_control)
-LDA_models <- select_LDA(LDAs, control$LDA_control)
-control <- control$TS_control
+LDAs <- LDA_set(document_term_table, topics, nseeds)
+LDA_models <- select_LDA(LDAs)
+
 
 test_that("check prep_TS_data", {
   mods <- expand_TS(LDA_models, formulas=c(~1, ~sin_year, ~newmoon), 
@@ -40,18 +39,18 @@ test_that("check prep_TS_data", {
 test_that("check select_TS", {
   mods <- TS_on_LDA(LDA_models, document_covariate_table, formulas, 
                   nchangepoints = 0:1, timename, weights, 
-                  control = TS_controls_list(nit = 1e3))
-  sel_mod <- select_TS(mods, control)
+                  control = list(nit = 1e3))
+  sel_mod <- select_TS(mods, list())
 
   expect_equal(length(mods), 2)
   expect_equal(length(sel_mod), 17)
   expect_is(mods, "TS_on_LDA")
   expect_is(sel_mod, "TS_fit")
-  expect_error(select_TS(mods, LDA_controls_list()))
-  expect_error(select_TS("ok", TS_controls_list()))
+  expect_error(select_TS(mods, "ok"))
+  expect_error(select_TS("ok", list()))
 
   xxfun <- function(x){x}
-  expect_warning(select_TS(mods, TS_controls_list(selector = xxfun)))
+  expect_warning(select_TS(mods, list(selector = xxfun)))
 })
 
 test_that("check package_TS_on_LDA", {
@@ -60,12 +59,12 @@ test_that("check package_TS_on_LDA", {
   TSmods <- vector("list", nmods)
 
   for(i in 1:nmods){
-    print_model_run_message(mods, i, LDA_models, control)
+    print_model_run_message(mods, i, LDA_models, list())
     formula_i <- mods$formula[[i]]
     nchangepoints_i <- mods$nchangepoints[i]
     data_i <- prep_TS_data(document_covariate_table, LDA_models, mods, i)
-    TSmods[[i]] <- TS(data_i, formula_i, nchangepoints_i, weights, timename,
[…]

### 2019-07-08 tests/testthat/test-11-utilities.R

@@ -67,19 +67,19 @@ test_that("check mirror_vcov", {
   formulas <- ~ newmoon
   nchangepoints <- 2
   weights <- document_weights(document_term_table)
-  control <- LDA_TS_controls_list()
+  control <- list()
   timename <- "newmoon"
-  LDAs <- LDA_set(document_term_table, topics, nseeds, control$LDA_control)
-  LDA_models <- select_LDA(LDAs, control$LDA_control)
-  control <- TS_controls_list(nit = 1e2, seed = 1)
+  LDAs <- LDA_set(document_term_table, topics, nseeds, list())
+  LDA_models <- select_LDA(LDAs, list())
+  control <- list(nit = 1e2, seed = 1)
   mods <- expand_TS(LDA_models, formulas, nchangepoints)
   formula <- mods$formula[[1]]
   nchangepoints <- mods$nchangepoints[1]
   data <- prep_TS_data(document_covariate_table, LDA_models, mods, 1)
-  rho_dist <- est_changepoints(data, formula, nchangepoints, weights, 
-                               timename, control)
-  mod <- multinom_TS(data, formula, changepoints = NULL, weights, 
-                     timename, control)
+  rho_dist <- est_changepoints(data, formula, nchangepoints, timename, 
+                               weights, control)
+  mod <- multinom_TS(data, formula, changepoints = NULL, timename, weights, 
+                     control)
   expect_equal(isSymmetric(vcov(mod[[1]][[1]])), FALSE)
   expect_equal(isSymmetric(mirror_vcov(mod[[1]][[1]])), TRUE)
 
@@ -103,12 +103,10 @@ test_that("check memoise_fun", {
 })
 
 test_that("check check_control", {
-  expect_silent(check_control(TS_controls_list(), "TS_controls"))
-  expect_silent(check_control(LDA_TS_controls_list(), "LDA_TS_controls"))
-  expect_silent(check_control(LDA_controls_list(), "LDA_controls"))
-  expect_error(check_control(TS_controls_list(), "LDA_controls"))
-  expect_error(check_control(LDA_controls_list(), "TS_controls"))
-  expect_error(check_control(LDA_TS_controls_list(), "LDA_controls"))
+  expect_silent(check_control(list(), "list"))
+  expect_silent(check_control(list()))
+  expect_error(check_control(list(), 1))
+  expect_error(check_control(1, "list"))
 })
 
 

### 2019-07-08 doc/rodents-example.Rmd
updating documentation to be clear about the time input
and strictly enforcing integer or date (and thus integer) timesteps

@@
 In this case, we have a column for `newmoon`, which is the time step for each of our sampling periods.
-This column can include nonconsecutive time steps (i.e., `2, 4, 5`) if sampling periods are skipped or at unequal intervals.
+This column can include nonconsecutive time steps (i.e., `2, 4, 5`) if sampling periods are skipped or at unequal intervals, but must be integer-conformable or a date.
+Although if a date, the assumption is that the timestep is 1 day, which is often not desired behavior.
 If there is no time column, `LDATS` will assume all time steps are equidistant.

### 2019-07-08 R/TS.R

@@ -35,7 +35,10 @@
 #'   \code{formula}.
 #'
 #' @param timename \code{character} element indicating the time variable
-#'   used in the time series. Defaults to \code{"time"}.
+#'   used in the time series. Defaults to \code{"time"}. The variable must be
+#'   integer-conformable or a \code{Date}. If the variable named
+#'   is a \code{Date}, the input is converted to an integer, resulting in the
+#'   timestep being 1 day, which is often not desired behavior. 
 #'
 #' @param weights Optional class \code{numeric} vector of weights for each 
 #'   document. Defaults to \code{NULL}, translating to an equal weight for

### 2019-07-08 R/TS_on_LDA.R

@@ -37,7 +37,10 @@
 #'   component of the TS model, for each selected LDA model.
 #'
 #' @param timename \code{character} element indicating the time variable
-#'   used in the time series. Defaults to \code{"time"}.
+#'   used in the time series. Defaults to \code{"time"}. The variable must be
+#'   integer-conformable or a \code{Date}. If the variable named
+#'   is a \code{Date}, the input is converted to an integer, resulting in the
+#'   timestep being 1 day, which is often not desired behavior.
 #'
 #' @param weights Optional class \code{numeric} vector of weights for each 
 #'   document. Defaults to \code{NULL}, translating to an equal weight for
@@ -454,7 +457,10 @@ check_document_covariate_table <- function(document_covariate_table,
 #' @title Check that the time vector is proper
 #' 
 #' @description Check that the vector of time values is included in the 
-#'   document covariate table and that it is either numeric or a date.
+#'   document covariate table and that it is either a integer-conformable or
+#'   a \code{date}. If it is a \code{date}, the input is converted to an 
+#'   integer, resulting in the timestep being 1 day, which is often not 
+#'   desired behavior. 
 #'   
 #' @param document_covariate_table Document covariate table used to query
 #'   for the time column.
@@ -475,8 +481,9 @@ check_timename <- function(document_covariate_table, timename){
     stop("timename not present in document covariate table")
   }
   time_covariate <- document_covariate_table[ , timename]
-  if (!(is.numeric(time_covariate)) & !(is.Date(time_covariate))){
-    stop("covariate indicated by timename is not numeric or temporal")
+  if (!(is.Date(time_covariate)) & 
+      !is.numeric(time_covariate) || !all(time_covariate %% 1 == 0)){
+    stop("covariate indicated by timename is not an integer or a date")
   }
 }
 

### 2019-07-08 R/ptMCMC.R

@@ -407,7 +407,10 @@ update_ids <- function(ids, swaps){
 #'   for each continuous model and each LDA model.
 #'
 #' @param timename \code{character} element indicating the time variable
-#'   used in the time series.
+#'   used in the time series. Defaults to \code{"time"}. The variable must be
+#'   integer-conformable or a \code{Date}. If the variable named
+#'   is a \code{Date}, the input is converted to an integer, resulting in the
+#'   timestep being 1 day, which is often not desired behavior.
 #'
 #' @param weights Optional class \code{numeric} vector of weights for each 
 #'   document. Defaults to \code{NULL}, translating to an equal weight for
@@ -622,7 +625,10 @@ process_saves <- function(saves, control = list()){
 #'   for each continuous model and each LDA model.
 #'
 #' @param timename \code{character} element indicating the time variable
-#'   used in the time series. 
+#'   used in the time series. Defaults to \code{"time"}. The variable must be
+#'   integer-conformable or a \code{Date}. If the variable named
+#'   is a \code{Date}, the input is converted to an integer, resulting in the
+#'   timestep being 1 day, which is often not desired behavior.
 #'
 #' @param weights Optional class \code{numeric} vector of weights for each 
 #'   document. Defaults to \code{NULL}, translating to an equal weight for

### 2019-07-08 tests/testthat/test-09-TS_on_LDA.R

@@ -189,6 +189,10 @@ test_that("check check_timename", {
                               rep(timename, 2)))
   expect_error(check_timename(document_covariate_table, 1))
   expect_error(check_timename(data.frame(letters), "letters"))
+
+  dct2 <- document_covariate_table
+  dct2[,timename] <- dct2[,timename] + 0.1
+  expect_error(check_timename(dct2, timename))
 })
 
 test_that("check check_formulas", {

### 2019-07-10 doc/paper-comparison.Rmd

@@
                              nchangepoints = 1:6,
-                             weights = NULL,
-                             control = TS_controls_list(timename = 'newmoon'))
+                             timename = 'newmoon',
+                             weights = NULL)
 
@@
                              nchangepoints = 1:6,
-                             weights = NULL,
-                              control = TS_controls_list(timename = 'newmoon'))
+                             timename = 'newmoon',
+                             weights = NULL)
 

### 2019-07-10 vignettes/christensenetal-comparison-files/christensenetal-2018-files/LDA-distance.R
for updated model output

@@ -1,117 +1,0 @@
-# This script assumes that `dat` is defined
-
-library(topicmodels)
-
-
-#' Compute hellinger distance
-#' 
-#' Hellinger distance is Euclidean distance between square roots, divided by
-#' sqrt(2)
-#' 
-#' @param a,b numeric vectors
-#' 
-Hellinger = function(a, b){
-  diff = sqrt(a) - sqrt(b)
-  sqrt(sum(diff^2) / 2)
-}
-
-#' Find the configuration with the minimum Hellinger distance between two
-#' probability matrices like those from `ps`.
-#' 
-#' @param p1,p2 numeric vectors: species composition
-#' @param k number of topics
-#' 
-min_H = function(p1, p2, k) {
-  # Find the cost associated with each pairwise topic assignment
-  costs = matrix(0, k, k)
-  for (i in 1:k) {
-    for (j in 1:k) {
-      costs[i, j] = Hellinger(p1[i, ], p2[j, ])
-    }
-  }
-  # Minimize the sum of the pairwise costs.
-  # Adding a constant doesn't change the results, but all costs must be strictly
-  # positive.
-  assignment = clue::solve_LSAP(1 + costs)
-  list(
-    min_cost = mean(costs[cbind(1:k, assignment)]),
-    assignment = assignment
-  )
-}
-
-
-
-#' Calculates multiple LDA models using different seeds
-#'
-#'
-#' @param ldas list of results from LDA model
-#' @param seeds vector of seeds for which LDA models will be fit and results compared
-#' 
-#' @return list including: seed associated with highest log likelihood
-#'                         mean distance between best model and all others
-#'                         max distuance between best model and worst
-#'
-calculate_LDA_distance = function(ldas,seeds) {
-  
-  # Calculate a bunch of LDAs with 4 topics
-  k = 4
-  ldas = purrr::map(seeds, 
-                    ~LDA(dat, k = k, method = "VEM", control = list(seed = .x)))
-  
-  # Log-likelihoods for each lda
-  lls = purrr:::map_dbl(ldas, logLik)
-  
-  # Pick the LDA model with the highest log-likelihood
-  best_lda = which.max(lls)
-  
-  # Topic allocations to each species (probabilities that sum to 1 in each row)
-  ps = purrr::map(ldas, ~exp(.x@beta))
-  best_ps = exp(ldas[[best_lda]]@beta)
[…]

### 2019-07-10 doc/paper-comparison.R
v0.2.2 for cran

@@ -1,472 +1,402 @@
+params <-
+list(run_models = FALSE)
+
 ## ----setup, include=FALSE------------------------------------------------
 knitr::opts_chunk$set(
   collapse = TRUE,
   comment = "#>"
 )
 
-
-## ---- eval=FALSE---------------------------------------------------------
-#  install.packages("devtools")
+## ---- eval = FALSE-------------------------------------------------------
+#  # install.packages("devtools")
 #  devtools::install_github("weecology/LDATS")
 
-## ---- eval = T-----------------------------------------------------------
+## ------------------------------------------------------------------------
 library(LDATS)
-library(dplyr)
-load(here::here('vignettes', 'christensenetal-comparison-files',  'data_files.Rdata'))
-load(here::here('vignettes', 'christensenetal-comparison-files', 'model-cache',  'ldas.RData'))
-load(here::here('vignettes', 'christensenetal-comparison-files', 'model-cache', 'ldats_changepoints.RData'))
-load(here::here('vignettes', 'christensenetal-comparison-files', 'model-cache', 'paper_changepoints.RData'))
+set.seed(42)
+nseeds <- 200
+nit <- 10000
 
-## ----download Christensen, eval = F--------------------------------------
-#  
-#  paper_filepath <- here::here('vignettes', 'christensenetal-comparison-files', 'christensenetal-2018-files')
-#  test_filepath <- here::here('vignettes', 'christensenetal-comparison-files', 'christensenetal-2018-files', 'rodent_LDA_analysis.r')
-#  
-#  if(!file.exists(test_filepath)) {
-#  
-#  files_to_download <- c('rodent_LDA_analysis.r', 'rodent_data_for_LDA.r', 'AIC_model_selection.R', 'changepointmodel.r', 'LDA-distance.R', 'Rodent_table_dat.csv', 'LDA_figure_scripts.R')
-#  
-#  for(i in 1:length(files_to_download)) {
-#    download.file(url = paste0("https://raw.githubusercontent.com/emchristensen/Extreme-events-LDA/master/", files_to_download[i]),
-#                  destfile = paste0(paper_filepath, '/', files_to_download[i]))
-#  }
-#  
-#  rm(files_to_download)
-#  rm(i)
-#  
-#  }
-#  
-#  rm(test_filepath)
-#  rm(paper_filepath)
+## ---- eval = FALSE-------------------------------------------------------
+#  rmarkdown::render("paper-comparison.Rmd", params = list(run_models = TRUE))
[…]

### 2019-07-10 doc/paper-comparison.R
vignette rebuilds

@@ -7,6 +7,9 @@ knitr::opts_chunk$set(
   comment = "#>"
 )
 
+## ---- include=FALSE------------------------------------------------------
+vers <- packageVersion("LDATS")
+
 ## ---- eval = FALSE-------------------------------------------------------
 #  # install.packages("devtools")
 #  devtools::install_github("weecology/LDATS")

### 2019-07-10 doc/paper-comparison.Rmd

@@
 title: "Comparison to Christensen et al. 2018"
-author: "Renata Diaz"
+author: "Renata Diaz and Hao Ye"
 output: rmarkdown::html_vignette
@@
   %\VignetteEncoding{UTF-8}
+params:
+  run_models: FALSE
 ---
@@
 )
-
+```
+```{r, include=FALSE}
+vers <- packageVersion("LDATS")
 ```
 
-This document prodives a side-by-side comparison of LDATS (v0.1.0) results with analysis from [Christensen et al. 2018](https://esajournals.onlinelibrary.wiley.com/doi/full/10.1002/ecy.2373).
+# Introduction
+
+This document provides a side-by-side comparison of **`LDATS`** (version `r vers`) results with analysis from [Christensen et al.
+2018](https://esajournals.onlinelibrary.wiley.com/doi/full/10.1002/ecy.2373).
 
@@
 
+# Setup
+
 ## LDATS Installation
 
-To obtain the most recent version of **LDATS**, install the most recent
-version from GitHub:
+To obtain the most recent version of **`LDATS`**, install the most recent version from GitHub:
 
-```{r, eval=FALSE}
-install.packages("devtools")
+```{r, eval = FALSE}
+# install.packages("devtools")
 devtools::install_github("weecology/LDATS")
@@
 
-```{r, eval = T}
+Load in the **`LDATS`** package.
+```{r}
 library(LDATS)
-library(dplyr)
-load(here::here('vignettes', 'christensenetal-comparison-files',  'data_files.Rdata'))
-load(here::here('vignettes', 'christensenetal-comparison-files', 'model-cache',  'ldas.RData'))
-load(here::here('vignettes', 'christensenetal-comparison-files', 'model-cache', 'ldats_changepoints.RData'))
-load(here::here('vignettes', 'christensenetal-comparison-files', 'model-cache', 'paper_changepoints.RData'))
+set.seed(42)
+nseeds <- 200
+nit <- 10000
 ```
 
-## Christensen 2018 analysis files
+## Running the Models
 
-Download Christensen 2018 analysis scripts & data files from [Extreme-events-LDA repo:](https://github.com/emchristensen/Extreme-events-LDA)
+Because both the Latent Dirichlet Allocation (LDA) and time series components of the analysis can take a long time to run (especially with the settings above for the number of seeds and iterations), we will use pre-generated model outputs and turn off certain code chunks that run the models using a global `rmarkdown` parameter, `run_models = FALSE`.
+
+To change this functionality, you can re-render this file with:
+```{r, eval = FALSE}
+rmarkdown::render("paper-comparison.Rmd", params = list(run_models = TRUE))
+```
+
+## Download Analysis Scripts and Data Files
+
+We're going to download analysis scripts, data files, and model objects, so we use a temporary location for storage:
+
+```{r set download location}
+vignette_files <- tempdir()
+```
+
+To replicate the Christensen et al.
+2018 analysis, we download some of the original scripts & data files from [Extreme-events-LDA repo](https://github.com/emchristensen/Extreme-events-LDA), as well as some raw data files from the [PortalData repo](https://github.com/weecology/PortalData):
 
@@
 
-  * rodent_LDA_analysis.R main script for analyzing rodent community change using LDA
+* `rodent_LDA_analysis.R`
+  - main script for analyzing rodent community change using LDA
 
-  * rodent_data_for_LDA.R contains a function that creates the rodent data table used in analyses
+* `rodent_data_for_LDA.R`
+  - contains a function that creates the rodent data table used in analyses
 
-  * AIC_model_selection.R contains functions for calculating AIC for different candidate LDA models
+* `AIC_model_selection.R`
+  - contains functions for calculating AIC for different candidate LDA models
 
-  * changepointmodel.r contains change-point model code
+* `changepointmodel.r`
+  - contains change-point model code
 
-  * LDA-distance.R function for computing Hellinger distance analyses
+* `LDA-distance.R`
+  - function for computing Hellinger distance analyses
 
@@
 
-  * Rodent_table_dat.csv table of rodent data, created by rodent_data_for_LDA.R
+* `Rodent_table_dat.csv`
+  - table of rodent data, created by rodent_data_for_LDA.R
+
+* `moon_dates.csv`
+  - table of census dates (downloaded from the PortalData repository)
+
+* `Portal_rodent_trapping.csv`
+  - table of trapping effort (downloaded from the PortalData repository)
 
@@
 
-  * LDA_figure_scripts.R contains functions for making main plots in manuscript (Fig 1). Called from rodent_LDA_analysis.R
+* `LDA_figure_scripts.R`
+- contains functions for making main plots in manuscript (Fig 1). Called from rodent_LDA_analysis.R
 
-```{r download Christensen, eval = F}
+```{r download scripts}
+test_file <- file.path(vignette_files, "rodent_LDA_analysis.r")
 
-paper_filepath <- here::here('vignettes', 'christensenetal-comparison-files', 'christensenetal-2018-files')
-test_filepath <- here::here('vignettes', 'christensenetal-comparison-files', 'christensenetal-2018-files', 'rodent_LDA_analysis.r')
+if (!file.exists(test_file))
+{
+  # from the Extreme-events-LDA repo
+  github_path <- "https://raw.githubusercontent.com/emchristensen/Extreme-events-LDA/master/"
+  files_to_download <- c("rodent_LDA_analysis.r", "rodent_data_for_LDA.r", 
+                         "AIC_model_selection.R", "changepointmodel.r", 
+                         "LDA-distance.R", "Rodent_table_dat.csv", 
+                         "LDA_figure_scripts.R")
 
-if(!file.exists(test_filepath)) {
-
-files_to_download <- c('rodent_LDA_analysis.r', 'rodent_data_for_LDA.r', 'AIC_model_selection.R', 'changepointmodel.r', 'LDA-distance.R', 'Rodent_table_dat.csv', 'LDA_figure_scripts.R')
-
-for(i in 1:length(files_to_download)) {
-  download.file(url = paste0("https://raw.githubusercontent.com/emchristensen/Extreme-events-LDA/master/", files_to_download[i]),
-                destfile = paste0(paper_filepath, '/', files_to_download[i]))
+  for (file in files_to_download)
+  {
+    download.file(url = paste0(github_path, file),
+                  destfile = file.path(vignette_files, file))
   }
 
-rm(files_to_download)
-rm(i)
+  # from the PortalData repo
+  github_path <- "https://raw.githubusercontent.com/weecology/PortalData/master/Rodents/"
+  files_to_download <- c("moon_dates.csv", "Portal_rodent_trapping.csv")
 
+  for (file in files_to_download)
+  {
+    download.file(url = paste0(github_path, file),
+                  destfile = file.path(vignette_files, file))
   }
-
-rm(test_filepath)
-rm(paper_filepath)
+}
 ```
 
-## Data
+## Download Model Outputs
 
-The Portal rodents control data is included in the LDATS package:
+We also have pre-generated model outputs that we download from the [LDATS-replications repo](https://github.com/weecology/LDATS-replications):
+
+LDA models:
+
+* `ldats_ldamodel.RDS`
+  - the best LDA model as selected by LDATS
+
+* `paper_ldamodel.RDS`
+  - the best LDA model as selected by the Christensen et al. analysis
+
+Changepoint outputs
+
+* `ldats_ldats.RDS`
+  - the posterior distribution of changepoints, using the LDATS LDA model and the LDATS changepoint selection
+
+* `ldats_paper.RDS`
+  - the posterior distribution of changepoints, using the LDATS LDA model and the paper's changepoint selection
+
+* `paper_ldats.RDS`
+  - the posterior distribution of changepoints, using the paper LDA model and the LDATS changepoint selection
+
+* `paper_paper.RDS`
+  - the posterior distribution of changepoints, using the paper LDA model and the paper's changepoint selection
+
+Figures
+
+* `lda_distances.png`
+  - figure showing the variance in the topics identified by the paper's LDA model code
+
+```{r download pre-generated model outputs, eval = !params$run_model}
+test_file <- file.path(vignette_files, "ldats_ldamodel.RDS")
+
+if (!file.exists(test_file))
+{
+  # from the Extreme-events-LDA repo
+  github_path <- "https://raw.githubusercontent.com/weecology/LDATS-replications/master/model_outputs/"
+  files_to_download <- c("ldats_ldamodel.RDS", "paper_ldamodel.RDS", 
+                         "ldats_ldats.RDS", "ldats_paper.RDS", 
+                         "paper_ldats.RDS", "paper_paper.RDS", 
+                         "lda_distances.png")
+
+  for (file in files_to_download)
+  {
+    download.file(url = paste0(github_path, file),
+                  destfile = file.path(vignette_files, file), 
+                  mode = "wb")
+  }
+}
+```
+
+# Data Comparison
+
+The dataset of Portal rodents on control plots is included in the LDATS package:
 
@@
 
-Load the data used in Christensen et al:
+We can compare this against the data used in Christensen et al:
 
-```{r Paper data, eval = F}
-source(here::here('vignettes', 'christensenetal-comparison-files', 'christensenetal-2018-files', 'rodent_data_for_LDA.r'))
+```{r Paper data}
+# parameters for subsetting the full Portal rodents data
+periods <- 1:436
+control_plots <- c(2, 4, 8, 11, 12, 14, 17, 22)
+species_list <- c("BA", "DM", "DO", "DS", "NA", "OL", "OT", "PB", "PE", "PF", 
+                  "PH", "PI", "PL", "PM", "PP", "RF", "RM", "RO", "SF", "SH", "SO")
 
-dat = create_rodent_table(period_first = 1,
-                          period_last = 436,
-                          selected_plots = c(2,4,8,11,12,14,17,22),
-                          selected_species = c('BA','DM','DO','DS','NA','OL','OT','PB','PE','PF','PH','PI','PL','PM','PP','RF','RM','RO','SF','SH','SO'))
+source(file.path(vignette_files, "rodent_data_for_LDA.r"))
 
-# dates to go with count data
-moondat = read.csv(text=getURL("https://raw.githubusercontent.com/weecology/PortalData/master/Rodents/moon_dates.csv"),stringsAsFactors = F)
-moondat$date = as.Date(moondat$censusdate)
+# assemble `paper_dat`, the data from Christensen et al. 2018
+paper_dat <- create_rodent_table(period_first = min(periods),
+                                 period_last = max(periods),
+                                 selected_plots = control_plots,
+                                 selected_species = species_list)
 
-period_dates = filter(moondat,period %in% rownames(dat)) %>% select(period,date)
-dates = period_dates$date
+# assemble `paper_covariates`, the associated dates and covariate data
+moondat <- read.csv(file.path(vignette_files, "moon_dates.csv"), stringsAsFactors = F)
 
-paper_dat <- dat
+paper_dates <- moondat %>%
+  dplyr::filter(period %>% dplyr::between(min(periods), max(periods))) %>% 
+  dplyr::pull(censusdate) %>%
+  as.Date()
 
-paper_dates <- dates
-
-rm(list = c('dat', 'dates', 'period_dates',
-            'create_rodent_table'))
-
+paper_covariates <- data.frame(
+  index = seq_along(paper_dates), 
+  date = paper_dates, 
+  year_continuous = lubridate::decimal_date(paper_dates)) %>%
+  dplyr::mutate( 
+    sin_year = sin(year_continuous * 2 * pi), 
+    cos_year = cos(year_continuous * 2 * pi)
+  )
 ```
 
-Compare paper data to LDATS data:
+## Compare the data from Christensen et al. with the included data in `LDATS`
 
 ```{r rodent data comparison}
-
 compare <- rodents[[1]] == paper_dat
@@
 There are 16 rows where the data included in LDATS differs from the paper data.
-This is because the LDATS data is not adjusted to account for trapping effort, but the paper data has divided all census counts by the actual number of plots trapped and multiplied by 8 to account for incompletely-trapped censuses.
+This is because the LDATS data is not adjusted to account for trapping effort, while the paper data does, by dividing all census counts by the actual number of plots trapped and multiplying by 8 to account for incompletely-trapped censuses.
 
@@
 
-```
+```{}
   # retrieve data on number of plots trapped per month
@@
   }
-
-    return(r_table_adjusted)
-
 ```
 
-Running the same procedure on the LDATS data, we modify 16 rows, and finish with a data frame that matches the one from the paper.
+We can run the same procedure on the LDATS data to verify that we obtain a data.frame that matches.
 
-```{r adjust LDATS data after Christensen et al, eval = F}
+```{r adjust LDATS data after Christensen et al, eval = TRUE}
+# get the trapping effort for each sample
+trap_table <- read.csv(file.path(vignette_files, "Portal_rodent_trapping.csv"))
+trap_table_controls <- dplyr::filter(trap_table, plot %in% control_plots)
+nplots_controls <- aggregate(trap_table_controls$sampled, 
+                            by = list(period = trap_table_controls$period), 
+                            FUN = sum)
 
-trap_table = read.csv('https://raw.githubusercontent.com/weecology/PortalData/master/Rodents/Portal_rodent_trapping.csv')
-
+# adjust species counts by number of plots trapped that month
+#   divide by number of control plots actually trapped (should be 8) and 
+#   multiply by 8 to estimate captures as if all plots were trapped
+ldats_rodents_adjusted <- as.data.frame.matrix(rodents[[1]])
+ldats_rodents_adjusted[periods, ] <- round(ldats_rodents_adjusted[periods, ] / nplots_controls$x[periods] * 8)
 ```
 
-```{r continue adjust LDATS after paper}
- trap_table_controls =dplyr::filter(trap_table, plot %in% c(2,4,8,11,12,14,17,22))
-  nplots_controls = aggregate(trap_table_controls$sampled,by=list(period = trap_table_controls$period),FUN=sum)
-   # adjust species counts by number of plots trapped that month
-  ldats_rodents_adjusted = as.data.frame.matrix(rodents[[1]])
-  for (n in 1:436) {
-    #divide by number of control plots actually trapped (should be 8) and multiply by 8 to estimate captures as if all plots were trapped
-    ldats_rodents_adjusted[n,] = round(ldats_rodents_adjusted[n,]/nplots_controls$x[n]*8)
-  }
+Now we can compare the adjusted ldats dataset with both the original ldats dataset and the dataset from the paper:
 
+```{r dataset comparisons}
 compare_raw <- rodents[[1]] == ldats_rodents_adjusted
@@
 length(which(rowSums(compare_adjusted) < ncol(compare_adjusted)))
-
-```
-
-```{r data adjustment cleanup, include = FALSE}
-rm(list = c('compare', 'compare_adjusted', 'compare_raw', 'n', 'ldats_rodents_adjusted', 'nplots_controls', 'trap_table', 'trap_table_controls'))
 ```
@@
 ```{r add dates to covariate table}
-
 head(rodents$document_covariate_table)
 
-moondat <- dplyr::select(moondat, newmoonnumber, censusdate)
-colnames(moondat) <- c('newmoon', 'censusdate')
-
-new_cov_table <- dplyr::left_join(rodents$document_covariate_table, moondat, by = 'newmoon')
+new_cov_table <- dplyr::left_join(rodents$document_covariate_table, 
+                                  dplyr::select(moondat, newmoonnumber, censusdate), 
+                                  by = c("newmoon" = "newmoonnumber")) %>%
+  dplyr::rename(date = censusdate)
 
 rodents$document_covariate_table <- new_cov_table
-
 ```
 
-```{r cleanup dates, include = F}
-rm(moondat)
-rm(new_cov_table)
-```
-
-## Stage 1: LDA
+# Identify community groups using LDA
 
@@
 
-```{r LDATS LDAs, eval = F}
+First, we run the LDA models from LDATS to identify the number of topics:
 
-ldats_ldas <- LDATS::LDA_set(document_term_table = rodents$document_term_table, topics = c(2:6), nseeds = 100)
-ldats_lda_selected <- LDATS::select_LDA(LDA_models = ldats_ldas)
+```{r LDATS LDAs, eval = params$run_models}
+ldats_ldas <- LDATS::LDA_set(document_term_table = rodents$document_term_table, 
+                             topics = 2:6, nseeds = nseeds)
+ldats_ldamodel <- LDATS::select_LDA(LDA_models = ldats_ldas)[[1]]
 
-save(ldats_ldas, file = here::here('vignettes', 'christensenetal-comparison-files', 'model-cache', 'ldats_ldas.Rds'))
-save(ldats_lda_selected, file = here::here('vignettes', 'christensenetal-comparison-files', 'model-cache', 'ldats_lda_selected.Rds'))
-
+saveRDS(ldats_ldamodel, file = file.path(vignette_files, "ldats_ldamodel.RDS"))
 ```
 
-Paper LDAS:
+Second, we run the LDA models from Christensen et al. to do the same task:
 
-```{r create dat for paper lda, include = F}
-dat = paper_dat
-```
+```{r paper LDAs, eval = params$run_models}
+source(file.path(vignette_files, "AIC_model_selection.R"))
+source(file.path(vignette_files, "LDA-distance.R"))
 
-```{r paper LDAs, eval = F}
-source(here::here('vignettes', 'christensenetal-comparison-files', 'christensenetal-2018-files', 'AIC_model_selection.R'))
-source(here::here('vignettes', 'christensenetal-comparison-files', 'christensenetal-2018-files', 'LDA-distance.R'))
+# Some of the functions require the data to be stored in the `dat` variable
+dat <- paper_dat
 
@@
 # Only use even numbers for seeds because consecutive seeds give identical results
-seeds = 2*seq(200)
+seeds <- 2 * seq(nseeds)
 
 # repeat LDA model fit and AIC calculation with a bunch of different seeds to test robustness of the analysis
-best_ntopic = repeat_VEM(paper_dat,
+best_ntopic <- repeat_VEM(paper_dat,
                           seeds,
-                         topic_min=2,
-                         topic_max=6)
-hist(best_ntopic$k,breaks=c(0.5,1.5,2.5,3.5,4.5,5.5,6.5,7.5,8.5,9.5),xlab='best # of topics', main='')
+                          topic_min = 2,
+                          topic_max = 6)
+hist(best_ntopic$k, breaks = seq(from = 0.5, to = 9.5), 
+     xlab = "best # of topics", main = "")
 
@@
 # get the best 100 seeds where 4 topics was the best LDA model
-seeds_4topics = best_ntopic %>%
+seeds_4topics <- best_ntopic %>%
   filter(k == 4) %>%
   arrange(aic) %>%
-  head(100) %>%
+  head(min(100, nseeds)) %>%
   pull(SEED)
@@
 # choose seed with highest log likelihood for all following analyses
-#    (also produces plot of community composition for 'best' run compared to 'worst')
-best_seed = calculate_LDA_distance(paper_dat,seeds_4topics)
-mean_dist = unlist(best_seed)[2]
-max_dist = unlist(best_seed)[3]
+#    (also produces plot of community composition for "best" run compared to "worst")
+
+png(file.path(vignette_files, "lda_distances.png"), width = 800, height = 400)
+dat <- paper_dat # calculate_LDA_distance has some required named variables
+best_seed <- calculate_LDA_distance(paper_dat, seeds_4topics)
+dev.off()
+mean_dist <- unlist(best_seed)[2]
+max_dist <- unlist(best_seed)[3]
 
@@
 # ==================================================================
-ntopics = 4
-SEED = unlist(best_seed)[1]  # For the paper, use seed 206
-ldamodel = LDA(paper_dat,ntopics, control = list(seed = SEED),method='VEM')
+ntopics <- 4
+SEED <- unlist(best_seed)[1]  # For the paper, use seed 206
+ldamodel <- LDA(paper_dat, ntopics, control = list(seed = SEED), method = "VEM")
 
-save(ldamodel,file = here::here('vignettes', 'christensenetal-comparison-files', 'model-cache', 'paper_ldamodel.Rds'))
+saveRDS(ldamodel, file = file.path(vignette_files, "paper_ldamodel.RDS"))
 ```
 
-### Plot LDAS
-
-#### Paper LDA
-```{r plot paper LDA, fig.width=7, fig.height=6}
-# Paper
-plot(ldamodel, cols = NULL, option = "D")
-
+```{r}
+knitr::include_graphics(file.path(vignette_files, "lda_distances.png"))
 ```
 
-#### LDATS LDA
-```{r plot LDATS LDA, fig.width=7, fig.height=6}
-# LDATS
-plot(ldats_lda_selected[[1]], cols = NULL, option = "D")
+## Plots
+
+To visualize the LDA assignment of species to topics, we load in the saved LDA models from previously:
+```{r}
+ldamodel <- readRDS(file.path(vignette_files, "paper_ldamodel.RDS"))
+ldats_ldamodel <- readRDS(file.path(vignette_files, "ldats_ldamodel.RDS"))
+```
+
+How the paper LDA model assigns species to topics:
+```{r plot paper LDA, fig.width = 7, fig.height = 6}
+plot(ldamodel, cols = NULL, option = "D")
+```
+
+How the LDATS LDA model assigns species to topics:
+```{r plot LDATS LDA, fig.width = 7, fig.height = 6}
+plot(ldats_ldamodel, cols = NULL, option = "D")
 ```
@@
 This is because of an update to the model selection procedure.
-The paper conservatively overestimates the number of parameters (by counting all of the variational parameteres) and therefore overpenalizes the AIC for models with more topics.
+The paper conservatively overestimates the number of parameters (by counting all of the variational parameters) and therefore overpenalizes the AIC for models with more topics.
 Comparatively, the LDATS method now uses the number of parameters remaining after the variational approximation, as returned by the LDA object.
@@
 
-## Changepoint models
+# Changepoint models
 
@@
 
-### Running paper changepoint models
+## Running paper changepoint models
 
-#### LDATS LDA and paper changepoint
+We define a few helper functions for running the changepoints model of Christensen et al. and processing the output to obtain the dates:
 
-```{r run LDATS LDA and paper cpt, eval = F}
+```{r paper changepoint models}
 #### Run changepoint ####
-source(here::here('vignettes', 'christensenetal-comparison-files', 'christensenetal-2018-files', 'changepointmodel.r'))
+source(file.path(vignette_files, "changepointmodel.r"))
 
+find_changepoints <- function(lda_model, paper_covariates, n_changepoints = 1:6)
+{
   # set up parameters for model
-year_continuous = 1970 + as.integer(julian(paper_dates)) / 365.25
-x = data.frame(
-  year_continuous = year_continuous,
-  sin_year = sin(year_continuous * 2 * pi),
-  cos_year = cos(year_continuous * 2 * pi)
-)
+  x <- dplyr::select(paper_covariates, 
+                     year_continuous, 
+                     sin_year, 
+                     cos_year)
 
   # run models with 1, 2, 3, 4, 5, 6 changepoints
-ldats_paper_cp_results_rodent = changepoint_model(ldats_lda_selected[[1]], x, 1, weights = rep(1,length(year_continuous)))
-save(ldats_paper_cp_results_rodent, file = here::here('vignettes', 'christensenetal-comparison-files', 'model-cache', 'ldats_paper_1.Rds'))
-rm(ldats_paper_cp_results_rodent)
+  cpt_results <- data.frame(n_changepoints = n_changepoints)
+  cpt_results$cpt_model <- lapply(cpt_results$n_changepoints,
+                                  function(n_changepoints)
+                                  {
+                                    changepoint_model(lda_model, x, n_changepoints, maxit = nit, 
+                                                      weights = rep(1, NROW(x)))
+                                  })
+  return(cpt_results)
+}
 
-ldats_paper_cp_results_rodent2 = changepoint_model(ldats_lda_selected[[1]], x, 2, weights = rep(1,length(year_continuous)))
-save(ldats_paper_cp_results_rodent2, file = here::here('vignettes', 'christensenetal-comparison-files', 'model-cache', 'ldats_paper_2.Rds'))
-rm(ldats_paper_cp_results_rodent2)
+# Among a selection of models with different # of changepoints, 
+#   - compute AIC
+#   - select the model with the best AIC
+#   - get the posterior distributions for the changepoints
+select_cpt_model <- function(cpt_results, ntopics)
+{
+  # compute log likelihood as the mean deviance
+  cpt_results$mean_deviances <- vapply(cpt_results$cpt_model, 
+                                       function(cpt_model) {mean(cpt_model$saved_lls)}, 
+                                       0)
 
-ldats_paper_cp_results_rodent3 = changepoint_model(ldats_lda_selected[[1]], x, 3, weights = rep(1,length(year_continuous)))
-save(ldats_paper_cp_results_rodent3, file = here::here('vignettes', 'christensenetal-comparison-files', 'model-cache', 'ldats_paper_3.Rds'))
-rm(ldats_paper_cp_results_rodent3)
+  # compute AIC = ( -2 * log likelihood) + 2 * (#parameters)
+  cpt_results$AIC <- cpt_results$mean_deviances * -2 + 
+    2 * (3 * (ntopics - 1) * (cpt_results$n_changepoints + 1) +
+           (cpt_results$n_changepoints))
 
-ldats_paper_cp_results_rodent4 = changepoint_model(ldats_lda_selected[[1]], x, 4, weights = rep(1,length(year_continuous)))
-save(ldats_paper_cp_results_rodent4, file = here::here('vignettes', 'christensenetal-comparison-files', 'model-cache', 'ldats_paper_4.Rds'))
-rm(ldats_paper_cp_results_rodent4)
+  # select the best model
+  cpt <- cpt_results$cpt_model[[which.min(cpt_results$AIC)]]
+  return(cpt)
+}
 
-ldats_paper_cp_results_rodent5 = changepoint_model(ldats_lda_selected[[1]], x, 5, weights = rep(1,length(year_continuous)))
-save(ldats_paper_cp_results_rodent5, file = here::here('vignettes', 'christensenetal-comparison-files', 'model-cache', 'ldats_paper_5.Rds'))
-rm(ldats_paper_cp_results_rodent5)
-
-ldats_paper_cp_results_rodent6 = changepoint_model(ldats_lda_selected[[1]], x, 6, weights = rep(1,length(year_continuous)))
-save(ldats_paper_cp_results_rodent6, file = here::here('vignettes', 'christensenetal-comparison-files', 'model-cache', 'ldats_paper_6.Rds'))
-rm(ldats_paper_cp_results_rodent6)
+# transform the output from `compute_cpt` and match up the time indices with 
+#   dates from the original data
+get_dates <- function(cpt, covariates = paper_covariates)
+{
+  cpt$saved[,1,] %>%
+    t() %>%
+    as.data.frame() %>%
+    reshape::melt() %>%
+    dplyr::left_join(covariates, by = c("value" = "index"))
+}
 ```
 
-```{r select LDATS LDA and paper cpt, eval = F}
-#### Changepoint model selection ####
-load(here::here('vignettes', 'christensenetal-comparison-files', 'model-cache', 'ldats_paper_1.Rds'))
-load(here::here('vignettes', 'christensenetal-comparison-files', 'model-cache', 'ldats_paper_2.Rds'))
-load(here::here('vignettes', 'christensenetal-comparison-files', 'model-cache', 'ldats_paper_3.Rds'))
-load(here::here('vignettes', 'christensenetal-comparison-files', 'model-cache', 'ldats_paper_4.Rds'))
-load(here::here('vignettes', 'christensenetal-comparison-files', 'model-cache', 'ldats_paper_5.Rds'))
-load(here::here('vignettes', 'christensenetal-comparison-files', 'model-cache', 'ldats_paper_6.Rds'))
+### LDATS LDA and paper changepoint
 
+Run the Christensen et al. time series model to identify changepoints on the LDA topics selected by LDATS:
+```{r run LDATS LDA and paper cpt, eval = params$run_models}
+ldats_paper_results <- find_changepoints(ldats_ldamodel, paper_covariates)
+
+saveRDS(ldats_paper_results, file = file.path(vignette_files, "ldats_paper.RDS"))
 ```
 
-```{r select LDATS-paper, eval = T}
-ntopics = ldats_lda_selected[[1]]@k
+Extract the dates of the changepoints:
+```{r compute changepoints for LDATS LDA and paper cpt}
+ldats_paper_results <- readRDS(file.path(vignette_files, "ldats_paper.RDS"))
 
-# change point model selection
-# mean deviance ( -2 * log likelihood) + 2*(#parameters)
-mean(ldats_paper_cp_results_rodent$saved_lls * -2) + 2*(3*(ntopics-1)*(1+1)+(1))
-mean(ldats_paper_cp_results_rodent2$saved_lls * -2)+ 2*(3*(ntopics-1)*(2+1)+(2))
-mean(ldats_paper_cp_results_rodent3$saved_lls * -2)+ 2*(3*(ntopics-1)*(3+1)+(3))
-mean(ldats_paper_cp_results_rodent4$saved_lls * -2)+ 2*(3*(ntopics-1)*(4+1)+(4))
-mean(ldats_paper_cp_results_rodent5$saved_lls * -2)+ 2*(3*(ntopics-1)*(5+1)+(5))
-mean(ldats_paper_cp_results_rodent6$saved_lls * -2)+ 2*(3*(ntopics-1)*(6+1)+(6))
-
-# The lowest deviance is for 4 changepoints.
-
-ldats_paper_cpt_selected = ldats_paper_cp_results_rodent4
-
-df_4 = as.data.frame(t(ldats_paper_cpt_selected$saved[,1,])) %>% reshape::melt()
-year_continuous = 1970 + as.integer(julian(paper_dates)) / 365.25
-df_4$value = year_continuous[df_4$value]
-
-ldats_paper_cpt_dates <- vector(length = 4)
-ldats_paper_cpt_dates[1] <- mean(df_4[df_4$variable=='V1','value']) %>% lubridate::date_decimal() %>% format('%Y-%m-%d')
-ldats_paper_cpt_dates[2] <-mean(df_4[df_4$variable=='V2','value']) %>% lubridate::date_decimal() %>% format('%Y-%m-%d')
-ldats_paper_cpt_dates[3] <-mean(df_4[df_4$variable=='V3','value']) %>% lubridate::date_decimal() %>% format('%Y-%m-%d')
-ldats_paper_cpt_dates[4] <-mean(df_4[df_4$variable=='V4','value']) %>% lubridate::date_decimal() %>% format('%Y-%m-%d')
-
-ldats_paper_cpt_dates
-
+ldats_paper_cpt <- select_cpt_model(ldats_paper_results, 
+                                    ntopics = ldats_ldamodel@k)
+ldats_paper_cpt_dates <- get_dates(ldats_paper_cpt)
 ```
 
-#### Paper LDA and paper changepoint
+### Paper LDA and paper changepoint
 
-```{r load paper LDA for paper cpt, eval = F}
+Run the Christensen et al. time series model to identify changepoints on the LDA topics selected by Christensen et al.:
+```{r run paper LDA and paper cpt, eval = params$run_models}
+paper_paper_results <- find_changepoints(ldamodel, paper_covariates)
 
-#### Load LDA ####
-load(here::here('vignettes', 'christensenetal-comparison-files', 'model-cache', 'paper_ldamodel.Rds'))
-
+saveRDS(paper_paper_results, file = file.path(vignette_files, "paper_paper.RDS"))
 ```
 
-```{r run paper LDA and paper cpt, eval = F}
-#### Run changepoint ####
-source(here::here('vignettes', 'christensenetal-comparison-files', 'christensenetal-2018-files', 'changepointmodel.r'))
+Extract the dates of the changepoints:
+```{r compute changepoints for paper LDA and paper cpt}
+paper_paper_results <- readRDS(file.path(vignette_files, "paper_paper.RDS"))
 
-# set up parameters for model
-year_continuous = 1970 + as.integer(julian(paper_dates)) / 365.25
-x = data.frame(
-  year_continuous = year_continuous,
-  sin_year = sin(year_continuous * 2 * pi),
-  cos_year = cos(year_continuous * 2 * pi)
-)
-
-# run models with 1, 2, 3, 4, 5, 6 changepoints
-paper_paper_cp_results_rodent = changepoint_model(ldamodel, x, 1, weights = rep(1,length(year_continuous)))
-save(paper_paper_cp_results_rodent, file = here::here('vignettes', 'christensenetal-comparison-files', 'model-cache', 'paper_paper_1.Rds'))
-rm(paper_paper_cp_results_rodent)
-
-paper_paper_cp_results_rodent2 = changepoint_model(ldamodel, x, 2, weights = rep(1,length(year_continuous)))
-save(paper_paper_cp_results_rodent2, file = here::here('vignettes', 'christensenetal-comparison-files', 'model-cache', 'paper_paper_2.Rds'))
-rm(paper_paper_cp_results_rodent2)
-
-paper_paper_cp_results_rodent3 = changepoint_model(ldamodel, x, 3, weights = rep(1,length(year_continuous)))
-save(paper_paper_cp_results_rodent3, file = here::here('vignettes', 'christensenetal-comparison-files', 'model-cache', 'paper_paper_3.Rds'))
-rm(paper_paper_cp_results_rodent3)
-
-paper_paper_cp_results_rodent4 = changepoint_model(ldamodel, x, 4, weights = rep(1,length(year_continuous)))
-save(paper_paper_cp_results_rodent4, file = here::here('vignettes', 'christensenetal-comparison-files', 'model-cache', 'paper_paper_4.Rds'))
-rm(paper_paper_cp_results_rodent4)
-
-paper_paper_cp_results_rodent5 = changepoint_model(ldamodel, x, 5, weights = rep(1,length(year_continuous)))
-save(paper_paper_cp_results_rodent5, file = here::here('vignettes', 'christensenetal-comparison-files', 'model-cache', 'paper_paper_5.Rds'))
-rm(paper_paper_cp_results_rodent5)
-
-paper_paper_cp_results_rodent6 = changepoint_model(ldamodel, x, 6, weights = rep(1,length(year_continuous)))
-save(paper_paper_cp_results_rodent6, file = here::here('vignettes', 'christensenetal-comparison-files', 'model-cache', 'paper_paper_6.Rds'))
-rm(paper_paper_cp_results_rodent6)
+paper_paper_cpt <- select_cpt_model(paper_paper_results, 
+                                    ntopics = ldamodel@k)
+paper_paper_cpt_dates <- get_dates(ldats_paper_cpt)
 ```
 
-```{r select paper LDA and paper cpt, eval = F}
-#### Changepoint model selection ####
-load(here::here('vignettes', 'christensenetal-comparison-files', 'model-cache', 'paper_paper_1.Rds'))
-load(here::here('vignettes', 'christensenetal-comparison-files', 'model-cache', 'paper_paper_2.Rds'))
-load(here::here('vignettes', 'christensenetal-comparison-files', 'model-cache', 'paper_paper_3.Rds'))
-load(here::here('vignettes', 'christensenetal-comparison-files', 'model-cache', 'paper_paper_4.Rds'))
-load(here::here('vignettes', 'christensenetal-comparison-files', 'model-cache', 'paper_paper_5.Rds'))
-load(here::here('vignettes', 'christensenetal-comparison-files', 'model-cache', 'paper_paper_6.Rds'))
-```
+## Running LDATS changepoint models
 
-```{r select paper-paper, eval = T}
-ntopics = ldamodel@k
+### LDATS LDA and LDATS changepoint
 
-# change point model selection
-# mean deviance ( -2 * log likelihood) + 2*(#parameters)
-mean(paper_paper_cp_results_rodent$saved_lls * -2) + 2*(3*(ntopics-1)*(1+1)+(1))
-mean(paper_paper_cp_results_rodent2$saved_lls * -2)+ 2*(3*(ntopics-1)*(2+1)+(2))
-mean(paper_paper_cp_results_rodent3$saved_lls * -2)+ 2*(3*(ntopics-1)*(3+1)+(3))
-mean(paper_paper_cp_results_rodent4$saved_lls * -2)+ 2*(3*(ntopics-1)*(4+1)+(4))
-mean(paper_paper_cp_results_rodent5$saved_lls * -2)+ 2*(3*(ntopics-1)*(5+1)+(5))
-mean(paper_paper_cp_results_rodent6$saved_lls * -2)+ 2*(3*(ntopics-1)*(6+1)+(6))
+Run the LDATS time series model to identify changepoints on the LDA topics selected by LDATS:
 
-# The lowest deviance is for 4 changepoints.
-
-paper_paper_cpt_selected = paper_paper_cp_results_rodent4
-
-df_4 = as.data.frame(t(paper_paper_cpt_selected$saved[,1,])) %>%  reshape::melt()
-df_4$value = year_continuous[df_4$value]
-
-paper_paper_cpt_dates <- vector(length = 4)
-paper_paper_cpt_dates[1] <- mean(df_4[df_4$variable=='V1','value']) %>% lubridate::date_decimal() %>% format('%Y-%m-%d')
-paper_paper_cpt_dates[2] <-mean(df_4[df_4$variable=='V2','value']) %>% lubridate::date_decimal() %>% format('%Y-%m-%d')
-paper_paper_cpt_dates[3] <-mean(df_4[df_4$variable=='V3','value']) %>% lubridate::date_decimal() %>% format('%Y-%m-%d')
-paper_paper_cpt_dates[4] <-mean(df_4[df_4$variable=='V4','value']) %>% lubridate::date_decimal() %>% format('%Y-%m-%d')
-
-paper_paper_cpt_dates
-
-```
-
-### Running LDATS changepoint models
-
-#### LDATS LDA and LDATS changepoint
-
-#### Run the TS models
-
-```{r run LDATS LDA and LDATS changepoint, eval = F}
-#### Run LDATS changepoint ####
-
-ldats_ldats_cpt <- TS_on_LDA(LDA_models = ldats_lda_selected,
+```{r run LDATS LDA and LDATS cpt, eval = params$run_models}
+ldats_ldats_results <- TS_on_LDA(LDA_models = ldats_ldamodel,
                                  document_covariate_table = rodents$document_covariate_table,
@@
                                  nchangepoints = 1:6,
-                             timename = 'newmoon',
-                             weights = NULL)
+                                 timename = "newmoon",
+                                 weights = NULL,
+                                 control = list(nit = nit))
 
-save(ldats_ldats_cpt, file = here::here('vignettes', 'christensenetal-comparison-files', 'model-cache', 'ldats_ldats.Rds'))
-rm(ldats_ldats_cpt)
-rm(ldats_lda_selected)
+saveRDS(ldats_ldats_results, file = file.path(vignette_files, "ldats_ldats.RDS"))
 ```
 
-### Select model
+Unlike the paper changepoint model, LDATS can recognize that sampling periods may not be equidistant, and can place changepoint estimates at new moons if they fall between nonconsecutive sampling periods.
+We can estimate the dates corresponding to those new moons, extrapolating from the census dates for adjacent census periods.
 
-```{r select ldats lda + ldats cpt}
+```{r construct lookup table for LDATS output for changepoint times}
+# make the full sequence of possible newmoon values
+full_index <- seq(min(rodents$document_covariate_table$newmoon), 
+                  max(rodents$document_covariate_table$newmoon))
 
-ldats_ldats_cpt_selected <- select_TS(ldats_ldats_cpt)
-
-ldats_ldats_cpt_selected$nchangepoints
-
+# generate a lookup table with dates for the newmoons, using `approx` to 
+#   linearly interpolate the missing values
+ldats_dates <- approx(rodents$document_covariate_table$newmoon, 
+                     as.Date(rodents$document_covariate_table$date), 
+                     full_index) %>%
+  as.data.frame() %>%
+  mutate(index = x, 
+         date = as.Date(y, origin = "1970-01-01")) %>%
+  select(index, date)
 ```
 
-### Get dates associated with changepoints
+Select the best time series model and extract the dates of the changepoints:
 
-```{r get ldats lda + ldats cpt dates}
-ldats_ldats_cpt_dates <- ldats_ldats_cpt_selected$rho_summary$Mean
-ldats_ldats_cpt_dates <- round(ldats_ldats_cpt_dates)
+```{r compute changepoints for LDATS LDA and LDATS cpt}
+ldats_ldats_results <- readRDS(file.path(vignette_files, "ldats_ldats.RDS"))
 
-ldats_ldats_cpt_dates <- as.data.frame(ldats_ldats_cpt_dates)
-colnames(ldats_ldats_cpt_dates) <- 'newmoon'
-ldats_ldats_cpt_dates <- dplyr::left_join(ldats_ldats_cpt_dates, rodents$document_covariate_table, by = 'newmoon')
-ldats_ldats_cpt_dates <- ldats_ldats_cpt_dates$censusdate
+ldats_ldats_cpt <- select_TS(ldats_ldats_results)
 
+ldats_ldats_cpt_dates <- ldats_ldats_cpt$rhos %>%
+  as.data.frame() %>%
+  reshape::melt() %>%
+  dplyr::left_join(ldats_dates, by = c("value" = "index"))
 ```
@@
 
-```{r load paper LDA for LDATS cpt, eval = F}
+Run the LDATS time series model to identify changepoints on the LDA topics selected by Christensen et al.:
 
-#### Load LDA ####
-load(here::here('vignettes', 'christensenetal-comparison-files', 'model-cache', 'paper_ldamodel.Rds'))
-
-```
-#### Run the TS models
-
-```{r run paper LDA and LDATS changepoint, eval = F}
-#### Run LDATS changepoint ####
-
-paper_ldats_cpt <- TS_on_LDA(LDA_models = ldamodel,
+```{r run paper LDA and LDATS cpt, eval = params$run_models}
+paper_ldats_results <- TS_on_LDA(LDA_models = ldamodel,
                              document_covariate_table = rodents$document_covariate_table,
@@
                              nchangepoints = 1:6,
-                             timename = 'newmoon',
-                             weights = NULL)
 
-save(paper_ldats_cpt, file = here::here('vignettes', 'christensenetal-comparison-files', 'model-cache', 'paper_ldats.Rds'))
-rm(paper_ldats_cpt)
-rm(ldamodel)
+                             timename = "newmoon",
+                             weights = NULL,
+                             control = list(nit = nit))
+
+saveRDS(paper_ldats_results, file = file.path(vignette_files, "paper_ldats.RDS"))
 ```
 
-### Select model
+Select the best time series model and extract the dates of the changepoints:
 
 ```{r select paper lda + ldats cpt}
+paper_ldats_results <- readRDS(file.path(vignette_files, "paper_ldats.RDS"))
 
-paper_ldats_cpt_selected <- select_TS(paper_ldats_cpt)
+paper_ldats_cpt <- select_TS(paper_ldats_results)
 
-paper_ldats_cpt_selected$nchangepoints
-
+paper_ldats_cpt_dates <- paper_ldats_cpt$rhos %>%
+  as.data.frame() %>%
+  reshape::melt() %>%
+  dplyr::left_join(ldats_dates, by = c("value" = "index"))
 ```
 
-#### Get dates
+## How many changepoints were identified?
 
-Unlike the paper changepoint model, LDATS can recognize that sampling periods may not be equidistant, and can place changepoint estimates at new moons if they fall between nonconsecutive sampling periods.
-We can estimate the dates corresponding to those new moons, extrapolating from the census dates for adjacent census periods.
-
-```{r get paper lda + ldats cpt dates}
-paper_ldats_cpt_dates <- paper_ldats_cpt_selected$rho_summary$Mean
-paper_ldats_cpt_dates <- round(paper_ldats_cpt_dates)
-
-paper_ldats_cpt_dates <- as.data.frame(paper_ldats_cpt_dates)
-colnames(paper_ldats_cpt_dates) <- 'newmoon'
-paper_ldats_cpt_dates <- dplyr::left_join(paper_ldats_cpt_dates, rodents$document_covariate_table, by = 'newmoon')
-
-halfinterval = as.integer((as.Date(rodents$document_covariate_table$censusdate[ which(rodents$document_covariate_table$newmoon == paper_ldats_cpt_dates$newmoon[1] + 1)]) - as.Date(rodents$document_covariate_table$censusdate[ which(rodents$document_covariate_table$newmoon == paper_ldats_cpt_dates$newmoon[1] - 1)]))/2)
-
-newdate <- lubridate::ymd(rodents$document_covariate_table$censusdate[ which(rodents$document_covariate_table$newmoon == paper_ldats_cpt_dates$newmoon[1] -1 )]) + lubridate::days(halfinterval)
-
-paper_ldats_cpt_dates$censusdate[1] <- as.character(newdate)
-
-halfinterval =as.integer((as.Date(rodents$document_covariate_table$censusdate[ which(rodents$document_covariate_table$newmoon == paper_ldats_cpt_dates$newmoon[3] + 2)]) - as.Date(rodents$document_covariate_table$censusdate[ which(rodents$document_covariate_table$newmoon == paper_ldats_cpt_dates$newmoon[3] - 1)]))/3)
-
-newdate <- lubridate::ymd(rodents$document_covariate_table$censusdate[ which(rodents$document_covariate_table$newmoon == paper_ldats_cpt_dates$newmoon[3] -2 )]) + lubridate::days(halfinterval)
-
-paper_ldats_cpt_dates$censusdate[3] <- as.character(newdate)
-
-paper_ldats_cpt_dates <- paper_ldats_cpt_dates$censusdate
+```{r}
+nlevels(ldats_paper_cpt_dates$variable)
+nlevels(paper_paper_cpt_dates$variable)
+nlevels(ldats_ldats_cpt_dates$variable)
+nlevels(paper_ldats_cpt_dates$variable)
 ```
@@
 
-### Plot changepoint models
+## Plot changepoint models
 
-```{r load cpts to plot, include = F}
+### Paper LDA and LDATS changepoint
 
-year_continuous = 1970 + as.integer(julian(paper_dates)) / 365.25
-x = data.frame(
-  year_continuous = year_continuous,
-  sin_year = sin(year_continuous * 2 * pi),
-  cos_year = cos(year_continuous * 2 * pi)
-)
-ntopics = 4
-
+```{r plot paper LDA and LDATS cpts, fig.width = 7, fig.height = 6}
+plot(paper_ldats_cpt)
 ```
 
-#### Paper LDA and LDATS changepoint
-
-```{r plot paper LDA and LDATS cpts, fig.width=7, fig.height=6}
-
-plot(paper_ldats_cpt_selected)
-
+### LDATS LDA and LDATS changepoint
+```{r plot ldats LDA and LDATS cpt, fig.width = 7, fig.height = 6}
+plot(ldats_ldats_cpt)
 ```
 
-#### LDATS LDA and LDATS changepoint
-```{r plot ldats LDA and LDATS cpt, fig.width=7, fig.height=6}
-plot(ldats_ldats_cpt_selected)
+### Paper LDA and paper changepoint
+```{r plot paper LDA and paper cpt, fig.width = 7, fig.height = 6}
+paper_cpts <- find_changepoint_location(paper_paper_cpt)
+ntopics <- ldamodel@k
+
+paper_cpt_plot <- get_ll_non_memoized_plot(ldamodel, paper_covariates, paper_cpts, make_plot = TRUE,
+                                           weights = rep(1, NROW(paper_covariates)))
+
+annual_hist(paper_paper_cpt, paper_covariates$year_continuous)
+paper_cpt_plot
 ```
 
-#### Paper LDA and paper changepoint
-```{r plot paper LDA and paper cpt, fig.width=7, fig.height=6}
+### LDATS LDA and paper changepoint
+```{r plot LDATS lda and paper cpt, fig.width = 7, fig.height = 6}
+ldats_cpts <- find_changepoint_location(ldats_paper_cpt)
+ntopics <- ldats_ldamodel@k
 
-source(here::here('vignettes', 'christensenetal-comparison-files', 'christensenetal-2018-files', 'LDA_figure_scripts.R'))
-source(here::here('vignettes', 'christensenetal-comparison-files', 'christensenetal-2018-files', 'changepointmodel.r'))
-paper_cpts = find_changepoint_location(paper_paper_cpt_selected)
-paper_cpt_plot = get_ll_non_memoized_plot(ldamodel,x,paper_cpts,make_plot=T,weights=rep(1,length(year_continuous)))
-annual_hist(paper_paper_cpt_selected, year_continuous)
-paper_cpt_plot
+ldats_cpt_plot <- get_ll_non_memoized_plot(ldats_ldamodel, paper_covariates, ldats_cpts, make_plot = TRUE,
+                                           weights = rep(1, NROW(paper_covariates)))
 
-```
-
-#### LDATS LDA and paper changepoint
-```{r plot LDATS lda and paper cpt, fig.width=7, fig.height=6}
-ntopics = 6
-ldats_cpts = find_changepoint_location(ldats_paper_cpt_selected)
-ldats_cpt_plot = get_ll_non_memoized_plot(ldats_lda_selected[[1]],x,ldats_cpts,make_plot=T,weights=rep(1,length(year_continuous)))
-annual_hist(ldats_paper_cpt_selected, year_continuous)
+annual_hist(ldats_paper_cpt, paper_covariates$year_continuous)
 ldats_cpt_plot
@@
 
-The results of the changepoint model appear extremely robust to both choice of LDA model and choice of changepoint model.
+The results of the changepoint model appear robust to both choice of LDA model and choice of changepoint model.
 
-### Report changepoint dates
+## Report changepoint dates
 ```{r report cpt dates, include = F}
-
-cpt_dates <- as.data.frame(paper_paper_cpt_dates)
-cpt_dates <- cbind(cpt_dates, ldats_paper_cpt_dates)
-cpt_dates <- cbind(cpt_dates, ldats_ldats_cpt_dates)
-cpt_dates <- cbind(cpt_dates, paper_ldats_cpt_dates)
-
-colnames(cpt_dates) <- c('paperLDA_papercpt', 'ldatsLDA_papercpt', 'ldatsLDA_ldatscpt', 'paperLDA_ldatscpt')
-
+cpt_dates <- dplyr::bind_rows("paperLDA_paperCPT" = paper_paper_cpt_dates, 
+                              "ldatsLDA_paperCPT" = ldats_paper_cpt_dates, 
+                              "ldatsLDA_ldatsCPT" = ldats_ldats_cpt_dates, 
+                              "paperLDA_ldatsCPT" = paper_ldats_cpt_dates, 
+                              .id = "analysis") %>%
+  dplyr::group_by(analysis, variable) %>%
+  dplyr::summarize(date = mean(date)) %>%
+  dplyr::ungroup() %>%
+  dplyr::rename(changepoint = variable) %>%
+  tidyr::spread(analysis, date)
 ```
@@
 ```{r print cpt dates}
-cpt_dates
+knitr::kable(cpt_dates)
 ```

### 2019-07-24 R/ptMCMC.R
documentation edits

@@ -661,7 +661,7 @@ prep_proposal_dist <- function(nchangepoints, control = list()){
 #' @description \code{prep_saves} creates the data structure used to save the 
 #'   output from each iteration of the ptMCMC algorithm, which is added via
 #'   \code{update_saves}. Once the ptMCMC is complete, the saved data objects
-#'   are then processed (burnin iterations are dropped and the remaining
+#'   are then processed (burn-in iterations are dropped and the remaining
 #'   iterations are thinned) via \code{process_saves}.
 #'   \cr \cr
 #'   This set of functions was designed to work within \code{\link{TS}} and 

### 2020-03-18 not.R
more work

@@ -1,0 +1,23 @@
+files to add to be downloaded
+
+data
+moon_dates.csv
+paper_dat.csv
+paper_dates.csv
+paper_covariates.csv
+
+"model_outputs",
+"ldats_rodents_adjusted.RDS",
+"rodents.RDS",
+"ldats_paper_cpt.RDS",
+"ldats_paper_cpt_dates.RDS",
+"ldats_ldats_cpt.RDS",
+"ldats_ldats_cpt_dates.RDS",
+"paper_paper_cpt.RDS",
+"paper_paper_cpt_dates.RDS",
+"paper_paper_cpt.RDS",
+"paper_paper_cpt_dates.RDS",
+"annual_hist"
+
+make model_outputs just output
+maybe use list files to download everything

### 2020-03-18 doc/rodents-example.Rmd
vignette rebuilds
more tidying

@@
 
-This vignette walks through an example of **LDATS** at the command line and
-was constructed using **LDATS** version `r vers` on `r today`.
+This vignette walks through an example of **`LDATS`** at the command line and
+was constructed using **`LDATS`** version `r vers` on `r today`.
 
@@
 Similarly, the TS models use the LDA compositional breakdown and weight the information from each period according to the total catch.
-We therefore give both models un-adjusted data.
+We therefore give both models unadjusted data.
 The values in the `document_term_table` must be integers.
@@
 
-```{r lda set not quiet}
+```{r lda set not quiet, eval =F}
 lda_model_set2 <- LDA_set(document_term_table = rodents$document_term_table,
@@
 ```{r load lda model set, include = F}
-load(here::here('vignettes', 'rodents-example-files', 'lda_model_set.Rds'))
+load(file.path('.', 'rodents-example-files', 'lda_model_set.Rds'))
 rm(lda_model_set2)
@@
 ```{r reload ts, include = F}
-load(here::here('vignettes', 'rodents-example-files', 'changepoint_models.Rds'))
+load(file.path('.', 'rodents-example-files', 'changepoint_models.Rds'))
 ```
@@
 ```{r load ldats results, include = F}
-load(here::here('vignettes', 'rodents-example-files', 'lda_ts_results.Rds'))
+load(file.path('.', 'rodents-example-files', 'lda_ts_results.Rds'))
 ```

### 2020-03-18 doc/paper-comparison.R

@@ -11,7 +11,7 @@ knitr::opts_chunk$set(
 vers <- packageVersion("LDATS")
 
 ## ---- eval = FALSE-------------------------------------------------------
-#  # install.packages("devtools")
+#  install.packages("devtools")
 #  devtools::install_github("weecology/LDATS")
 
 ## ------------------------------------------------------------------------
@@ -20,57 +20,74 @@ set.seed(42)
 nseeds <- 200
 nit <- 10000
 
+## ---- eval = params$run_models-------------------------------------------
+#  install.packages(c("dplyr", "gridExtra", "multipanel", "RColorBrewer", "RCurl", "reshape2"))
+
 ## ---- eval = FALSE-------------------------------------------------------
 #  rmarkdown::render("paper-comparison.Rmd", params = list(run_models = TRUE))
 
 ## ----set download location-----------------------------------------------
 vignette_files <- tempdir()
 
-## ----download scripts----------------------------------------------------
-test_file <- file.path(vignette_files, "rodent_LDA_analysis.r")
+## ----download scripts and data-------------------------------------------
+test_file <- file.path(vignette_files, "scripts", "rodent_LDA_analysis.r")
 
-if (!file.exists(test_file))
-{
-  # from the Extreme-events-LDA repo
-  github_path <- "https://raw.githubusercontent.com/emchristensen/Extreme-events-LDA/master/"
+if (!file.exists(test_file)){
+
+  # scripts
+  dir.create(file.path(vignette_files, "scripts"), showWarnings = FALSE)
+  github_path <- "https://raw.githubusercontent.com/weecology/LDATS-replications/master/scripts/"
   files_to_download <- c("rodent_LDA_analysis.r", "rodent_data_for_LDA.r", 
                          "AIC_model_selection.R", "changepointmodel.r", 
-                         "LDA-distance.R", "Rodent_table_dat.csv", 
-                         "LDA_figure_scripts.R")
+                         "LDA-distance.R", "LDA_figure_scripts.R")
   
-  for (file in files_to_download)
-  {
+  for (file in files_to_download)  {
     download.file(url = paste0(github_path, file),
-                  destfile = file.path(vignette_files, file))
+                  destfile = file.path(vignette_files, "scripts", file))
   }
+
+    
+  # data
+  dir.create(file.path(vignette_files, "data"), showWarnings = FALSE)
[…]

### 2020-03-18 doc/rodents-example.R

@@ -32,7 +32,7 @@ head(rodents$document_covariate_table, 10)
 #                           nseeds = 2)
 
 ## ----load lda model set, include = F-------------------------------------
-load(here::here('vignettes', 'rodents-example-files', 'lda_model_set.Rds'))
+load(file.path('.', 'rodents-example-files', 'lda_model_set.Rds'))
 rm(lda_model_set2)
 
 ## ----select LDA----------------------------------------------------------
@@ -63,7 +63,7 @@ plot(selected_lda_model[[1]])
 #  
 
 ## ----reload ts, include = F----------------------------------------------
-load(here::here('vignettes', 'rodents-example-files', 'changepoint_models.Rds'))
+load(file.path('.', 'rodents-example-files', 'changepoint_models.Rds'))
 
 ## ----select ts-----------------------------------------------------------
 selected_changepoint_model <- select_TS(changepoint_models)
@@ -93,7 +93,7 @@ plot(selected_changepoint_model)
 #                           control = list(nit = 1000))
 
 ## ----load ldats results, include = F-------------------------------------
-load(here::here('vignettes', 'rodents-example-files', 'lda_ts_results.Rds'))
+load(file.path('.', 'rodents-example-files', 'lda_ts_results.Rds'))
 
 ## ----LDA_TS results------------------------------------------------------
 names(lda_ts_results)

### 2020-03-18 not.R
pkg updating

@@ -1,23 +1,0 @@
-files to add to be downloaded
-
-data
-moon_dates.csv
-paper_dat.csv
-paper_dates.csv
-paper_covariates.csv
-
-"model_outputs",
-"ldats_rodents_adjusted.RDS",
-"rodents.RDS",
-"ldats_paper_cpt.RDS",
-"ldats_paper_cpt_dates.RDS",
-"ldats_ldats_cpt.RDS",
-"ldats_ldats_cpt_dates.RDS",
-"paper_paper_cpt.RDS",
-"paper_paper_cpt_dates.RDS",
-"paper_paper_cpt.RDS",
-"paper_paper_cpt_dates.RDS",
-"annual_hist"
-
-make model_outputs just output
-maybe use list files to download everything

### 2020-03-23 R/LDA.R
organizationals

@@ -7,6 +7,9 @@
 #'     \emph{et al.} 2018). \cr \cr
 #'   \code{prepare_LDA} pre-prepares the LDA model objects for simpler 
 #'     use within the subfunctions. \cr \cr 
+#'   \code{check_LDA} ensures that the inputs are proper. 
+#'     See \code{\link{check_topics}}, \code{\link{check_replicates}},
+#'     and \code{\link{check_control}} for specifics. \cr \cr
 #'   \code{LDA_control} defines and creates the control list used to fit 
 #'     the LDA model. \cr \cr 
 #'   \code{run_LDA} runs (via \code{\link{LDA_call}}) all LDA models
@@ -94,6 +97,8 @@
 #'     used to fit the model. \cr \cr
 #'   \code{prepare_LDA}: \code{list} of \code{list}s, each of which is a
 #'     preliminary model object for an LDA model fit. \cr \cr
+#'   \code{check_LDA}: an error message is thrown if any input is improper, 
+#'     otherwise \code{NULL}.
 #'   \code{LDA_control}: \code{list} of controls for the LDA model, with
 #'     named elements corresponding to the arguments.
 #'   \code{run_LDA}: \code{LDA_set} \code{list} of model results from all
@@ -103,7 +108,7 @@
 #'     run of a \code{<model>} function, such as 
 #'     \code{\link{topicmodels_LDA}}. \cr \cr
 #'   \code{measure_LDA}: \code{vector} of values corresponding to the model
-#'     evaluations.
+#'     evaluations. \cr \cr
 #'   \code{select_LDA}: \code{list} of selected models' \code{list}s.
 #' 
 #' @references 
@@ -143,7 +148,18 @@ LDA <- function(data, topics = 2, replicates = 1, control = list()){
 #'
 #' @export
 #'
+check_LDA <- function(topics = 2, replicates = 1, control = list()){
+  check_topics(topics = topics)
+  check_replicates(replicates = replicates)
+  check_control(control = control)
+}
+
+#' @rdname LDA
+#'
+#' @export
+#'
 prepare_LDA <- function(data, topics = 2, replicates = 1, control = list()){
+  check_LDA(topics = topics, replicates = replicates, control = control)
   control <- do.call("LDA_control", control)
   messageq("----- Linguistic Decomposition Analyses -----", control$quiet)
   data <- conform_data(data = data, control = control)
[…]

### 2020-03-23 R/TS.R

@@ -8,6 +8,12 @@
 #'     \emph{et al.} 2018). \cr \cr
 #'   \code{prepare_TS} pre-prepares the TS model objects for simpler 
 #'     use within the subfunctions. \cr \cr
+#'   \code{check_TS} ensures that the inputs are proper. 
+#'     See \code{\link{check_LDAs}}, 
+#'     \code{\link{check_document_covariate_table}}, 
+#'     code{\link{check_formulas}}, \code{\link{check_nchangepoints}}, 
+#'     \code{\link{check_timename}}, \code{\link{check_weights}}, 
+#'     and \code{\link{check_control}} for specifics. \cr \cr
 #'   \code{TS_control} defines and creates the control \code{list} for the TS
 #'     model running. \cr \cr
 #'   \code{run_TS} runs (via \code{\link{TS_call}}) all TS models
@@ -36,15 +42,6 @@
 #'       the whole model, and
 #'     \item Package the results.
 #'   }
-#'
-#' @param data \code{data.frame} including [1] the time variable (indicated 
-#'   in \code{timename}), [2] the predictor variables (required by
-#'   \code{formula}) and [3], the compositional response variable (indicated
-#'   in \code{formula}). \cr \cr
-#'   Note that the response variables should be formatted as a 
-#'   \code{data.frame} object named as indicated by the \code{response} entry
-#'   in the \code{control} list, such as \code{gamma} for a standard TS 
-#'   analysis on LDA output. 
 #'
 #' @param formulas Vector of \code{\link[stats]{formula}}(s) defining the 
 #'   regression between the change points. Any predictor variable included 
@@ -141,6 +138,8 @@
 #'     used to fit the model. \cr \cr
 #'   \code{prepare_TS}: \code{list} of \code{list}s, each of which is a
 #'     preliminary model object for a Time Series model fit. \cr \cr
+#'   \code{check_TS}: an error message is thrown if any input is improper, 
+#'     otherwise \code{NULL}.
 #'   \code{TS_control}: \code{list} of named control elements for
 #'     model fitting.
 #'   \code{measure_TS}: \code{vector} of values corresponding to the model
@@ -195,9 +194,9 @@
 #'
 #' @export
 #'
-TS <- function(LDAs, data, formulas = ~ 1, nchangepoints = 0, 
+TS <- function(LDAs, formulas = ~ 1, nchangepoints = 0, 
[…]

### 2020-03-23 R/TS_methods.R

@@ -347,7 +347,7 @@ prep_cpts <- function(TS){
     args <- list(data = data, formula = TS$formula, changepoints = cps[ , i], 
                  timename = TS$timename, weights = TS$weights, 
                  control = TS$control$response_args$control)
-    modfit <- soft_call(fun, args, TRUE)
+    modfit <- soft_call(what = fun, args = args, soften = TRUE)
     lls[i] <- modfit$logLik
   }  
   cps <- cps[ , order(lls, decreasing = TRUE), drop = FALSE]
@@ -575,7 +575,7 @@ proposed_step_mods <- function(TS, prop_changepts, inputs){
                  changepoints = prop_changepts[ , i], 
                  timename = TS$timename, weights = TS$weights, 
                  control = control)
-    out[[i]] <- soft_call(fun, args, TRUE)
+    out[[i]] <- soft_call(what = fun, args = args, soften = TRUE)
   }
   out
 }

### 2020-03-23 R/TS_models.R

@@ -193,7 +193,7 @@ package_sequential_TS <- function(TS, rho_dist, eta_dist){
     args <- list(data = data, formula = TS$formula, changepoints = NULL, 
                  timename = TS$timename, weights = TS$weights, 
                  control = TS$control$response_args$control)
-    mod <- soft_call(fun, args, TRUE)
+    mod <- soft_call(what = fun, args = args, soften = TRUE)
     lls <- as.numeric(logLik(mod))
 
   } else{
@@ -236,7 +236,7 @@ est_changepoints <- function(TS){
   }
   fun <- TS$control$method
   args <- update_list(TS$control$method_args, TS = TS)
-  soft_call(fun = fun, args = args, soften = TS$control$soften)
+  soft_call(what = fun, args = args, soften = TS$control$soften)
 }
 
 #' @rdname sequential_TS
@@ -254,7 +254,7 @@ est_regressors <- function(rho_dist, TS){
     args <- list(data = data, formula = TS$formula, changepoints = NULL, 
                  timename = TS$timename, weights = TS$weights, 
                  control = TS$control$response_args$control)
-    mod <- soft_call(fun, args, TRUE)
+    mod <- soft_call(what = fun, args = args, soften = TRUE)
 
     mod <- mod[[1]][[1]]
     mv <- as.vector(t(coef(mod)))
@@ -299,7 +299,7 @@ est_regressors <- function(rho_dist, TS){
     args <- list(data = data, formula = TS$formula, changepoints = cpts, 
                  timename = TS$timename, weights = TS$weights, 
                  control = TS$control$response_args$control)
-    mod <- soft_call(fun, args, TRUE)
+    mod <- soft_call(what = fun, args = args, soften = TRUE)
 
 
     ndraws <- freq_r[i]

### 2020-03-23 R/utilities.R

@@ -1,30 +1,67 @@
-soft_call <- function(fun = function(x){invisible(NULL)}, args = list(NULL), 
-                      soften = FALSE){
+#' @title Soften a flexible function call to errors 
+#'
+#' @description Wrapping around \code{\link{do.call}}, this function allows
+#'   for a simple "softening" where errors are returned as text, rather
+#'   than causing a break in the encompassing pipeline.
+#'
+#' @details As this is a basic wrapper on \code{\link{do.call}}, the basic 
+#'   rules of its usage still apply: \cr
+#'   If \code{quote = FALSE} (the default), then the arguments are evaluated
+#'     in the calling environment and not in \code{envir}. \cr
+#'   If \code{quote = TRUE}, each argument is \code{\link{quote}}d 
+#'     so that the effect of argument evaluation is to remove the quotes
+#'     leaving the original arguments unevaluated when the call is built. \cr
+#'   The behavior of some functions, such as \code{\link{substitute}}, will 
+#'     not be the same for functions evaluated using \code{\link{do.call}}
+#'     and thus \code{soft_call} as if they were evaluated from the 
+#'     interpreter. The precise semantics are currently undefined and subject 
+#'     to change.
+#'
+#' @param what \code{function} or a non-empty \code{character} string 
+#'   naming the \code{function} to be called. \cr
+#'   See \code{\link{do.call}}.
+#'
+#' @param args \code{list} of arguments to the \code{what} call. The names
+#'   attribute of the \code{list} gives the argument names.\cr
+#'   See \code{\link{do.call}}.
+#'
+#' @param quote \code{logical} value indicating whether to quote the 
+#'   arguments. \cr
+#'   See \code{\link{do.call}}.
+#'
+#' @param envir \code{environment} within which to evaluate the call. This 
+#'   argument will be most useful if \code{what} is a \code{character} 
+#'   string and the arguments are symbols or \code{quote}d expressions.\cr
+#'   See \code{\link{do.call}}.
+#'
+#' @param soften \code{logical} value indicating whether to soften any errors 
+#'   in the running of \code{what}. \cr
+#'
+#' @return Either the result of the (evaluated) call of \code{what} or a
+#'   \code{list} of the error message that resulted (if \code{soften = TRUE}). 
+#'
+#' @export
+#'
[…]

### 2020-03-23 R/LDATS.R
documents

@@ -12,7 +12,7 @@
 #' @importFrom mvtnorm rmvnorm
 #' @importFrom nnet multinom
 #' @importFrom progress progress_bar
-#' @importFrom stats acf AIC as.formula coef ecdf logLik median rgeom rnorm
+#' @importFrom stats acf AIC as.formula coef ecdf lm logLik median rgeom rnorm
 #'   runif sd terms var vcov
 #' @importFrom viridis viridis
 #'

### 2020-03-23 R/LDA_TS_plots.R

@@ -52,7 +52,7 @@
 #'
 #' @param cols \code{list} of elements used to define the colors for the two
 #'   panels of the summary plot, as generated simply using
-#'   \code{\link{set_LDA_TS_plot_cols}}. \code{cols} has two elements:
+#'   \code{\link{set_LDA_TS_plot_colors}}. \code{cols} has two elements:
 #'   \code{LDA} and \code{TS}, each corresponding the set of plots for
 #'   its stage in the full model. \code{LDA} contains entries \code{cols}
 #'   and \code{option} (see \code{\link{set_LDA_plot_colors}}). \code{TS}

### 2020-03-23 R/LDA_models.R

@@ -3,26 +3,24 @@
 #'   as conducted via the topicmodels package
 #'
 #' @description Fit the standard LDATS LDA model (a true Latent Dirichlet
-#'   Allocation) using \code{\link[topicmodels]{topicmodels::LDA}}
-#'   (Grun and Hornik 2011). \cr 
-#'   Default methodology is the Variational
-#'   Expectation Maximization routine (VEM) as described by 
-#'   Blei \emph{et al.} (2003) and implemented by Grun and Hornik (2011). \cr
+#'   Allocation) using \code{\link[topicmodels]{LDA}} (Grun and Hornik 2011).
+#'   Default methodology is the Variational Expectation Maximization routine 
+#'   (VEM) as described by Blei \emph{et al.} (2003) and implemented by 
+#'   Grun and Hornik (2011). \cr \cr
 #'   If the model is defined to only fit one topic, \code{\link{identity_LDA}}
 #'   is used by default.
 #'
-#' @param LDA A prepared (via \code{\link{prep_LDA_models}} LDA model
+#' @param LDA A prepared (via \code{\link{prepare_LDA}} LDA model
 #'   \code{list}.
 #'
 #' @param ... Additional arguments to be passed to 
-#'   \code{\link[topicmodels]{topicmodels::LDA}} as a \code{control} input.
+#'   \code{\link[topicmodels]{LDA}} as a \code{control} input.
 #'
 #' @param seeded \code{logical} indicator of if the LDA should be a seeded
 #'   replicate. 
 #'
-#' @param method Fitting routine used in 
-#'   \code{\link[topicmodels]{topicmodels::LDA}}. Currenlty, only
-#'   \code{"VEM"} and \code{"Gibbs"} are supported.
+#' @param method Fitting routine used in \code{\link[topicmodels]{LDA}}.
+#'   Currenlty, only \code{"VEM"} and \code{"Gibbs"} are supported.
 #'
 #' @return \code{LDA} \code{list} with components
 #'  \describe{
@@ -90,7 +88,7 @@ topicmodels_LDA <- function(LDA, method = "VEM", seeded = TRUE, ...){
 #'   output is functionally the input. This allows for "single-topic" models
 #'   that do not actually decompose the data to be included in the model set.
 #'
-#' @param LDA A prepared (via \code{\link{prep_LDA_models}} LDA model
+#' @param LDA A prepared (via \code{\link{prepare_LDA}} LDA model
 #'   \code{list}.
 #'
 #' @return \code{LDA} \code{list} with components (many of which are 

### 2020-03-23 R/LDA_plots.R

@@ -6,7 +6,7 @@
 #'     showing the topic proportions for each word and a bottom panel showing
 #'     the topic proportions of each document/over time. \cr \cr
 #'   \code{plot.LDA_set} plots either the \code{selected} model results or
-#'     all of them from a \code{LDA_set} of \code{LDA} models.
+#'     all of them from a \code{LDA_set} of \code{LDA} models. \cr \cr
 #'   \code{LDA_plot_top_panel} creates an LDATS LDA summary plot 
 #'     top panel showing the topic proportions word-by-word. \cr \cr
 #'   \code{LDA_plot_bottom_panel} creates an LDATS LDA summary plot

### 2020-03-23 R/TS.R

@@ -80,10 +80,11 @@
 #'   Time Series model. Values not input assume defaults set by 
 #'   \code{\link{TS_control}}.
 #'  
+#' @param LDAs Class \code{LDA_set} \code{list} of both selected and all 
+#'   results from \code{\link{LDA}}.
+#'
 #' @param TS,TSs time series model \code{list} (\code{TS}) or a \code{list} 
 #'   of many time series model \code{list}s (\code{TSs}).
-#'
-#' @param selected_TSs \code{list} of selected time series model \code{list}s.
 #'
 #' @param response \code{character} element indicating the response variable 
 #'   used in the time series. \cr \cr
@@ -104,6 +105,11 @@
 #'
 #' @param soften \code{logical} indicator of whether the model should error 
 #'   softly or if errors should trigger a full-stop to the pipeline.
+#'
+#' @param model Main Time Series \code{function}.
+#'
+#' @param model_args \code{list} of (named) arguments to be used in 
+#'   \code{model} via \code{\link{TS_call}}. 
 #'
 #' @param method \code{function} used to drive the sampler of the TS
 #'   models; \code{method} defines and operates the computational procedure.
@@ -147,7 +153,7 @@
 #'   \code{select_TS}: \code{list} of selected models' \code{list}s. \cr \cr
 #'   \code{run_TS}: \code{TS_set} \code{list} of model results from all
 #'     runs of a \code{<model>} function, such as 
-#'     \code{\link{topicmodels_TS}}. \cr \cr
+#'     \code{\link{topicmodels_LDA}}. \cr \cr
 #'   \code{TS_call}: \code{TS} \code{list} of model results from a single
 #'     run of a \code{<model>} function, such as 
 #'     \code{\link{sequential_TS}}. \cr \cr

### 2020-03-23 R/TS_methods.R

@@ -331,8 +331,8 @@ prep_cpts <- function(TS){
   data <- TS$data$train$ts_data
   temps <- prep_temp_sequence(TS = TS)
   ntemps <- length(temps)
-  min_time <- min(data[ , timename])
-  max_time <- max(data[ , timename])
+  min_time <- min(data[ , TS$timename])
+  max_time <- max(data[ , TS$timename])
   times <- seq(min_time, max_time, 1)
   avail_times <- times[-c(1, length(times))]
   cps <- matrix(NA, nrow = TS$nchangepoints, ncol = ntemps)
@@ -502,6 +502,11 @@ prep_ids <- function(TS){
 update_ids <- function(ids, swaps){
   swaps$ids
 }
+
+#' @rdname ldats_classic
+#'
+#' @export
+#'
 step_chains <- function(TS, i, cpts, inputs){
   prop_step <- propose_step(TS = TS, i = i, cpts = cpts, inputs = inputs)
   accept_step <- eval_step(i = i, cpts = cpts, prop_step = prop_step, 

### 2020-03-23 R/TS_models.R

@@ -48,13 +48,9 @@
 #'
 #' @param rhos \code{matrix} of change point locations (columns) across
 #'   iterations of the sampler (rows) or \code{NULL} if no change points are 
-#'   in the model, as returned from \code{\link{est_change points}}.
+#'   in the model, as returned from \code{\link{est_changepoints}}.
 #'
 #' @param TS Time series model \code{list}.
-#'
-#' @param response \code{character} element indicating the response variable 
-#'   used in the time series. \cr \cr
-#'   Must have a corresponding \code{<response>_TS} function.
 #'
 #' @param summary_prob Probability used for summarizing the posterior 
 #'   distributions (via the highest posterior density interval, see
@@ -121,7 +117,7 @@
 #'       \item{timename}{\code{timename} input to the function.}
 #'       \item{control}{\code{control} input to the function.}
 #'       \item{lls}{Iteration-by-iteration 
-#'                  \link[=logLik.multinom_TS_fit]{logLik} values for the
+#'                  \link[=logLik.TS_fit]{logLik} values for the
 #'                   full time series fit by \code{\link{multinom_TS}}.}
 #'       \item{rhos}{Iteration-by-iteration change point estimates from
 #'                   \code{\link{est_changepoints}} and diagnostics.}
@@ -160,7 +156,7 @@
 #'     each regressor. \cr \cr
 #'   \code{measure_eta_vcov}: variance-covariance matrix for chunk-level
 #'     regressors. \cr \cr
-#'  \code{summarize_rhos}: table of summary statistics for change point
+#'   \code{summarize_rhos}: table of summary statistics for change point
 #'     locations including mean, median, mode, posterior interval, standard
 #'     deviation, MCMC error, autocorrelation, and effective sample size for 
 #'     each change point location. \cr \cr
@@ -259,7 +255,7 @@ est_regressors <- function(rho_dist, TS){
     mod <- mod[[1]][[1]]
     mv <- as.vector(t(coef(mod)))
     vcv <- mirror_vcov(mod)
-    eta <- rmvnorm(control$method_args$nit, mv, vcv)
+    eta <- rmvnorm(TS$control$method_args$nit, mv, vcv)
     seg_names <- rep(1, ncol(vcv))
     coef_names <- colnames(vcv)
     colnames(eta) <- paste(seg_names, coef_names, sep = "_")

### 2020-03-23 R/TS_plots.R

@@ -83,6 +83,9 @@
 #'   summary plot. Currently only defined for \code{"median"} and 
 #'   \code{"mode"}.
 #'
+#' @param together \code{logical} indicating if the subplots are part of a 
+#'   larger plot output. 
+#'
 #' @param cols,rho_cols,gamma_cols 
 #'   In \code{plot.TS}, \code{cols} is a \code{list} of elements used to
 #'     define the colors for the two panels of the summary plot, as generated 
@@ -352,16 +355,14 @@ pred_gamma_TS_plot <- function(x, selection = "median",
   }
   x$control$timename <- NULL # to remove from v0.1.0 model fits
 
-#
-# NEEDS TO BE GENERALIZED
-#
-  seg_mods <- multinom_TS(x$data$train$ts_data, x$formula, spec_rhos,  
-                          x$timename, x$weights, x$control)
-#
-#
-#
+  fun <- x$control$response
+  args <- list(data = x$data$train$ts_data, formula = x$formula,
+               changepoints = spec_rhos, timename = x$timename, 
+               weights = x$weights, 
+               control = x$control$response_args$control)
+  mod <- soft_call(what = fun, args = args, soften = TRUE)
 
-  nsegs <- length(seg_mods[[1]])
+  nsegs <- length(mod[[1]])
   t1 <- min(x$data$train$ts_data[, x$timename])
   t2 <- max(x$data$train$ts_data[, x$timename])
 
@@ -383,7 +384,7 @@ pred_gamma_TS_plot <- function(x, selection = "median",
   pred_vals <- matrix(NA, nrow(x$data$train$ts_data), ntopics)
   sp1 <- 1
   for (i in 1:nsegs){
-    mod_i <- seg_mods[[1]][[i]]
+    mod_i <- mod[[1]][[i]]
     spec_vals <- sp1:(sp1 + nrow(mod_i$fitted.values) - 1)
     pred_vals[spec_vals, ] <- mod_i$fitted.values
     time_obs[spec_vals] <- mod_i$timevals

### 2020-03-23 R/simulate.R

@@ -1,10 +1,25 @@
 
-#' @title Simulate LDA data from an LDA structure given parameters
+#' @title Simulate LDA_TS data from LDA and TS model structures and parameters
 #'
-#' @description For a given set of parameters \code{alpha} and \code{Beta} and
-#'   document-specific total word counts, simulate a document-by-term matrix.
-#'   Additional structuring variables (the numbers of topics (k),
-#'   documents (M), terms (V)) are inferred from input objects.
+#' @description 
+#'   \code{simulate_LDA_data} for a given set of parameters \code{alpha} and
+#'     \code{Beta} and document-specific total word counts, simulate a 
+#'     document-by-term matrix. \cr
+#'     Additional structuring variables (the numbers of topics (k),
+#'     documents (M), terms (V)) are inferred from input objects. \cr \cr
+#'   \code{simulate_TS_data} for a given set of covariates \code{X};
+#'     parameters \code{Eta}, \code{rho}, and \code{err}; and
+#'     document-specific time stamps \code{tD}, simulate a document-by-topic
+#'     matrix. Additional structuring variables (numbers of topics (k), 
+#'     documents (M), segments (S), and covariates per segment (C)) are 
+#'     inferred from input objects.
+#'   \code{simulate_LDA_TS_data} for a given set of covariates \code{X}; 
+#'     parameters  \code{Beta}, \code{Eta}, \code{rho}, and \code{err}; and 
+#'     document-specific time stamps \code{tD} and lengths \code{N}),
+#'     simulate a document-by-topic matrix. \cr
+#'     Additional structuring variables (the numbers of topics (k), terms (V), 
+#'     documents (M), segments (S), and covariates per segment (C))
+#'     are inferred from input objects.
 #'
 #' @param N A vector of document sizes (total word counts). Must be integer
 #'   conformable. Is used to infer the total number of documents.
@@ -13,7 +28,28 @@
 #'   terms within topics. Dimension: k x V (number of topics x number of 
 #'   terms). Used to infer both (k) and (V). Must be non-negative and sum to 
 #'   1 within topics.
+#'
[…]

### 2020-03-23 R/LDA.R
rodents example vignette

@@ -163,7 +163,7 @@ prepare_LDA <- function(data, topics = 2, replicates = 1, control = list()){
   data <- conform_data(data = data, control = control)
   subsets <- names(data)
   if(length(replicates) < length(topics)){
-    reps <- rep(replicates, length(topics))
+    replicates <- rep(replicates, length(topics))
   }
   LDA_topics <- rep(topics, replicates)
   LDA_reps <- sequence(replicates)
@@ -172,14 +172,20 @@ prepare_LDA <- function(data, topics = 2, replicates = 1, control = list()){
   LDA_topics <- rep(LDA_topics, length(subsets))
   nLDA <- length(LDA_topics)
   LDAs <- vector("list", length = nLDA)
+  topicword <- rep(NA, nLDA)
   for(i in 1:nLDA){
-    LDAs[[i]] <- list(data = data[[LDA_subsets[[i]]]], 
-                      data_subset = LDA_subsets[[i]],
-                      topics = LDA_topics[[i]], 
-                      replicate = LDA_reps[[i]],
+    LDAs[[i]] <- list(data = data[[LDA_subsets[i]]], 
+                      data_subset = LDA_subsets[i],
+                      topics = LDA_topics[i], 
+                      replicate = LDA_reps[i],
                       control = control)
+    topicword[i] <- ifelse(LDA_topics[i] == 1, "topic", "topics")
   }
-  names(LDAs) <- paste0("model_", 1:nLDA)
+  name_tab <- data.frame(paste("data subset", LDA_subsets), 
+                         paste(",", LDA_topics, topicword),
+                         paste(", replicate", LDA_reps))
+  names(LDAs) <- apply(name_tab, 1, paste0, collapse = "")
+
   LDAs
 }
 

### 2020-03-23 R/TS.R

@@ -298,8 +298,8 @@ prepare_TS <- function(LDAs, formulas = ~ 1, nchangepoints = 0,
   formulas <- formulas2
   nmods <- length(LDAs[[1]])
   mods <- 1:nmods
-  tab <- expand.grid(mods, formulas, nchangepoints, stringsAsFactors = FALSE)
-  colnames(tab) <- c("LDA", "formula", "nchangepoints") 
+  tab <- expand.grid(LDA = mods, formulas = formulas, 
+                     nchangepoints = nchangepoints, stringsAsFactors = FALSE)
   
   nTSs <- NROW(tab)
   TSs <- vector("list", length = nTSs)

### 2020-04-21 doc/LDATS_codebase.Rmd
codebase vignette
updated to the current codebase

@@
 
-This vignette outlines the code base for the **LDATS** package. It was
-constructed using **LDATS** version `r vers` on `r today`.
+This vignette outlines the code base for the **LDATS** package. 
+It was constructed using **LDATS** version `r vers` on `r today`.
 
@@
 
-To obtain the most recent version of **LDATS**, install the most recent 
-version from GitHub:
+To obtain the most recent version of **LDATS**, install the most recent version from GitHub:
 
@@
 
-The full LDATS analysis can be executed using the `LDA_TS` function,
-which is the top-level of the coding pipeline's hierarchy of
-functions:
+The full LDATS analysis can be executed using the `LDA_TS` function, which is the top-level of the coding pipeline's hierarchy of functions:
 
 * `LDA_TS()`
-  * `check_LDA_TS_inputs()`
-    * `check_timename()`
-    * `check_formulas()`
-    * `check_nchangepoints()`
-    * `check_weights()`
-    * `check_control()`
-    * `check_document_term_table()`
-    * `check_document_covariate_table()`	
-    * `check_topics()`
-    * `check_seeds()`
-  * `LDA_set()`
-    * `check_LDA_TS_inputs()`
-      * `check_control()`
-      * `check_document_term_table()`
-      * `check_topics()`
-      * `check_seeds()`
-    * loop over models
-      * `prep_LDA_control()`
+  * `LDA_TS_control()`
+    * `LDA_control()`
+    * `TS_control()`
   * `LDA()`
+    * `prepare_LDA()`
+      * `check_LDA()`
+      * `LDA_control()`
+      * `conform_data()`
+    * `run_LDA()`
+      * `LDA_call()` (replicated for each model)
+        * `LDA_msg()`
+        * `soft_call()` using `LDA$control$model` with `LDA$control$model_args`
+          * `topicmodels::LDA()` (*default*) 
+      * `package_LDA()`
         * `select_LDA()`
-    * applies the `measurer` and `selector` functions supplied via `LDA_controls_list()`
-  * `TS_on_LDA()`
-    * `check_LDA_TS_inputs()`
-      * `check_LDA_models()`
-      * `check_formulas()`
-      * `check_nchangepoints()`
-      * `check_timename()`
-      * `check_control()`
-      * `check_weights()`
-      * `check_document_covariate_table()`	
-    * `expand_TS()`
-    * loop over models
-      * `prep_TS_data()`
+          * `soft_call()` using `LDA$control$selector with LDA$control$selector_args
+            * `min()` (*default*) 
+            * `measure_LDA()` (replicated for each model)
+              * `soft_call()` using `LDA$control$measurer` with `LDA$control$measurer_args`
+            * `AIC()` (*default*)
+
   * `TS()`
-        * `check_TS_inputs()`
-          * `check_formulas()`
-          * `check_nchangepoints()`
-          * `check_control()`
-          * `check_weights()`
+    * `prepare_TS()`
+      * `TS_control()`
+      * `run_TS()`
+        * `TS_call()` (replicated for each model)
+          * `TS_msg()`
+          * `TS$control$model with TS$control$model_args
+            * `sequential_TS_control()`
             * `est_changepoints()`
+              * `TS$control$method` with `TS$control$method_args`
+                * `ldats_classic_control()`
                   * `prep_saves()`
                   * `prep_ptMCMC_inputs()`
+                    * `prep_temp_sequence()`
+                    * `prep_proposal_dist()`
                   * `prep_cpts()`
-            * loop over chains
-              * `multinom_TS()`
+                    * `prep_temp_sequence()`
+                    * `TS$control$response` with `TS$control$response_args`
                   * `prep_ids()`
                   * `prep_pbar()`
-          * loop over iterations
                   * `update_pbar()`
@@
                       * `proposed_step_mods()`
-                  * loop over chains
-                    * `multinom_TS()`
+                      * `TS$control$response` with `TS$control$response_args`
                     * `eval_step()`
-              * `take_step()`
+                    * `accept_step()`
                   * `swap_chains()`
-                * loop over chain neighbors
                   * `update_saves()`
@@
                   * `update_ids()`
+                  * `process_saves()`
+                    * `count_trips()`
              * `est_regressors()`
-          * loop over unique realizations of change point locations
-            * `multinom_TS()`
-            * loop over chunks
-              * `mirror_vcov()`
-              * `rmvnorm()`
-        * `summarize_TS()`
-          * `diagnose_ptMCMC()`
-            * `count_trips()`
+               * `TS$control$response` with `TS$control$response_args`
+             * `package_sequential_TS()`
                * `summarize_rhos()`
                * `measure_rho_vcov()`
-          * `summarize_etas()`
+               * `summarize_eta()`
                * `measure_eta_vcov()`
+       * `package_TS()`
          * `select_TS()`
-      * applies the `measurer` and `selector` functions supplied via `TS_controls_list()`
+           * `TS$control$selector` with `TS$control$selector_args`
+           * `measure_TS()` (replicated for each model)
+             * `TS$control$measurer` with `TS$control$measurer_args`
   * `package_LDA_TS()`
 
-Each component model's function (`LDA_set()` and `TS()`) can be run
-independently, as well.
+Each component model's function (`LDA()` and `TS()`) can be run independently, as well.
 
@@
 
-To minimize the length of argument lists and facilitate simple default usage
-throughout the pipeline, we implement an options/controls list approach, where
-each of the main functions (`LDA_TS`, `LDA_set`, and `TS`) has a controls 
-list (although `LDA_set` is referenced as `LDA` in the naming scheme)
+To minimize the length of argument lists and facilitate simple default usage throughout the pipeline, we implement an options/controls list approach, where each of the main functions (`LDA_TS`, `LDA`, and `TS`) and their subfunctions have a `control` argument that takes a `list` to replace the defaults returned by its `_control` function:
 
-* `LDA_TS_controls_list()`
-* `LDA_controls_list()`
-* `TS_controls_list()`
+* `LDA_TS_control()`
+* `LDA_control()`
+* `TS_control()`

### 2020-04-21 R/TS.R

@@ -353,6 +353,8 @@ package_TS <- function(TSs){
 #'
 select_TS <- function(TSs){
 
+# use softcall!
+
   vals <- measure_TS(TSs = TSs)
   fun <- TSs[[1]]$control$selector
   args <- update_list(TSs[[1]]$control$selector_args, x = vals)
@@ -366,6 +368,8 @@ select_TS <- function(TSs){
 #' @export
 #'
 measure_TS <- function(TSs){
+
+# use softcall!
 
   nTSs <- length(TSs)
   vals <- rep(NA, nTSs)

### 2020-04-21 notes.R

@@ -4,8 +4,11 @@ predict
 tidying check functions and general usage
 examples
 tests
+softcall message handling
 tempering
 vignettes
+   data 
+   simplex response
   
 
 devtools::load_all()
@@ -13,21 +16,16 @@ devtools::load_all()
 LDAs <- LDA(data = rodents, topics = 2:3, replicates = c(1,4))
 
 
-TS(LDAs = LDAs, formulas = ~ newmoon, nchangepoints = 1, 
+TSs <- TS(LDAs = LDAs, formulas = ~ newmoon, nchangepoints = 1, 
           timename = "newmoon", weights = TRUE,
           control = list(response = simplex_TS, 
-                         method_args = 
-                         list(control = ldats_classic_control(nit = 100)),
-                                response_args = 
-                           list(control = 
-                            simplex_TS_control(transformation = "ilr"))))
+                         response_args = list(control = 
+                                              list(transformation = "ilr")),
+                         method_args = list(control = list(nit = 100))))
 
-TS(LDAs = LDAs, formulas = ~ newmoon, nchangepoints = 1, 
+TSs <- TS(LDAs = LDAs, formulas = ~ newmoon, nchangepoints = 1, 
           timename = "newmoon", weights = TRUE,
-          control = list(method_args = 
-                         list(control = ldats_classic_control(nit = 100))))
-
-
+          control = list(method_args = list(control = list(nit = 100))))
 
 plot(LDAs)
 plot(TSs)

### 2020-04-24 R/TS_methods.R
patch for bad starting values

@@ -350,6 +350,42 @@ prep_cpts <- function(TS){
     modfit <- soft_call(what = fun, args = args, soften = TRUE)
     lls[i] <- modfit$logLik
   }  
+
+  if(any(lls == -Inf)){
+    iter <- 1
+    while(all(lls == -Inf)){
+
+      cps <- matrix(NA, nrow = TS$nchangepoints, ncol = ntemps)
+      for (i in 1:ntemps){
+        cp_times <- sort(sample(avail_times, TS$nchangepoints, 
+                         replace = FALSE))
+        cps[ , i] <- cp_times
+      }
+      lls <- rep(NA, ntemps)
+      for (i in 1:ntemps){
+        fun <- TS$control$response
+        fun <- memoise_fun(fun, TS$control$memoise)
+        args <- list(data = data, formula = TS$formula, 
+                     changepoints = cps[ , i], 
+                     timename = TS$timename, weights = TS$weights, 
+                     control = TS$control$response_args$control)
+        modfit <- soft_call(what = fun, args = args, soften = TRUE)
+        lls[i] <- modfit$logLik
+      }  
+      iter <- iter + 1
+      if(iter > 10){
+        stop("max number of starts tried with all -Inf LogLiks")
+      }
+    }
+    to_replace <- which(lls == -Inf)
+    to_use <- which(lls != -Inf)
+    for(i in 1:length(to_replace)){
+      to_use_i <- sample(to_use, 1)
+      lls[to_replace[i]] <- lls[to_use_i]
+      cps[, to_replace[i]] <- cps[ , to_use_i]
+    }
+  }
+
   cps <- cps[ , order(lls, decreasing = TRUE), drop = FALSE]
   lls <- sort(lls, decreasing = TRUE)
 

### 2023-09-11 doc/paper-comparison.R
precompiling the paper comparison
[WIP]

@@ -1,494 +1,4 @@
 params <-
 list(run_models = FALSE)
 
-## ----setup, include=FALSE-----------------------------------------------------
-knitr::opts_chunk$set(
-  collapse = TRUE,
-  comment = "#>"
-)
-
-## ---- include=FALSE-----------------------------------------------------------
-vers <- packageVersion("LDATS")
-
-## ---- eval = FALSE------------------------------------------------------------
-#  install.packages("devtools")
-#  devtools::install_github("weecology/LDATS")
-
-## -----------------------------------------------------------------------------
-library(LDATS)
-set.seed(42)
-nseeds <- 200
-nit <- 10000
-
-## ---- eval = params$run_models------------------------------------------------
-#  install.packages(c("dplyr", "gridExtra", "multipanel", "RColorBrewer", "RCurl", "reshape2"))
-
-## ---- eval = FALSE------------------------------------------------------------
-#  rmarkdown::render("paper-comparison.Rmd", params = list(run_models = TRUE))
-
-## ----set download location----------------------------------------------------
-vignette_files <- tempdir()
-
-## ----download scripts and data------------------------------------------------
-test_file <- file.path(vignette_files, "scripts", "rodent_LDA_analysis.r")
-
-if (!file.exists(test_file)){
-
-  # scripts
-  dir.create(file.path(vignette_files, "scripts"), showWarnings = FALSE)
-  github_path <- "https://raw.githubusercontent.com/weecology/LDATS-replications/master/scripts/"
-  files_to_download <- c("rodent_LDA_analysis.r", "rodent_data_for_LDA.r", 
-                         "AIC_model_selection.R", "changepointmodel.r", 
-                         "LDA-distance.R", "LDA_figure_scripts.R")
-  
-  for (file in files_to_download)  {
-    download.file(url = paste0(github_path, file),
-                  destfile = file.path(vignette_files, "scripts", file))
-  }
-
-    
-  # data
-  dir.create(file.path(vignette_files, "data"), showWarnings = FALSE)
-  github_path <- "https://raw.githubusercontent.com/weecology/LDATS-replications/master/data/"
-  files_to_download <- c("moon_dates.csv", "Portal_rodent_trapping.csv", 
-                         "Rodent_table_dat.csv", "paper_dat.csv",
-                         "paper_dates.csv", "paper_covariates.csv")
-  
[…]

### 2023-09-11 vignettes/precompile.R

@@ -1,0 +1,6 @@
+#Vignettes that depend on internet access have been precompiled:
+
+
+knitr::knit("vignettes/paper-comparison.Rmd.remote_needs", output = "vignettes/paper-comparison.Rmd")
+
+devtools::build_vignettes( )

### 2023-09-16 doc/paper-comparison.Rmd
no mark

@@
 title: "Comparison to Christensen et al. 2018"
-author: "Renata Diaz and Hao Ye"
+author: "Renata Diaz, Juniper Simonis, and Hao Ye"
 output: rmarkdown::html_vignette
@@
 
-```{r setup, include=FALSE}
-knitr::opts_chunk$set(
-  collapse = TRUE,
-  comment = "#>"
-)
-```
-```{r, include=FALSE}
-vers <- packageVersion("LDATS")
-```
-
 # Introduction
 
-This document provides a side-by-side comparison of **`LDATS`** (version `r vers`) results with analysis from [Christensen et al.
+This document provides a side-by-side comparison of **`LDATS`** (version 0.3.0) results with analysis from [Christensen et al.
 2018](https://esajournals.onlinelibrary.wiley.com/doi/full/10.1002/ecy.2373).
+Due to the size and duration of model runs, we use pre-generated model output from the [LDATS-replications repo](https://github.com/weecology/LDATS-replications).
 
@@
 
-```{r, eval = FALSE}
-# install.packages("devtools")
+```r
+install.packages("devtools")
 devtools::install_github("weecology/LDATS")
@@
 Load in the **`LDATS`** package.
-```{r}
+
+```r
 library(LDATS)
@@
 nit <- 10000
+```
+
+To run the analyses here, you will also need to download **`dplyr`**, **`gridExtra`**, **`multipanel`**, **`RColorBrewer`**, **`RCurl`**, and **`reshape2`** as the manuscript's code relies on these packages.
+
+```r
+install.packages(c("dplyr", "gridExtra", "multipanel", "RColorBrewer", "RCurl", "reshape2"))
 ```
@@
 To change this functionality, you can re-render this file with:
-```{r, eval = FALSE}
+
+```r
 rmarkdown::render("paper-comparison.Rmd", params = list(run_models = TRUE))
@@
 
-```{r set download location}
-vignette_files <- tempdir()
+```r
+vignette_files <- "."
 ```
@@
 To replicate the Christensen et al.
-2018 analysis, we download some of the original scripts & data files from [Extreme-events-LDA repo](https://github.com/emchristensen/Extreme-events-LDA), as well as some raw data files from the [PortalData repo](https://github.com/weecology/PortalData):
+2018 analysis, we download some of the original scripts & data files from [Extreme-events-LDA repo](https://github.com/emchristensen/Extreme-events-LDA), as well as some raw data files from the [PortalData repo](https://github.com/weecology/PortalData), which are stored in the [LDATS-replications repo](https://github.com/weecology/LDATS-replications):
 
@@
 
+* `paper_dat.csv`
+  - rodent data table from Christensen et al. 2018
+
+* `paper_dates.csv`
+  - dates used in Christensen et al. 2018
+
+* `paper_covariates.csv`
+  - table of dates and covariate data
+
 Figure scripts:
@@
 
-```{r download scripts}
-test_file <- file.path(vignette_files, "rodent_LDA_analysis.r")
+```r
+test_file <- file.path(vignette_files, "scripts", "rodent_LDA_analysis.r")
 
-if (!file.exists(test_file))
-{
-  # from the Extreme-events-LDA repo
-  github_path <- "https://raw.githubusercontent.com/emchristensen/Extreme-events-LDA/master/"
+if (!file.exists(test_file)){
+
+  # scripts
+  dir.create(file.path(vignette_files, "scripts"), showWarnings = FALSE)
+  github_path <- "https://raw.githubusercontent.com/weecology/LDATS-replications/master/scripts/"
   files_to_download <- c("rodent_LDA_analysis.r", "rodent_data_for_LDA.r", 
                          "AIC_model_selection.R", "changepointmodel.r", 
-                         "LDA-distance.R", "Rodent_table_dat.csv", 
-                         "LDA_figure_scripts.R")
+                         "LDA-distance.R", "LDA_figure_scripts.R")
 
-  for (file in files_to_download)
-  {
+  for (file in files_to_download)  {
     download.file(url = paste0(github_path, file),
-                  destfile = file.path(vignette_files, file))
+                  destfile = file.path(vignette_files, "scripts", file))
   }
 
-  # from the PortalData repo
-  github_path <- "https://raw.githubusercontent.com/weecology/PortalData/master/Rodents/"
-  files_to_download <- c("moon_dates.csv", "Portal_rodent_trapping.csv")
+  # data
+  dir.create(file.path(vignette_files, "data"), showWarnings = FALSE)
+  github_path <- "https://raw.githubusercontent.com/weecology/LDATS-replications/master/data/"
+  files_to_download <- c("moon_dates.csv", "Portal_rodent_trapping.csv", 
+                         "Rodent_table_dat.csv", "paper_dat.csv",
+                         "paper_dates.csv", "paper_covariates.csv")
 
-  for (file in files_to_download)
-  {
+  for (file in files_to_download)  {
     download.file(url = paste0(github_path, file),
-                  destfile = file.path(vignette_files, file))
+                  destfile = file.path(vignette_files, "data", file))
   }
@@
 
-Changepoint outputs
+Changepoint outputs:
 
-* `ldats_ldats.RDS`
-  - the posterior distribution of changepoints, using the LDATS LDA model and the LDATS changepoint selection
+* `ldats_ldats.RDS`, `ldats_ldats_cpt.RDS`, `ldats_ldats_cpt_dates.RDS`
+  - the posterior distribution, count, and dates of changepoints, using the LDATS LDA model and the LDATS changepoint selection
 
-* `ldats_paper.RDS`
-  - the posterior distribution of changepoints, using the LDATS LDA model and the paper's changepoint selection
+* `ldats_paper.RDS`, `ldats_paper_cpt.RDS`, `ldats_paper_cpt_dates.RDS
+  - the posterior distribution, count, and dates of changepoints, using the LDATS LDA model and the paper's changepoint selection
 
-* `paper_ldats.RDS`
-  - the posterior distribution of changepoints, using the paper LDA model and the LDATS changepoint selection
+* `paper_ldats.RDS`, `paper_ldats_cpt.RDS`, `paper_ldats_cpt_dates.RDS
+  - the posterior distribution, count, and dates of changepoints, using the paper LDA model and the LDATS changepoint selection
 
-* `paper_paper.RDS`
-  - the posterior distribution of changepoints, using the paper LDA model and the paper's changepoint selection
+* `paper_paper.RDS`, `paper_paper_cpt.RDS`, `paper_paper_cpt_dates.RDS
+  - the posterior distribution, count, and dates of changepoints, using the paper LDA model and the paper's changepoint selection
+
+* `cpt_dates.RDS`
+  - summary table of changepoint results across models
 
@@
 
-```{r download pre-generated model outputs, eval = !params$run_model}
-test_file <- file.path(vignette_files, "ldats_ldamodel.RDS")
+* `paper_paper_cpt_plot.png`
+  - figure showing the time series results for the paper analysis of the paper LDA
 
-if (!file.exists(test_file))
-{
-  # from the Extreme-events-LDA repo
-  github_path <- "https://raw.githubusercontent.com/weecology/LDATS-replications/master/model_outputs/"
+* `ldats_paper_cpt_plot.png`
+  - figure showing the time series results for the paper analysis of the LDATS LDA
+
+* `annual_hist.RDS`
+  - function for making histogram of change points over years
+
+```r
+test_file <- file.path(vignette_files, "output", "ldats_ldamodel.RDS")
+
+if (!file.exists(test_file)){
+
+  dir.create(file.path(vignette_files, "output"), showWarnings = FALSE)
+  github_path <- "https://raw.githubusercontent.com/weecology/LDATS-replications/master/output/"
   files_to_download <- c("ldats_ldamodel.RDS", "paper_ldamodel.RDS", 
@@
                          "paper_ldats.RDS", "paper_paper.RDS", 
-                         "lda_distances.png")
+                         "ldats_rodents_adjusted.RDS", "rodents.RDS",
+                         "ldats_paper_cpt.RDS", "ldats_paper_cpt_dates.RDS",
+                         "ldats_ldats_cpt.RDS", "ldats_ldats_cpt_dates.RDS",
+                         "paper_paper_cpt.RDS", "paper_paper_cpt_dates.RDS",
+                         "paper_ldats_cpt.RDS", "paper_ldats_cpt_dates.RDS",
+                         "annual_hist.RDS", "cpt_dates.RDS",
+                         "lda_distances.png", "paper_paper_cpt_plot.png",
+                         "ldats_paper_cpt_plot.png")
 
-  for (file in files_to_download)
-  {
+  for (file in files_to_download){
     download.file(url = paste0(github_path, file),
-                  destfile = file.path(vignette_files, file), 
+                  destfile = file.path(vignette_files, "output", file), 
                   mode = "wb")
@@
 
-```{r LDATS data}
+```r
 data(rodents)
@@
 head(rodents[[1]])
+#>   BA DM DO DS NA. OL OT PB PE PF PH PI PL PM PP RF RM RO SF SH SO
+#> 1  0 13  0  2   2  0  0  0  1  1  0  0  0  0  3  0  0  0  0  0  0
+#> 2  0 20  1  3   2  0  0  0  0  4  0  0  0  0  2  0  0  0  0  0  0
+#> 3  0 21  0  8   4  0  0  0  1  2  0  0  0  0  1  0  0  0  0  0  0
+#> 4  0 21  3 12   4  2  3  0  1  1  0  0  0  0  0  0  0  0  0  0  0
+#> 5  0 16  1  9   5  2  1  0  0  2  0  0  0  0  0  0  1  0  0  0  0
+#> 6  0 17  1 13   5  1  5  0  0  3  0  0  0  0  0  0  0  0  0  0  0
 ```
@@
 
-```{r Paper data}
+```r
 # parameters for subsetting the full Portal rodents data
@@
 
-source(file.path(vignette_files, "rodent_data_for_LDA.r"))
+source(file.path(vignette_files, "scripts", "rodent_data_for_LDA.r"))
 
@@
 # assemble `paper_covariates`, the associated dates and covariate data
-moondat <- read.csv(file.path(vignette_files, "moon_dates.csv"), stringsAsFactors = F)
+moondat <- read.csv(file.path(vignette_files, "data", "moon_dates.csv"), stringsAsFactors = F)
 
@@
 
-```{r rodent data comparison}
+```r
 compare <- rodents[[1]] == paper_dat
@@
 length(which(rowSums(compare) < ncol(compare)))
+#> [1] 16
 ```
@@
 There are 16 rows where the data included in LDATS differs from the paper data.
-This is because the LDATS data is not adjusted to account for trapping effort, while the paper data does, by dividing all census counts by the actual number of plots trapped and multiplying by 8 to account for incompletely-trapped censuses.
+This is because the LDATS data is not adjusted to account for trapping effort, while the paper data is, by dividing all census counts by the actual number of plots trapped and multiplying by 8 to account for incompletely-trapped censuses.
 
@@
 
-```{}
+```r
   # retrieve data on number of plots trapped per month
@@
 
-```{r adjust LDATS data after Christensen et al, eval = TRUE}
+```r
 # get the trapping effort for each sample
-trap_table <- read.csv(file.path(vignette_files, "Portal_rodent_trapping.csv"))
+trap_table <- read.csv(file.path(vignette_files, "data", "Portal_rodent_trapping.csv"))
 trap_table_controls <- dplyr::filter(trap_table, plot %in% control_plots)
@@
 
-Now we can compare the adjusted ldats dataset with both the original ldats dataset and the dataset from the paper:
+Now we can compare the adjusted LDATS dataset with both the original ldats dataset and the dataset from the paper:
 
-```{r dataset comparisons}
+```r
 compare_raw <- rodents[[1]] == ldats_rodents_adjusted
@@
 
-```{r switch to adjusted rodents}
+```r
 rodents[[1]] <- paper_dat
@@
 
-```{r add dates to covariate table}
+```r
 head(rodents$document_covariate_table)
+#>   newmoon sin_year cos_year
+#> 1       1  -0.2470  -0.9690
+#> 2       2  -0.6808  -0.7325
+#> 3       3  -0.9537  -0.3008
+#> 4       4  -0.9813   0.1925
+#> 5       5  -0.7583   0.6519
+#> 6       6  -0.3537   0.9354
+```
 
+```r
 new_cov_table <- dplyr::left_join(rodents$document_covariate_table, 
@@
 
-```{r LDATS LDAs, eval = params$run_models}
+```r
 ldats_ldas <- LDATS::LDA_set(document_term_table = rodents$document_term_table, 
@@
 
-```{r paper LDAs, eval = params$run_models}
-source(file.path(vignette_files, "AIC_model_selection.R"))
-source(file.path(vignette_files, "LDA-distance.R"))
+```r
+source(file.path(vignette_files, "scripts", "AIC_model_selection.R"))
+source(file.path(vignette_files, "scripts", "LDA-distance.R"))
 
@@
 
-png(file.path(vignette_files, "lda_distances.png"), width = 800, height = 400)
+png(file.path(vignette_files, "output", "lda_distances.png"), width = 800, height = 400)
 dat <- paper_dat # calculate_LDA_distance has some required named variables
@@
 
-```{r}
-knitr::include_graphics(file.path(vignette_files, "lda_distances.png"))
+```r
+knitr::include_graphics(file.path(vignette_files, "output", "lda_distances.png"))
 ```
+
+![plot of chunk unnamed-chunk-10](./output/lda_distances.png)
 
@@
 To visualize the LDA assignment of species to topics, we load in the saved LDA models from previously:
-```{r}
-ldamodel <- readRDS(file.path(vignette_files, "paper_ldamodel.RDS"))
-ldats_ldamodel <- readRDS(file.path(vignette_files, "ldats_ldamodel.RDS"))
+
+```r
+ldamodel <- readRDS(file.path(vignette_files, "output", "paper_ldamodel.RDS"))
+ldats_ldamodel <- readRDS(file.path(vignette_files, "output", "ldats_ldamodel.RDS"))
 ```
@@
 How the paper LDA model assigns species to topics:
-```{r plot paper LDA, fig.width = 7, fig.height = 6}
+
+```r
 plot(ldamodel, cols = NULL, option = "D")
@@
 
+![plot of chunk plot-paper-LDA](figure/plot-paper-LDA-1.png)
+
 How the LDATS LDA model assigns species to topics:
-```{r plot LDATS LDA, fig.width = 7, fig.height = 6}
+
+```r
 plot(ldats_ldamodel, cols = NULL, option = "D")
 ```
+
+![plot of chunk plot-LDATS-LDA](figure/plot-LDATS-LDA-1.png)
 
@@
 
-```{r paper changepoint models}
+```r
 #### Run changepoint ####
-source(file.path(vignette_files, "changepointmodel.r"))
+source(file.path(vignette_files, "scripts", "changepointmodel.r"))
 
-find_changepoints <- function(lda_model, paper_covariates, n_changepoints = 1:6)
-{
+find_changepoints <- function(lda_model, paper_covariates, n_changepoints = 1:6){
   # set up parameters for model
@@
   cpt_results$cpt_model <- lapply(cpt_results$n_changepoints,
-                                  function(n_changepoints)
-                                  {
+                                  function(n_changepoints){
                                     changepoint_model(lda_model, x, n_changepoints, maxit = nit, 
@@
 #   - get the posterior distributions for the changepoints
-select_cpt_model <- function(cpt_results, ntopics)
-{
+select_cpt_model <- function(cpt_results, ntopics){
   # compute log likelihood as the mean deviance
@@
 #   dates from the original data
-get_dates <- function(cpt, covariates = paper_covariates)
-{
+get_dates <- function(cpt, covariates = paper_covariates){
   cpt$saved[,1,] %>%
@@
 Run the Christensen et al. time series model to identify changepoints on the LDA topics selected by LDATS:
-```{r run LDATS LDA and paper cpt, eval = params$run_models}
+
+```r
 ldats_paper_results <- find_changepoints(ldats_ldamodel, paper_covariates)
 
-saveRDS(ldats_paper_results, file = file.path(vignette_files, "ldats_paper.RDS"))
+saveRDS(ldats_paper_results, file = file.path(vignette_files, "output", "ldats_paper.RDS"))
 ```
@@
 Extract the dates of the changepoints:
-```{r compute changepoints for LDATS LDA and paper cpt}
-ldats_paper_results <- readRDS(file.path(vignette_files, "ldats_paper.RDS"))
+
+```r
+ldats_paper_results <- readRDS(file.path(vignette_files, "output", "ldats_paper.RDS"))
 
@@
 Run the Christensen et al. time series model to identify changepoints on the LDA topics selected by Christensen et al.:
-```{r run paper LDA and paper cpt, eval = params$run_models}
+
+```r
 paper_paper_results <- find_changepoints(ldamodel, paper_covariates)
@@
 Extract the dates of the changepoints:
-```{r compute changepoints for paper LDA and paper cpt}
-paper_paper_results <- readRDS(file.path(vignette_files, "paper_paper.RDS"))
+
+```r
+paper_paper_results <- readRDS(file.path(vignette_files, "output", "paper_paper.RDS"))
 
@@
 
-```{r run LDATS LDA and LDATS cpt, eval = params$run_models}
+```r
 ldats_ldats_results <- TS_on_LDA(LDA_models = ldats_ldamodel,
@@
 
-saveRDS(ldats_ldats_results, file = file.path(vignette_files, "ldats_ldats.RDS"))
+saveRDS(ldats_ldats_results, file = file.path(vignette_files, "output", "ldats_ldats.RDS"))
 ```
@@
 
-```{r construct lookup table for LDATS output for changepoint times}
+```r
 # make the full sequence of possible newmoon values
@@
 
-```{r compute changepoints for LDATS LDA and LDATS cpt}
-ldats_ldats_results <- readRDS(file.path(vignette_files, "ldats_ldats.RDS"))
+```r
+ldats_ldats_results <- readRDS(file.path(vignette_files, "output", "ldats_ldats.RDS"))
 
@@
 
-```{r run paper LDA and LDATS cpt, eval = params$run_models}
+```r
 paper_ldats_results <- TS_on_LDA(LDA_models = ldamodel,
@@
 
-saveRDS(paper_ldats_results, file = file.path(vignette_files, "paper_ldats.RDS"))
+saveRDS(paper_ldats_results, file = file.path(vignette_files, "output", "paper_ldats.RDS"))
 ```
@@
 
-```{r select paper lda + ldats cpt}
-paper_ldats_results <- readRDS(file.path(vignette_files, "paper_ldats.RDS"))
+```r
+paper_ldats_results <- readRDS(file.path(vignette_files, "output", "paper_ldats.RDS"))
 
@@
 
-```{r}
+```r
 nlevels(ldats_paper_cpt_dates$variable)
+#> [1] 4
 nlevels(paper_paper_cpt_dates$variable)
+#> [1] 4
 nlevels(ldats_ldats_cpt_dates$variable)
+#> [1] 4
 nlevels(paper_ldats_cpt_dates$variable)
+#> [1] 4
 ```
@@
 
-```{r plot paper LDA and LDATS cpts, fig.width = 7, fig.height = 6}
+```r
 plot(paper_ldats_cpt)
@@
 
+![plot of chunk plot paper LDA and LDATS cpts](figure/plot paper LDA and LDATS cpts-1.png)
+
 ### LDATS LDA and LDATS changepoint
-```{r plot ldats LDA and LDATS cpt, fig.width = 7, fig.height = 6}
+
+```r
 plot(ldats_ldats_cpt)
@@
 
+![plot of chunk plot ldats LDA and LDATS cpt](figure/plot ldats LDA and LDATS cpt-1.png)
+
+```r
+annual_hist <- readRDS(file.path(vignette_files, "output", "annual_hist.RDS"))
+```
+
 ### Paper LDA and paper changepoint
-```{r plot paper LDA and paper cpt, fig.width = 7, fig.height = 6}
+
+```r
 paper_cpts <- find_changepoint_location(paper_paper_cpt)
@@
 
-paper_cpt_plot <- get_ll_non_memoized_plot(ldamodel, paper_covariates, paper_cpts, make_plot = TRUE,
+png(file.path(vignette_files, "output", "paper_paper_cpt_plot.png"), width = 800, height = 600)
+get_ll_non_memoized_plot(ldamodel, paper_covariates, paper_cpts, make_plot = TRUE,
                                            weights = rep(1, NROW(paper_covariates)))
-
-annual_hist(paper_paper_cpt, paper_covariates$year_continuous)
-paper_cpt_plot
+dev.off()
 ```
 
+```r
+paper_paper_hist <- annual_hist(paper_paper_cpt, paper_covariates$year_continuous)
+```
+
+![plot of chunk plot paper LDA and LDATS cpts2](figure/plot paper LDA and LDATS cpts2-1.png)
+
+```r
+knitr::include_graphics(file.path(vignette_files, "output", "paper_paper_cpt_plot.png"))
+```
+
+![plot of chunk unnamed-chunk-19](./output/paper_paper_cpt_plot.png)
+
 ### LDATS LDA and paper changepoint
-```{r plot LDATS lda and paper cpt, fig.width = 7, fig.height = 6}
+
+```r
 ldats_cpts <- find_changepoint_location(ldats_paper_cpt)
@@
 
-ldats_cpt_plot <- get_ll_non_memoized_plot(ldats_ldamodel, paper_covariates, ldats_cpts, make_plot = TRUE,
+png(file.path(vignette_files, "output", "ldats_paper_cpt_plot.png"), width = 800, height = 600)
+get_ll_non_memoized_plot(ldats_ldamodel, paper_covariates, ldats_cpts, make_plot = TRUE,
                                            weights = rep(1, NROW(paper_covariates)))
+dev.off()
+```
 
-annual_hist(ldats_paper_cpt, paper_covariates$year_continuous)
-ldats_cpt_plot
+```r
+ldats_paper_hist <- annual_hist(ldats_paper_cpt, paper_covariates$year_continuous)
 ```
+
+![plot of chunk plot LDATS lda and paper cpt2](figure/plot LDATS lda and paper cpt2-1.png)
+
+```r
+knitr::include_graphics(file.path(vignette_files, "output", "ldats_paper_cpt_plot.png"))
+```
+
+![plot of chunk unnamed-chunk-20](./output/ldats_paper_cpt_plot.png)
 
@@
 ## Report changepoint dates
-```{r report cpt dates, include = F}
-cpt_dates <- dplyr::bind_rows("paperLDA_paperCPT" = paper_paper_cpt_dates, 
-                              "ldatsLDA_paperCPT" = ldats_paper_cpt_dates, 
-                              "ldatsLDA_ldatsCPT" = ldats_ldats_cpt_dates, 
-                              "paperLDA_ldatsCPT" = paper_ldats_cpt_dates, 
-                              .id = "analysis") %>%
-  dplyr::group_by(analysis, variable) %>%
-  dplyr::summarize(date = mean(date)) %>%
-  dplyr::ungroup() %>%
-  dplyr::rename(changepoint = variable) %>%
-  tidyr::spread(analysis, date)
-```
 
-```{r print cpt dates}
+```r
 knitr::kable(cpt_dates)
@@
 
+|changepoint |ldatsLDA_ldatsCPT |ldatsLDA_paperCPT |paperLDA_ldatsCPT |paperLDA_paperCPT |
+|:-----------|:-----------------|:-----------------|:-----------------|:-----------------|
+|V1          |1984-01-30        |1984-01-11        |1984-04-27        |1984-01-11        |
+|V2          |1991-05-30        |1991-05-29        |1993-04-19        |1991-05-29        |
+|V3          |1999-01-25        |1998-12-21        |1999-07-07        |1998-12-21        |
+|V4          |2009-11-15        |2009-09-29        |2010-01-20        |2009-09-29        |
+
 The choice of LDA has more influence on the changepoint locations than the choice of changepoint model - probably because the LDATS LDA has 6 topics, and the paper LDA has 4.

### 2023-09-16 doc/paper-comparison.Rmd
l

@@
 vignette: >
-  %\VignetteIndexEntry{paper-comparison}
+  %\VignetteIndexEntry{Comparison to Christensen et al. 2018}
   %\VignetteEngine{knitr::rmarkdown}
@@
   run_models: FALSE
+  rmarkdown.html_vignette.check_title: FALSE
 ---
@@
 
-This document provides a side-by-side comparison of **`LDATS`** (version 0.3.0) results with analysis from [Christensen et al.
+This document provides a side-by-side comparison of **`LDATS`** (version 0.2.7) results with analysis from [Christensen et al.
 2018](https://esajournals.onlinelibrary.wiley.com/doi/full/10.1002/ecy.2373).
@@
 
-![plot of chunk plot-paper-LDA](figure/plot-paper-LDA-1.png)
+![plot of chunk plot_paper_LDA](figure/plot_paper_LDA-1.png)
 
@@
 
-![plot of chunk plot-LDATS-LDA](figure/plot-LDATS-LDA-1.png)
+![plot of chunk plot_LDATS_LDA](figure/plot_LDATS_LDA-1.png)
 
@@
 
-![plot of chunk plot paper LDA and LDATS cpts](figure/plot paper LDA and LDATS cpts-1.png)
+![plot of chunk plot_paper_LDA_and_LDATS_cpts](figure/plot_paper_LDA_and_LDATS_cpts-1.png)
 
@@
 
-![plot of chunk plot ldats LDA and LDATS cpt](figure/plot ldats LDA and LDATS cpt-1.png)
+![plot of chunk plot_ldats_LDA_and_LDATS_cpt](figure/plot_ldats_LDA_and_LDATS_cpt-1.png)
 
@@
 
-![plot of chunk plot paper LDA and LDATS cpts2](figure/plot paper LDA and LDATS cpts2-1.png)
+![plot of chunk plot_paper_LDA_and_LDATS_cpts2](figure/plot_paper_LDA_and_LDATS_cpts2-1.png)
 
@@
 
-![plot of chunk plot LDATS lda and paper cpt2](figure/plot LDATS lda and paper cpt2-1.png)
+![plot of chunk plot_LDATS_lda_and_paper_cpt2](figure/plot_LDATS_lda_and_paper_cpt2-1.png)
 

### 2023-09-16 doc/paper-comparison.R

@@ -1,4 +1,4 @@
 params <-
-list(run_models = FALSE)
+list(run_models = FALSE, rmarkdown.html_vignette.check_title = FALSE)
 
 

### 2023-09-16 paper_comparison/paper-comparison.Rmd
temporary move of paper comparison
move vignette out of folder and rbuildignore it

--- 2023-09-16 doc/paper-comparison.Rmd
@@
   run_models: FALSE
-  rmarkdown.html_vignette.check_title: FALSE
 ---
@@
 
-![plot of chunk unnamed-chunk-10](./output/lda_distances.png)
+<div class="figure">
+<img src="./output/lda_distances.png" alt="plot of chunk unnamed-chunk-10" width="1067" />
+<p class="caption">plot of chunk unnamed-chunk-10</p>
+</div>
 
@@
 
-![plot of chunk plot_paper_LDA](figure/plot_paper_LDA-1.png)
+![plot of chunk plot_paper_LDA](plot_paper_LDA-1.png)
 
@@
 
-![plot of chunk plot_LDATS_LDA](figure/plot_LDATS_LDA-1.png)
+![plot of chunk plot_LDATS_LDA](plot_LDATS_LDA-1.png)
 
@@
 
-![plot of chunk plot_paper_LDA_and_LDATS_cpts](figure/plot_paper_LDA_and_LDATS_cpts-1.png)
+![plot of chunk plot_paper_LDA_and_LDATS_cpts](plot_paper_LDA_and_LDATS_cpts-1.png)
 
@@
 
-![plot of chunk plot_ldats_LDA_and_LDATS_cpt](figure/plot_ldats_LDA_and_LDATS_cpt-1.png)
+![plot of chunk plot_ldats_LDA_and_LDATS_cpt](plot_ldats_LDA_and_LDATS_cpt-1.png)
 
@@
 
-![plot of chunk plot_paper_LDA_and_LDATS_cpts2](figure/plot_paper_LDA_and_LDATS_cpts2-1.png)
+![plot of chunk plot_paper_LDA_and_LDATS_cpts2](plot_paper_LDA_and_LDATS_cpts2-1.png)
 
@@
 
-![plot of chunk unnamed-chunk-19](./output/paper_paper_cpt_plot.png)
+<div class="figure">
+<img src="./output/paper_paper_cpt_plot.png" alt="plot of chunk unnamed-chunk-19" width="1067" />
+<p class="caption">plot of chunk unnamed-chunk-19</p>
+</div>
 
@@
 
-![plot of chunk plot_LDATS_lda_and_paper_cpt2](figure/plot_LDATS_lda_and_paper_cpt2-1.png)
+![plot of chunk plot_LDATS_lda_and_paper_cpt2](plot_LDATS_lda_and_paper_cpt2-1.png)
 
@@
 
-![plot of chunk unnamed-chunk-20](./output/ldats_paper_cpt_plot.png)
+<div class="figure">
+<img src="./output/ldats_paper_cpt_plot.png" alt="plot of chunk unnamed-chunk-20" width="1067" />
+<p class="caption">plot of chunk unnamed-chunk-20</p>
+</div>
 

### 2023-09-16 doc/paper-comparison.R

@@ -1,4 +1,0 @@
-params <-
-list(run_models = FALSE)
-
-

### 2023-09-16 vignettes/scripts/rodent_data_for_LDA.r

@@ -1,51 +1,0 @@
-library(RCurl)
-library(dplyr)
-
-
-#' @title Create rodent species table
-#'
-#' @description Processes rodent capture data so it can be used for LDA analysis
-#' 
-#'
-#' @param period_first first trapping period desired
-#' @param period_last last trapping period desired
-#' @param selected_plots vector of plot numbers to be included
-#' @param selected_species vector of species codes to be included
-#' 
-#' @return Table of species counts per period
-#' 
-#' @example
-#' r_table = create_rodent_table(1,436,c(2,4,8,11,12,14,17,22),
-#                              selected_species = c('BA','DM','DO','DS','NA','OL','OT','PB','PE','PF','PH','PI','PL','PM','PP','RF','RM','RO','SF','SH','SO'))
-#'
-
-create_rodent_table = function(period_first,period_last,selected_plots,selected_species) {
-  
-  # retrieve current version of rodent data
-  rodents = read.csv(text=getURL("https://raw.githubusercontent.com/weecology/PortalData/master/Rodents/Portal_rodent.csv"),
-                     na.strings=c(""), colClasses=c('tag'='character'), stringsAsFactors = FALSE)
-  
-  # extract desired data by period, plot, and species
-  rod = filter(rodents, period>=period_first, period<=period_last,
-               plot %in% selected_plots,
-               species %in% selected_species)
-  
-  # create table of species counts by period
-  r_table = table(rod$period,rod$species)
-  
-  # retrieve data on number of plots trapped per month
-  trap_table = read.csv('https://raw.githubusercontent.com/weecology/PortalData/master/Rodents/Portal_rodent_trapping.csv')
-  trap_table_controls = filter(trap_table, plot %in% selected_plots)
-  nplots_controls = aggregate(trap_table_controls$sampled,by=list(period = trap_table_controls$period),FUN=sum)
-  
-  # adjust species counts by number of plots trapped that month
-  r_table_adjusted = as.data.frame.matrix(r_table)
-  for (n in 1:436) {
-    #divide by number of control plots actually trapped (should be 8) and multiply by 8 to estimate captures as if all plots were trapped
-    r_table_adjusted[n,] = round(r_table_adjusted[n,]/nplots_controls$x[n]*8)
-  }
-  
-  #write.table(r_table_adjusted,'Rodent_table_dat.csv',sep=',',row.names = F)
-  return(r_table_adjusted)
-}
-

### 2023-09-17 paper_comparison/precompile.R
precompile paper comparison vignette
@diazrenata this is my attempt to get the precompiling approach to work following https://ropensci.org/blog/2019/12/08/precompute-vignettes/.

the vignette needs to be pre built by collowing the precompile.R script in the paper_comparison folder, which is ignored for the R build

@@ -1,12 +1,37 @@
-# Vignettes that depend on internet access have been precompiled:
+# Vignettes that depend on internet access are precompiled following
+#    https://ropensci.org/blog/2019/12/08/precompute-vignettes/.
+
+# temporarily change working directory 
 
 wd <- getwd()
-setwd("vignettes/")
+setwd("paper_comparison/")
 
-knitr::knit("paper-comparison.Rmd.remote_needs", output = "paper-comparison.Rmd")
+knitr::knit(input  = "paper-comparison.Rmd.remote_needs", 
+            output = "../vignettes/paper-comparison.Rmd")
+
+
+# Move the output figure files to the vignettes folder
+
+from_files <- list.files(path       = "output", 
+                         full.names = TRUE)[grep(".png", list.files("output"))]
+
+file.copy(from      = from_files,
+          to        = file.path("..", "vignettes", "output"),
+          recursive = TRUE,
+          overwrite = TRUE)
+
+file.copy(from      = file.path("figure"),
+          to        = file.path("..", "vignettes"),
+          recursive = TRUE,
+          overwrite = TRUE)
+
+
+# revert the working directory
+
 setwd(wd)
+
 
 # Build checking
 
-devtools::build_vignettes( )
+# devtools::build_vignettes( )
 

### 2023-09-17 paper_comparison/scripts/rodent_LDA_analysis.r

@@ -1,0 +1,508 @@
+# LDA and changepoint analysis pipeline on rodent data -- uses VEM method
+#
+#  1. prepare data
+#  2a. run LDA with different number of topics and use AIC to select best model
+#  2b. how much does seed choice affect species composition of component communities?
+#  3. run LDA with best number of topics (determined by 2a)
+#  4. run changepoint model
+#  5. produce figures
+#  6. appendix figures
+
+
+library(topicmodels)
+library(RCurl)
+library(multipanelfigure)
+library(reshape2)
+
+
+source('rodent_data_for_LDA.r')
+source('AIC_model_selection.R')
+source('LDA_figure_scripts.R')
+source('changepointmodel.r')
+source('LDA-distance.R')
+
+# ===================================================================
+# 1. prepare rodent data
+# ===================================================================
+dat = create_rodent_table(period_first = 1,
+                          period_last = 436,
+                          selected_plots = c(2,4,8,11,12,14,17,22),
+                          selected_species = c('BA','DM','DO','DS','NA','OL','OT','PB','PE','PF','PH','PI','PL','PM','PP','RF','RM','RO','SF','SH','SO'))
+
+# dates to go with count data
+moondat = read.csv(text=getURL("https://raw.githubusercontent.com/weecology/PortalData/master/Rodents/moon_dates.csv"),stringsAsFactors = F)
+moondat$date = as.Date(moondat$censusdate)
+
+period_dates = filter(moondat,period %in% rownames(dat)) %>% select(period,date)
+dates = period_dates$date
+
+
+# ==================================================================
+# 2a. select number of topics
+# ==================================================================
+
+# Fit a bunch of LDA models with different seeds
+# Only use even numbers for seeds because consecutive seeds give identical results
+seeds = 2*seq(200)
+
+# repeat LDA model fit and AIC calculation with a bunch of different seeds to test robustness of the analysis
+best_ntopic = repeat_VEM(dat,
+                         seeds,
[…]

### 2023-09-17 doc/LDATS_codebase.Rmd

--- 2019-07-04 doc/LDATS_codebase.Rmd
@@
 ---
-title: "Latent Dirichlet Allocation Time Series (LDATS)"
+title: "LDATS Codebase"
 author: "Juniper L. Simonis"
@@
 vignette: >
-  %\VignetteIndexEntry{LDATScodebase}
+  %\VignetteIndexEntry{LDATS Codebase}
   %\VignetteEngine{knitr::rmarkdown}
@@
 throughout the pipeline, we implement an options/controls list approach, where
-each of the main functions (`LDA_TS`, `LDA_set`, and `TS`) has a controls 
-list (although `LDA_set` is referenced as `LDA` in the naming scheme)
+each of the main functions (`LDA_TS`, `LDA_set`, and `TS`) and its subfunctions
+have a `control` argument that takes a `list` to replace the defaults returned 
+by its `_control` function:
 
-* `LDA_TS_controls_list()`
-* `LDA_controls_list()`
-* `TS_controls_list()`
+* `LDA_TS_control()`
+* `LDA_set_control()`
+* `TS_control()`
